### 2015-07-02 paper.tex
Initial template files

% THIS IS SIGPROC-SP.TEX - VERSION 3.1
% WORKS WITH V3.2SP OF ACM_PROC_ARTICLE-SP.CLS
% APRIL 2009
%
% It is an example file showing how to use the 'acm_proc_article-sp.cls' V3.2SP
% LaTeX2e document class file for Conference Proceedings submissions.
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V3.2SP) *DOES NOT* produce:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) Page numbering
% ---------------------------------------------------------------------------------------------------------------
% It is an example which *does* use the .bib file (from which the .bbl file
% is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission,
% you need to 'insert'  your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% Questions regarding SIGS should be sent to
% Adrienne Griscti ---> griscti@acm.org
%
% Questions/suggestions regarding the guidelines, .tex and .cls files, etc. to
% Gerald Murray ---> murray@hq.acm.org
%
% For tracking purposes - this is V3.1SP - APRIL 2009

\documentclass{acm_proc_article-sp}

\begin{document}

\title{A Sample {\ttlit ACM} SIG Proceedings Paper in LaTeX
Format\titlenote{(Does NOT produce the permission block, copyright information nor page numbering). For use with ACM\_PROC\_ARTICLE-SP.CLS. Supported by ACM.}}
\subtitle{[Extended Abstract]
\titlenote{A full version of this paper is available as
\textit{Author's Guide to Preparing ACM SIG Proceedings Using
\LaTeX$2_\epsilon$\ and BibTeX} at
\texttt{www.acm.org/eaddress.htm}}}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{8} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Ben Trovato\titlenote{Dr.~Trovato insisted his name be first.}\\
       \affaddr{Institute for Clarity in Documentation}\\
       \affaddr{1932 Wallamaloo Lane}\\
       \affaddr{Wallamaloo, New Zealand}\\
       \email{trovato@corporation.com}
% 2nd. author
\alignauthor
G.K.M. Tobin\titlenote{The secretary disavows
any knowledge of this author's actions.}\\
       \affaddr{Institute for Clarity in Documentation}\\
       \affaddr{P.O. Box 1212}\\
       \affaddr{Dublin, Ohio 43017-6221}\\
       \email{webmaster@marysville-ohio.com}
% 3rd. author
\alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
one who did all the really hard work.}\\
       \affaddr{The Th{\o}rv{\"a}ld Group}\\
       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
       \affaddr{Hekla, Iceland}\\
       \email{larst@affiliation.org}
\and  % use '\and' if you need 'another row' of author names
% 4th. author
\alignauthor Lawrence P. Leipuner\\
       \affaddr{Brookhaven Laboratories}\\
       \affaddr{Brookhaven National Lab}\\
       \affaddr{P.O. Box 5000}\\
       \email{lleipuner@researchlabs.org}
% 5th. author
\alignauthor Sean Fogarty\\
       \affaddr{NASA Ames Research Center}\\
       \affaddr{Moffett Field}\\
       \affaddr{California 94035}\\
       \email{fogartys@amesres.org}
% 6th. author
\alignauthor Charles Palmer\\
       \affaddr{Palmer Research Laboratories}\\
       \affaddr{8600 Datapoint Drive}\\
       \affaddr{San Antonio, Texas 78229}\\
       \email{cpalmer@prl.com}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle
\begin{abstract}
This paper provides a sample of a \LaTeX\ document which conforms to
the formatting guidelines for ACM SIG Proceedings.
It complements the document \textit{Author's Guide to Preparing
ACM SIG Proceedings Using \LaTeX$2_\epsilon$\ and Bib\TeX}. This
source file has been written with the intention of being
compiled under \LaTeX$2_\epsilon$\ and BibTeX.

The developers have tried to include every imaginable sort
of ``bells and whistles", such as a subtitle, footnotes on
title, subtitle and authors, as well as in the text, and
every optional component (e.g. Acknowledgments, Additional
Authors, Appendices), not to mention examples of
equations, theorems, tables and figures.

To make best use of this sample document, run it through \LaTeX\
and BibTeX, and compare this source code with the printed
output produced by the dvi file.
\end{abstract}

% A category with the (minimum) three required fields
\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

\terms{Theory}

\keywords{ACM proceedings, \LaTeX, text tagging} % NOT required for Proceedings

\section{Introduction}
The \textit{proceedings} are the records of a conference.
ACM seeks to give these conference by-products a uniform,
high-quality appearance.  To do this, ACM has some rigid
requirements for the format of the proceedings documents: there
is a specified format (balanced  double columns), a specified
set of fonts (Arial or Helvetica and Times Roman) in
certain specified sizes (for instance, 9 point for body copy),
a specified live area (18 $\times$ 23.5 cm [7" $\times$ 9.25"]) centered on
the page, specified size of margins (1.9 cm [0.75"]) top, (2.54 cm [1"]) bottom
and (1.9 cm [.75"]) left and right; specified column width
(8.45 cm [3.33"]) and gutter size (.83 cm [.33"]).

The good news is, with only a handful of manual
settings\footnote{Two of these, the {\texttt{\char'134 numberofauthors}}
and {\texttt{\char'134 alignauthor}} commands, you have
already used; another, {\texttt{\char'134 balancecolumns}}, will
be used in your very last run of \LaTeX\ to ensure
balanced column heights on the last page.}, the \LaTeX\ document
class file handles all of this for you.

The remainder of this document is concerned with showing, in
the context of an ``actual'' document, the \LaTeX\ commands
specifically available for denoting the structure of a
proceedings paper, rather than with giving rigorous descriptions
or explanations of such commands.

\section{The {\secit Body} of The Paper}
Typically, the body of a paper is organized
into a hierarchical structure, with numbered or unnumbered
headings for sections, subsections, sub-subsections, and even
smaller sections.  The command \texttt{{\char'134}section} that
precedes this paragraph is part of such a
hierarchy.\footnote{This is the second footnote.  It
starts a series of three footnotes that add nothing
informational, but just give an idea of how footnotes work
and look. It is a wordy one, just so you see
how a longish one plays out.} \LaTeX\ handles the numbering
and placement of these headings for you, when you use
the appropriate heading commands around the titles
of the headings.  If you want a sub-subsection or
smaller part to be unnumbered in your output, simply append an
asterisk to the command name.  Examples of both
numbered and unnumbered headings will appear throughout the
balance of this sample document.

Because the entire article is contained in
the \textbf{document} environment, you can indicate the
start of a new paragraph with a blank line in your
input file; that is why this sentence forms a separate paragraph.

\subsection{Type Changes and {\subsecit Special} Characters}
We have already seen several typeface changes in this sample.  You
can indicate italicized words or phrases in your text with
the command \texttt{{\char'134}textit}; emboldening with the
command \texttt{{\char'134}textbf}
and typewriter-style (for instance, for computer code) with
\texttt{{\char'134}texttt}.  But remember, you do not
have to indicate typestyle changes when such changes are
part of the \textit{structural} elements of your
article; for instance, the heading of this subsection will
be in a sans serif\footnote{A third footnote, here.
Let's make this a rather short one to
see how it looks.} typeface, but that is handled by the
document class file. Take care with the use
of\footnote{A fourth, and last, footnote.}
the curly braces in typeface changes; they mark
the beginning and end of
the text that is to be in the different typeface.

You can use whatever symbols, accented characters, or
non-English characters you need anywhere in your document;
you can find a complete list of what is
available in the \textit{\LaTeX\
User's Guide}\cite{Lamport:LaTeX}.

\subsection{Math Equations}
You may want to display math equations in three distinct styles:
inline, numbered or non-numbered display.  Each of
the three are discussed in the next sections.

\subsubsection{Inline (In-text) Equations}
A formula that appears in the running text is called an
inline or in-text formula.  It is produced by the
\textbf{math} environment, which can be
invoked with the usual \texttt{{\char'134}begin. . .{\char'134}end}
construction or with the short form \texttt{\$. . .\$}. You
can use any of the symbols and structures,
from $\alpha$ to $\omega$, available in
\LaTeX\cite{Lamport:LaTeX}; this section will simply show a
few examples of in-text equations in context. Notice how
this equation: \begin{math}\lim_{n\rightarrow \infty}x=0\end{math},
set here in in-line math style, looks slightly different when
set in display style.  (See next section).

\subsubsection{Display Equations}
A numbered display equation -- one set off by vertical space
from the text and centered horizontally -- is produced
by the \textbf{equation} environment. An unnumbered display
equation is produced by the \textbf{displaymath} environment.

Again, in either environment, you can use any of the symbols
and structures available in \LaTeX; this section will just
give a couple of examples of display equations in context.
First, consider the equation, shown as an inline equation above:
\begin{equation}\lim_{n\rightarrow \infty}x=0\end{equation}
Notice how it is formatted somewhat differently in
the \textbf{displaymath}
environment.  Now, we'll enter an unnumbered equation:
\begin{displaymath}\sum_{i=0}^{\infty} x + 1\end{displaymath}
and follow it with another numbered equation:
\begin{equation}\sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f\end{equation}
just to demonstrate \LaTeX's able handling of numbering.

\subsection{Citations}
Citations to articles \cite{bowman:reasoning, clark:pct, braams:babel, herlihy:methodology},
conference
proceedings \cite{clark:pct} or books \cite{salas:calculus, Lamport:LaTeX} listed
in the Bibliography section of your
article will occur throughout the text of your article.
You should use BibTeX to automatically produce this bibliography;
you simply need to insert one of several citation commands with
a key of the item cited in the proper location in
the \texttt{.tex} file \cite{Lamport:LaTeX}.
The key is a short reference you invent to uniquely
identify each work; in this sample document, the key is
the first author's surname and a
word from the title.  This identifying key is included
with each item in the \texttt{.bib} file for your article.

The details of the construction of the \texttt{.bib} file
are beyond the scope of this sample document, but more
information can be found in the \textit{Author's Guide},
and exhaustive details in the \textit{\LaTeX\ User's
Guide}\cite{Lamport:LaTeX}.

This article shows only the plainest form
of the citation command, using \texttt{{\char'134}cite}.
This is what is stipulated in the SIGS style specifications.
No other citation format is endorsed.

\subsection{Tables}
Because tables cannot be split across pages, the best
placement for them is typically the top of the page
nearest their initial cite.  To
ensure this proper ``floating'' placement of tables, use the
environment \textbf{table} to enclose the table's contents and
the table caption.  The contents of the table itself must go
in the \textbf{tabular} environment, to
be aligned properly in rows and columns, with the desired
horizontal and vertical rules.  Again, detailed instructions
on \textbf{tabular} material
is found in the \textit{\LaTeX\ User's Guide}.

Immediately following this sentence is the point at which
Table 1 is included in the input file; compare the
placement of the table here with the table in the printed
dvi output of this document.

\begin{table}
\centering
\caption{Frequency of Special Characters}
\begin{tabular}{|c|c|l|} \hline
Non-English or Math&Frequency&Comments\\ \hline
\O & 1 in 1,000& For Swedish names\\ \hline
$\pi$ & 1 in 5& Common in math\\ \hline
\$ & 4 in 5 & Used in business\\ \hline
$\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
\hline\end{tabular}
\end{table}

To set a wider table, which takes up the whole width of
the page's live area, use the environment
\textbf{table*} to enclose the table's contents and
the table caption.  As with a single-column table, this wide
table will ``float" to a location deemed more desirable.
Immediately following this sentence is the point at which
Table 2 is included in the input file; again, it is
instructive to compare the placement of the
table here with the table in the printed dvi
output of this document.

\begin{table*}
\centering
\caption{Some Typical Commands}
\begin{tabular}{|c|c|l|} \hline
Command&A Number&Comments\\ \hline
\texttt{{\char'134}alignauthor} & 100& Author alignment\\ \hline
\texttt{{\char'134}numberofauthors}& 200& Author enumeration\\ \hline
\texttt{{\char'134}table}& 300 & For tables\\ \hline
\texttt{{\char'134}table*}& 400& For wider tables\\ \hline\end{tabular}
\end{table*}
% end the environment with {table*}, NOTE not {table}!

\subsection{Figures}
Like tables, figures cannot be split across pages; the
best placement for them
is typically the top or the bottom of the page nearest
their initial cite.  To ensure this proper ``floating'' placement
of figures, use the environment
\textbf{figure} to enclose the figure and its caption.

This sample document contains examples of \textbf{.eps}
and \textbf{.ps} files to be displayable with \LaTeX.  More
details on each of these is found in the \textit{Author's Guide}.

\begin{figure}
\centering
\epsfig{file=fly.eps}
\caption{A sample black and white graphic (.eps format).}
\end{figure}

\begin{figure}
\centering
\epsfig{file=fly.eps, height=1in, width=1in}
\caption{A sample black and white graphic (.eps format)
that has been resized with the \texttt{epsfig} command.}
\end{figure}

As was the case with tables, you may want a figure
that spans two columns.  To do this, and still to
ensure proper ``floating'' placement of tables, use the environment
\textbf{figure*} to enclose the figure and its caption.

Note that either {\textbf{.ps}} or {\textbf{.eps}} formats are
used; use
the \texttt{{\char'134}epsfig} or \texttt{{\char'134}psfig}
commands as appropriate for the different file types.

\subsection{Theorem-like Constructs}
Other common constructs that may occur in your article are
the forms for logical constructs like theorems, axioms,
corollaries and proofs.  There are
two forms, one produced by the
command \texttt{{\char'134}newtheorem} and the
other by the command \texttt{{\char'134}newdef}; perhaps
the clearest and easiest way to distinguish them is
to compare the two in the output of this sample document:

This uses the \textbf{theorem} environment, created by
the\linebreak\texttt{{\char'134}newtheorem} command:
\newtheorem{theorem}{Theorem}
\begin{theorem}
Let $f$ be continuous on $[a,b]$.  If $G$ is
an antiderivative for $f$ on $[a,b]$, then
\begin{displaymath}\int^b_af(t)dt = G(b) - G(a).\end{displaymath}
\end{theorem}

The other uses the \textbf{definition} environment, created
by the \texttt{{\char'134}newdef} command:
\newdef{definition}{Definition}
\begin{definition}
If $z$ is irrational, then by $e^z$ we mean the
unique number which has
logarithm $z$: \begin{displaymath}{\log e^z = z}\end{displaymath}
\end{definition}

Two lists of constructs that use one of these
forms is given in the
\textit{Author's  Guidelines}.

\begin{figure*}
\centering
\epsfig{file=flies.eps}
\caption{A sample black and white graphic (.eps format)
that needs to span two columns of text.}
\end{figure*}
and don't forget to end the environment with
{figure*}, not {figure}!

There is one other similar construct environment, which is
already set up
for you; i.e. you must \textit{not} use
a \texttt{{\char'134}newdef} command to
create it: the \textbf{proof} environment.  Here
is a example of its use:
\begin{proof}
Suppose on the contrary there exists a real number $L$ such that
\begin{displaymath}
\lim_{x\rightarrow\infty} \frac{f(x)}{g(x)} = L.
\end{displaymath}
Then
\begin{displaymath}
l=\lim_{x\rightarrow c} f(x)
= \lim_{x\rightarrow c}
\left[ g{x} \cdot \frac{f(x)}{g(x)} \right ]
= \lim_{x\rightarrow c} g(x) \cdot \lim_{x\rightarrow c}
\frac{f(x)}{g(x)} = 0\cdot L = 0,
\end{displaymath}
which contradicts our assumption that $l\neq 0$.
\end{proof}

Complete rules about using these environments and using the
two different creation commands are in the
\textit{Author's Guide}; please consult it for more
detailed instructions.  If you need to use another construct,
not listed therein, which you want to have the same
formatting as the Theorem
or the Definition\cite{salas:calculus} shown above,
use the \texttt{{\char'134}newtheorem} or the
\texttt{{\char'134}newdef} command,
respectively, to create it.

\subsection*{A {\secit Caveat} for the \TeX\ Expert}
Because you have just been given permission to
use the \texttt{{\char'134}newdef} command to create a
new form, you might think you can
use \TeX's \texttt{{\char'134}def} to create a
new command: \textit{Please refrain from doing this!}
Remember that your \LaTeX\ source code is primarily intended
to create camera-ready copy, but may be converted
to other forms -- e.g. HTML. If you inadvertently omit
some or all of the \texttt{{\char'134}def}s recompilation will
be, to say the least, problematic.

\section{Conclusions}
This paragraph will end the body of this sample document.
Remember that you might still have Acknowledgments or
Appendices; brief samples of these
follow.  There is still the Bibliography to deal with; and
we will make a disclaimer about that here: with the exception
of the reference to the \LaTeX\ book, the citations in
this paper are to articles which have nothing to
do with the present subject and are used as
examples only.
%\end{document}  % This is where a 'short' article might terminate

%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}
This section is optional; it is a location for you
to acknowledge grants, funding, editing assistance and
what have you.  In the present case, for example, the
authors would like to thank Gerald Murray of ACM for
his help in codifying this \textit{Author's Guide}
and the \textbf{.cls} and \textbf{.tex} files that it describes.

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{dateparse}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns
\appendix
%Appendix A
\section{Headings in Appendices}
The rules about hierarchical headings discussed above for
the body of the article are different in the appendices.
In the \textbf{appendix} environment, the command
\textbf{section} is used to
indicate the start of each Appendix, with alphabetic order
designation (i.e. the first is A, the second B, etc.) and
a title (if you include one).  So, if you need
hierarchical structure
\textit{within} an Appendix, start with \textbf{subsection} as the
highest level. Here is an outline of the body of this
document in Appendix-appropriate form:
\subsection{Introduction}
\subsection{The Body of the Paper}
\subsubsection{Type Changes and  Special Characters}
\subsubsection{Math Equations}
\paragraph{Inline (In-text) Equations}
\paragraph{Display Equations}
\subsubsection{Citations}
\subsubsection{Tables}
\subsubsection{Figures}
\subsubsection{Theorem-like Constructs}
\subsubsection*{A Caveat for the \TeX\ Expert}
\subsection{Conclusions}
\subsection{Acknowledgments}
\subsection{Additional Authors}
This section is inserted by \LaTeX; you do not insert it.
You just add the names and information in the
\texttt{{\char'134}additionalauthors} command at the start
of the document.
\subsection{References}
Generated by bibtex from your ~.bib file.  Run latex,
then bibtex, then latex twice (to resolve references)
to create the ~.bbl file.  Insert that ~.bbl file into
the .tex source file and comment out
the command \texttt{{\char'134}thebibliography}.
% This next section command marks the start of
% Appendix B, and does not continue the present hierarchy
\section{More Help for the Hardy}
The acm\_proc\_article-sp document class file itself is chock-full of succinct
and helpful comments.  If you consider yourself a moderately
experienced to expert user of \LaTeX, you may find reading
it useful but please remember not to change it.
\balancecolumns
% That's all folks!
\end{document}

### 2015-07-02 paper.tex
added title, authors and abstract

@@
 
-\title{A Sample {\ttlit ACM} SIG Proceedings Paper in LaTeX
-Format\titlenote{(Does NOT produce the permission block, copyright information nor page numbering). For use with ACM\_PROC\_ARTICLE-SP.CLS. Supported by ACM.}}
-\subtitle{[Extended Abstract]
-\titlenote{A full version of this paper is available as
-\textit{Author's Guide to Preparing ACM SIG Proceedings Using
-\LaTeX$2_\epsilon$\ and BibTeX} at
-\texttt{www.acm.org/eaddress.htm}}}
+\title{Does Anybody Really Know What Time It Is?}
+\subtitle{Automating the Extraction of Date Scalars}
 %
@@
 
-\numberofauthors{8} %  in this sample file, there are a *total*
+\numberofauthors{3} %  in this sample file, there are a *total*
 % of EIGHT authors. SIX appear on the 'first-page' (for formatting
@@
 \alignauthor
-Ben Trovato\titlenote{Dr.~Trovato insisted his name be first.}\\
-       \affaddr{Institute for Clarity in Documentation}\\
-       \affaddr{1932 Wallamaloo Lane}\\
-       \affaddr{Wallamaloo, New Zealand}\\
-       \email{trovato@corporation.com}
+Richard Wesley \\
+%       \affaddr{Institute for Clarity in Documentation}\\
+%       \affaddr{1932 Wallamaloo Lane}\\
+%       \affaddr{Wallamaloo, New Zealand}\\
+%       \email{trovato@corporation.com}
 % 2nd. author
 \alignauthor
-G.K.M. Tobin\titlenote{The secretary disavows
-any knowledge of this author's actions.}\\
-       \affaddr{Institute for Clarity in Documentation}\\
-       \affaddr{P.O. Box 1212}\\
-       \affaddr{Dublin, Ohio 43017-6221}\\
-       \email{webmaster@marysville-ohio.com}
+Vidya Setlur\\
+%       \affaddr{Institute for Clarity in Documentation}\\
+%       \affaddr{P.O. Box 1212}\\
+%       \affaddr{Dublin, Ohio 43017-6221}\\
+%       \email{webmaster@marysville-ohio.com}
 % 3rd. author
-\alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
-one who did all the really hard work.}\\
-       \affaddr{The Th{\o}rv{\"a}ld Group}\\
-       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
-       \affaddr{Hekla, Iceland}\\
-       \email{larst@affiliation.org}
-\and  % use '\and' if you need 'another row' of author names
-% 4th. author
-\alignauthor Lawrence P. Leipuner\\
-       \affaddr{Brookhaven Laboratories}\\
-       \affaddr{Brookhaven National Lab}\\
-       \affaddr{P.O. Box 5000}\\
-       \email{lleipuner@researchlabs.org}
-% 5th. author
-\alignauthor Sean Fogarty\\
-       \affaddr{NASA Ames Research Center}\\
-       \affaddr{Moffett Field}\\
-       \affaddr{California 94035}\\
-       \email{fogartys@amesres.org}
-% 6th. author
-\alignauthor Charles Palmer\\
-       \affaddr{Palmer Research Laboratories}\\
-       \affaddr{8600 Datapoint Drive}\\
-       \affaddr{San Antonio, Texas 78229}\\
-       \email{cpalmer@prl.com}
+\alignauthor
+Dan Cory\\
+%       \affaddr{The Th{\o}rv{\"a}ld Group}\\
+%       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
+%       \affaddr{Hekla, Iceland}\\
+%       \email{larst@affiliation.org}
 }
@@
 \begin{abstract}
-This paper provides a sample of a \LaTeX\ document which conforms to
-the formatting guidelines for ACM SIG Proceedings.
-It complements the document \textit{Author's Guide to Preparing
-ACM SIG Proceedings Using \LaTeX$2_\epsilon$\ and Bib\TeX}. This
-source file has been written with the intention of being
-compiled under \LaTeX$2_\epsilon$\ and BibTeX.
+With the advent of modern data visualization tools, data preparation has become a bottleneck for analytic workflows.
+Users who are already engaged in data analysis prefer to stay in the visualization environment for cleaning and preparation tasks whenever possible, both to preserve analytic flow and to take advantage of the visualization environment to spot inconsistent data.
+This has led many visualization environments to include simple data preparation functions such as scalar parsing, pattern matching and categorical binning in their analytic toolkits.
 
-The developers have tried to include every imaginable sort
-of ``bells and whistles", such as a subtitle, footnotes on
-title, subtitle and authors, as well as in the text, and
-every optional component (e.g. Acknowledgments, Additional
-Authors, Appendices), not to mention examples of
-equations, theorems, tables and figures.
+One of the most common parsing tasks is extracting date and time data from string representations.
+Several databases include date parsing “mini-languages” to cover the wide range of possible formats.
+Tableau has recently provided a DATEPARSE function with a single syntax that is translated into the nearest equivalent in the underlying database.
+Subsequent analysis of customer usage of this function shows that the parsing language syntax was difficult for users to master.
 
-To make best use of this sample document, run it through \LaTeX\
-and BibTeX, and compare this source code with the printed
-output produced by the dvi file.
+In this paper, we present two algorithms for automatically deriving date format strings from a column of data in our syntax, one based on minimum entropy and one based on natural language modeling.
+Both have similar accuracies of over 90\% on a large corpus of date columns extracted from Tableau Public and are in substantial agreement with each other.
+The minimal entropy approach can also produce results below the user’s perceptual threshold, making it suitable for interactive work.
 \end{abstract}

### 2015-07-02 intro.tex
added keywords, general terms and classification

In recent years, there has been a growth of interest in data visualization technologies for human-assisted data analysis using systems such as [1,10,11]. While computers can provide high-speed and high-volume data processing, humans have domain knowledge and the ability to process data in parallel, notably by using our visual systems. Most importantly, humans provide the definition of what is valuable in an analysis. Accordingly, human/computer analytic systems are essential to extracting knowledge of value to humans and their societies from the large amounts of data being generated today.

\subsection{Interactivity}
Visualization systems are most effective when they are interactive, thereby allowing a user to explore data and connect it to their domain knowledge and sense of what is important without breaking cognitive flow. In recent years, a number of such systems have been developed, both by the academic community and by the commercial sector. Exploration of data consists not only in creating visual displays, but also in creating and modifying domain-specific computations. Consequently, most data visualization systems include facilities for defining such calculations as part of the data model being analyzed. The most effective systems allow users to define these calculations as part of the analytic interaction, which permits the user to stay in the flow of analysis [9].

During the analytic process, a user may discover that parts of the data are not yet suitable for analysis. Solutions to this problem are often provided by data preparation tools external to the visual analysis environment, which requires the user to break their analytic flow, launch another tool and reprocess their data before returning to their analysis. If the user does not own this process (e.g. it is the responsibility of another department), then there can be significant delays (including ``never.”) More subtly, the result of updated external processing may not be compatible with the user’s existing work, which can lead to more time lost reconciling the new data model with the existing analysis.

From the user’s perspective, the boundary between preparation and analysis is not nearly so clean cut. Bad data is often discovered using visual analysis techniques (e.g. histograms or scatter plots) and it is most natural for the user to “clean what she sees” instead of switching to a second tool. This leads to an ``adaptive” process whereby users will prefer leveraging existing tools in the analytics environment (no matter how well suited to the task) over switching to another application. Thus a well-designed interactive visual analysis environment will provide tools that enable users to perform such cleaning tasks as interactively as possible.

### 2015-07-02 paper.tex
added keywords, general terms and classification

@@
 One of the most common parsing tasks is extracting date and time data from string representations.
-Several databases include date parsing “mini-languages” to cover the wide range of possible formats.
-Tableau has recently provided a DATEPARSE function with a single syntax that is translated into the nearest equivalent in the underlying database.
+Several databases include date parsing ``mini-languages" to cover the wide range of possible formats.
+Tableau has recently provided a \\DATEPARSE function with a single syntax that is translated into the nearest equivalent in the underlying database.
 Subsequent analysis of customer usage of this function shows that the parsing language syntax was difficult for users to master.
@@
 Both have similar accuracies of over 90\% on a large corpus of date columns extracted from Tableau Public and are in substantial agreement with each other.
-The minimal entropy approach can also produce results below the user’s perceptual threshold, making it suitable for interactive work.
+The minimal entropy approach can also produce results below the user's perceptual threshold, making it suitable for interactive work.
 \end{abstract}
 
+%D.3.3 [Data Processing]: Data Cleaning, Natural Language Processing.
 % A category with the (minimum) three required fields
-\category{H.4}{Information Systems Applications}{Miscellaneous}
+\category{D.3.3}{Data Processing}{Data Cleaning, Natural Language Processing}
 %A category including the fourth, optional field follows...
-\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
+%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
 
-\terms{Theory}
+\terms{Algorithms, Performance}
 
-\keywords{ACM proceedings, \LaTeX, text tagging} % NOT required for Proceedings
+\keywords{Date parsing, automated cleaning, structure extraction} % NOT required for Proceedings
 
 \section{Introduction}
-The \textit{proceedings} are the records of a conference.
-ACM seeks to give these conference by-products a uniform,
-high-quality appearance.  To do this, ACM has some rigid
-requirements for the format of the proceedings documents: there
-is a specified format (balanced  double columns), a specified
-set of fonts (Arial or Helvetica and Times Roman) in
-certain specified sizes (for instance, 9 point for body copy),
-a specified live area (18 $\times$ 23.5 cm [7" $\times$ 9.25"]) centered on
-the page, specified size of margins (1.9 cm [0.75"]) top, (2.54 cm [1"]) bottom
-and (1.9 cm [.75"]) left and right; specified column width
-(8.45 cm [3.33"]) and gutter size (.83 cm [.33"]).
-
-The good news is, with only a handful of manual
-settings\footnote{Two of these, the {\texttt{\char'134 numberofauthors}}
-and {\texttt{\char'134 alignauthor}} commands, you have
-already used; another, {\texttt{\char'134 balancecolumns}}, will
-be used in your very last run of \LaTeX\ to ensure
-balanced column heights on the last page.}, the \LaTeX\ document
-class file handles all of this for you.
-
-The remainder of this document is concerned with showing, in
-the context of an ``actual'' document, the \LaTeX\ commands
-specifically available for denoting the structure of a
-proceedings paper, rather than with giving rigorous descriptions
-or explanations of such commands.
+\input{intro}
 

### 2015-07-02 intro.tex
Quote fixing, added placeholder sections

@@
 If the user does not own this process (e.g.
-it is the responsibility of another department), then there can be significant delays (including ``never.”) More subtly, the result of updated external processing may not be compatible with the user’s existing work, which can lead to more time lost reconciling the new data model with the existing analysis.
+it is the responsibility of another department), then there can be significant delays (including ``never.") More subtly, the result of updated external processing may not be compatible with the user's existing work, which can lead to more time lost reconciling the new data model with the existing analysis.
 
-From the user’s perspective, the boundary between preparation and analysis is not nearly so clean cut.
+From the user's perspective, the boundary between preparation and analysis is not nearly so clean cut.
 Bad data is often discovered using visual analysis techniques (e.g.
-histograms or scatter plots) and it is most natural for the user to “clean what she sees” instead of switching to a second tool.
-This leads to an ``adaptive” process whereby users will prefer leveraging existing tools in the analytics environment (no matter how well suited to the task) over switching to another application.
+histograms or scatter plots) and it is most natural for the user to ``clean what she sees" instead of switching to a second tool.
+This leads to an ``adaptive" process whereby users will prefer leveraging existing tools in the analytics environment (no matter how well suited to the task) over switching to another application.
 Thus a well-designed interactive visual analysis environment will provide tools that enable users to perform such cleaning tasks as interactively as possible.

### 2015-07-02 nlp.tex
Quote fixing, added placeholder sections

\subsection{Context-Free Grammar}
ICU date-time formats are well defined both structurally and semantically, and can be defined by a context-free grammar (CFG). A CFG is commonly defined as a set of productions or rules of the form A ? ? where A is a variable, ? is a sequence of variables and terminal symbols (the tokens that make up the alphabet of the language) plus null (? ), and the production symbol (?) indicates that the variable A can be expanded into ? . A CFG can be formally specified with four components: V, T, P, and S, where V is the set of variables, T the set of terminal symbols, P the set of productions, and S the set of available start symbols (a non-empty subset of V) [cite].

Pattern-recognition problems such as parsing date and time formats initiate from observations generated by some structured stochastic process. In other words, even if the initial higher-level production rule of the grammar is known (i.e. date, time or date-time), there could be several directions that the parser resolve to. For example, in a date string 5/6/2015, the pattern could either be M/d/yyyy or d/M/yyyy.

Probabilistic context-free grammars (PCFGs) have provided a useful method for modeling such uncertainty [cite]. Once we have created a PCFG model of a process, we can apply existing PCFG parsing algorithms to identify a variety of date-time formats. However, the parser’s success is often limited in the types of the dominant patterns that it can identify. In addition, the standard parsing techniques generally require specification of a complete observation sequence. In many contexts, we may have only a partial sequence available (e.g. an incomplete entry). Finally, we may be interested in computing the probabilities of date-time patterns that the grammar may not explicitly define. To extend the forms of evidence, inferences, and pattern distributions supported, we need a flexible and expressive representation for the distribution of structures generated by the grammar. We adopt Bayesian networks for this purpose, and define an algorithm to generate a probabilistic distribution of possible parse trees corresponding to a set of date-time patterns as opposed to individual ones.

%•	Motivate the use of context-free grammar, particularly Backus-Normal-Form
%•	Explain why we use a probabilistic version of the grammar
%•	Explain how the production rules are created based on the ICU format
%•	Extensions to the parser to account for columnar context for maximizing the probabilistic occurrence of the dominant pattern
%•	Other semantic extensions – looking at the file name, adding rules for leap year, using external dictionaries for non-English terms

### 2015-07-02 paper.tex
Quote fixing, added placeholder sections

@@
 
-\section{The {\secit Body} of The Paper}
-Typically, the body of a paper is organized
-into a hierarchical structure, with numbered or unnumbered
-headings for sections, subsections, sub-subsections, and even
-smaller sections.  The command \texttt{{\char'134}section} that
-precedes this paragraph is part of such a
-hierarchy.\footnote{This is the second footnote.  It
-starts a series of three footnotes that add nothing
-informational, but just give an idea of how footnotes work
-and look. It is a wordy one, just so you see
-how a longish one plays out.} \LaTeX\ handles the numbering
-and placement of these headings for you, when you use
-the appropriate heading commands around the titles
-of the headings.  If you want a sub-subsection or
-smaller part to be unnumbered in your output, simply append an
-asterisk to the command name.  Examples of both
-numbered and unnumbered headings will appear throughout the
-balance of this sample document.
+\section{Parameters}
+\input{parameters}
 
-Because the entire article is contained in
-the \textbf{document} environment, you can indicate the
-start of a new paragraph with a blank line in your
-input file; that is why this sentence forms a separate paragraph.
+\section{Minimal Descriptive Length}
+\input{mdl}
 
-\subsection{Type Changes and {\subsecit Special} Characters}
-We have already seen several typeface changes in this sample.  You
-can indicate italicized words or phrases in your text with
-the command \texttt{{\char'134}textit}; emboldening with the
-command \texttt{{\char'134}textbf}
-and typewriter-style (for instance, for computer code) with
-\texttt{{\char'134}texttt}.  But remember, you do not
-have to indicate typestyle changes when such changes are
-part of the \textit{structural} elements of your
-article; for instance, the heading of this subsection will
-be in a sans serif\footnote{A third footnote, here.
-Let's make this a rather short one to
-see how it looks.} typeface, but that is handled by the
-document class file. Take care with the use
-of\footnote{A fourth, and last, footnote.}
-the curly braces in typeface changes; they mark
-the beginning and end of
-the text that is to be in the different typeface.
-
-You can use whatever symbols, accented characters, or
-non-English characters you need anywhere in your document;
-you can find a complete list of what is
-available in the \textit{\LaTeX\
-User's Guide}\cite{Lamport:LaTeX}.
-
-\subsection{Math Equations}
-You may want to display math equations in three distinct styles:
-inline, numbered or non-numbered display.  Each of
-the three are discussed in the next sections.
-
-\subsubsection{Inline (In-text) Equations}
-A formula that appears in the running text is called an
-inline or in-text formula.  It is produced by the
-\textbf{math} environment, which can be
-invoked with the usual \texttt{{\char'134}begin. . .{\char'134}end}
-construction or with the short form \texttt{\$. . .\$}. You
-can use any of the symbols and structures,
-from $\alpha$ to $\omega$, available in
-\LaTeX\cite{Lamport:LaTeX}; this section will simply show a
-few examples of in-text equations in context. Notice how
-this equation: \begin{math}\lim_{n\rightarrow \infty}x=0\end{math},
-set here in in-line math style, looks slightly different when
-set in display style.  (See next section).
-
-\subsubsection{Display Equations}
-A numbered display equation -- one set off by vertical space
-from the text and centered horizontally -- is produced
-by the \textbf{equation} environment. An unnumbered display
-equation is produced by the \textbf{displaymath} environment.
-
-Again, in either environment, you can use any of the symbols
-and structures available in \LaTeX; this section will just
-give a couple of examples of display equations in context.
-First, consider the equation, shown as an inline equation above:
-\begin{equation}\lim_{n\rightarrow \infty}x=0\end{equation}
-Notice how it is formatted somewhat differently in
-the \textbf{displaymath}
-environment.  Now, we'll enter an unnumbered equation:
-\begin{displaymath}\sum_{i=0}^{\infty} x + 1\end{displaymath}
-and follow it with another numbered equation:
-\begin{equation}\sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f\end{equation}
-just to demonstrate \LaTeX's able handling of numbering.
-
-\subsection{Citations}
-Citations to articles \cite{bowman:reasoning, clark:pct, braams:babel, herlihy:methodology},
-conference
-proceedings \cite{clark:pct} or books \cite{salas:calculus, Lamport:LaTeX} listed
-in the Bibliography section of your
-article will occur throughout the text of your article.
-You should use BibTeX to automatically produce this bibliography;
-you simply need to insert one of several citation commands with
-a key of the item cited in the proper location in
-the \texttt{.tex} file \cite{Lamport:LaTeX}.
-The key is a short reference you invent to uniquely
-identify each work; in this sample document, the key is
-the first author's surname and a
-word from the title.  This identifying key is included
-with each item in the \texttt{.bib} file for your article.
-
-The details of the construction of the \texttt{.bib} file
-are beyond the scope of this sample document, but more
-information can be found in the \textit{Author's Guide},
-and exhaustive details in the \textit{\LaTeX\ User's
-Guide}\cite{Lamport:LaTeX}.
-
-This article shows only the plainest form
-of the citation command, using \texttt{{\char'134}cite}.
-This is what is stipulated in the SIGS style specifications.
-No other citation format is endorsed.
-
-\subsection{Tables}
-Because tables cannot be split across pages, the best
-placement for them is typically the top of the page
-nearest their initial cite.  To
-ensure this proper ``floating'' placement of tables, use the
-environment \textbf{table} to enclose the table's contents and
-the table caption.  The contents of the table itself must go
-in the \textbf{tabular} environment, to
-be aligned properly in rows and columns, with the desired
-horizontal and vertical rules.  Again, detailed instructions
-on \textbf{tabular} material
-is found in the \textit{\LaTeX\ User's Guide}.
-
-Immediately following this sentence is the point at which
-Table 1 is included in the input file; compare the
-placement of the table here with the table in the printed
-dvi output of this document.
-
-\begin{table}
-\centering
-\caption{Frequency of Special Characters}
-\begin{tabular}{|c|c|l|} \hline
-Non-English or Math&Frequency&Comments\\ \hline
-\O & 1 in 1,000& For Swedish names\\ \hline
-$\pi$ & 1 in 5& Common in math\\ \hline
-\$ & 4 in 5 & Used in business\\ \hline
-$\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
-\hline\end{tabular}
-\end{table}
-
-To set a wider table, which takes up the whole width of
-the page's live area, use the environment
-\textbf{table*} to enclose the table's contents and
-the table caption.  As with a single-column table, this wide
-table will ``float" to a location deemed more desirable.
-Immediately following this sentence is the point at which
-Table 2 is included in the input file; again, it is
-instructive to compare the placement of the
-table here with the table in the printed dvi
-output of this document.
-
-\begin{table*}
-\centering
-\caption{Some Typical Commands}
-\begin{tabular}{|c|c|l|} \hline
-Command&A Number&Comments\\ \hline
-\texttt{{\char'134}alignauthor} & 100& Author alignment\\ \hline
-\texttt{{\char'134}numberofauthors}& 200& Author enumeration\\ \hline
-\texttt{{\char'134}table}& 300 & For tables\\ \hline
-\texttt{{\char'134}table*}& 400& For wider tables\\ \hline\end{tabular}
-\end{table*}
-% end the environment with {table*}, NOTE not {table}!
-
-\subsection{Figures}
-Like tables, figures cannot be split across pages; the
-best placement for them
-is typically the top or the bottom of the page nearest
-their initial cite.  To ensure this proper ``floating'' placement
-of figures, use the environment
-\textbf{figure} to enclose the figure and its caption.
-
-This sample document contains examples of \textbf{.eps}
-and \textbf{.ps} files to be displayable with \LaTeX.  More
-details on each of these is found in the \textit{Author's Guide}.
-
-\begin{figure}
-\centering
-\epsfig{file=fly.eps}
-\caption{A sample black and white graphic (.eps format).}
-\end{figure}
-
-\begin{figure}
-\centering
-\epsfig{file=fly.eps, height=1in, width=1in}
-\caption{A sample black and white graphic (.eps format)
-that has been resized with the \texttt{epsfig} command.}
-\end{figure}
-
-As was the case with tables, you may want a figure
-that spans two columns.  To do this, and still to
-ensure proper ``floating'' placement of tables, use the environment
-\textbf{figure*} to enclose the figure and its caption.
-
-Note that either {\textbf{.ps}} or {\textbf{.eps}} formats are
-used; use
-the \texttt{{\char'134}epsfig} or \texttt{{\char'134}psfig}
-commands as appropriate for the different file types.
-
-\subsection{Theorem-like Constructs}
-Other common constructs that may occur in your article are
-the forms for logical constructs like theorems, axioms,
-corollaries and proofs.  There are
-two forms, one produced by the
-command \texttt{{\char'134}newtheorem} and the
-other by the command \texttt{{\char'134}newdef}; perhaps
-the clearest and easiest way to distinguish them is
-to compare the two in the output of this sample document:
-
-This uses the \textbf{theorem} environment, created by
-the\linebreak\texttt{{\char'134}newtheorem} command:
-\newtheorem{theorem}{Theorem}
-\begin{theorem}
-Let $f$ be continuous on $[a,b]$.  If $G$ is
-an antiderivative for $f$ on $[a,b]$, then
-\begin{displaymath}\int^b_af(t)dt = G(b) - G(a).\end{displaymath}
-\end{theorem}
-
-The other uses the \textbf{definition} environment, created
-by the \texttt{{\char'134}newdef} command:
-\newdef{definition}{Definition}
-\begin{definition}
-If $z$ is irrational, then by $e^z$ we mean the
-unique number which has
-logarithm $z$: \begin{displaymath}{\log e^z = z}\end{displaymath}
-\end{definition}
-
-Two lists of constructs that use one of these
-forms is given in the
-\textit{Author's  Guidelines}.
-
-\begin{figure*}
-\centering
-\epsfig{file=flies.eps}
-\caption{A sample black and white graphic (.eps format)
-that needs to span two columns of text.}
-\end{figure*}
-and don't forget to end the environment with
-{figure*}, not {figure}!
-
-There is one other similar construct environment, which is
-already set up
-for you; i.e. you must \textit{not} use
-a \texttt{{\char'134}newdef} command to
-create it: the \textbf{proof} environment.  Here
-is a example of its use:
-\begin{proof}
-Suppose on the contrary there exists a real number $L$ such that
-\begin{displaymath}
-\lim_{x\rightarrow\infty} \frac{f(x)}{g(x)} = L.
-\end{displaymath}
-Then
-\begin{displaymath}
-l=\lim_{x\rightarrow c} f(x)
-= \lim_{x\rightarrow c}
-\left[ g{x} \cdot \frac{f(x)}{g(x)} \right ]
-= \lim_{x\rightarrow c} g(x) \cdot \lim_{x\rightarrow c}
-\frac{f(x)}{g(x)} = 0\cdot L = 0,
-\end{displaymath}
-which contradicts our assumption that $l\neq 0$.
-\end{proof}
-
-Complete rules about using these environments and using the
-two different creation commands are in the
-\textit{Author's Guide}; please consult it for more
-detailed instructions.  If you need to use another construct,
-not listed therein, which you want to have the same
-formatting as the Theorem
-or the Definition\cite{salas:calculus} shown above,
-use the \texttt{{\char'134}newtheorem} or the
-\texttt{{\char'134}newdef} command,
-respectively, to create it.
-
-\subsection*{A {\secit Caveat} for the \TeX\ Expert}
-Because you have just been given permission to
-use the \texttt{{\char'134}newdef} command to create a
-new form, you might think you can
-use \TeX's \texttt{{\char'134}def} to create a
-new command: \textit{Please refrain from doing this!}
-Remember that your \LaTeX\ source code is primarily intended
-to create camera-ready copy, but may be converted
-to other forms -- e.g. HTML. If you inadvertently omit
-some or all of the \texttt{{\char'134}def}s recompilation will
-be, to say the least, problematic.
-
-\section{Conclusions}
-This paragraph will end the body of this sample document.
-Remember that you might still have Acknowledgments or
-Appendices; brief samples of these
-follow.  There is still the Bibliography to deal with; and
-we will make a disclaimer about that here: with the exception
-of the reference to the \LaTeX\ book, the citations in
-this paper are to articles which have nothing to
-do with the present subject and are used as
-examples only.
-%\end{document}  % This is where a 'short' article might terminate
+\section{Natural Language Processing}
+\input{nlp}
 

### 2015-07-02 intro.tex
Adding more into the intro from the Word doc

@@
 Thus a well-designed interactive visual analysis environment will provide tools that enable users to perform such cleaning tasks as interactively as possible.
+
+\subsection{The Tableau Ecosystem}
+In this paper, we shall be looking at an example of this problem in the context of the Tableau system.
+Tableau is a commercial visual analysis environment derived from the Polaris [1] system developed at Stanford.
+In addition to an interactive desktop application for creating data visualizations, the Tableau ecosystem also includes servers for publishing and sharing interactive visualizations.
+These servers can be privately operated by individual customers or hosted in the cloud.
+
+\subsubsection{Tableau Public}
+Of particular relevance in the present work is a free version of the server environment called Tableau Public [12].
+Tableau Public (or ``Public") allows authors to publish data visualizations that can be shared with the general public.
+The published data sets are limited to 100K rows and must be stored using the Tableau Data Engine (or ``TDE") [2, 3], but are otherwise free to use the full analytic capabilities of the Tableau system.
+In particular, the TDE provides native support for all analytic calculations defined by the authoring and publishing components.
+Moreover, visualizations published to Public are uploaded as complete Tableau workbooks, including the data model developed as part of the analysis.
+This makes it a rich source of data on the analytic habits and frustrations of Tableau users.
+
+\subsubsection{Functions}
+While the existing standards for query languages are helpful for defining a core set of row-level functions, there are many useful functions beyond the standards that are only explicitly supported by a subset of RDBMSes, often with different names.
+And even when a database does not provide an implementation of a particular function, it is usually possible to generate it as an inline combination of existing functionality.
+Tableau's calculation language is designed to be a lingua franca that tames this Babel of dialects by providing a common syntax for as many of these functions as possible.
+
+\subsection{DATEPARSE}
+Among the recent additions to the Tableau function library is a function called DATEPARSE, the usability of which is the focus of the work in this paper.
+
+\subsubsection{	Scalar Dates}
+The SQL-99 standard defines three temporal scalar types: DATE, TIMESTAMP and TIME.
+They are typically implemented as fixed-point types containing an offset from some epoch (\textit{e.g.} Julian Days.) This makes them compact to store and allows some temporal operations to be implemented very efficiently using arithmetic.
+Thus from the RDBMS perspective, representing dates in scalar form provides numerous benefits for users, both in terms of available operations and query performance.
+
+Tableau models the first two of these types as ``Date" and ``Date \& Time"; the third (pure time) is folded into Date \& Time by appending it to a fixed date of 1899-12-30.
+From the analytic perspective, date types are dimensional (\textit{i.e.} independent variables) and can be used as either categorical (simply ordered) or quantitative (ordered with a distance metric) fields.
+Categorical dates have a natural hierarchy associated with them generated by calendar binning.
+The visualization in Figure I1 shows an example of a bar chart employing binned categorical dates in a year/quarter hierarchy.

### 2015-07-04 intro.tex
More transferring from Word into LaTeX

@@
 The visualization in Figure I1 shows an example of a bar chart employing binned categorical dates in a year/quarter hierarchy.
+
+\subsubsection{Parsing}
+One of the oldest data preparation problems observed in Tableau is the parsing of date scalars so that users can perform these kinds of analyses.
+Training materials from the early days of the company included examples of how to convert columns of integers in the form yyyyMMdd to date scalars.
+The documented solution at the time was to convert the integer to a string and perform some locale-dependent string operations before casting the string back to a date.
+Unfortunately, this approach had a number of problems:
+\begin{itemize}
+\item String operations are notoriously slow compared to scalar operations (typically 10-100x slower)
+\item Default parsing of date formats is locale-dependent, and may not work when the workbook is shared across an international organization (e.g. between the US and European offices)
+\item The expression code was hard to understand and maintain because it used a verbose, general-purpose string-handling syntax instead of a domain language.
+\end{itemize}
+
+This is but a single format.
+Our studies of the workbooks on Public suggest that there are hundreds of distinct temporal date formats in user data sets.
+Some are common, but others can be quite idiosyncratic.
+Table 1 shows a selection of unusual date formats found on Public.
+
+% INSERT TABLE HERE
+
+Our first solution to this problem was to add a new function to the Tableau calculation language called DATEPARSE.
+This function would take a string column and convert it to a datetime using a special purpose domain language for describing dates.
+Such functions exist in a number of databases (e.g.
+Oracle, Postgres and MySQL) and adding it to the TDE was straightforward, so it was a natural addition to the function library.
+We chose the date parsing syntax defined by the International Components for Unicode (or ICU) project [8] for the common domain language because it mirrored what was available in the Tableau code base, and translating this syntax into the formats used by various database vendors (e.g.
+MySQL, Oracle, Postgres) was relatively painless.
+There are a few patches of non-overlapping functionality, but they tend to be obscure and we provide warnings when functionality is missing.
+
+\subsubsection{	Usability}
+After shipping this new function to our user base, we investigated how it was being applied in the field.
+We examined a few months of workbooks that had been uploaded to Public after the release to see if and how it was being used.
+We found that although there were a number of new calculations using DATEPARSE, the error rate for the domain language syntax was about 15\%.
+DATEPARSE was capable of solving the problem, and some users were able to discover it, but the syntax did not appear to be easy to use reliably.
+
+\subsection{Automated Format Extraction}
+One solution might have been to design a graphical environment that enabled users to construct valid patterns.
+This would have involved a substantial development investment with no guarantee that the result would be correct if the user misunderstood the environment.
+Instead, we have developed two algorithms for automating the derivation of the format string, each of which uses a different machine learning technique.
+Both algorithms result in over 90\% parsing accuracy – and 100\% syntactic correctness because they are machine generated.
+
+% vidya: this probably should be its own section
+
+\subsection{Related Work}
+Data preparation has been a known analytic bottleneck since at least the description of the Potter’s Wheel system [4].
+Since then, several other interactive data preparation systems have been proposed [5,6].
+While effective, these systems all make assumptions about possible date formats (e.g.
+the domain system in [4]), which we suggest are too restrictive for real world data.
+The Minimum Descriptive Length technique was first proposed in [7]. We have built on the version described in [4].
+Related work on parsing languages, outlier detection.
+
+\subsection{Overview}
+The rest of this paper is organized as follows: The next section introduces the parameters of the problem space.
+The following two sections describe the two different algorithms, one using Minimum Descriptive Length and the other using Natural Language Processing.
+In section 5, we evaluate the algorithms on a corpus of 30K columns, both by sampling the outputs manually and then by using the algorithms to validate each other.
+We then discuss future work in section 6 and conclude in section 7.

### 2015-07-04 mdl.tex
More transferring from Word into LaTeX

The first algorithm is a Minimum Descriptive Length [7] approach derived from the domain system presented in [4]. We describe a number of extensions to their structure extraction system to support more complex redundancy, non-English locales, improved performance and date-specific pruning.

\subsection{Domains}
[4] presents an algorithm for deriving a common structure for a set of strings by breaking each string down into a sequence of domains. These domains are described by an interface that includes:
\begin{itemize}
\item A required inclusion function to test for membership in the domain (match)
\item An optional function to compute the number of values in the domain with a given length (cardinality)
\item An optional function to update statistics for the domain based on a given value (updateStatistics)
\item An optional function to prevent consideration of a domain that is redundant (isRedundantAfter).
\end{itemize}

In our approach, we implement all of these functions, but with significant changes to the last one, which we will describe below.

With this interface, we can now define a set of domains for each date part that we wish to be able to parse. These are mostly straightforward enumerations and numeric ranges, each tagged with the ICU format code. Since the ICU parser is flexible about parsing single or double digit formats, we use double-digit formats, but accept one or two digits. One important exception to this rule is for years, which are fixed width fields (2 or 4).

We also included some enumerated domains for handling constant strings and some simple regular expression domains for delimiters such as spaces, punctuation or alphabetic characters. We found that the inclusion of arbitrary numeric domains caused the run time to grow exponentially as the number of possible matches could not be pruned intelligently. This restriction extends to domains that can contain arbitrary digit sequences (such as any). Because of this restriction, the algorithm cannot extract non-date numeric fields.

\subsection{Redundancy Extensions}
A difficulty in using this kind of structure extraction is that the algorithm for enumerating structures is exponential in the number of domains. This is especially true in the date format problem because there are identical domains (e.g. months and meridian hours), nearly identical domains (e.g. days and hours) and there are often no field delimiters (e.g. 2012Mar06134427). To handle this, we have extended the existing pruning API with two other sets of domain identifiers:
\begin{itemize}
\item A set of \textit{prunable} identifiers, which are not allowed to precede the domain. For example, once we have a month field, no other month fields should be generated. Each month domain therefore lists all the month domains in its prunable set.
\item A set of \textit{context} identifiers, one of which must have been previously generated before the domain is considered. For example, a meridian domain can only be generated once an hour field has been found, but there may be other intervening fields.
\end{itemize}

\subsection{Performance}
Structure enumeration is computationally expensive, so we added a number of enhancements to the original domain extraction algorithm to keep the run time low enough for interactivity.

\subsubsection{Domain Characteristics}
Date domains typically have small widths, so we found it advantageous to provide the shortest and longest match sizes for use in structure enumeration and matching.

Date domains are also often uniform in that adding more characters to a mismatch will not help. For example, a 2-digit day domain that does not match a 1-letter substring will not be able to generate a match by adding more characters.

\subsubsection{Parallel Evaluation}
We have also identified two opportunities for parallel computation during structure enumeration.

The first computationally expensive operation is the enumeration of structures for a given sample. To parallelize this step, each thread is given a subset of the samples and independently produces a set of candidate domains. When all threads have completed, the duplicates are removed to produce a single list of candidate structures.

The second computationally expensive operation is the evaluation of each generated structure over the entire sample set. This includes computation of the MDL, recording of domain statistics and parameterizing the structure. These tasks are simple to parallelize, because there are no overlaps between the data for each structure.

\subsection{Unparameterization}
Domain parameterization is an important part of generating compact representations via MDL, but it creates problems for date recognition. For example, if a set of dates contains a constant month string (e.g. all values are in September) it is important to keep track of the month name domain. Consequently, when we parameterize a constant generic <Word> domain, we tag it with the date part domain that it matches (if any). We then need to apply an additional pruning step to remove any structures that also found an equivalent domain (e.g. two-digit month). These rules are equivalent to the context-based redundancy rules above, but have to be applied again after parameterization of generics.

\subsection{Global Pruning}
The pruning rules used for the structure extraction reduce the search space dramatically, but they are also contextual and can only look backwards. The domains also contain a fair amount of ambiguity that requires the application of domain knowledge. We therefore found it necessary to add some post-generation global pruning rules:
\begin{itemize}
\item The set of date parts cannot contain place value gaps (e.g. structures that have year and day without month are removed.)
\item Similarly, the set of time parts cannot contain place value gaps and must also be in place value order (times are never written in orders such as mhs.)
\item The existence of time parts cannot make dates incomplete (e.g. patterns like year-day-hour are removed.)
\item Two digit years require special handling. In particular, they cannot appear adjacent to a two-digit field if the structure contains punctuation. (This can come up in some small early 21st century year domains where a two-digit year can masquerade as almost any numeric field.)
\end{itemize}

These global pruning rules are simple and intuitive, but are essential to further reducing the search space.

\subsection{Locale Sensitivity}
Providing an acceptable international user experience requires correctly handling the locale of the column text. Accordingly, at the start of the structure extraction we use the column locale to create a set of domains containing locale-sensitive strings such as month names. We also use the locale to map these strings to upper- and lower-case in addition to the ICU mixed case strings. This enables us to accurately compute MDL statistics without having to map the candidate strings at runtime (which would be slow).

Knowing the locale of a string is not always helpful. In our data set, we found numerous cases where the locale was specified (e.g. Sweden) but the data was actually in English. Accordingly, we test both locales and rank the combined results.
Tableau currently does not support non-Gregorian calendars. Therefore, we have built this system for the Gregorian calendar only. ICU does support non-Gregorian calendars and we expect this system could be applied to them as well.

\subsection{Ranking}
MDL structure analysis naturally produces a ranked list of format candidates, but we have found that a number of other properties of the formats should be preferred over simple compactness:
\begin{itemize}
\item Since we have a set of samples, we can apply the candidate format to the strings to see how well it performs. Formats with fewer parse errors are preferred.
\item Date parts can be considered a place-value system, so we prefer “more significant” components (e.g. month-day-year over hour-minute-second).
\item If two formats from different locales give the same results, prefer the original column locale. The sample set may have missed an example where this could be important.
\item If the format has an ambiguous date order (e.g. all days are less than 12), then prefer the default date order of the locale. Again, the sample set may have missed a counterexample, so this is the best option.
\item Once these semantic preferences have been considered, we then prefer the more compact (MDL) representation.
The output of the algorithm is now an ordered list of formats and associated locales. These can then be used to drive a user interface that allows the user to choose between the possibilities or the top-ranking format can simply be used automatically.
\end{itemize}

### 2015-07-04 parameters.tex
More transferring from Word into LaTeX

Before delving into detailed descriptions of the two algorithms, we describe the common elements of the problem they are intended to solve. These are the input data, the output language and the interpretation of the output language for partial dates.

\subsection{Input Data}
The training data and test data are taken from Public data sets. We selected a large number of string columns whose names suggested that they might contain temporal data. We included words from multiple languages besides English and extracted the column’s locale (actually the collation) along with the data.

\subsubsection{NULL Filtering}
Each column was then reduced to a maximum sample set of 32. We excluded domain values that appeared to be representations of NULL:
\begin{itemize}
\item Values containing the substring NULL
\item Values containing no digits and at most one non-whitespace character (e.g. " / / ")
\item Values on a list of empirically determined common NULL values (e.g. 0000-00-00, NaN)
\end{itemize}

Columns that had no remaining valid samples were discarded.

\subsubsection{Sampling}
The remaining samples were hashed and sorted on the hash value, with the top 32 values being retained as the column’s sample. We can increase the sample size if needed, but for dates, it appears to be an adequate number. In any case, the median number of non-null rows per domain in our test data is 50, so increasing the sample size would have little benefit.

\subsubsection{Numeric Timestamps}
After NULL filtering, there remained one common class of date representation that was unsuited for parsing via our date format syntax, namely numeric timestamps. This included Unix epoch timestamps (expressed as second, millisecond or microsecond counts from 1970-01-01) and Microsoft Excel timestamps (expressed fractional days since 1900-01-01). A simple test to check that the column was numeric and inside a specific range of values representing recent dates allowed us to tag these columns to avoid analyzing them further (and incidentally to identify them for generating a simple date extraction calculation for the user.)

\subsection{The ICU Date Format Language}
The date format syntax we selected is the one provided by the open-source project. We chose it because we were already using ICU in the code base, we had access to the source code and it provides localized date part data for a large number of languages.

The syntax is documented at the ICU web site [8]. While fairly complete, it has a few limitations that we ran into when evaluating the algorithms:
\begin{itemize}
\item No support for 4 letter year abbreviations (e.g. Sept.)
\item No support for ordinal days (e.g. July 4th)
\item No support for quarter postfix notation (e.g. 2Q)
\item No support for variant meridian markers (e.g. a.m.)
\end{itemize}

These limitations did not affect the results significantly, and in the future we hope to submit ICU extensions to handle some of these issues.

One other quirk of the ICU syntax may be a contributing factor to the user confusion around writing correct ICU date formats. The use of lexicographical case in the meta-symbols of the format language can be confusing (e.g. y is used for years, but M is used for months while m is used for minutes.) Another advantage of an automated algorithm is that it hides such problems from most users, significantly improving the usability of the function.

\subsection{Partial Dates}
Many of the date formats that we encountered were incomplete dates, which necessitated creating rules for what date scalar they represented.

ICU’s date parsing APIs allow the specification of default values for parts when a format does not contain them. In our implementation, all time fields are set to 0 (midnight) and the date fields are set based on whether the format contains any date part specifications. When date parts are present, we use 2000-01-01 as the set of default date parts as it is the start of a leap year. When dealing with pure time formats, we use 1899-12-30, (which is the date used internally by Tableau to signal pure time values.)

ICU will also parse Time Zones (which we recognize but do not use in Tableau) and Quarters (which we interpret as the first month of the period.) RDBMSes such as Oracle and Postgres that support time zones will be able to take advantage of the generated time zone field.

### 2015-07-04 conclusion.tex
Added remaining sections

In this paper, we have described two effective algorithms for extracting date format strings from a small set of samples, one using a minimum descriptive length approach and one using natural language techniques. Both algorithms are accurate enough to be used automatically without user involvement. The MDL algorithm is also fast enough to deploy in an interactive environment.

While validating the algorithms on a large corpus, we also found that the number of distinct formats in the wild is surprisingly high, and demonstrates the wisdom of including general-purpose date parsing functions in data visualization tools, data cleaning tools and RDBMSes. In particular, it is interesting to note that the most prominent open source RDBMSes (MySQL and Postgres) both have a built-in version of DATEPARSE, possibly reflecting that this is a common need that gets implemented when users are empowered to add to a code base.

### 2015-07-04 futurework.tex
Added remaining sections

While we are satisfied with the results so far, we found a number of date formats that we were not able to handle with these techniques. Many columns contained multiple formats, and we would like to recognize this situation and generate predicate calculations to increase the accuracy of the results and thereby make the results even more seamless for the user. Other columns contain date ranges, which we would like to handle by generating multiple calculations, possibly by combining regular expressions with date parsing.

In this paper, we have only considered the parsing of strings, but dates are often formatted as integers (e.g. 201507016). It is significantly faster to decompose integers into date parts using arithmetic operations (e.g. mod and div) than by using slower, locale-sensitive string parsing functions. Moreover, the number of possible formats is low enough that we may be able to enumerate them. Timestamps are another form of numeric date for which we intend to automate the preparation.
In the course of our research, we have also identified a number of date part variants (e.g. ordinal dates, 4 letter month abbreviations, alternate meridian markers and postfix quarter syntax) that we would like to commend to the ICU project and maybe provide implementations for.

### 2015-07-04 paper.tex
Added remaining sections

@@
 
+\section{Testing}
+\input{testing}
+
+\section{Future Work}
+\input{futurework}
+
+\section{Conclusion}
+\input{conclusion}
+
 %ACKNOWLEDGMENTS are optional
 \section{Acknowledgments}
-This section is optional; it is a location for you
-to acknowledge grants, funding, editing assistance and
-what have you.  In the present case, for example, the
-authors would like to thank Gerald Murray of ACM for
-his help in codifying this \textit{Author's Guide}
-and the \textbf{.cls} and \textbf{.tex} files that it describes.
+We would like to thank to Pat Hanrahan for encouraging this work; Roger Hau for providing the resources to productize it; and Jason King, whose harrowing live demonstration of manual DATEPARSE in front of several thousand people inspired this project.
 
@@
 %
-%APPENDICES are optional
-%\balancecolumns
-\appendix
-%Appendix A
-\section{Headings in Appendices}
-The rules about hierarchical headings discussed above for
-the body of the article are different in the appendices.
-In the \textbf{appendix} environment, the command
-\textbf{section} is used to
-indicate the start of each Appendix, with alphabetic order
-designation (i.e. the first is A, the second B, etc.) and
-a title (if you include one).  So, if you need
-hierarchical structure
-\textit{within} an Appendix, start with \textbf{subsection} as the
-highest level. Here is an outline of the body of this
-document in Appendix-appropriate form:
-\subsection{Introduction}
-\subsection{The Body of the Paper}
-\subsubsection{Type Changes and  Special Characters}
-\subsubsection{Math Equations}
-\paragraph{Inline (In-text) Equations}
-\paragraph{Display Equations}
-\subsubsection{Citations}
-\subsubsection{Tables}
-\subsubsection{Figures}
-\subsubsection{Theorem-like Constructs}
-\subsubsection*{A Caveat for the \TeX\ Expert}
-\subsection{Conclusions}
-\subsection{Acknowledgments}
-\subsection{Additional Authors}
-This section is inserted by \LaTeX; you do not insert it.
-You just add the names and information in the
-\texttt{{\char'134}additionalauthors} command at the start
-of the document.
-\subsection{References}
-Generated by bibtex from your ~.bib file.  Run latex,
-then bibtex, then latex twice (to resolve references)
-to create the ~.bbl file.  Insert that ~.bbl file into
-the .tex source file and comment out
-the command \texttt{{\char'134}thebibliography}.
-% This next section command marks the start of
-% Appendix B, and does not continue the present hierarchy
-\section{More Help for the Hardy}
-The acm\_proc\_article-sp document class file itself is chock-full of succinct
-and helpful comments.  If you consider yourself a moderately
-experienced to expert user of \LaTeX, you may find reading
-it useful but please remember not to change it.
+
 \balancecolumns

### 2015-07-04 testing.tex
Added remaining sections

\subsection{Data}

\subsubsection{Training Data}
All files published to Tableau Public through February 2014 were scanned. We collected the contents of columns with names containing any of the following strings: Date, month, created, dt (abbreviation for date), mes (month in Spanish), datum (date in German), fecha (date in Spanish), data (date in Portuguese), ? (day in Chinese and Japanese). Roughly 95\% of the data in Tableau Public is in English, but we attempted to include non-English data that was available. Fields of any data type other than date or datetime were analyzed, including strings, integers, and floats. One file was created for each scanned column, containing the unique non-null values in the column.

The resulting set of files were de-duplicated. This handles any duplicate files on Tableau Public including sample data. Files over 1MB were manually reviewed and those that did not contain dates removed. There were 30968 files in the resulting set.
Most database and spreadsheet systems already detect a limited set of date formats. For instance, typing the string “12/31/1999” into Microsoft Excel, is automatically interpreted as the date 1999-12-31. The Microsoft Jet library that Tableau used to read these text files detects a few date formats as well. Any column already converted by Excel or Jet was not included in this study.

\textbf{Verification Data}\\
Verification used another set of data from Tableau Public, collected through April 2015. This set was collected similarly, but also added columns named time. We limited this set to new Tableau Excel and text connectors that were released in May 2014. It is likely that some of these files were older files republished to Tableau Public. There were 31546 files in the resulting set.

The new Tableau text connector is able to automatically interpret a somewhat different limited set of date formats. Any column already converted by Excel or the new Tableau text connector was not included in this study.

\subsection{Evaluation}
Once the training data was analyzed, it was grouped by date format. A sample of each produced date format was manually labeled. This allowed us to quickly skip over very common formats like MM/dd/yyyy and focus our efforts on much less common formats. The samples were judged as to whether the produced format was reasonable and were tagged with correct formats if the produced format was unreasonable.

Further manual tagging focused on the files most likely to represent dates. Many of the files in the collection are not actually dates, such as fields named “updated by” (which contain “date”). A random sample of 850 columns named exactly ``date", ``time" or ``month" (case-insensitive) were manually judged.

\subsubsection{	Minimum Descriptive Length}
Testing of the MDL algorithm was performed on a 24-core Dell T7610 running Windows 7 with the data stored on a 250GB SSD.

%Figure M1. MDL Parsing Statistics
To test the MDL algorithm, we ran it over the set of samples from each validation file to generate a ranked list of formats for the file. Each format was then applied to the entire file’s data set, recording both the number of errors and the elapsed times. In cases where we generated multiple formats and the main format produced errors, we applied the second format to the unparsed strings. The summary statistics from this processing are presented in Figure M1.

The analysis speed is the average time needed per sample for structure extraction. At 2.5ms, this is well below most human perceptual thresholds for a set of 32 samples, so any latency in command execution would be restricted to the ability of the underlying database to provide the samples for analysis in a timely manner. (As a column store, the TDE can often supply such domains without a full table scan, further improving responsiveness.) 
The validation speed is the average time needed to parse a value, and provides an estimate of how fast an ICU-based implementation (such as the TDE) can process string values into scalars and works out to 620K values per core per second.

%Figure M2. MDL Error Rate
The error rate reflects the fact that only about 40\% of the files have an associated format that parses the non-null values without error. To examine the error rate in greater detail, we turn now to Figure M2.
On the left hand side, we can see that the algorithm found 744 distinct formats that parsed 13,424 files with no error. This is a remarkable number of valid formats and underscores the need for this kind of algorithm. Raising the error rate threshold to 5\% results in about 2500 formats found in 15,000 files, or nearly half of the files.

What do these formats look like? Figure M3 shows a histogram of the 25 most common formats containing a year format code at the 5\% threshold, color-coded by error rate. (A sample value is provided to the right of each bar for illustrative purposes.) The formats have also been filtered to files with at least 5 samples. Most of the samples are clearly dates with a wide range of formats (the format where the time zone is between the time and the year is surprisingly common.)

%Figure M3. MDL Output
Some of the dates are clearly just numbers, but our approach is to assume that when the user tells us that the column contains dates we should find the best fit. The samples include dates from a wide range of historical sources (e.g. Roman pottery dates) so we have elected to defer the date identification task to the user.

\subsubsection{Natural Language Processing}

\subsubsection{	Cross-Checking}

### 2015-07-06 conclusion.tex
Word => TeX formatting
Authors, fixed-width fonts, citations, dumb quotes.

@@
 While validating the algorithms on a large corpus, we also found that the number of distinct formats in the wild is surprisingly high, and demonstrates the wisdom of including general-purpose date parsing functions in data visualization tools, data cleaning tools and RDBMSes.
-In particular, it is interesting to note that the most prominent open source RDBMSes (MySQL and Postgres) both have a built-in version of DATEPARSE, possibly reflecting that this is a common need that gets implemented when users are empowered to add to a code base.
+In particular, it is interesting to note that the most prominent open source RDBMSes (MySQL and Postgres) both have a built-in version of \texttt{DATEPARSE}, possibly reflecting that this is a common need that gets implemented when users are empowered to add to a code base.

### 2015-07-06 intro.tex
Word => TeX formatting
Authors, fixed-width fonts, citations, dumb quotes.

@@
-In recent years, there has been a growth of interest in data visualization technologies for human-assisted data analysis using systems such as [1,10,11].
+In recent years, there has been a growth of interest in data visualization technologies for human-assisted data analysis using systems such as \cite{Polaris,Qlik,Spotfire}.
 While computers can provide high-speed and high-volume data processing, humans have domain knowledge and the ability to process data in parallel, notably by using our visual systems.
@@
 Consequently, most data visualization systems include facilities for defining such calculations as part of the data model being analyzed.
-The most effective systems allow users to define these calculations as part of the analytic interaction, which permits the user to stay in the flow of analysis [9].
+The most effective systems allow users to define these calculations as part of the analytic interaction, which permits the user to stay in the flow of analysis \cite{Morton}.
 
@@
 If the user does not own this process (e.g.
-it is the responsibility of another department), then there can be significant delays (including ``never.") More subtly, the result of updated external processing may not be compatible with the user's existing work, which can lead to more time lost reconciling the new data model with the existing analysis.
+it is the responsibility of another department), then there can be significant delays (including "never.") More subtly, the result of updated external processing may not be compatible with the user's existing work, which can lead to more time lost reconciling the new data model with the existing analysis.
 
@@
 Bad data is often discovered using visual analysis techniques (e.g.
-histograms or scatter plots) and it is most natural for the user to ``clean what she sees" instead of switching to a second tool.
-This leads to an ``adaptive" process whereby users will prefer leveraging existing tools in the analytics environment (no matter how well suited to the task) over switching to another application.
+histograms or scatter plots) and it is most natural for the user to "clean what she sees" instead of switching to a second tool.
+This leads to an "adaptive" process whereby users will prefer leveraging existing tools in the analytics environment (no matter how well suited to the task) over switching to another application.
 Thus a well-designed interactive visual analysis environment will provide tools that enable users to perform such cleaning tasks as interactively as possible.
@@
 In this paper, we shall be looking at an example of this problem in the context of the Tableau system.
-Tableau is a commercial visual analysis environment derived from the Polaris [1] system developed at Stanford.
+Tableau is a commercial visual analysis environment derived from the Polaris \cite{Polaris} system developed at Stanford.
 In addition to an interactive desktop application for creating data visualizations, the Tableau ecosystem also includes servers for publishing and sharing interactive visualizations.
@@
 \subsubsection{Tableau Public}
-Of particular relevance in the present work is a free version of the server environment called Tableau Public [12].
-Tableau Public (or ``Public") allows authors to publish data visualizations that can be shared with the general public.
-The published data sets are limited to 100K rows and must be stored using the Tableau Data Engine (or ``TDE") [2, 3], but are otherwise free to use the full analytic capabilities of the Tableau system.
+Of particular relevance in the present work is a free version of the server environment called Tableau Public \cite{Public}.
+Tableau Public (or "Public") allows authors to publish data visualizations that can be shared with the general public.
+The published data sets are limited to 100K rows and must be stored using the Tableau Data Engine (or "TDE")\cite{TDE1,TDE2}, but are otherwise free to use the full analytic capabilities of the Tableau system.
 In particular, the TDE provides native support for all analytic calculations defined by the authoring and publishing components.
@@
 \subsection{DATEPARSE}
-Among the recent additions to the Tableau function library is a function called DATEPARSE, the usability of which is the focus of the work in this paper.
+Among the recent additions to the Tableau function library is a function called \texttt{DATEPARSE}, the usability of which is the focus of the work in this paper.
 
 \subsubsection{	Scalar Dates}
-The SQL-99 standard defines three temporal scalar types: DATE, TIMESTAMP and TIME.
+The SQL-99 standard defines three temporal scalar types: \texttt{DATE}, \texttt{TIMESTAMP} and \texttt{TIME}.
 They are typically implemented as fixed-point types containing an offset from some epoch (\textit{e.g.} Julian Days.) This makes them compact to store and allows some temporal operations to be implemented very efficiently using arithmetic.
@@
 
-Tableau models the first two of these types as ``Date" and ``Date \& Time"; the third (pure time) is folded into Date \& Time by appending it to a fixed date of 1899-12-30.
+Tableau models the first two of these types as "Date" and "Date \& Time"; the third (pure time) is folded into Date \& Time by appending it to a fixed date of \texttt{1899-12-30}.
 From the analytic perspective, date types are dimensional (\textit{i.e.} independent variables) and can be used as either categorical (simply ordered) or quantitative (ordered with a distance metric) fields.
+
 Categorical dates have a natural hierarchy associated with them generated by calendar binning.
@@
 One of the oldest data preparation problems observed in Tableau is the parsing of date scalars so that users can perform these kinds of analyses.
-Training materials from the early days of the company included examples of how to convert columns of integers in the form yyyyMMdd to date scalars.
+Training materials from the early days of the company included examples of how to convert columns of integers in the form \texttt{yyyyMMdd} to date scalars.
 The documented solution at the time was to convert the integer to a string and perform some locale-dependent string operations before casting the string back to a date.
@@
 
-Our first solution to this problem was to add a new function to the Tableau calculation language called DATEPARSE.
+Our first solution to this problem was to add a new function to the Tableau calculation language called \texttt{DATEPARSE}.
 This function would take a string column and convert it to a datetime using a special purpose domain language for describing dates.
@@
 Oracle, Postgres and MySQL) and adding it to the TDE was straightforward, so it was a natural addition to the function library.
-We chose the date parsing syntax defined by the International Components for Unicode (or ICU) project [8] for the common domain language because it mirrored what was available in the Tableau code base, and translating this syntax into the formats used by various database vendors (e.g.
+
+We chose the date parsing syntax defined by the International Components for Unicode (or ICU) project \cite{icu} for the common domain language because it mirrored what was available in the Tableau code base, and translating this syntax into the formats used by various database vendors (e.g.
 MySQL, Oracle, Postgres) was relatively painless.
@@
 We examined a few months of workbooks that had been uploaded to Public after the release to see if and how it was being used.
-We found that although there were a number of new calculations using DATEPARSE, the error rate for the domain language syntax was about 15\%.
-DATEPARSE was capable of solving the problem, and some users were able to discover it, but the syntax did not appear to be easy to use reliably.
+We found that although there were a number of new calculations using \texttt{DATEPARSE}, the error rate for the domain language syntax was about 15\%.
+\texttt{DATEPARSE} was capable of solving the problem, and some users were able to discover it, but the syntax did not appear to be easy to use reliably.
 
@@
 Instead, we have developed two algorithms for automating the derivation of the format string, each of which uses a different machine learning technique.
-Both algorithms result in over 90\% parsing accuracy – and 100\% syntactic correctness because they are machine generated.
+Both algorithms result in over 90\% parsing accuracy -- and 100\% syntactic correctness because they are machine generated.
 
 % vidya: this probably should be its own section
+% hawkfish: Not sure it is big enough, but I can see the logic.
 
 \subsection{Related Work}
-Data preparation has been a known analytic bottleneck since at least the description of the Potter’s Wheel system [4].
-Since then, several other interactive data preparation systems have been proposed [5,6].
+Data preparation has been a known analytic bottleneck since at least the description of the Potter's Wheel system \cite{PottersWheel}.
+Since then, several other interactive data preparation systems have been proposed \cite{Wrangler,Refine}.
 While effective, these systems all make assumptions about possible date formats (e.g.
-the domain system in [4]), which we suggest are too restrictive for real world data.
-The Minimum Descriptive Length technique was first proposed in [7]. We have built on the version described in [4].
+the domain system in \cite{PottersWheel}), which we suggest are too restrictive for real world data.
+
+The Minimum Descriptive Length technique was first proposed in \cite{mdl}.  We have built on the version described in \cite{PottersWheel}.
+
 Related work on parsing languages, outlier detection.

### 2015-07-06 mdl.tex
Word => TeX formatting
Authors, fixed-width fonts, citations, dumb quotes.

@@
-The first algorithm is a Minimum Descriptive Length [7] approach derived from the domain system presented in [4].
+The first algorithm is a Minimum Descriptive Length \cite{mdl} approach derived from the domain system presented in \cite{PottersWheel}.
 We describe a number of extensions to their structure extraction system to support more complex redundancy, non-English locales, improved performance and date-specific pruning.
@@
 \subsection{Domains}
-[4] presents an algorithm for deriving a common structure for a set of strings by breaking each string down into a sequence of domains. These domains are described by an interface that includes:
+\cite{PottersWheel} presents an algorithm for deriving a common structure for a set of strings by breaking each string down into a sequence of domains.
+These domains are described by an interface that includes:
 \begin{itemize}
-\item A required inclusion function to test for membership in the domain (match)
-\item An optional function to compute the number of values in the domain with a given length (cardinality)
-\item An optional function to update statistics for the domain based on a given value (updateStatistics)
-\item An optional function to prevent consideration of a domain that is redundant (isRedundantAfter).
+\item A required inclusion function to test for membership in the domain (\texttt{match})
+\item An optional function to compute the number of values in the domain with a given length (\texttt{cardinality})
+\item An optional function to update statistics for the domain based on a given value (\texttt{updateStatistics})
+\item An optional function to prevent consideration of a domain that is redundant (\texttt{isRedundantAfter}).
 \end{itemize}
@@
 days and hours) and there are often no field delimiters (e.g.
-2012Mar06134427).
+\texttt{2012Mar06134427}).
 To handle this, we have extended the existing pruning API with two other sets of domain identifiers:
@@
 \item The set of date parts cannot contain place value gaps (e.g. structures that have year and day without month are removed.)
-\item Similarly, the set of time parts cannot contain place value gaps and must also be in place value order (times are never written in orders such as mhs.)
+\item Similarly, the set of time parts cannot contain place value gaps and must also be in place value order (times are never written in orders such as \texttt{mhs}.)
 \item The existence of time parts cannot make dates incomplete (e.g. patterns like year-day-hour are removed.)
@@
 \item Since we have a set of samples, we can apply the candidate format to the strings to see how well it performs. Formats with fewer parse errors are preferred.
-\item Date parts can be considered a place-value system, so we prefer “more significant” components (e.g. month-day-year over hour-minute-second).
+\item Date parts can be considered a place-value system, so we prefer "more significant" components (e.g. month-day-year over hour-minute-second).
 \item If two formats from different locales give the same results, prefer the original column locale. The sample set may have missed an example where this could be important.

### 2015-07-06 paper.tex
Word => TeX formatting
Authors, fixed-width fonts, citations, dumb quotes.

@@
 
+%
+\def\sharedaffiliation{%
+\end{tabular}
+\begin{tabular}{c}}
+%
+
 \begin{document}
 
-\title{Does Anybody Really Know What Time It Is?}
-\subtitle{Automating the Extraction of Date Scalars}
+\conferenceinfo{SIGMOD}{'16 SanFrancisco, CA USA}
+
+\title{Does Anybody Really Know What Time It Is?\\
+Automating the Extraction of Date Scalars}
 %
@@
 Richard Wesley \\
-%       \affaddr{Institute for Clarity in Documentation}\\
-%       \affaddr{1932 Wallamaloo Lane}\\
-%       \affaddr{Wallamaloo, New Zealand}\\
-%       \email{trovato@corporation.com}
+%       \affaddr{Tableau Software}\\
+%       \affaddr{837 North 34th Street}\\
+%       \affaddr{Seattle, WA 98103}\\
+%      \email{hawkfish@tableau.com}
 % 2nd. author
@@
 Vidya Setlur\\
-%       \affaddr{Institute for Clarity in Documentation}\\
-%       \affaddr{P.O. Box 1212}\\
-%       \affaddr{Dublin, Ohio 43017-6221}\\
-%       \email{webmaster@marysville-ohio.com}
+%       \affaddr{Tableau Software}\\
+ %      \affaddr{837 North 34th Street}\\
+%       \affaddr{Seattle, WA 98103}\\
+ %      \email{vsetlur@tableau.com}
 % 3rd. author
@@
 Dan Cory\\
-%       \affaddr{The Th{\o}rv{\"a}ld Group}\\
-%       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
-%       \affaddr{Hekla, Iceland}\\
-%       \email{larst@affiliation.org}
+%       \affaddr{Tableau Software}\\
+%       \affaddr{837 North 34th Street}\\
+%       \affaddr{Seattle, WA 98103}\\
+%       \email{dcory@tableau.com}
+%       \email{\{hawkfish,vsetlur,dcory\}@tableau.com}
+\sharedaffiliation
+       \affaddr{Tableau Software}\\
+       \affaddr{837 North 34th Street}\\
+       \affaddr{Seattle, WA 98103}\\
+       \email{\{hawkfish,vsetlur,dcory\}@tableau.com}
 }
-% There's nothing stopping you putting the seventh, eighth, etc.
-% author on the opening page (as the 'third row') but we ask,
-% for aesthetic reasons that you place these 'additional authors'
-% in the \additional authors block, viz.
-\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
-email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
-(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
-\date{30 July 1999}
 % Just remember to make sure that the TOTAL number of authors
@@
 Several databases include date parsing ``mini-languages" to cover the wide range of possible formats.
-Tableau has recently provided a \\DATEPARSE function with a single syntax that is translated into the nearest equivalent in the underlying database.
+Tableau has recently provided a function called \texttt{DATEPARSE}  with a single syntax that is translated into the nearest equivalent in the underlying database.
 Subsequent analysis of customer usage of this function shows that the parsing language syntax was difficult for users to master.
@@
 \section{Acknowledgments}
-We would like to thank to Pat Hanrahan for encouraging this work; Roger Hau for providing the resources to productize it; and Jason King, whose harrowing live demonstration of manual DATEPARSE in front of several thousand people inspired this project.
+We would like to thank to Pat Hanrahan for encouraging this work; Roger Hau for providing the resources to productize it; and Jason King, whose harrowing live demonstration of manual \texttt{DATEPARSE} in front of several thousand people inspired this project.
 

### 2015-07-06 parameters.tex
Word => TeX formatting
Authors, fixed-width fonts, citations, dumb quotes.

@@
 We selected a large number of string columns whose names suggested that they might contain temporal data.
-We included words from multiple languages besides English and extracted the column’s locale (actually the collation) along with the data.
+We included words from multiple languages besides English and extracted the column's locale (actually the collation) along with the data.
 
 \subsubsection{NULL Filtering}
-Each column was then reduced to a maximum sample set of 32. We excluded domain values that appeared to be representations of NULL:
+Each column was then reduced to a maximum sample set of 32. We excluded domain values that appeared to be representations of \texttt{NULL}:
 \begin{itemize}
-\item Values containing the substring NULL
-\item Values containing no digits and at most one non-whitespace character (e.g. " / / ")
-\item Values on a list of empirically determined common NULL values (e.g. 0000-00-00, NaN)
+\item Values containing the substring \texttt{NULL}
+\item Values containing no digits and at most one non-whitespace character (e.g. \texttt{" / / "})
+\item Values on a list of empirically determined common \texttt{NULL} values (e.g. \texttt{0000-00-00}, \texttt{NaN})
 \end{itemize}
-
 Columns that had no remaining valid samples were discarded.
@@
 \subsubsection{Sampling}
-The remaining samples were hashed and sorted on the hash value, with the top 32 values being retained as the column’s sample.
+The remaining samples were hashed and sorted on the hash value, with the top 32 values being retained as the column's sample.
 We can increase the sample size if needed, but for dates, it appears to be an adequate number.
@@
 After NULL filtering, there remained one common class of date representation that was unsuited for parsing via our date format syntax, namely numeric timestamps.
-This included Unix epoch timestamps (expressed as second, millisecond or microsecond counts from 1970-01-01) and Microsoft Excel timestamps (expressed fractional days since 1900-01-01).
+This included Unix epoch timestamps (expressed as second, millisecond or microsecond counts from \texttt{1970-01-01}) and Microsoft Excel timestamps (expressed fractional days since \texttt{1900-01-01}).
 A simple test to check that the column was numeric and inside a specific range of values representing recent dates allowed us to tag these columns to avoid analyzing them further (and incidentally to identify them for generating a simple date extraction calculation for the user.)
@@
 
-The syntax is documented at the ICU web site [8]. While fairly complete, it has a few limitations that we ran into when evaluating the algorithms:
+The syntax is documented at the ICU web site \cite{icu}. While fairly complete, it has a few limitations that we ran into when evaluating the algorithms:
 \begin{itemize}
-\item No support for 4 letter year abbreviations (e.g. Sept.)
-\item No support for ordinal days (e.g. July 4th)
-\item No support for quarter postfix notation (e.g. 2Q)
-\item No support for variant meridian markers (e.g. a.m.)
+\item No support for 4 letter year abbreviations (e.g. \texttt{Sept.})
+\item No support for ordinal days (e.g. \texttt{July 4th})
+\item No support for quarter postfix notation (e.g. \texttt{2Q})
+\item No support for variant meridian markers (e.g. \texttt{a.m.})
 \end{itemize}
@@
 The use of lexicographical case in the meta-symbols of the format language can be confusing (e.g.
-y is used for years, but M is used for months while m is used for minutes.) Another advantage of an automated algorithm is that it hides such problems from most users, significantly improving the usability of the function.
+\texttt{y} is used for years, but \texttt{M} is used for months while \texttt{m} is used for minutes.) Another advantage of an automated algorithm is that it hides such problems from most users, significantly improving the usability of the function.
 
@@
 
-ICU’s date parsing APIs allow the specification of default values for parts when a format does not contain them.
-In our implementation, all time fields are set to 0 (midnight) and the date fields are set based on whether the format contains any date part specifications.
-When date parts are present, we use 2000-01-01 as the set of default date parts as it is the start of a leap year.
-When dealing with pure time formats, we use 1899-12-30, (which is the date used internally by Tableau to signal pure time values.)
+ICU's date parsing APIs allow the specification of default values for parts when a format does not contain them.
+In our implementation, all time fields are set to \texttt{0} (midnight) and the date fields are set based on whether the format contains any date part specifications.
+When date parts are present, we use \texttt{2000-01-01} as the set of default date parts as it is the start of a leap year.
+When dealing with pure time formats, we use \texttt{1899-12-30}, (which is the date used internally by Tableau to signal pure time values.)
 

### 2015-07-06 testing.tex
Word => TeX formatting
Authors, fixed-width fonts, citations, dumb quotes.

@@
 Files over 1MB were manually reviewed and those that did not contain dates removed.
-There were 30968 files in the resulting set.
+There were 30,968 files in the resulting set.
 Most database and spreadsheet systems already detect a limited set of date formats.
-For instance, typing the string “12/31/1999” into Microsoft Excel, is automatically interpreted as the date 1999-12-31.
+For instance, typing the string \texttt{"12/31/1999"} into Microsoft Excel, is automatically interpreted as the date \texttt{1999-12-31}.
 The Microsoft Jet library that Tableau used to read these text files detects a few date formats as well.
@@
 It is likely that some of these files were older files republished to Tableau Public.
-There were 31546 files in the resulting set.
+There were 31,546 files in the resulting set.
 
@@
 A sample of each produced date format was manually labeled.
-This allowed us to quickly skip over very common formats like MM/dd/yyyy and focus our efforts on much less common formats.
+This allowed us to quickly skip over very common formats like \texttt{MM/dd/yyyy} and focus our efforts on much less common formats.
 The samples were judged as to whether the produced format was reasonable and were tagged with correct formats if the produced format was unreasonable.
@@
 Further manual tagging focused on the files most likely to represent dates.
-Many of the files in the collection are not actually dates, such as fields named “updated by” (which contain “date”).
-A random sample of 850 columns named exactly ``date", ``time" or ``month" (case-insensitive) were manually judged.
+Many of the files in the collection are not actually dates, such as fields named "updated by" (which contain "date").
+A random sample of 850 columns named exactly "date", "time" or "month" (case-insensitive) were manually judged.
 
@@
 (As a column store, the TDE can often supply such domains without a full table scan, further improving responsiveness.)
+
 The validation speed is the average time needed to parse a value, and provides an estimate of how fast an ICU-based implementation (such as the TDE) can process string values into scalars and works out to 620K values per core per second.
@@
 To examine the error rate in greater detail, we turn now to Figure M2.
+
 On the left hand side, we can see that the algorithm found 744 distinct formats that parsed 13,424 files with no error.

### 2015-07-06 futurework.tex
Word => TeX formatting
Authors, fixed-width fonts, citations, dumb quotes.
More formatting
Redo all the quotes (I’d forgotten how annoying TeX’s quote mark
conventions are). Format e.g. everywhere (next time I’ll define a
macro…)

@@
 
-In this paper, we have only considered the parsing of strings, but dates are often formatted as integers (e.g.
-201507016).
-It is significantly faster to decompose integers into date parts using arithmetic operations (e.g.
-mod and div) than by using slower, locale-sensitive string parsing functions.
+In this paper, we have only considered the parsing of strings, but dates are often formatted as integers (\textit{e.g.} \texttt{201507016}).
+It is significantly faster to decompose integers into date parts using arithmetic operations (\textit{e.g.} mod and div) than by using slower, locale-sensitive string parsing functions.
 Moreover, the number of possible formats is low enough that we may be able to enumerate them.
 Timestamps are another form of numeric date for which we intend to automate the preparation.
-In the course of our research, we have also identified a number of date part variants (e.g.
-ordinal dates, 4 letter month abbreviations, alternate meridian markers and postfix quarter syntax) that we would like to commend to the ICU project and maybe provide implementations for.
+
+In the course of our research, we have also identified a number of date part variants (\textit{e.g.} ordinal dates, four-letter month abbreviations, alternate meridian markers and postfix quarter syntax) that we would like to commend to the ICU project and maybe provide implementations for.

### 2015-07-06 intro.tex
More formatting
Redo all the quotes (I’d forgotten how annoying TeX’s quote mark
conventions are). Format e.g. everywhere (next time I’ll define a
macro…)

@@
 Solutions to this problem are often provided by data preparation tools external to the visual analysis environment, which requires the user to break their analytic flow, launch another tool and reprocess their data before returning to their analysis.
-If the user does not own this process (e.g.
-it is the responsibility of another department), then there can be significant delays (including "never.") More subtly, the result of updated external processing may not be compatible with the user's existing work, which can lead to more time lost reconciling the new data model with the existing analysis.
+If the user does not own this process (\textit{e.g.} it is the responsibility of another department), then there can be significant delays (including ``never.'') More subtly, the result of updated external processing may not be compatible with the user's existing work, which can lead to more time lost reconciling the new data model with the existing analysis.
 
@@
 Bad data is often discovered using visual analysis techniques (e.g.
-histograms or scatter plots) and it is most natural for the user to "clean what she sees" instead of switching to a second tool.
-This leads to an "adaptive" process whereby users will prefer leveraging existing tools in the analytics environment (no matter how well suited to the task) over switching to another application.
+histograms or scatter plots) and it is most natural for the user to ``clean what she sees'' instead of switching to a second tool.
+This leads to an ``adaptive'' process whereby users will prefer leveraging existing tools in the analytics environment (no matter how well suited to the task) over switching to another application.
 Thus a well-designed interactive visual analysis environment will provide tools that enable users to perform such cleaning tasks as interactively as possible.
@@
 Of particular relevance in the present work is a free version of the server environment called Tableau Public \cite{Public}.
-Tableau Public (or "Public") allows authors to publish data visualizations that can be shared with the general public.
-The published data sets are limited to 100K rows and must be stored using the Tableau Data Engine (or "TDE")\cite{TDE1,TDE2}, but are otherwise free to use the full analytic capabilities of the Tableau system.
+Tableau Public (or ``Public'') allows authors to publish data visualizations that can be shared with the general public.
+The published data sets are limited to 100K rows and must be stored using the Tableau Data Engine (or ``TDE'')\cite{TDE1,TDE2}, but are otherwise free to use the full analytic capabilities of the Tableau system.
 In particular, the TDE provides native support for all analytic calculations defined by the authoring and publishing components.
@@
 
-Tableau models the first two of these types as "Date" and "Date \& Time"; the third (pure time) is folded into Date \& Time by appending it to a fixed date of \texttt{1899-12-30}.
+Tableau models the first two of these types as ``Date'' and ``Date \& Time''; the third (pure time) is folded into Date \& Time by appending it to a fixed date of \texttt{1899-12-30}.
 From the analytic perspective, date types are dimensional (\textit{i.e.} independent variables) and can be used as either categorical (simply ordered) or quantitative (ordered with a distance metric) fields.
@@
 \item String operations are notoriously slow compared to scalar operations (typically 10-100x slower)
-\item Default parsing of date formats is locale-dependent, and may not work when the workbook is shared across an international organization (e.g. between the US and European offices)
+\item Default parsing of date formats is locale-dependent, and may not work when the workbook is shared across an international organization (\textit{e.g.} between the US and European offices)
 \item The expression code was hard to understand and maintain because it used a verbose, general-purpose string-handling syntax instead of a domain language.
@@
 This function would take a string column and convert it to a datetime using a special purpose domain language for describing dates.
-Such functions exist in a number of databases (e.g.
-Oracle, Postgres and MySQL) and adding it to the TDE was straightforward, so it was a natural addition to the function library.
+Such functions exist in a number of databases (\textit{e.g.} Oracle, Postgres and MySQL) and adding it to the TDE was straightforward, so it was a natural addition to the function library.
 
-We chose the date parsing syntax defined by the International Components for Unicode (or ICU) project \cite{icu} for the common domain language because it mirrored what was available in the Tableau code base, and translating this syntax into the formats used by various database vendors (e.g.
-MySQL, Oracle, Postgres) was relatively painless.
+We chose the date parsing syntax defined by the International Components for Unicode (or ICU) project \cite{icu} for the common domain language because it mirrored what was available in the Tableau code base, and translating this syntax into the formats used by various database vendors (\textit{e.g.} MySQL, Oracle, Postgres) was relatively painless.
 There are a few patches of non-overlapping functionality, but they tend to be obscure and we provide warnings when functionality is missing.
@@
 Since then, several other interactive data preparation systems have been proposed \cite{Wrangler,Refine}.
-While effective, these systems all make assumptions about possible date formats (e.g.
-the domain system in \cite{PottersWheel}), which we suggest are too restrictive for real world data.
+While effective, these systems all make assumptions about possible date formats (\textit{e.g.} the domain system in \cite{PottersWheel}), which we suggest are too restrictive for real world data.
 

### 2015-07-06 mdl.tex
More formatting
Redo all the quotes (I’d forgotten how annoying TeX’s quote mark
conventions are). Format e.g. everywhere (next time I’ll define a
macro…)

@@
 A difficulty in using this kind of structure extraction is that the algorithm for enumerating structures is exponential in the number of domains.
-This is especially true in the date format problem because there are identical domains (e.g.
-months and meridian hours), nearly identical domains (e.g.
-days and hours) and there are often no field delimiters (e.g.
+This is especially true in the date format problem because there are identical domains (\textit{e.g.} months and meridian hours), nearly identical domains (\textit{e.g.} days and hours) and there are often no field delimiters (e.g.
 \texttt{2012Mar06134427}).
@@
 Domain parameterization is an important part of generating compact representations via MDL, but it creates problems for date recognition.
-For example, if a set of dates contains a constant month string (e.g.
-all values are in September) it is important to keep track of the month name domain.
+For example, if a set of dates contains a constant month string (\textit{e.g.} all values are in September) it is important to keep track of the month name domain.
 Consequently, when we parameterize a constant generic <Word> domain, we tag it with the date part domain that it matches (if any).
-We then need to apply an additional pruning step to remove any structures that also found an equivalent domain (e.g.
-two-digit month).
+We then need to apply an additional pruning step to remove any structures that also found an equivalent domain (\textit{e.g.} two-digit month).
 These rules are equivalent to the context-based redundancy rules above, but have to be applied again after parameterization of generics.
@@
 \begin{itemize}
-\item The set of date parts cannot contain place value gaps (e.g. structures that have year and day without month are removed.)
+\item The set of date parts cannot contain place value gaps (\textit{e.g.} structures that have year and day without month are removed.)
 \item Similarly, the set of time parts cannot contain place value gaps and must also be in place value order (times are never written in orders such as \texttt{mhs}.)
-\item The existence of time parts cannot make dates incomplete (e.g. patterns like year-day-hour are removed.)
+\item The existence of time parts cannot make dates incomplete (\textit{e.g.} patterns like year-day-hour are removed.)
 \item Two digit years require special handling.
@@
 Knowing the locale of a string is not always helpful.
-In our data set, we found numerous cases where the locale was specified (e.g.
-Sweden) but the data was actually in English.
+In our data set, we found numerous cases where the locale was specified (\textit{e.g.} Sweden) but the data was actually in English.
 Accordingly, we test both locales and rank the combined results.
@@
 \item Since we have a set of samples, we can apply the candidate format to the strings to see how well it performs. Formats with fewer parse errors are preferred.
-\item Date parts can be considered a place-value system, so we prefer "more significant" components (e.g. month-day-year over hour-minute-second).
+\item Date parts can be considered a place-value system, so we prefer ``more significant'' components (\textit{e.g.} month-day-year over hour-minute-second).
 \item If two formats from different locales give the same results, prefer the original column locale. The sample set may have missed an example where this could be important.
-\item If the format has an ambiguous date order (e.g.
-all days are less than 12), then prefer the default date order of the locale.
+\item If the format has an ambiguous date order (\textit{e.g.} all days are less than 12), then prefer the default date order of the locale.
 Again, the sample set may have missed a counterexample, so this is the best option.

### 2015-07-06 parameters.tex
More formatting
Redo all the quotes (I’d forgotten how annoying TeX’s quote mark
conventions are). Format e.g. everywhere (next time I’ll define a
macro…)

@@
 \item Values containing the substring \texttt{NULL}
-\item Values containing no digits and at most one non-whitespace character (e.g. \texttt{" / / "})
-\item Values on a list of empirically determined common \texttt{NULL} values (e.g. \texttt{0000-00-00}, \texttt{NaN})
+\item Values containing no digits and at most one non-whitespace character (\textit{e.g.} \texttt{" / / "})
+\item Values on a list of empirically determined common \texttt{NULL} values (\textit{e.g.} \texttt{0000-00-00}, \texttt{NaN})
 \end{itemize}
@@
 \begin{itemize}
-\item No support for 4 letter year abbreviations (e.g. \texttt{Sept.})
-\item No support for ordinal days (e.g. \texttt{July 4th})
-\item No support for quarter postfix notation (e.g. \texttt{2Q})
-\item No support for variant meridian markers (e.g. \texttt{a.m.})
+\item No support for 4 letter year abbreviations (\textit{e.g.} \texttt{Sept.})
+\item No support for ordinal days (\textit{e.g.} \texttt{July 4th})
+\item No support for quarter postfix notation (\textit{e.g.} \texttt{2Q})
+\item No support for variant meridian markers (\textit{e.g.} \texttt{a.m.})
 \end{itemize}

### 2015-07-06 testing.tex
More formatting
Redo all the quotes (I’d forgotten how annoying TeX’s quote mark
conventions are). Format e.g. everywhere (next time I’ll define a
macro…)

@@
 Further manual tagging focused on the files most likely to represent dates.
-Many of the files in the collection are not actually dates, such as fields named "updated by" (which contain "date").
-A random sample of 850 columns named exactly "date", "time" or "month" (case-insensitive) were manually judged.
+Many of the files in the collection are not actually dates, such as fields named ``updated by'' (which contain ``date'').
+A random sample of 850 columns named exactly ``date", ``time'' or ``month'' (case-insensitive) were manually judged.
 

### 2015-07-06 intro.tex
Added first two figures

@@
 \subsubsection{	Scalar Dates}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=\columnwidth]{figures/FigureI1}
+\caption{Categorical Date Scalars.}
+\label{fig:I1}
+\end{figure}
+
 The SQL-99 standard defines three temporal scalar types: \texttt{DATE}, \texttt{TIMESTAMP} and \texttt{TIME}.
@@
 Categorical dates have a natural hierarchy associated with them generated by calendar binning.
-The visualization in Figure I1 shows an example of a bar chart employing binned categorical dates in a year/quarter hierarchy.
+The visualization in Figure \ref{fig:I1} shows an example of a bar chart employing binned categorical dates in a year/quarter hierarchy.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=\columnwidth]{figures/FigureI2}
+\caption{Quantitative Date Scalars.}
+\label{fig:I2}
+\end{figure}
+
+Quantitative dates are typically used for time series on an axis that maps the underlying distance measure to display pixel distance.
+The quantitative analog to categorical binning is truncation whereby all the dates in a bin are represented by the first date in the bin.
+Truncation accurately preserves the distance semantics while enabling roll-up to coarser levels of detail.
+The visualization in Figure \ref{fig:I2} shows the same sales data in a time series rolled up to the quarter level.
+
+These types of visualizations are difficult to specify with simple categorical string representations of dates.
+Thus the conversion of unparsed strings to date scalars is an important component of data modeling for analytics.
 

### 2015-07-06 intro.tex
Added table 1.

@@
 Table 1 shows a selection of unusual date formats found on Public.
+The first example shows a time zone in the middle of the date and a year after the time; the second shows a leading unmatched bracket and a colon between the date and time components; the third shows confusion between the seconds' decimal point and the time part delimiter; the fourth shows a two digit year apostrophe on a four digit year and the fifth shows a dash separating the date and time components.
 
-% INSERT TABLE HERE
+\begin{table}[ht]
+\centering
+%\begin{tabular}{ |l|l|l| }
+\begin{tabular}{|p{0.498\linewidth}| p{0.485\linewidth}|}
+\hline
+\centering
+\textbf{ICU Format} & \textbf{Example}\\ \hline
+\scriptsize{EEE MMM dd HH:mm:ss zzz yyyy} & \scriptsize{Fri Apr 01 02:09:27 EDT 2011}\\ \hline
+\scriptsize{[dd/MMM/yyyy:HH:mm:ss} & \scriptsize{[10/Aug/2014:09:30:40}\\ \hline
+\scriptsize{dd-MMM-yy hh.mm.ss.SSSSSS a} & \scriptsize{01-OCT-13 01.09.00.000000 PM}\\ \hline
+\scriptsize{MM ''yyyy} & \scriptsize{01 '2013}\\ \hline
+\scriptsize{MM/dd/yyyy - HH:mm} & \scriptsize{04/09/2014 - 23:47}\\ \hline
+\end{tabular}
+\label{tab:dateformats}
+\caption{Unusual Date Formats.}
+\end{table}
 

### 2015-07-06 testing.tex
Added MDL Parsing Statistics Table
Added figures M2 and M3

@@
 
-%Figure M1. MDL Parsing Statistics
+\begin{table}[ht]
+\begin{tabular}{|p{0.48\linewidth}| p{0.48\linewidth}|}
+\hline
+\textbf{Number of Records} & 31,546\\ \hline
+\textbf{Error Rate} & 27.95\% \\ \hline
+\textbf{Analysis Speed ($\mu s$)} & 2,245.04 \\ \hline
+\textbf{Validation Speed ($\mu s$)} & 1.65 \\ \hline
+\textbf{Median Not Null} & 50 \\ \hline
+\end{tabular}
+\label{tab:mdlstats}
+\caption{MDL Parsing Statistics.}
+\end{table}
+
 To test the MDL algorithm, we ran it over the set of samples from each validation file to generate a ranked list of formats for the file.
@@
 In cases where we generated multiple formats and the main format produced errors, we applied the second format to the unparsed strings.
-The summary statistics from this processing are presented in Figure M1.
+The summary statistics from this processing are presented in Table 2.
 
@@
 
-%Figure M2. MDL Error Rate
+\begin{figure}[ht]
+\centering
+\includegraphics[width=\columnwidth]{figures/FigureM2}
+\caption{MDL Error Rate}
+\label{fig:M2}
+\end{figure}
+
 The error rate reflects the fact that only about 40\% of the files have an associated format that parses the non-null values without error.
-To examine the error rate in greater detail, we turn now to Figure M2.
+To examine the error rate in greater detail, we turn now to Figure \ref{fig:M2}.
 
@@
 What do these formats look like?
-Figure M3 shows a histogram of the 25 most common formats containing a year format code at the 5\% threshold, color-coded by error rate.
+Figure \ref{fig:M3} shows a histogram of the 25 most common formats containing a year format code at the 5\% threshold, color-coded by error rate.
 (A sample value is provided to the right of each bar for illustrative purposes.) The formats have also been filtered to files with at least 5 samples.
@@
 
-%Figure M3. MDL Output
+\begin{figure}[ht]
+\centering
+\includegraphics[width=\columnwidth]{figures/FigureM3}
+\caption{MDL Output}
+\label{fig:M3}
+\end{figure}
+
 Some of the dates are clearly just numbers, but our approach is to assume that when the user tells us that the column contains dates we should find the best fit.

### 2015-07-06 intro.tex
Fixed Polaris citation
Few more citation reference fixes

@@
 Consequently, most data visualization systems include facilities for defining such calculations as part of the data model being analyzed.
-The most effective systems allow users to define these calculations as part of the analytic interaction, which permits the user to stay in the flow of analysis \cite{Morton}.
+The most effective systems allow users to define these calculations as part of the analytic interaction, which permits the user to stay in the flow of analysis \cite{Morton:2012}.
 
@@
 In this paper, we shall be looking at an example of this problem in the context of the Tableau system.
-Tableau is a commercial visual analysis environment derived from the Polaris \cite{Polaris} system developed at Stanford.
+Tableau is a commercial visual analysis environment derived from the Polaris \cite{Stolte:2008} system developed at Stanford.
 In addition to an interactive desktop application for creating data visualizations, the Tableau ecosystem also includes servers for publishing and sharing interactive visualizations.
@@
 \subsection{Related Work}
-Data preparation has been a known analytic bottleneck since at least the description of the Potter's Wheel system \cite{PottersWheel}.
-Since then, several other interactive data preparation systems have been proposed \cite{Wrangler,Refine}.
+Data preparation has been a known analytic bottleneck since at least the description of the Potter's Wheel system \cite{Raman:2001}.
+Since then, several other interactive data preparation systems have been proposed \cite{Kandel:2011,Refine}.
 While effective, these systems all make assumptions about possible date formats (\textit{e.g.} the domain system in \cite{PottersWheel}), which we suggest are too restrictive for real world data.

### 2015-07-06 futurework.tex
Add some consistency macros

--- 2015-07-04 futurework.tex
@@
 
-In this paper, we have only considered the parsing of strings, but dates are often formatted as integers (e.g.
-201507016).
-It is significantly faster to decompose integers into date parts using arithmetic operations (e.g.
-mod and div) than by using slower, locale-sensitive string parsing functions.
+In this paper, we have only considered the parsing of strings, but dates are often formatted as integers (\eg \texttt{201507016}).
+It is significantly faster to decompose integers into date parts using arithmetic operations (\eg \texttt{mod} and \texttt{div}) than by using slower, locale-sensitive string parsing functions.
 Moreover, the number of possible formats is low enough that we may be able to enumerate them.
 Timestamps are another form of numeric date for which we intend to automate the preparation.
-In the course of our research, we have also identified a number of date part variants (e.g.
-ordinal dates, 4 letter month abbreviations, alternate meridian markers and postfix quarter syntax) that we would like to commend to the ICU project and maybe provide implementations for.
+
+In the course of our research, we have also identified a number of date part variants (\eg ordinal dates, four-letter month abbreviations, alternate meridian markers and postfix quarter syntax) that we would like to commend to the ICU project and maybe provide implementations for.

### 2015-07-06 intro.tex
Add some consistency macros

--- 2015-07-06 intro.tex
@@
-In recent years, there has been a growth of interest in data visualization technologies for human-assisted data analysis using systems such as \cite{Polaris,Qlik,Spotfire}.
+In recent years, there has been a growth of interest in data visualization technologies for human-assisted data analysis using systems such as \cite{Stolte:2008,Qlik,Ahlberg:1996}.
 While computers can provide high-speed and high-volume data processing, humans have domain knowledge and the ability to process data in parallel, notably by using our visual systems.
@@
 Consequently, most data visualization systems include facilities for defining such calculations as part of the data model being analyzed.
-The most effective systems allow users to define these calculations as part of the analytic interaction, which permits the user to stay in the flow of analysis \cite{Morton}.
+The most effective systems allow users to define these calculations as part of the analytic interaction, which permits the user to stay in the flow of analysis~\cite{Morton:2012}.
 
@@
 Solutions to this problem are often provided by data preparation tools external to the visual analysis environment, which requires the user to break their analytic flow, launch another tool and reprocess their data before returning to their analysis.
-If the user does not own this process (\textit{e.g.} it is the responsibility of another department), then there can be significant delays (including ``never.'') More subtly, the result of updated external processing may not be compatible with the user's existing work, which can lead to more time lost reconciling the new data model with the existing analysis.
+If the user does not own this process (\eg it is the responsibility of another department), then there can be significant delays (including ``never.'') More subtly, the result of updated external processing may not be compatible with the user's existing work, which can lead to more time lost reconciling the new data model with the existing analysis.
 
@@
 In this paper, we shall be looking at an example of this problem in the context of the Tableau system.
-Tableau is a commercial visual analysis environment derived from the Polaris \cite{Polaris} system developed at Stanford.
+Tableau is a commercial visual analysis environment derived from the Polaris~\cite{Stolte:2008} system developed at Stanford.
 In addition to an interactive desktop application for creating data visualizations, the Tableau ecosystem also includes servers for publishing and sharing interactive visualizations.
@@
 \subsubsection{Tableau Public}
-Of particular relevance in the present work is a free version of the server environment called Tableau Public \cite{Public}.
+Of particular relevance in the present work is a free version of the server environment called Tableau Public.
 Tableau Public (or ``Public'') allows authors to publish data visualizations that can be shared with the general public.
-The published data sets are limited to 100K rows and must be stored using the Tableau Data Engine (or ``TDE'')\cite{TDE1,TDE2}, but are otherwise free to use the full analytic capabilities of the Tableau system.
+The published data sets are limited to 100K rows and must be stored using the Tableau Data Engine (or ``TDE'')~\cite{Wesley:2011,Wesley:2014}, but are otherwise free to use the full analytic capabilities of the Tableau system.
 In particular, the TDE provides native support for all analytic calculations defined by the authoring and publishing components.
@@
 \subsection{DATEPARSE}
-Among the recent additions to the Tableau function library is a function called \texttt{DATEPARSE}, the usability of which is the focus of the work in this paper.
+Among the recent additions to the Tableau function library is a function called \dateparse, the usability of which is the focus of the work in this paper.
 
@@
 The SQL-99 standard defines three temporal scalar types: \texttt{DATE}, \texttt{TIMESTAMP} and \texttt{TIME}.
-They are typically implemented as fixed-point types containing an offset from some epoch (\textit{e.g.} Julian Days.) This makes them compact to store and allows some temporal operations to be implemented very efficiently using arithmetic.
+They are typically implemented as fixed-point types containing an offset from some epoch (\eg Julian Days.) This makes them compact to store and allows some temporal operations to be implemented very efficiently using arithmetic.
 Thus from the RDBMS perspective, representing dates in scalar form provides numerous benefits for users, both in terms of available operations and query performance.
@@
 Tableau models the first two of these types as ``Date'' and ``Date \& Time''; the third (pure time) is folded into Date \& Time by appending it to a fixed date of \texttt{1899-12-30}.
-From the analytic perspective, date types are dimensional (\textit{i.e.} independent variables) and can be used as either categorical (simply ordered) or quantitative (ordered with a distance metric) fields.
+From the analytic perspective, date types are dimensional (\ie independent variables) and can be used as either categorical (simply ordered) or quantitative (ordered with a distance metric) fields.
 
@@
 \item String operations are notoriously slow compared to scalar operations (typically 10-100x slower)
-\item Default parsing of date formats is locale-dependent, and may not work when the workbook is shared across an international organization (\textit{e.g.} between the US and European offices)
+\item Default parsing of date formats is locale-dependent, and may not work when the workbook is shared across an international organization (\eg between the US and European offices)
 \item The expression code was hard to understand and maintain because it used a verbose, general-purpose string-handling syntax instead of a domain language.
@@
 
-Our first solution to this problem was to add a new function to the Tableau calculation language called \texttt{DATEPARSE}.
+Our first solution to this problem was to add a new function to the Tableau calculation language called \dateparse.
 This function would take a string column and convert it to a datetime using a special purpose domain language for describing dates.
-Such functions exist in a number of databases (\textit{e.g.} Oracle, Postgres and MySQL) and adding it to the TDE was straightforward, so it was a natural addition to the function library.
+Such functions exist in a number of databases (\eg Oracle, Postgres and MySQL) and adding it to the TDE was straightforward, so it was a natural addition to the function library.
 
-We chose the date parsing syntax defined by the International Components for Unicode (or ICU) project \cite{icu} for the common domain language because it mirrored what was available in the Tableau code base, and translating this syntax into the formats used by various database vendors (\textit{e.g.} MySQL, Oracle, Postgres) was relatively painless.
+We chose the date parsing syntax defined by the International Components for Unicode (or ICU) project~\cite{ICU} for the common domain language because it mirrored what was available in the Tableau code base, and translating this syntax into the formats used by various database vendors (\eg MySQL, Oracle, Postgres) was relatively painless.
 There are a few patches of non-overlapping functionality, but they tend to be obscure and we provide warnings when functionality is missing.
@@
 We examined a few months of workbooks that had been uploaded to Public after the release to see if and how it was being used.
-We found that although there were a number of new calculations using \texttt{DATEPARSE}, the error rate for the domain language syntax was about 15\%.
-\texttt{DATEPARSE} was capable of solving the problem, and some users were able to discover it, but the syntax did not appear to be easy to use reliably.
+We found that although there were a number of new calculations using \dateparse, the error rate for the domain language syntax was about 15\%.
+\dateparse was capable of solving the problem, and some users were able to discover it, but the syntax did not appear to be easy to use reliably.
 
@@
 \subsection{Related Work}
-Data preparation has been a known analytic bottleneck since at least the description of the Potter's Wheel system \cite{PottersWheel}.
-Since then, several other interactive data preparation systems have been proposed \cite{Wrangler,Refine}.
-While effective, these systems all make assumptions about possible date formats (\textit{e.g.} the domain system in \cite{PottersWheel}), which we suggest are too restrictive for real world data.
+Data preparation has been a known analytic bottleneck since at least the description of the Potter's Wheel system~\cite{Ramen:2001}.
+Since then, several other interactive data preparation systems have been proposed, including Data Wrangler~\cite{Kandel:2001} and Google refine~\cite{Refine}.
+While effective, these systems all make assumptions about possible date formats (\eg the domain system in~\cite{Ramen:2001}), which we suggest are too restrictive for real world data.
 
-The Minimum Descriptive Length technique was first proposed in \cite{mdl}.  We have built on the version described in \cite{PottersWheel}.
+The Minimum Descriptive Length technique was first proposed in Rissanen~\cite{Rissanen:1978}.  We have built on the version described in~\cite{Ramen:2001}.
 

### 2015-07-06 paper.tex
More formatting
Redo all the quotes (I’d forgotten how annoying TeX’s quote mark
conventions are). Format e.g. everywhere (next time I’ll define a
macro…)
Richard Bibliography
Add some consistency macros

@@
 %
-
+\def\eg{\textit{e.g.} }
+\def\ie{\textit{i.e.} }
+\def\dateparse{\texttt{DATEPARSE} }
+%
 \begin{document}
@@
 % NOTE: You are NOT restricted in how many 'rows' of
-% "name/affiliations" may appear. We just ask that you restrict
+% ``name/affiliations'' may appear. We just ask that you restrict
 % the number of 'columns' to three.
@@
 One of the most common parsing tasks is extracting date and time data from string representations.
-Several databases include date parsing ``mini-languages" to cover the wide range of possible formats.
-Tableau has recently provided a function called \texttt{DATEPARSE}  with a single syntax that is translated into the nearest equivalent in the underlying database.
+Several databases include date parsing ``mini-languages'' to cover the wide range of possible formats.
+Tableau has recently provided a function called \dateparse with a single syntax that is translated into the nearest equivalent in the underlying database.
 Subsequent analysis of customer usage of this function shows that the parsing language syntax was difficult for users to master.
@@
 \terms{Algorithms, Performance}
-
-\keywords{Date parsing, automated cleaning, structure extraction} % NOT required for Proceedings
 
@@
 \section{Acknowledgments}
-We would like to thank to Pat Hanrahan for encouraging this work; Roger Hau for providing the resources to productize it; and Jason King, whose harrowing live demonstration of manual \texttt{DATEPARSE} in front of several thousand people inspired this project.
+We would like to thank to Pat Hanrahan for encouraging this work; Roger Hau for providing the resources to productize it; and Jason King, whose harrowing live demonstration of manual \dateparse in front of several thousand people inspired this project.
 
@@
 \bibliography{dateparse}  % sigproc.bib is the name of the Bibliography in this case
-% You must have a proper ".bib" file
+% You must have a proper ``.bib'' file
 %  and remember to run:

### 2015-07-06 parameters.tex
Add some consistency macros

--- 2015-07-06 parameters.tex
@@
 \item Values containing the substring \texttt{NULL}
-\item Values containing no digits and at most one non-whitespace character (e.g. \texttt{" / / "})
-\item Values on a list of empirically determined common \texttt{NULL} values (e.g. \texttt{0000-00-00}, \texttt{NaN})
+\item Values containing no digits and at most one non-whitespace character (\eg \texttt{" / / "})
+\item Values on a list of empirically determined common \texttt{NULL} values (\eg \texttt{0000-00-00}, \texttt{NaN})
 \end{itemize}
@@
 
-The syntax is documented at the ICU web site \cite{icu}. While fairly complete, it has a few limitations that we ran into when evaluating the algorithms:
+The syntax is documented at the ICU web site~\cite{ICU}. While fairly complete, it has a few limitations that we ran into when evaluating the algorithms:
 \begin{itemize}
-\item No support for 4 letter year abbreviations (e.g. \texttt{Sept.})
-\item No support for ordinal days (e.g. \texttt{July 4th})
-\item No support for quarter postfix notation (e.g. \texttt{2Q})
-\item No support for variant meridian markers (e.g. \texttt{a.m.})
+\item No support for 4 letter year abbreviations (\eg \texttt{Sept.})
+\item No support for ordinal days (\eg \texttt{July 4th})
+\item No support for quarter postfix notation (\eg \texttt{2Q})
+\item No support for variant meridian markers (\eg \texttt{a.m.})
 \end{itemize}
@@
 One other quirk of the ICU syntax may be a contributing factor to the user confusion around writing correct ICU date formats.
-The use of lexicographical case in the meta-symbols of the format language can be confusing (e.g.
-\texttt{y} is used for years, but \texttt{M} is used for months while \texttt{m} is used for minutes.) Another advantage of an automated algorithm is that it hides such problems from most users, significantly improving the usability of the function.
+The use of lexicographical case in the meta-symbols of the format language can be confusing (\eg \texttt{y} is used for years, but \texttt{M} is used for months while \texttt{m} is used for minutes.) Another advantage of an automated algorithm is that it hides such problems from most users, significantly improving the usability of the function.
 

### 2015-07-06 testing.tex
Add some consistency macros

@@
 \begin{table}[ht]
-\begin{tabular}{|p{0.48\linewidth}| p{0.48\linewidth}|}
+\centering
+\begin{tabular}{|p{0.48\linewidth}| p{0.24\linewidth}|}
 \hline
@@
 Some of the dates are clearly just numbers, but our approach is to assume that when the user tells us that the column contains dates we should find the best fit.
-The samples include dates from a wide range of historical sources (e.g.
-Roman pottery dates) so we have elected to defer the date identification task to the user.
+The samples include dates from a wide range of historical sources (\eg Roman pottery dates) so we have elected to defer the date identification task to the user.
 

### 2015-07-06 intro.tex
Added refine reference

--- 2015-07-06 intro.tex
@@
-In recent years, there has been a growth of interest in data visualization technologies for human-assisted data analysis using systems such as \cite{Polaris,Qlik,Spotfire}.
+In recent years, there has been a growth of interest in data visualization technologies for human-assisted data analysis using systems such as \cite{Stolte:2008,Qlik,Spotfire}.
 While computers can provide high-speed and high-volume data processing, humans have domain knowledge and the ability to process data in parallel, notably by using our visual systems.
@@
 Data preparation has been a known analytic bottleneck since at least the description of the Potter's Wheel system \cite{Raman:2001}.
-Since then, several other interactive data preparation systems have been proposed \cite{Kandel:2011,Refine}.
+Since then, several other interactive data preparation systems have been proposed \cite{Kandel:2011,refine}.
 While effective, these systems all make assumptions about possible date formats (\textit{e.g.} the domain system in \cite{PottersWheel}), which we suggest are too restrictive for real world data.

### 2015-07-06 conclusion.tex

--- 2015-07-04 conclusion.tex
@@
 While validating the algorithms on a large corpus, we also found that the number of distinct formats in the wild is surprisingly high, and demonstrates the wisdom of including general-purpose date parsing functions in data visualization tools, data cleaning tools and RDBMSes.
-In particular, it is interesting to note that the most prominent open source RDBMSes (MySQL and Postgres) both have a built-in version of DATEPARSE, possibly reflecting that this is a common need that gets implemented when users are empowered to add to a code base.
+In particular, it is interesting to note that the most prominent open source RDBMSes (\eg MySQL and Postgres) both have a built-in version of \dateparse, possibly reflecting that this is a common need that gets implemented when users are empowered to add to a code base.

### 2015-07-06 intro.tex

--- 2015-07-06 intro.tex
@@
 \subsection{Related Work}
+
 Data preparation has been a known analytic bottleneck since at least the description of the Potter's Wheel system~\cite{Ramen:2001}.
-Since then, several other interactive data preparation systems have been proposed, including Data Wrangler~\cite{Kandel:2001} and Google refine~\cite{Refine}.
+Since then, several other interactive data preparation systems have been proposed, including Data Wrangler~\cite{Kandel:2001} and Google Refine~\cite{refine}.
 While effective, these systems all make assumptions about possible date formats (\eg the domain system in~\cite{Ramen:2001}), which we suggest are too restrictive for real world data.

### 2015-07-06 mdl.tex

--- 2015-07-06 mdl.tex
@@
-The first algorithm is a Minimum Descriptive Length \cite{mdl} approach derived from the domain system presented in \cite{PottersWheel}.
+The first algorithm is a Minimum Descriptive Length~\cite{Rissanen:1978} approach derived from the domain system presented in the Potter's Wheel system of Raman et al.~\cite{Raman:2001}.
 We describe a number of extensions to their structure extraction system to support more complex redundancy, non-English locales, improved performance and date-specific pruning.
@@
 \subsection{Domains}
-\cite{PottersWheel} presents an algorithm for deriving a common structure for a set of strings by breaking each string down into a sequence of domains.
+Potter's Wheel presents an algorithm for deriving a common structure for a set of strings by breaking each string down into a sequence of domains.
 These domains are described by an interface that includes:
@@
 A difficulty in using this kind of structure extraction is that the algorithm for enumerating structures is exponential in the number of domains.
-This is especially true in the date format problem because there are identical domains (e.g.
-months and meridian hours), nearly identical domains (e.g.
-days and hours) and there are often no field delimiters (e.g.
-\texttt{2012Mar06134427}).
+This is especially true in the date format problem because there are identical domains (\eg months and meridian hours), nearly identical domains (\eg days and hours) and there are often no field delimiters (\eg \texttt{2012Mar06134427}).
 To handle this, we have extended the existing pruning API with two other sets of domain identifiers:
@@
 Domain parameterization is an important part of generating compact representations via MDL, but it creates problems for date recognition.
-For example, if a set of dates contains a constant month string (e.g.
-all values are in September) it is important to keep track of the month name domain.
-Consequently, when we parameterize a constant generic <Word> domain, we tag it with the date part domain that it matches (if any).
-We then need to apply an additional pruning step to remove any structures that also found an equivalent domain (e.g.
-two-digit month).
+For example, if a set of dates contains a constant month string (\eg all values are in September) it is important to keep track of the month name domain.
+Consequently, when we parameterize a constant generic \texttt{<Word>} domain, we tag it with the date part domain that it matches (if any).
+We then need to apply an additional pruning step to remove any structures that also found an equivalent domain (\eg two-digit month).
 These rules are equivalent to the context-based redundancy rules above, but have to be applied again after parameterization of generics.
@@
 \begin{itemize}
-\item The set of date parts cannot contain place value gaps (e.g. structures that have year and day without month are removed.)
+\item The set of date parts cannot contain place value gaps (\eg structures that have year and day without month are removed.)
 \item Similarly, the set of time parts cannot contain place value gaps and must also be in place value order (times are never written in orders such as \texttt{mhs}.)
-\item The existence of time parts cannot make dates incomplete (e.g. patterns like year-day-hour are removed.)
+\item The existence of time parts cannot make dates incomplete (\eg patterns like year-day-hour are removed.)
 \item Two digit years require special handling.
@@
 Knowing the locale of a string is not always helpful.
-In our data set, we found numerous cases where the locale was specified (e.g.
-Sweden) but the data was actually in English.
+In our data set, we found numerous cases where the locale was specified (\eg Sweden) but the data was actually in English.
 Accordingly, we test both locales and rank the combined results.
@@
 \item Since we have a set of samples, we can apply the candidate format to the strings to see how well it performs. Formats with fewer parse errors are preferred.
-\item Date parts can be considered a place-value system, so we prefer "more significant" components (e.g. month-day-year over hour-minute-second).
+\item Date parts can be considered a place-value system, so we prefer ``more significant'' components (\eg month-day-year over hour-minute-second).
 \item If two formats from different locales give the same results, prefer the original column locale. The sample set may have missed an example where this could be important.
-\item If the format has an ambiguous date order (e.g.
-all days are less than 12), then prefer the default date order of the locale.
+\item If the format has an ambiguous date order (\eg all days are less than 12), then prefer the default date order of the locale.
 Again, the sample set may have missed a counterexample, so this is the best option.

### 2015-07-06 intro.tex
URL citations

@@
-In recent years, there has been a growth of interest in data visualization technologies for human-assisted data analysis using systems such as \cite{Stolte:2008,Qlik,Ahlberg:1996}.
+In recent years, there has been a growth of interest in data visualization technologies for human-assisted data analysis using systems such as \cite{Stolte:2008,Ahlberg:1996}.
 While computers can provide high-speed and high-volume data processing, humans have domain knowledge and the ability to process data in parallel, notably by using our visual systems.
@@
 \subsection{Related Work}
+Data preparation has been a known analytic bottleneck since at least the description of the Potter's Wheel system~\cite{Raman:2001}.
+Since then, several other interactive data preparation systems have been proposed, including Data Wrangler~\cite{Kandel:2011} and Google refine~\cite{Refine}.
+While effective, these systems all make assumptions about possible date formats (\eg the domain system in~\cite{Raman:2001}), which we suggest are too restrictive for real world data.
 
-Data preparation has been a known analytic bottleneck since at least the description of the Potter's Wheel system~\cite{Ramen:2001}.
-Since then, several other interactive data preparation systems have been proposed, including Data Wrangler~\cite{Kandel:2001} and Google Refine~\cite{refine}.
-While effective, these systems all make assumptions about possible date formats (\eg the domain system in~\cite{Ramen:2001}), which we suggest are too restrictive for real world data.
-
-The Minimum Descriptive Length technique was first proposed in Rissanen~\cite{Rissanen:1978}.  We have built on the version described in~\cite{Ramen:2001}.
+The Minimum Descriptive Length technique was first proposed in Rissanen~\cite{Rissanen:1978}.  We have built on the version described in~\cite{Raman:2001}.
 

### 2015-07-06 paper.tex
Added refine reference
Resolved a conflict
URL citations
Anonymise abstract

@@
 \documentclass{acm_proc_article-sp}
-
+%
+\usepackage{url}
 %
@@
 Several databases include date parsing ``mini-languages'' to cover the wide range of possible formats.
-Tableau has recently provided a function called \dateparse with a single syntax that is translated into the nearest equivalent in the underlying database.
-Subsequent analysis of customer usage of this function shows that the parsing language syntax was difficult for users to master.
+Analysis of the usage of such functions in one visualization system shows that the parsing language syntax was difficult for users to master.
 
-In this paper, we present two algorithms for automatically deriving date format strings from a column of data in our syntax, one based on minimum entropy and one based on natural language modeling.
-Both have similar accuracies of over 90\% on a large corpus of date columns extracted from Tableau Public and are in substantial agreement with each other.
+In this paper, we present two algorithms for automatically deriving date format strings from a column of data, one based on minimum entropy and one based on natural language modeling.
+Both have similar accuracies of over 90\% on a large corpus of date columns extracted from an online data repository and are in substantial agreement with each other.
 The minimal entropy approach can also produce results below the user's perceptual threshold, making it suitable for interactive work.

### 2015-07-07 nlp.tex
Some more points and outline to fill out

@@
 
+While there are several functionally equivalent notations for representing CFG, we use the \textit{Backus-Naur Form} (BNF) for defining the grammar rules for \texttt{DATEPARSE} formats.
+
+%
+%   date-fullyear   = 4DIGIT
+%   date-month      = 2DIGIT  ; 01-12
+%   date-mday       = 2DIGIT  ; 01-28, 01-29, 01-30, 01-31 based on month/year
+%   time-hour       = 2DIGIT  ; 00-23
+%   time-minute     = 2DIGIT  ; 00-59
+%   time-second     = 2DIGIT  ; 00-58, 00-59, 00-60 based on leap second
+%                             ; rules
+%   time-secfrac    = "." 1*DIGIT
+%   time-numoffset  = ("+" / "-") time-hour ":" time-minute
+%   time-offset     = "Z" / time-numoffset
+%
+%   partial-time    = time-hour ":" time-minute ":" time-second
+%                     [time-secfrac]
+%   full-date       = date-fullyear "-" date-month "-" date-mday
+%   full-time       = partial-time time-offset
+%
+%   date-time       = full-date | full-time
+
+\subsection{Production Rule Constraints}
+
+\subsection{Probabilistic Context-Free Grammar}
 Pattern-recognition problems such as parsing date and time formats initiate from observations generated by some structured stochastic process.

### 2015-07-07 paper.tex
Create anonymisation files
Doctor Who theme - sorry!

@@
 Automating the Extraction of Date Scalars}
-%
-% You need the command \numberofauthors to handle the 'placement
-% and alignment' of the authors beneath the title.
-%
-% For aesthetic reasons, we recommend 'three authors at a time'
-% i.e. three 'name/affiliation blocks' be placed beneath the title.
-%
-% NOTE: You are NOT restricted in how many 'rows' of
-% ``name/affiliations'' may appear. We just ask that you restrict
-% the number of 'columns' to three.
-%
-% Because of the available 'opening page real-estate'
-% we ask you to refrain from putting more than six authors
-% (two rows with three columns) beneath the article title.
-% More than six makes the first-page appear very cluttered indeed.
-%
-% Use the \alignauthor commands to handle the names
-% and affiliations for an 'aesthetic maximum' of six authors.
-% Add names, affiliations, addresses for
-% the seventh etc. author(s) as the argument for the
-% \additionalauthors command.
-% These 'additional authors' will be output/set for you
-% without further effort on your part as the last section in
-% the body of your article BEFORE References or any Appendices.
-
-\numberofauthors{3} %  in this sample file, there are a *total*
-% of EIGHT authors. SIX appear on the 'first-page' (for formatting
-% reasons) and the remaining two appear in the \additionalauthors section.
-%
-\author{
-% You can go ahead and credit any number of authors here,
-% e.g. one 'row of three' or two rows (consisting of one row of three
-% and a second row of one, two or three).
-%
-% The command \alignauthor (no curly braces needed) should
-% precede each author name, affiliation/snail-mail address and
-% e-mail address. Additionally, tag each line of
-% affiliation/address with \affaddr, and tag the
-% e-mail address with \email.
-%
-% 1st. author
-\alignauthor
-Richard Wesley \\
-%       \affaddr{Tableau Software}\\
-%       \affaddr{837 North 34th Street}\\
-%       \affaddr{Seattle, WA 98103}\\
-%      \email{hawkfish@tableau.com}
-% 2nd. author
-\alignauthor
-Vidya Setlur\\
-%       \affaddr{Tableau Software}\\
- %      \affaddr{837 North 34th Street}\\
-%       \affaddr{Seattle, WA 98103}\\
- %      \email{vsetlur@tableau.com}
-% 3rd. author
-\alignauthor
-Dan Cory\\
-%       \affaddr{Tableau Software}\\
-%       \affaddr{837 North 34th Street}\\
-%       \affaddr{Seattle, WA 98103}\\
-%       \email{dcory@tableau.com}
-%       \email{\{hawkfish,vsetlur,dcory\}@tableau.com}
-\sharedaffiliation
-       \affaddr{Tableau Software}\\
-       \affaddr{837 North 34th Street}\\
-       \affaddr{Seattle, WA 98103}\\
-       \email{\{hawkfish,vsetlur,dcory\}@tableau.com}
-}
-% Just remember to make sure that the TOTAL number of authors
-% is the number that will appear on the first page PLUS the
-% number that will appear in the \additionalauthors section.
+\input{anonymous}
 
@@
 \section{Acknowledgments}
-We would like to thank to Pat Hanrahan for encouraging this work; Roger Hau for providing the resources to productize it; and Jason King, whose harrowing live demonstration of manual \dateparse in front of several thousand people inspired this project.
+\input{credits}
 

### 2015-07-07 conclusion.tex
End of Richard's anonymisation.

--- 2015-07-04 conclusion.tex
@@
 While validating the algorithms on a large corpus, we also found that the number of distinct formats in the wild is surprisingly high, and demonstrates the wisdom of including general-purpose date parsing functions in data visualization tools, data cleaning tools and RDBMSes.
-In particular, it is interesting to note that the most prominent open source RDBMSes (MySQL and Postgres) both have a built-in version of DATEPARSE, possibly reflecting that this is a common need that gets implemented when users are empowered to add to a code base.
+In particular, it is interesting to note that the most prominent open source RDBMSes (\eg MySQL and Postgres) both have a built-in version of \dateparse, possibly reflecting that this is a common need that gets implemented when users are empowered to extend the function library of an RDBMS.

### 2015-07-07 futurework.tex

@@
-While we are satisfied with the results so far, we found a number of date formats that we were not able to handle with these techniques.
-Many columns contained multiple formats, and we would like to recognize this situation and generate predicate calculations to increase the accuracy of the results and thereby make the results even more seamless for the user.
+While we are satisfied with the results so far, we found a number of column that we were not able to process with these techniques.
+Many columns contained multiple date formats, and we would like to recognize this situation and generate predicate-based calculations to increase the accuracy of the results and thereby make the experience even more seamless for the user.
 Other columns contain date ranges, which we would like to handle by generating multiple calculations, possibly by combining regular expressions with date parsing.
@@
 Moreover, the number of possible formats is low enough that we may be able to enumerate them.
-Timestamps are another form of numeric date for which we intend to automate the preparation.
+Timestamp preparation from numeric representations is another form of preparation which we would like to automate.
 
-In the course of our research, we have also identified a number of date part variants (\eg ordinal dates, four-letter month abbreviations, alternate meridian markers and postfix quarter syntax) that we would like to commend to the ICU project and maybe provide implementations for.
+In the course of our research, we have also identified a number of date part variants (\eg ordinal dates, four-letter month abbreviations, alternate meridian markers and postfix quarter syntax) that we would like to commend to the ICU project and maybe provide implementations of.

### 2015-07-07 intro.tex

@@
-In recent years, there has been a growth of interest in data visualization technologies for human-assisted data analysis using systems such as \cite{Stolte:2008,Ahlberg:1996}.
+In recent years, there has been a growth of interest in data visualization technologies for human-assisted data analysis using systems such as Polaris~\cite{Stolte:2008} and Spotfire~\cite{Ahlberg:1996}.
 While computers can provide high-speed and high-volume data processing, humans have domain knowledge and the ability to process data in parallel, notably by using our visual systems.
@@
 
-\subsection{The Tableau Ecosystem}
-In this paper, we shall be looking at an example of this problem in the context of the Tableau system.
-Tableau is a commercial visual analysis environment derived from the Polaris~\cite{Stolte:2008} system developed at Stanford.
-In addition to an interactive desktop application for creating data visualizations, the Tableau ecosystem also includes servers for publishing and sharing interactive visualizations.
-These servers can be privately operated by individual customers or hosted in the cloud.
-
-\subsubsection{Tableau Public}
-Of particular relevance in the present work is a free version of the server environment called Tableau Public.
-Tableau Public (or ``Public'') allows authors to publish data visualizations that can be shared with the general public.
-The published data sets are limited to 100K rows and must be stored using the Tableau Data Engine (or ``TDE'')~\cite{Wesley:2011,Wesley:2014}, but are otherwise free to use the full analytic capabilities of the Tableau system.
-In particular, the TDE provides native support for all analytic calculations defined by the authoring and publishing components.
-Moreover, visualizations published to Public are uploaded as complete Tableau workbooks, including the data model developed as part of the analysis.
-This makes it a rich source of data on the analytic habits and frustrations of Tableau users.
-
 \subsubsection{Functions}
@@
 And even when a database does not provide an implementation of a particular function, it is usually possible to generate it as an inline combination of existing functionality.
-Tableau's calculation language is designed to be a lingua franca that tames this Babel of dialects by providing a common syntax for as many of these functions as possible.
+The calculation languages of visualization systems can thus be designed to provide a \textit{lingua franca} that tames this Babel of dialects by providing a common syntax for as many of these functions as possible.
 
-\subsection{DATEPARSE}
-Among the recent additions to the Tableau function library is a function called \dateparse, the usability of which is the focus of the work in this paper.
+\subsection{\dateparse}
+The focus of this work is the usability of one such function from our visualization system that we will call \dateparse.
 
 \subsubsection{	Scalar Dates}
+
+The SQL-99 standard defines three temporal scalar types: \texttt{DATE}, \texttt{TIMESTAMP} and \texttt{TIME}.
+They are typically implemented as fixed-point types containing an offset from some epoch (\eg Julian Days.) This makes them compact to store using column store coompression techniques such as those in C-Store~\cite{Stonebraker:2005} and MonetDB/X100~\cite{Zukowski:2006}, and further allows some temporal operations to be implemented very efficiently using simple arithmetic.
+Thus from the RDBMS perspective, representing dates in scalar form provides  benefits for users, both in terms of available operations and query performance.
 
@@
 
-The SQL-99 standard defines three temporal scalar types: \texttt{DATE}, \texttt{TIMESTAMP} and \texttt{TIME}.
-They are typically implemented as fixed-point types containing an offset from some epoch (\eg Julian Days.) This makes them compact to store and allows some temporal operations to be implemented very efficiently using arithmetic.
-Thus from the RDBMS perspective, representing dates in scalar form provides numerous benefits for users, both in terms of available operations and query performance.
+Our visualization system models the first two of these SQL-99 types natively; the third (pure time) is folded into \texttt{TIMESTAMP} by appending it to a fixed date of \texttt{1899-12-30} (a convention derived from Microsoft Excel).
+From the analytic perspective, date types are dimensional (\ie independent variables) and can be used as either \textit{categorical} (simply ordered) or \textit{quantitative} (ordered with a distance metric) fields.
 
-Tableau models the first two of these types as ``Date'' and ``Date \& Time''; the third (pure time) is folded into Date \& Time by appending it to a fixed date of \texttt{1899-12-30}.
-From the analytic perspective, date types are dimensional (\ie independent variables) and can be used as either categorical (simply ordered) or quantitative (ordered with a distance metric) fields.
-
-Categorical dates have a natural hierarchy associated with them generated by calendar binning.
+Categorical dates have a natural hierarchy associated with them generated by calendar \textit{binning}.
 The visualization in Figure \ref{fig:I1} shows an example of a bar chart employing binned categorical dates in a year/quarter hierarchy.
@@
 Quantitative dates are typically used for time series on an axis that maps the underlying distance measure to display pixel distance.
-The quantitative analog to categorical binning is truncation whereby all the dates in a bin are represented by the first date in the bin.
+The quantitative analog to categorical binning is \textit{truncation} whereby all the dates in a bin are represented by the first date in the bin.
 Truncation accurately preserves the distance semantics while enabling roll-up to coarser levels of detail.
@@
 These types of visualizations are difficult to specify with simple categorical string representations of dates.
-Thus the conversion of unparsed strings to date scalars is an important component of data modeling for analytics.
+Thus the conversion of unparsed strings to date scalars is also an important component of data modeling for analytics.
 
 \subsubsection{Parsing}
-One of the oldest data preparation problems observed in Tableau is the parsing of date scalars so that users can perform these kinds of analyses.
-Training materials from the early days of the company included examples of how to convert columns of integers in the form \texttt{yyyyMMdd} to date scalars.
-The documented solution at the time was to convert the integer to a string and perform some locale-dependent string operations before casting the string back to a date.
-Unfortunately, this approach had a number of problems:
+This parsing of date representations is one of the oldest data preparation problems we have observed.
+A common version of this problem is how to convert columns of integers of the form \texttt{yyyyMMdd} to date scalars.
+Naive users often solve this problem by converting the integer to a string and performing some locale-dependent string operations before casting the string back to a date.
+Unfortunately, this approach has a number of problems:
 \begin{itemize}
-\item String operations are notoriously slow compared to scalar operations (typically 10-100x slower)
-\item Default parsing of date formats is locale-dependent, and may not work when the workbook is shared across an international organization (\eg between the US and European offices)
-\item The expression code was hard to understand and maintain because it used a verbose, general-purpose string-handling syntax instead of a domain language.
+\item String operations are notoriously slow compared to scalar operations (typically 10-100x slower in modern RDBMSes)
+\item Default parsing of date formats is locale-dependent, and may not work when the analysis is shared across an international organization (\eg between the US and European offices)
+\item The parsing code was hard to understand and maintain because it used a verbose, general-purpose string-handling syntax instead of a specialized domain language.
 \end{itemize}
 
-This is but a single format.
-Our studies of the workbooks on Public suggest that there are hundreds of distinct temporal date formats in user data sets.
+And this is but a single date format.
+Our studies of online data collections suggest that there are hundreds of distinct temporal date formats in user data sets.
 Some are common, but others can be quite idiosyncratic.
-Table 1 shows a selection of unusual date formats found on Public.
+Table 1 shows a selection of unusual date formats found in our test data.
 The first example shows a time zone in the middle of the date and a year after the time; the second shows a leading unmatched bracket and a colon between the date and time components; the third shows confusion between the seconds' decimal point and the time part delimiter; the fourth shows a two digit year apostrophe on a four digit year and the fifth shows a dash separating the date and time components.
@@
 
-Our first solution to this problem was to add a new function to the Tableau calculation language called \dateparse.
+Our first attempt to help users with this problem was to add a new function to the the calculation language called \dateparse.
 This function would take a string column and convert it to a datetime using a special purpose domain language for describing dates.
-Such functions exist in a number of databases (\eg Oracle, Postgres and MySQL) and adding it to the TDE was straightforward, so it was a natural addition to the function library.
+Such functions exist in a number of databases (\eg MySQL, Oracle and Postgres), so it was a natural addition to the function library.
 
-We chose the date parsing syntax defined by the International Components for Unicode (or ICU) project~\cite{ICU} for the common domain language because it mirrored what was available in the Tableau code base, and translating this syntax into the formats used by various database vendors (\eg MySQL, Oracle, Postgres) was relatively painless.
+We chose the date parsing syntax defined by the International Components for Unicode (or ICU) project~\cite{ICU} for the common domain language because it mirrored what was already available in our code base, and translating this syntax into the formats used by various database vendors was relatively painless.
 There are a few patches of non-overlapping functionality, but they tend to be obscure and we provide warnings when functionality is missing.
@@
 \subsubsection{	Usability}
-After shipping this new function to our user base, we investigated how it was being applied in the field.
-We examined a few months of workbooks that had been uploaded to Public after the release to see if and how it was being used.
+After providing this new function to our user base, we waited a few months and then investigated how it was being applied in a free, cloud-based version of the system.
 We found that although there were a number of new calculations using \dateparse, the error rate for the domain language syntax was about 15\%.
@@
 \subsection{Automated Format Extraction}
-One solution might have been to design a graphical environment that enabled users to construct valid patterns.
-This would have involved a substantial development investment with no guarantee that the result would be correct if the user misunderstood the environment.
+One solution to this problem might have been to design a graphical environment that enabled users to construct valid patterns.
+This would have involved a substantial development effort with no guarantee that the result would be correct if the user misunderstood the environment.
 Instead, we have developed two algorithms for automating the derivation of the format string, each of which uses a different machine learning technique.
@@
 \subsection{Related Work}
-Data preparation has been a known analytic bottleneck since at least the description of the Potter's Wheel system~\cite{Raman:2001}.
-Since then, several other interactive data preparation systems have been proposed, including Data Wrangler~\cite{Kandel:2011} and Google refine~\cite{Refine}.
-While effective, these systems all make assumptions about possible date formats (\eg the domain system in~\cite{Raman:2001}), which we suggest are too restrictive for real world data.
+Data preparation has been considered an analytic bottleneck since at least the description of the Potter's Wheel system~\cite{Raman:2001}.
+Since then, several other interactive data preparation systems have been proposed, including Data Wrangler~\cite{Kandel:2011} and Google Refine~\cite{Refine}.
+While effective, these systems all make assumptions about possible date formats, which we suggest are too restrictive for real world data.
 
-The Minimum Descriptive Length technique was first proposed in Rissanen~\cite{Rissanen:1978}.  We have built on the version described in~\cite{Raman:2001}.
+Various approaches have been described for deriving regular expressions, and a good overview is provided by Li et al.
+in their paper on the ReLIE system for deriving regular expressions given a starting expression provided by a domain expert~\cite{Li:2008}.
+The Minimum Descriptive Length technique first described in Rissanen~\cite{Rissanen:1978} was used in~\cite{Raman:2001} to generate regular expressions.
 
-Related work on parsing languages, outlier detection.
+Related work on parsing languages, outlier detection??
 

### 2015-07-07 mdl.tex

@@
 Accordingly, we test both locales and rank the combined results.
-Tableau currently does not support non-Gregorian calendars.
-Therefore, we have built this system for the Gregorian calendar only.
-ICU does support non-Gregorian calendars and we expect this system could be applied to them as well.
+
+We have built this system for the Gregorian calendar only as our visualization system does not support non-Gregorian calendars and we have little evidence of other calendars (\eg Hebrew, Islamic Civil) being used for analytics.
+ICU supports non-Gregorian calendars and we expect the algorithm could be extended to them as well if needed.
 

### 2015-07-07 parameters.tex

@@
 \subsection{Input Data}
-The training data and test data are taken from Public data sets.
+The training data and test data are taken from data sets published to our free, online visualization service.
+The published data includes both the raw data and how it was used for analysis.
+The service limits data sets to a maximum of 100K rows stored in a database which supports an implementation of \dateparse.
 We selected a large number of string columns whose names suggested that they might contain temporal data.
@@
 \subsection{The ICU Date Format Language}
-The date format syntax we selected is the one provided by the open-source project.
-We chose it because we were already using ICU in the code base, we had access to the source code and it provides localized date part data for a large number of languages.
+The date format syntax we selected is the one provided by the ICU open-source project.
+We chose it because we were already using ICU in the code base, we had access to the source code, and it provides localized date part data for a large number of languages.
 
@@
 When date parts are present, we use \texttt{2000-01-01} as the set of default date parts as it is the start of a leap year.
-When dealing with pure time formats, we use \texttt{1899-12-30}, (which is the date used internally by Tableau to signal pure time values.)
+When dealing with pure time formats, we model the output as a date/time and use \texttt{1899-12-30} for the date parts.
 
-ICU will also parse Time Zones (which we recognize but do not use in Tableau) and Quarters (which we interpret as the first month of the period.) RDBMSes such as Oracle and Postgres that support time zones will be able to take advantage of the generated time zone field.
+ICU will also parse Time Zones (which the algorithms recognize but the visualization system does not model) and Quarters (which we interpret as the first month of the period.) RDBMSes such as Oracle and Postgres that support time zones will be able to take advantage of  identified time zone fields.

### 2015-07-08 nlp.tex
Added NLP book citation
Added grammar definition

--- 2015-07-02 nlp.tex
@@
 
+While there are several functionally equivalent notations for representing CFG, we use the \textit{Backus-Naur Form} (BNF) for defining the grammar rules for \dateparse formats.
+In particular, we use the \textit{Extended Backus-Naur Form} (EBNF) as the notation is more compact and readable for frequently used constructions~\cite{Grune:1990}.
+
+We define a grammar for identifying date-time strings based on other EBNF based date-time formats [cite] as a reference.
+A partial definition of the grammar is found below (see supplementary material for complete definition):
+
+\begin{grammar}
+<TimeGrammar> ::= <Hours> ':' <Minutes> ':' <Seconds> <AMPM>;\\
+
+<DateGrammar> ::= <BigEndianDate> 
+				\alt <MiddleEndianDate> 
+				\alt <LittleEndianDate>;
+
+<DateTimeGrammar>  ::= <DateGrammar> 
+					\alt <TimeGrammar>;
+
+<AMPM>  ::=  WORD('a.m.') 
+			 \alt WORD('p.m.') ;
+
+<Year> ::= <TwoYear> | <FourYear>;
+
+<Month>  ::= <MonthFullForm> | <MonthAbbrForm> | <MonthLetterForm> | <MonthNumber>;
+
+<BigEndianDate> ::= <Year> <Month>  <Day> ;
+
+<MiddleEndianDate> ::= <Month> <Day> <Year>;
+
+<LittleEndianDate> ::= <Day> <Month> <Year>;
+
+\end{grammar}
+
+\subsection{Translation of Grammar Output to ICU Format}
+
+\subsection{Production Rule Constraints and Variants}
+
+\subsection{Probabilistic Context-Free Grammar}
 Pattern-recognition problems such as parsing date and time formats initiate from observations generated by some structured stochastic process.

### 2015-07-08 nlp.tex
started a section on semantic constraints

@@
 
+The EBNF date-time grammar includes a large number of syntactically correct but semantically invalid
+date-time expressions. For example, there is no  `November 31, 2015', `February 29, 2013',
+or `Sunday, May  5, 1965'. November only as 30 days in any year; 2013 was not a leap year; and May 5, 1965 was a Wednesday.
+
+While custom production rules can be added to the existing grammar to exclude such expressions, this approach is not optimal as it leads to a rather large grammar that needs to account for every single semantically valid date-time sequence of terminal symbols.
+Rather, we define a grammar that excludes such expressions.
+By intersecting this constraint
+grammar with the Date-Time grammar, we can define a language that contains only semantically valid dates.
+
 \subsection{Probabilistic Context-Free Grammar}

### 2015-07-08 intro.tex
Abstract and intro tidying

@@
 
+\subsubsection{Clean as You Go}
 During the analytic process, a user may discover that parts of the data are not yet suitable for analysis.
@@
 
-\subsubsection{Functions}
+\subsubsection{Function Libraries}
 While the existing standards for query languages are helpful for defining a core set of row-level functions, there are many useful functions beyond the standards that are only explicitly supported by a subset of RDBMSes, often with different names.
@@
 The SQL-99 standard defines three temporal scalar types: \texttt{DATE}, \texttt{TIMESTAMP} and \texttt{TIME}.
-They are typically implemented as fixed-point types containing an offset from some epoch (\eg Julian Days.) This makes them compact to store using column store coompression techniques such as those in C-Store~\cite{Stonebraker:2005} and MonetDB/X100~\cite{Zukowski:2006}, and further allows some temporal operations to be implemented very efficiently using simple arithmetic.
+They are typically implemented as fixed-point types containing an offset from some epoch (\eg Julian Days.) This makes them compact to store using column store compression techniques such as those in C-Store~\cite{Stonebraker:2005} and MonetDB/X100~\cite{Zukowski:2006}, and further allows some temporal operations to be implemented very efficiently using simple arithmetic.
 Thus from the RDBMS perspective, representing dates in scalar form provides  benefits for users, both in terms of available operations and query performance.
@@
 A common version of this problem is how to convert columns of integers of the form \texttt{yyyyMMdd} to date scalars.
-Naive users often solve this problem by converting the integer to a string and performing some locale-dependent string operations before casting the string back to a date.
+Na\"{i}ve users often solve this problem by converting the integer to a string and performing some locale-dependent string operations before casting the string back to a date.
 Unfortunately, this approach has a number of problems:

### 2015-07-08 paper.tex
End of Richard's anonymisation.
Added grammar definition
Abstract and intro tidying

@@
 \usepackage{url}
+\usepackage{syntax}
+
 %
@@
 
-One of the most common parsing tasks is extracting date and time data from string representations.
-Several databases include date parsing ``mini-languages'' to cover the wide range of possible formats.
-Analysis of the usage of such functions in one visualization system shows that the parsing language syntax was difficult for users to master.
+One of the most common scalar parsing tasks is extracting date and time data from string representations.
+Several RDBMSes include date parsing ``mini-languages'' to cover the wide range of possible formats.
+Analysis of the usage of such functions in one visualization system shows, however, that the date parsing language syntax is difficult for users to master.
 
@@
 Both have similar accuracies of over 90\% on a large corpus of date columns extracted from an online data repository and are in substantial agreement with each other.
-The minimal entropy approach can also produce results below the user's perceptual threshold, making it suitable for interactive work.
+The minimal entropy approach is also fast enough to produce results within the user's perceptual threshold, making it suitable for interactive work.
 \end{abstract}

### 2015-07-08 testing.tex
Anonymized source data

@@
 \subsubsection{Training Data}
-All files published to Tableau Public through February 2014 were scanned.
-We collected the contents of columns with names containing any of the following strings: Date, month, created, dt (abbreviation for date), mes (month in Spanish), datum (date in German), fecha (date in Spanish), data (date in Portuguese), ?
-(day in Chinese and Japanese).
-Roughly 95\% of the data in Tableau Public is in English, but we attempted to include non-English data that was available.
+We scanned all files published to a free, online data analysis website through February 2014.
+We collected the contents of columns with names containing any of the following strings: Date, month, created, dt (abbreviation for date), mes (month in Spanish), datum (date in German), fecha (date in Spanish), data (date in Portuguese), and the day character used in Chinese and Japanese.
+Roughly 95\% of the data on the website is in English, but we attempted to include non-English data that was available.
 Fields of any data type other than date or datetime were analyzed, including strings, integers, and floats.
@@
 The resulting set of files were de-duplicated.
-This handles any duplicate files on Tableau Public including sample data.
+This handles any duplicate files such as sample data.
 Files over 1MB were manually reviewed and those that did not contain dates removed.
@@
 For instance, typing the string \texttt{"12/31/1999"} into Microsoft Excel, is automatically interpreted as the date \texttt{1999-12-31}.
-The Microsoft Jet library that Tableau used to read these text files detects a few date formats as well.
+The Microsoft Jet library used to read these text files detects a few date formats as well.
 Any column already converted by Excel or Jet was not included in this study.
@@
 \textbf{Verification Data}\\
-Verification used another set of data from Tableau Public, collected through April 2015.
+Verification used another set of data from the same website, collected from May 2014 to April 2015.
 This set was collected similarly, but also added columns named time.
-We limited this set to new Tableau Excel and text connectors that were released in May 2014.
-It is likely that some of these files were older files republished to Tableau Public.
+There may be some duplication between the training and verification data.
 There were 31,546 files in the resulting set.
-
-The new Tableau text connector is able to automatically interpret a somewhat different limited set of date formats.
-Any column already converted by Excel or the new Tableau text connector was not included in this study.
 

### 2015-07-09 nlp.tex
Text for grammar constraints

@@
 \begin{grammar}
-<TimeGrammar> ::= <Hours> ':' <Minutes> ':' <Seconds> <AMPM>;\\
+<TimeGrammar> ::= <Hours> ':' <Minutes> ':' <Seconds> ;\\
 
@@
 
-<AMPM>  ::=  WORD('a.m.') 
-			 \alt WORD('p.m.') ;
-
-<Year> ::= <TwoYear> | <FourYear>;
-
-<Month>  ::= <MonthFullForm> | <MonthAbbrForm> | <MonthLetterForm> | <MonthNumber>;
-
 <BigEndianDate> ::= <Year> <Month>  <Day> ;
@@
 
+<Year> ::= <TwoYear> | <FourYear>;
+
+<Month>  ::= <MonthFullForm> | <MonthAbbrForm> | <MonthLetterForm> | <MonthNumber>;
+
+<Day>     ::= NUMBER(1-31);
+
+<Hour> ::= <TwelveHour> | <TwentyFourHour>
+
+<TwelveHour> ::= NUMBER(1-12);
+
+<TwentyFourHour> ::= NUMBER(1-24);
+
 \end{grammar}
 
-\subsection{Translation of Grammar Output to ICU Format}
+\subsection{Translation to ICU Format}
 
@@
 The EBNF date-time grammar includes a large number of syntactically correct but semantically invalid
-date-time expressions. For example, there is no  `November 31, 2015', `February 29, 2013',
-or `Sunday, May  5, 1965'. November only as 30 days in any year; 2013 was not a leap year; and May 5, 1965 was a Wednesday.
+date-time expressions.
+While we have added range restrictions to symbols such as \texttt{Hour} (1--12 for 12-hour format and 1--24 for 24-hour format), \texttt{Days} (1--7), \texttt{Month} (1--12), there are special cases that need to be accounted for.
+For example, there is no  `November 31, 2015', `February 29, 2013', or `Sunday, May  5, 1965'.
+November only as 30 days in any year; 2013 was not a leap year; and May 5, 1965 was a Wednesday.
 
 While custom production rules can be added to the existing grammar to exclude such expressions, this approach is not optimal as it leads to a rather large grammar that needs to account for every single semantically valid date-time sequence of terminal symbols.
-Rather, we define a grammar that excludes such expressions.
-By intersecting this constraint
-grammar with the Date-Time grammar, we can define a language that contains only semantically valid dates.
+Rather, we modify the existing grammar with the following additional constraints to the \texttt{Day} terminal symbol for excluding such expressions:
+\begin{itemize}
+\item Restriction on the distribution of $30$ and $31$:
+Months usually alternate between lengths of 30 and 31 days. We can use x mod 2 to get an alternating pattern of 1 and 0, then just add our constant base number of days:
+
+\begin{equation}
+\texttt{Day} = 30 + x \Mod 2
+\end{equation}
+
+where $x \in [1..12]$ for each of the $12$ months.
+
+Except February, Equation $1$ addresses January and March through July. After July, the pattern should skip one, and the rest of the months should follow the alternating pattern inversely.
+
+To obtain an inverse pattern of alternating 0 and 1, we add 1 to the dividend:
+
+\begin{equation}
+\texttt{Day} = 30 + (x + 1) \Mod 2
+\end{equation}
+
+In Equation $2$ the number of days for August through December is correct ($x \in [8..12]$), but not for the remaining months.
+We hence introduce a bit-masking function so that the equation is equal to 1 over the desired domain and 0 otherwise.
+Multiplying a term by this expression will result in the term being cancelled out outside its domain.
+To mask the latter piece of our function, we need an expression equal to $1$ where $8 \le x\le 12$.
+Floor division by 8 works well, since $x < 16$.
+
+Now if we substitute this expression in the x + 1 dividend of Equarion $2$, we can invert the pattern using our mask:
+
+\begin{equation}
+\texttt{Day} = 30 + (x + \floor{\frac{x}{8}}) \Mod 2
+\end{equation}
+
+Equation $3$ is now added as a restriction to the production rule \begin{grammar} <Days> ::= NUMBER(1-31) \end{grammar}.
+\item LeapYear Restriction on February 29.
+
+\end{itemize}
 

### 2015-07-09 nlp.tex

@@
+(add technical detail on programming environment and parser used for grammar)
+
 \subsection{Context-Free Grammar}
@@
 Rather, we modify the existing grammar with the following additional constraints to the \texttt{Day} terminal symbol for excluding such expressions:
+
 \begin{itemize}
@@
 
-Now if we substitute this expression in the x + 1 dividend of Equarion $2$, we can invert the pattern using our mask:
+Now if we substitute this expression in the x + 1 dividend of Equation $2$, we can invert the pattern using our mask:
 
@@
 
-Equation $3$ is now added as a restriction to the production rule \begin{grammar} <Days> ::= NUMBER(1-31) \end{grammar}.
-\item LeapYear Restriction on February 29.
+\item LeapYear Restriction for February:
+While the above restriction applies to all months barring February, we also apply a constraint to the number of days for February, based on whether the year is leap year or not.
+For this, we define a new symbol in the grammar called \texttt{LeapYear}.
+If an expression containing the month `February' or any such variant (\textit{e.g.} `Feb', `2') with the day `29' and a year, would need to resolve the \texttt{Year} symbol to be a \texttt{LeapYear}, defined as:
+
+\begin{equation}
+\texttt{Year} \Mod 4 == 0
+\end{equation}
 
 \end{itemize}
+
+Equations $3$ and $4$ are now added as a constraint to the production rule \begin{grammar} <Days> ::= NUMBER(1-31) \end{grammar}.
 

### 2015-07-10 nlp.tex
Text tweaks

@@
 
-Equations $3$ and $4$ are now added as a constraint to the production rule \begin{grammar} <Days> ::= NUMBER(1-31) \end{grammar}.
+Equations $3$ and $4$ are now added as constraints to the \texttt{Days} symbol in the grammar.
 
@@
 In other words, even if the initial higher-level production rule of the grammar is known (i.e.
-date, time or date-time), there could be several directions that the parser resolve to.
+date, time or date-time), there could be several directions that the parser resolves to.
 For example, in a date string 5/6/2015, the pattern could either be M/d/yyyy or d/M/yyyy.
@@
 Once we have created a PCFG model of a process, we can apply existing PCFG parsing algorithms to identify a variety of date-time formats.
-However, the parser’s success is often limited in the types of the dominant patterns that it can identify.
+However, the parserÕs success is often limited in the types of the dominant patterns that it can identify.
 In addition, the standard parsing techniques generally require specification of a complete observation sequence.
@@
 
-%•	Motivate the use of context-free grammar, particularly Backus-Normal-Form
-%•	Explain why we use a probabilistic version of the grammar
-%•	Explain how the production rules are created based on the ICU format
-%•	Extensions to the parser to account for columnar context for maximizing the probabilistic occurrence of the dominant pattern
-%•	Other semantic extensions – looking at the file name, adding rules for leap year, using external dictionaries for non-English terms
+%¥	Motivate the use of context-free grammar, particularly Backus-Normal-Form
+%¥	Explain why we use a probabilistic version of the grammar
+%¥	Explain how the production rules are created based on the ICU format
+%¥	Extensions to the parser to account for columnar context for maximizing the probabilistic occurrence of the dominant pattern
+%¥	Other semantic extensions Ð looking at the file name, adding rules for leap year, using external dictionaries for non-English terms

### 2015-07-10 testing.tex
Tweak section formatting.

@@
 
-\subsubsection{Training Data}
-We scanned all files published to a free, online data analysis website through February 2014.
+We scanned all files published to a free, online data analysis website.
 We collected the contents of columns with names containing any of the following strings: Date, month, created, dt (abbreviation for date), mes (month in Spanish), datum (date in German), fecha (date in Spanish), data (date in Portuguese), and the day character used in Chinese and Japanese.
@@
 
-The resulting set of files were de-duplicated.
-This handles any duplicate files such as sample data.
-Files over 1MB were manually reviewed and those that did not contain dates removed.
-There were 30,968 files in the resulting set.
+The resulting set of files were de-duplicated. This handles any duplicate files from repeated publishing events. Files over 1MB were manually reviewed and those that did not contain dates removed.
+
 Most database and spreadsheet systems already detect a limited set of date formats.
@@
 
-\textbf{Verification Data}\\
+% Should we quantify the number of formats excluded here?
+
+\subsubsection{Training Data}
+
+Training used a subset of the online data uploaded through February 2014. There were 30,968 files in the resulting set.
+
+\subsubsection{Verification Data}
 Verification used another set of data from the same website, collected from May 2014 to April 2015.
@@
 There may be some duplication between the training and verification data.
-There were 31,546 files in the resulting set.
+There were 31,546 files in the resulting verification set.
 

### 2015-07-10 nlp.tex
More writing on PCFG and bibtex updated.

@@
 date, time or date-time), there could be several directions that the parser resolves to.
-For example, in a date string 5/6/2015, the pattern could either be M/d/yyyy or d/M/yyyy.
+For example, in a date string `5/6/2015', the pattern could either be \texttt{M/d/yyyy} or \texttt{d/M/yyyy}.
 
-Probabilistic context-free grammars (PCFGs) have provided a useful method for modeling such uncertainty [cite].
+In the context of CFGs, probabilities have been used to define a probability distribution over a set of parse trees defined by the CFG, and are a useful method for modeling such ambiguity~\cite{Collins:2003,Manning:1999}.
+The resulting formalism called Probabilistic Context-Free Grammar (PCFG), extends the CFG by assigning probabilities to the production rules of the grammar.
+During the process of parsing the date-time pattern, the probabilities are used as a filtering mechanism to rank the pattern(s) that a given string resolves to in the grammar.
+
+(add the rules for how probabilistic weights are assigned to the CFG)
+
 Once we have created a PCFG model of a process, we can apply existing PCFG parsing algorithms to identify a variety of date-time formats.

### 2015-07-10 nlp.tex
more updates

@@
-(add technical detail on programming environment and parser used for grammar)
+(motivate why an NLP approach was considered. Also add technical detail on programming environment and parser used for grammar)
 
@@
 
-<Day>     ::= NUMBER(1-31);
+<Day>     ::= d d (between 01 and 28-31, depending on month/year);
 
@@
 
-<TwelveHour> ::= NUMBER(1-12);
+<TwelveHour> ::= d d (between 00-12);
 
-<TwentyFourHour> ::= NUMBER(1-24);
+<TwentyFourHour> ::= d d (between 00 and 23);
+
+<d> ::= `0' | `1' | `2' | `3' | `4' | `5' | `6' | `7' | `8' | `9'
 
@@
 date, time or date-time), there could be several directions that the parser resolves to.
-For example, in a date string `5/6/2015', the pattern could either be \texttt{M/d/yyyy} or \texttt{d/M/yyyy}.
+For example, in a date string \texttt{5/6/2015}, the pattern could either be \texttt{M/d/yyyy} or \texttt{d/M/yyyy}.
 
@@
 During the process of parsing the date-time pattern, the probabilities are used as a filtering mechanism to rank the pattern(s) that a given string resolves to in the grammar.
+The parser then chooses the parse tree with the maximum probability.
 
@@
 
+\subsection{Columnar Context}
+
 Once we have created a PCFG model of a process, we can apply existing PCFG parsing algorithms to identify a variety of date-time formats.
-However, the parserÕs success is often limited in the types of the dominant patterns that it can identify.
+However, the parser's success is often limited in the types of the dominant patterns that it can identify.
 In addition, the standard parsing techniques generally require specification of a complete observation sequence.
-In many contexts, we may have only a partial sequence available (e.g.
-an incomplete entry).
+In many contexts, we may have only a partial sequence available (\textit{e.g.} an incomplete entry).
 Finally, we may be interested in computing the probabilities of date-time patterns that the grammar may not explicitly define.
@@
 
-%¥	Motivate the use of context-free grammar, particularly Backus-Normal-Form
+\subsection{Semantic Extensions}
+
 %¥	Explain why we use a probabilistic version of the grammar

### 2015-07-10 nlp.tex
More structural changes to section

@@
 
-\subsection{Columnar Context}
+\subsection{Extensions to PCFG}
 
+\subsubsection{Columnar Context}
 Once we have created a PCFG model of a process, we can apply existing PCFG parsing algorithms to identify a variety of date-time formats.
@@
 
-\subsection{Semantic Extensions}
+\subsubsection{Locale and File Name Context}
 

### 2015-07-13 nlp.tex
More writing

@@
 \subsection{Probabilistic Context-Free Grammar}
+
+%Statistical parsing uses a probabilistic model of syntax in order to assign probabilities to each parse tree.
+%Provides principled approach to resolving syntactic ambiguity.
+%Allows supervised learning of parsers from tree-banks of parse trees provided by human linguists.
+%Also allows unsupervised learning of parsers from corpora.
+
 Pattern-recognition problems such as parsing date and time formats initiate from observations generated by some structured stochastic process.
@@
 (add the rules for how probabilistic weights are assigned to the CFG)
+(show cfg with prob values)
+(examples -> one ambiguous, one not)
+(parse tree with final probability weights)
+
+\subsection{Training the PCFG}
+
+(used a subset of the public data)
+
+Having defined initial probabilistic weights to the CFG, we train the PCFG with a known training set of date time formats to enhance the coverage of the weights to the various production rules of the grammar.
+
+\subsection{Parsing the PCFG}
+Given a date-time string $s$, as input, we determine the highest scoring parse tree $t$ as  $max_{t \in T(s)} p(t)$.
 

### 2015-07-13 nlp.tex
some math notation added.

@@
 During the process of parsing the date-time pattern, the probabilities are used as a filtering mechanism to rank the pattern(s) that a given string resolves to in the grammar.
-The parser then chooses the parse tree with the maximum probability.
+
+Given a CFG grammar $G$, we define  $\tau_G(s)$ as the set of all possible parse trees for input date-time string $s$.
+For any $t \in \tau(G)$, the probability $p(t) \ge 1$.
+In addition, $\sum_{t \in \tau_{G}} = 1$.
+The parser then chooses the parse tree with the maximum probability, \textit{i.e.} given a date-time string $s$, as input, we determine the highest scoring parse tree $t$ as  $max_{t \in \tau(s)} p(t)$.
 
@@
 \subsection{Parsing the PCFG}
-Given a date-time string $s$, as input, we determine the highest scoring parse tree $t$ as  $max_{t \in T(s)} p(t)$.
 

### 2015-07-13 testing.tex
Filled out section 5.2.3

@@
 Many of the files in the collection are not actually dates, such as fields named ``updated by'' (which contain ``date'').
-A random sample of 850 columns named exactly ``date", ``time'' or ``month'' (case-insensitive) were manually judged.
+A random sample of 850 columns named exactly ``date'', ``time'' or ``month'' (case-insensitive) were manually judged.
 
-\subsubsection{	Minimum Descriptive Length}
+\subsubsection{Minimum Descriptive Length}
 Testing of the MDL algorithm was performed on a 24-core Dell T7610 running Windows 7 with the data stored on a 250GB SSD.
@@
 
-\subsubsection{	Cross-Checking}
+\subsubsection{Cross-Checking}
+We compared the results of the minimum description length and natural language processing algorithms.
+The two algorithms match for 97.9% of our validation set.
+The main differences in the results are due to small differences in the implementations.
+The minimum description length implementation recognizes formats with leading plus and minus signs, but these are almost certainly numbers, not dates.
+The minimum description length implementation also recognizes the Excel date formats, but this was not supported by the natural language implementation.
+There was one file that contained formats such as ``Fall 2000'' and ``Spring 2000'', and the two algorithms picked different seasons in their format.
+One other file contained a set of integers that are not actually dates, and one algorithm picked MMyyyy format, while the other picked HHmmss format.

### 2015-07-14 nlp.tex
added rules that define initial probability weight assignment

@@
 
-<Day>     ::= d d (between 01 and 28-31, depending on month/year);
+<Day>     ::= dd (between 01 and 28-31, depending on month/year);
 
@@
 
-<TwelveHour> ::= d d (between 00-12);
+<TwelveHour> ::= dd (between 00-12);
 
-<TwentyFourHour> ::= d d (between 00 and 23);
+<TwentyFourHour> ::= dd (between 00 and 23);
 
@@
 
+(add notation for how rules are defined)
+
 Given a CFG grammar $G$, we define  $\tau_G(s)$ as the set of all possible parse trees for input date-time string $s$.
@@
 
-(add the rules for how probabilistic weights are assigned to the CFG)
+In order to bootstrap the probability weights in the \texttt{PCFG}, we manually set initial probability weights to various rules in the grammar as follows:
+\begin{itemize}
+\item Similar to the MDL ranking properties described in Section 3.7, rules involving the non-terminal \texttt{DateGrammar} on the left-side of the rule, are assigned probability weights higher than those rules with the non-terminal being \texttt{TimeGrammar}.
+In practice, assigning $p(\texttt{DateGrammar} \rightarrow \beta_{1}) = 0.9$ and $p(\texttt{TimeGrammar} \rightarrow \beta_{2}) = 0.7$ leads to optimal ranking of the parse trees.
+
+\item For the non-terminal symbol \texttt{Day}, the constraint specified is a range between 01 and 28-31, depending on month/year.
+However, for \texttt{dd} $> 12$, $p(\texttt{Day} \rightarrow d d) = 1.0$, and $0.5$ otherwise.
+This assignment helps disambiguate days from months.
+
+\end{itemize}
+
 (show cfg with prob values)

### 2015-07-14 nlp.tex
More PCFG writing.

@@
 ICU date-time formats are well defined both structurally and semantically, and can be defined by a context-free grammar (CFG).
-A CFG is commonly defined as a set of productions or rules of the form A ?
-?
-where A is a variable, ?
-is a sequence of variables and terminal symbols (the tokens that make up the alphabet of the language) plus null (?
-), and the production symbol (?) indicates that the variable A can be expanded into ?
-.
-A CFG can be formally specified with four components: V, T, P, and S, where V is the set of variables, T the set of terminal symbols, P the set of productions, and S the set of available start symbols (a non-empty subset of V) [cite].
+A CFG consists of a set of non-terminal symbols $X$, a set of terminal symbols $\beta$, a start non-terminal symbol $S \in X$ from which the grammar generates the strings, and a set of production rules $\tau$, with each rule of the form $X \rightarrow \beta$ [cite].
 
@@
 
-%Statistical parsing uses a probabilistic model of syntax in order to assign probabilities to each parse tree.
-%Provides principled approach to resolving syntactic ambiguity.
-%Allows supervised learning of parsers from tree-banks of parse trees provided by human linguists.
-%Also allows unsupervised learning of parsers from corpora.
-
 Pattern-recognition problems such as parsing date and time formats initiate from observations generated by some structured stochastic process.
@@
 
-(add notation for how rules are defined)
+The production rules are of the form $ X \rightarrow \beta$ where $X$ is the left-hand side of the rule, while $\beta$ is the right-hand side.
 
 Given a CFG grammar $G$, we define  $\tau_G(s)$ as the set of all possible parse trees for input date-time string $s$.
-For any $t \in \tau(G)$, the probability $p(t) \ge 1$.
-In addition, $\sum_{t \in \tau_{G}} = 1$.
-The parser then chooses the parse tree with the maximum probability, \textit{i.e.} given a date-time string $s$, as input, we determine the highest scoring parse tree $t$ as  $max_{t \in \tau(s)} p(t)$.
+For any $X \rightarrow \beta \in \tau(G)$, the probability $p(X \rightarrow \beta) \ge 1$.
+In addition, $\sum_{(X \rightarrow \beta) \in \tau_{G}} = 1$.
+The parser then chooses the parse tree with the maximum probability, \textit{i.e.} given a date-time string $s$, as input, we determine the highest scoring parse tree $X \rightarrow \beta$ as  $max_{(X \rightarrow \beta) \in \tau(s)} p(X \rightarrow \beta)$.
 
@@
 
-\subsection{Training the PCFG}
+\subsection{Supervised Learning}
 
-(used a subset of the public data)
+Having defined initial probabilistic weights to the \texttt{PCFG}, we employ supervised learning with a known training set of date time formats to estimate the rule probabilities.
+The training corpus is a set of files obtained from an online data analysis corpus, described in Section 5.
 
-Having defined initial probabilistic weights to the CFG, we train the PCFG with a known training set of date time formats to enhance the coverage of the weights to the various production rules of the grammar.
+The occurrence frequencies of the rules in the correct
+(disambiguated) parse trees can be determined from the corpus using the maximum-likelihood parameter estimates.
+\begin{equation}
+p(X \rightarrow \beta) = \frac{Count(X \rightarrow \beta)}{Count(X)}
+\end{equation}
 
-\subsection{Parsing the PCFG}
+where $Count(X \rightarrow \beta)$ is the number of times that the rule $X \rightarrow \beta$ is seen in
+the parse trees $\tau$, and $Count(X)$ is the number of times the non-terminal $X$ is seen in $\tau$.
+These frequencies can then be normalized into rule probabilities.
+This method produces accurate probability estimates when trained on a sufficiently large corpus of disambiguated parse trees.
 
-\subsection{Extensions to PCFG}
+\subsection{Extensions}
 

### 2015-07-14 nlp.tex
Added pcfg reference

@@
 ICU date-time formats are well defined both structurally and semantically, and can be defined by a context-free grammar (CFG).
-A CFG consists of a set of non-terminal symbols $X$, a set of terminal symbols $\beta$, a start non-terminal symbol $S \in X$ from which the grammar generates the strings, and a set of production rules $\tau$, with each rule of the form $X \rightarrow \beta$ [cite].
+A CFG consists of a set of non-terminal symbols $X$, a set of terminal symbols $\beta$, a start non-terminal symbol $S \in X$ from which the grammar generates the strings, and a set of production rules $\tau$, with each rule of the form $X \rightarrow \beta$~\cite{Hopcroft:1990}.
 

### 2015-07-14 intro.tex
Starting to fill out related work section.

@@
 Both algorithms result in over 90\% parsing accuracy -- and 100\% syntactic correctness because they are machine generated.
-
-% vidya: this probably should be its own section
-% hawkfish: Not sure it is big enough, but I can see the logic.
-
-\subsection{Related Work}
-Data preparation has been considered an analytic bottleneck since at least the description of the Potter's Wheel system~\cite{Raman:2001}.
-Since then, several other interactive data preparation systems have been proposed, including Data Wrangler~\cite{Kandel:2011} and Google Refine~\cite{Refine}.
-While effective, these systems all make assumptions about possible date formats, which we suggest are too restrictive for real world data.
-
-Various approaches have been described for deriving regular expressions, and a good overview is provided by Li et al.
-in their paper on the ReLIE system for deriving regular expressions given a starting expression provided by a domain expert~\cite{Li:2008}.
-The Minimum Descriptive Length technique first described in Rissanen~\cite{Rissanen:1978} was used in~\cite{Raman:2001} to generate regular expressions.
-
-Related work on parsing languages, outlier detection??
-
-\subsection{Overview}
-The rest of this paper is organized as follows: The next section introduces the parameters of the problem space.
-The following two sections describe the two different algorithms, one using Minimum Descriptive Length and the other using Natural Language Processing.
-In section 5, we evaluate the algorithms on a corpus of 30K columns, both by sampling the outputs manually and then by using the algorithms to validate each other.
-We then discuss future work in section 6 and conclude in section 7.

### 2015-07-14 relatedwork.tex
Starting to fill out related work section.

% vidya: this probably should be its own section
% hawkfish: Not sure it is big enough, but I can see the logic.

Data preparation has been considered an analytic bottleneck since at least the description of the Potter's Wheel system~\cite{Raman:2001}. Since then, several other interactive data preparation systems have been proposed, including Data Wrangler~\cite{Kandel:2011} and Google Refine~\cite{Refine}. While effective, these systems all make assumptions about possible date formats, which we suggest are too restrictive for real world data.

Various approaches have been described for deriving regular expressions, and a good overview is provided by Li et al. in their paper on the ReLIE system for deriving regular expressions given a starting expression provided by a domain expert~\cite{Li:2008}. The Minimum Descriptive Length technique first described in Rissanen~\cite{Rissanen:1978} was used in~\cite{Raman:2001} to generate regular expressions.

There are several bodies of research on developing semantic parsers and grammars for interpreting time and date expressions. Lee \textit{et al} use a combinatory categorical grammar by combining the TIMEX3 standard~\cite{timex3} with contextual cues such as document creation time to determine the reference for parsing time expressions such as `2nd Friday of July'~\cite{LeeADZ14}.

\vidya{Should this be preceded by a bulleted list of contributions, under the section Contributions?}
%\subsection{Overview}
The rest of this paper is organized as follows: The next section introduces the parameters of the problem space. The following two sections describe the two different algorithms, one using Minimum Descriptive Length and the other using Natural Language Processing. In section 5, we evaluate the algorithms on a corpus of 30K columns, both by sampling the outputs manually and then by using the algorithms to validate each other. We then discuss future work in section 6 and conclude in section 7.

### 2015-07-14 relatedwork.tex
More text added to related work section.

@@
-% vidya: this probably should be its own section
-% hawkfish: Not sure it is big enough, but I can see the logic.
-
 Data preparation has been considered an analytic bottleneck since at least the description of the Potter's Wheel system~\cite{Raman:2001}.
@@
 Lee \textit{et al} use a combinatory categorical grammar by combining the TIMEX3 standard~\cite{timex3} with contextual cues such as document creation time to determine the reference for parsing time expressions such as `2nd Friday of July'~\cite{LeeADZ14}.
+Related work by Gabor \textit{et al.} employs a probabilistic approach for learning to interpret temporal phrases~\cite{Angeli:2012}.
+Han and Lavie developed a formalism called the Time Calculus for Natural Language (TCNL), designed to capture the meaning of temporal expressions in natural language.
+In this formalism, each temporal expression is converted to a formula in TCNL, which then can be processed to calculate the value of a temporal expression.
+Their temporal focus tracking mechanism allows correct interpretation of cases like `I am free next week.
+How
+about Friday?', where the TCNL formula for Friday, reflects the occurrence of `next' in the preceding sentence~\cite{Han:2004}.
+In all this previous research using natural language techniques, the interpretation of temporal expressions are done individually, using the presence of lexical tokens such as `next' and `past' along with the tense of the verb tokens.
+
+Ou work differs from these previous methods for parsing date-time expressions as we are using an entire column of data as context without the necessary presence of rich lexical identifiers to determine temporal context.
+This sort of data tends to be prevalent in datasets that are used in visualizations.
 

### 2015-07-15 relatedwork.tex
Contributions

@@
+\subsection{Previous Work}
+
 Data preparation has been considered an analytic bottleneck since at least the description of the Potter's Wheel system~\cite{Raman:2001}.
@@
 
-Ou work differs from these previous methods for parsing date-time expressions as we are using an entire column of data as context without the necessary presence of rich lexical identifiers to determine temporal context.
+\subsection{Contributions}
+
+Our work differs from these previous methods for parsing date-time expressions as we are using an entire column of data as context without the necessary presence of rich lexical identifiers to determine temporal context.
 This sort of data tends to be prevalent in datasets that are used in visualizations.
 
-\vidya{Should this be preceded by a bulleted list of contributions, under the section Contributions?}
-%\subsection{Overview}
+In particular, our contributions are:
+\begin{itemize}
+\item By analysing an on line corpus, we provide evidence that practical date parsing requires the ability to recognise hundreds of formats;
+\item We show how to extend prior work on Minimum Descriptive Length structure extraction to generate a freely available date format domain language with over 90\% accuracy;
+\item We describe a second Natural Language Processing technique for generating the same date format domain language with similar accuracy;
+\item We describe some limitations of this domain language that would improve its utility.
+\end{itemize}
+
+\subsection{Overview}
 The rest of this paper is organized as follows: The next section introduces the parameters of the problem space.
 The following two sections describe the two different algorithms, one using Minimum Descriptive Length and the other using Natural Language Processing.
-In section 5, we evaluate the algorithms on a corpus of 30K columns, both by sampling the outputs manually and then by using the algorithms to validate each other.
-We then discuss future work in section 6 and conclude in section 7.
+In section 6, we evaluate the algorithms on a corpus of 30K columns, both by sampling the outputs manually and then by using the algorithms to validate each other.
+We then discuss future work in section 7 and conclude in section 8.

### 2015-07-15 testing.tex
Fixed percentage notation.
Fixed some bibtex formatting and initial text for section 6.2.2

@@
 To test the MDL algorithm, we ran it over the set of samples from each validation file to generate a ranked list of formats for the file.
-Each format was then applied to the entire file’s data set, recording both the number of errors and the elapsed times.
+Each format was then applied to the entire file's data set, recording both the number of errors and the elapsed times.
 In cases where we generated multiple formats and the main format produced errors, we applied the second format to the unparsed strings.
@@
 \subsubsection{Natural Language Processing}
+We implemented the grammar, training and parsing in Python 3.4.1 using the Natural Language Toolkit libraries~\cite{nltk} on a quad core Dell T7600 running Windows 7.
 
@@
 We compared the results of the minimum description length and natural language processing algorithms.
-The two algorithms match for 97.9% of our validation set.
+The two algorithms match for 97.9\% of our validation set.
 The main differences in the results are due to small differences in the implementations.

### 2015-07-15 testing.tex
More content

@@
 We implemented the grammar, training and parsing in Python 3.4.1 using the Natural Language Toolkit libraries~\cite{nltk} on a quad core Dell T7600 running Windows 7.
+For parsing the PCFG, we use a CKY parser, a dynamic programming algorithm that identifies the highest probable syntactic parse trees, given the grammar and an input date-time string.
+The complexity of the parsing algorithm is $O(n^{3}|G|)$, where $n$ is the length of the input and $|G|$ is the size of the grammar~\cite{Younger67}.
+The PCFG grammar uses a set of X phrase-level non-terminals and Y terminals to classify the various constituents in the corpus.
+Since the CKY algorithm finds the maximum likelihood solution, there are no search errors, and a modification in the grammar is required to improve accuracy.
+
+Similar to the MDL algorithm described in the previous subsection, The CKY parser is run over the same set of samples from each validation file to generate a ranked list of parse trees representing the date-time formats.
+Starting from the most probable parse tree, each parse tree is then applied to the entire file's columnar data to determine the dominant pattern.
+
+%% talk about data formats it recognizes and doesn't recognize.
 

### 2015-07-16 nlp.tex
updated bnf to be comprehensive

@@
 
-We define a grammar for identifying date-time strings based on other EBNF based date-time formats [cite] as a reference.
-A partial definition of the grammar is found below (see supplementary material for complete definition):
+We define a grammar for identifying date-time strings based on other EBNF based date-time formats [cite] as a reference. A partial definition of the grammar is found below:
 
 \begin{grammar}
-<TimeGrammar> ::= <Hours> ':' <Minutes> ':' <Seconds> ;\\
+<TimeGrammar> ::= <Hours> ':' <Minutes> ':' <Seconds> <TimeZone> <AMPM> (for 12-hour formats)
 
@@
 
-<Month>  ::= <MonthFullForm> | <MonthAbbrForm> | <MonthLetterForm> | <MonthNumber>;
+<QuarterYear> := <Quarter> <Year>;
 
-<Day>     ::= dd (between 01 and 28-31, depending on month/year);
+<Day>     ::= dd (where $dd \in [01-31]$, depending on month/year);
 
-<Hour> ::= <TwelveHour> | <TwentyFourHour>
+<Month> := <MonthWord> | <MonthNumber>;
 
-<TwelveHour> ::= dd (between 00-12);
+<MonthWord> := `January' | `February' | `March' | `April' | `May' | `June' | `July' | `August' | `September' | `October' | `November' | `December';
 
-<TwentyFourHour> ::= dd (between 00 and 23);
+<MonthNumber> := dd (where $dd \in [01-12]$);
 
-<d> ::= `0' | `1' | `2' | `3' | `4' | `5' | `6' | `7' | `8' | `9'
+<DayOfWeek> ::= `Monday' | `Tuesday' | `Wednesday' | `Thursday' | `Friday' | `Saturday' | `Sunday';
+
+<Hour> ::= <TwelveHour> | <TwentyFourHour>;
+
+<Quarter> ::= `Quarter' dd (where $dd \in [01 - 04]$);
+
+<TwoYear> ::= dd;
+
+<FourYear> := dddd;
+
+<TwelveHour> ::= dd (where $dd \in [00-12]$);
+
+<TwentyFourHour> ::= dd (where $dd \in [00 -23]$);
+
+<AMPM> := `a.m.' | `p.m.';
+
+<d> ::= `0' | `1' | `2' | `3' | `4' | `5' | `6' | `7' | `8' | `9';
 
 \end{grammar}
+
+%% how do we handle abbreviations, inflections and spelling variants, 01 -> 1
 
@@
 
-\subsection{Extensions}
+\subsection{Grammar Extensions}
 
@@
 
+\subsubsection{PCFG Parser}
+We use a Cocke-Kasami-Younger (CKY) parser implemented in Python 3.4.1 for
+syntactically parsing a column of date-time strings.
+
 %¥	Explain why we use a probabilistic version of the grammar

### 2015-07-16 testing.tex
updated bnf to be comprehensive

@@
 The PCFG grammar uses a set of X phrase-level non-terminals and Y terminals to classify the various constituents in the corpus.
-Since the CKY algorithm finds the maximum likelihood solution, there are no search errors, and a modification in the grammar is required to improve accuracy.
+Since the CKY algorithm finds the maximum likelihood solution, there are no search errors (rather probability $p = 0$, and a modification in the grammar is required to improve accuracy.
 
 Similar to the MDL algorithm described in the previous subsection, The CKY parser is run over the same set of samples from each validation file to generate a ranked list of parse trees representing the date-time formats.
-Starting from the most probable parse tree, each parse tree is then applied to the entire file's columnar data to determine the dominant pattern.
+As a second pass of the parser, starting from the most probable parse tree, each parse tree is then applied to the entire file's columnar data to determine the dominant pattern(s).
+For the second pass, we can parallelize the CKY parsing because the dependencies between reapplying the ranked list of parse trees are very sparse.
+
+The initial average parsing speed to compute the ranked set of probable parse trees is $0.93s$, while the average time taken to compute the overall dominant pattern(s) for the entire column of data is $1.4s$.
+While the natural language parsing implementation is in Python as the NLTK package is easily configurable, we could expect greater speed-up in parsing performance by employing C/C++ CKY parsing libraries~\cite{}.
 

### 2015-07-16 nlp.tex
Completed writing.

@@
-(motivate why an NLP approach was considered. Also add technical detail on programming environment and parser used for grammar)
+The main characteristic of formal grammar is the presence of production rules that are used to rewrite an expression, and a starting symbol which is the origin of all forms of a given expression.
+In the production rules, we find non-terminals and terminals, with parsed expressions containing only terminals.
+A formal grammar is considered ``context free'' when its production rules can be applied regardless of the context of a nonterminal.
+This has several advantages.
+First, it allows for modular syntax definitions, which simplifies grammar development and enables reuse.
+Second, it grants total freedom
+in structuring a grammar to best fit its intended use.
+The power of such a grammar is its expressiveness, and is not constrained by pattern rigidities that are prevalent with regular expressions and text processing~\cite{Grune:1990}.
 
 \subsection{Context-Free Grammar}
-ICU date-time formats are well defined both structurally and semantically, and can be defined by a context-free grammar (CFG).
+ICU date formats are well defined both structurally and semantically, and can be defined by a context-free grammar (CFG).
 A CFG consists of a set of non-terminal symbols $X$, a set of terminal symbols $\beta$, a start non-terminal symbol $S \in X$ from which the grammar generates the strings, and a set of production rules $\tau$, with each rule of the form $X \rightarrow \beta$~\cite{Hopcroft:1990}.
@@
 In particular, we use the \textit{Extended Backus-Naur Form} (EBNF) as the notation is more compact and readable for frequently used constructions~\cite{Grune:1990}.
+We use a Cocke-Younger-Kasami (CYK) parser for syntactically parsing a column of date-time strings based on the grammar~\cite{Cocke:1969,Younger67,Kasami:1965}.
+The parser is a dynamic programming algorithm that identifies the highest probable syntactic parse trees, given the grammar and an input date-time string.
 
-We define a grammar for identifying date-time strings based on other EBNF based date-time formats [cite] as a reference. A partial definition of the grammar is found below:
+We define a grammar for identifying date-time strings based on other EBNF based date-time formats as a reference  [cite] . The definition of the grammar is found below:
 
@@
 
-%% how do we handle abbreviations, inflections and spelling variants, 01 -> 1
+The parsed output is then converted to the ICU Date Format Language described in Section 3.2 in order to effectively cross-validate results from the MDL algorithm.
 
-\subsection{Translation to ICU Format}
+\subsection{Grammar Variants and Constraints}
 
-\subsection{Production Rule Constraints and Variants}
+A classical CFG performs poorly on inflected expressions resulting from misapplication of morphological inflection and syntactic rules.
+Large numbers of non-terminals are necessary for representation of all variations of such features, and the parser output would consist of many parse trees.
+We extend the date-time grammar by adding morpho-syntactic variants.
+These variants allows the parser to correct for syntactic errors (\textit{e.g.} usage of whitespace characters and punctuation marks), capitalization errors and abbreviations (\textit{e.g.} `Mon' for `Monday').
+We use several various external corpora for such corrections~\cite{nltkcorpora}.
 
-The EBNF date-time grammar includes a large number of syntactically correct but semantically invalid
+The date-time grammar also includes a large number of syntactically correct but semantically invalid
 date-time expressions.
@@
 
-(show cfg with prob values)
-(examples -> one ambiguous, one not)
-(parse tree with final probability weights)
+s
 
@@
 
-\subsection{Grammar Extensions}
+\subsection{Context Extensions to the Grammar}
 
-\subsubsection{Columnar Context}
-Once we have created a PCFG model of a process, we can apply existing PCFG parsing algorithms to identify a variety of date-time formats.
-However, the parser's success is often limited in the types of the dominant patterns that it can identify.
-In addition, the standard parsing techniques generally require specification of a complete observation sequence.
-In many contexts, we may have only a partial sequence available (\textit{e.g.} an incomplete entry).
-Finally, we may be interested in computing the probabilities of date-time patterns that the grammar may not explicitly define.
-To extend the forms of evidence, inferences, and pattern distributions supported, we need a flexible and expressive representation for the distribution of structures generated by the grammar.
-We adopt Bayesian networks for this purpose, and define an algorithm to generate a probabilistic distribution of possible parse trees corresponding to a set of date-time patterns as opposed to individual ones.
+The grammar parser's success is often limited by data ambiguity, and incomplete expressions.
+For example an input string `3/7/2005' could either be interpreted as as the 7th day of March or the 3rd day of July depending on the date format used.
+While a single entry cannot resolve the parser's ambiguity, a column of data provides more context for determining the dominant pattern prevalent in the dataset.
 
-\subsubsection{Locale and File Name Context}
+The CKY parser is run over the same set of samples from each validation file to generate a probabilistic distribution of possible parse trees corresponding to a set of date-time patterns.
+As a second pass of the parser, starting from the most probable parse tree, each parse tree is then applied to the entire file's columnar data to determine the dominant pattern.
+For example, another entry in the same dataset could be `25/3/2007' thus increasing the probability that the format is \texttt{dd/mm/yyyy} as opposed to \texttt{mm/dd/yyyy}.
 
-\subsubsection{PCFG Parser}
-We use a Cocke-Kasami-Younger (CKY) parser implemented in Python 3.4.1 for
-syntactically parsing a column of date-time strings.
-
-%¥	Explain why we use a probabilistic version of the grammar
-%¥	Explain how the production rules are created based on the ICU format
-%¥	Extensions to the parser to account for columnar context for maximizing the probabilistic occurrence of the dominant pattern
-%¥	Other semantic extensions Ð looking at the file name, adding rules for leap year, using external dictionaries for non-English terms
+In addition to the  columnar data context, the locale is are also used to help maximize the probabilistic likelihood of a particular date-time pattern.
+The locale of the data column is used to retrieve the corresponding corpora for helping with the parsing and for correcting syntactical errors.
+However, similar to the locale sensitivity issue described in Section 4.6, the locale and the data may not always match.
+The dominant pattern is simply computed over the data column based on the ranked set of parse tree results.

### 2015-07-16 testing.tex
Completed writing.

--- 2015-07-15 testing.tex
@@
 We implemented the grammar, training and parsing in Python 3.4.1 using the Natural Language Toolkit libraries~\cite{nltk} on a quad core Dell T7600 running Windows 7.
+The complexity of the CYK parsing algorithm is $O(n^{3}|G|)$, where $n$ is the length of the input and $|G|$ is the size of the grammar~\cite{Younger67}.
+The PCFG grammar uses a set of $22$ phrase-level non-terminals and $30$ terminals to classify the various constituents in the corpus.
+Since the CYK algorithm finds the maximum likelihood solution, there are no search errors (rather probability $p = 0$, and a modification in the grammar is required to improve accuracy.
+
+Similar to the MDL algorithm described in the previous subsection, the CYK parser is run over the same set of samples from each validation file to generate a ranked list of parse trees representing the date-time formats.
+As a second pass of the parser, starting from the most probable parse tree, each parse tree is then applied to the entire file's columnar data to determine the dominant pattern(s).
+For the second pass, we can parallelize the CYK parsing because the dependencies between reapplying the ranked list of parse trees are very sparse.
+
+The initial average parsing speed to compute the ranked set of probable parse trees is $0.93s$, while the average time taken to compute the overall dominant pattern(s) for the entire column of data is $1.4s$.
+While the natural language parsing implementation is in Python as the NLTK package is easily configurable, we could expect greater speed-up in parsing performance by employing C/C++ CYK parsing libraries.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=\columnwidth]{figures/FigureNLP1}
+\caption{Most common date formats identified by the NLP algorithm.}
+\label{fig:NLP1}
+\end{figure}
+
+Out of the $31,546$ files used for testing, the NLP parser identified a dominant format for $26, 534 (84.11\%)$ where $p >= 0.5$.
+$1634$ unique formats were identified.
+Figure \ref{fig:NLP1} shows a histogram of the most common formats identified.
 

### 2015-07-16 intro.tex
Fix typos

@@
 \item Default parsing of date formats is locale-dependent, and may not work when the analysis is shared across an international organization (\eg between the US and European offices)
-\item The parsing code was hard to understand and maintain because it used a verbose, general-purpose string-handling syntax instead of a specialized domain language.
+\item The parsing code is hard to understand and maintain because it uses a verbose, general-purpose string-handling syntax instead of a specialized domain language.
 \end{itemize}
@@
 \scriptsize{dd-MMM-yy hh.mm.ss.SSSSSS a} & \scriptsize{01-OCT-13 01.09.00.000000 PM}\\ \hline
-\scriptsize{MM ''yyyy} & \scriptsize{01 '2013}\\ \hline
+\scriptsize{MM 'yyyy} & \scriptsize{01 '2013}\\ \hline
 \scriptsize{MM/dd/yyyy - HH:mm} & \scriptsize{04/09/2014 - 23:47}\\ \hline

### 2015-07-16 nlp.tex
Remove extraneous s
Fix typos

@@
 For example, there is no  `November 31, 2015', `February 29, 2013', or `Sunday, May  5, 1965'.
-November only as 30 days in any year; 2013 was not a leap year; and May 5, 1965 was a Wednesday.
+November only has 30 days in any year; 2013 was not a leap year; and May 5, 1965 was a Wednesday.
 
@@
 
-s
-
 \subsection{Supervised Learning}
@@
 
-In addition to the  columnar data context, the locale is are also used to help maximize the probabilistic likelihood of a particular date-time pattern.
+In addition to the columnar data context, the locale is also used to help maximize the probabilistic likelihood of a particular date-time pattern.
 The locale of the data column is used to retrieve the corresponding corpora for helping with the parsing and for correcting syntactical errors.

### 2015-07-16 testing.tex
Fix typos

@@
 At 2.5ms, this is well below most human perceptual thresholds for a set of 32 samples, so any latency in command execution would be restricted to the ability of the underlying database to provide the samples for analysis in a timely manner.
-(As a column store, the TDE can often supply such domains without a full table scan, further improving responsiveness.)
+(As a column store, the database can often supply such domains without a full table scan, further improving responsiveness.)
 
-The validation speed is the average time needed to parse a value, and provides an estimate of how fast an ICU-based implementation (such as the TDE) can process string values into scalars and works out to 620K values per core per second.
+The validation speed is the average time needed to parse a value, and provides an estimate of how fast an ICU-based implementation can process string values into scalars and works out to 620K values per core per second.
 

### 2015-07-16 futurework.tex
Fix typos
Some typo fixes

@@
 
-In this paper, we have only considered the parsing of strings, but dates are often formatted as integers (\eg \texttt{201507016}).
-It is significantly faster to decompose integers into date parts using arithmetic operations (\eg \texttt{mod} and \texttt{div}) than by using slower, locale-sensitive string parsing functions.
+In this paper, we have only considered the parsing of strings, but dates are often formatted as integers (\textit{e.g.} \texttt{201507016}).
+It is significantly faster to decompose integers into date parts using arithmetic operations (\textit{e.g.} \texttt{mod} and \texttt{div}) than by using slower, locale-sensitive string parsing functions.
 Moreover, the number of possible formats is low enough that we may be able to enumerate them.
@@
 
-In the course of our research, we have also identified a number of date part variants (\eg ordinal dates, four-letter month abbreviations, alternate meridian markers and postfix quarter syntax) that we would like to commend to the ICU project and maybe provide implementations of.
+In the course of our research, we have also identified a number of date part variants (\textit{e.g.} ordinal dates, four-letter month abbreviations, alternate meridian markers and postfix quarter syntax) that we would like to commend to the ICU project and maybe provide implementations of.

### 2015-07-16 intro.tex
Some typo fixes

--- 2015-07-14 intro.tex
@@
 \item String operations are notoriously slow compared to scalar operations (typically 10-100x slower in modern RDBMSes)
-\item Default parsing of date formats is locale-dependent, and may not work when the analysis is shared across an international organization (\eg between the US and European offices)
+\item Default parsing of date formats is locale-dependent, and may not work when the analysis is shared across an international organization (\textit{e.g.} between the US and European offices)
 \item The parsing code was hard to understand and maintain because it used a verbose, general-purpose string-handling syntax instead of a specialized domain language.

### 2015-07-16 mdl.tex
Some typo fixes

@@
-The first algorithm is a Minimum Descriptive Length~\cite{Rissanen:1978} approach derived from the domain system presented in the Potter's Wheel system of Raman et al.~\cite{Raman:2001}.
+The first algorithm is a Minimum Descriptive Length~\cite{Rissanen:1978} approach derived from the domain system presented in the Potter's Wheel system of Raman \textit{et al.}~\cite{Raman:2001}.
 We describe a number of extensions to their structure extraction system to support more complex redundancy, non-English locales, improved performance and date-specific pruning.

### 2015-07-16 nlp.tex
Some typo fixes

--- 2015-07-16 nlp.tex
@@
 
-We define a grammar for identifying date-time strings based on other EBNF based date-time formats as a reference  [cite] . The definition of the grammar is found below:
+We define a \textit{EBNF} grammar for identifying date-time strings below:
 
@@
 We extend the date-time grammar by adding morpho-syntactic variants.
-These variants allows the parser to correct for syntactic errors (\textit{e.g.} usage of whitespace characters and punctuation marks), capitalization errors and abbreviations (\textit{e.g.} `Mon' for `Monday').
+These variants allows the parser to correct for syntactic errors (\textit{e.g.} usage of whitespace characters and punctuation marks), capitalization errors and abbreviations (\textit{e.g.} \texttt{`Mon'} for \texttt{Monday}).
 We use several various external corpora for such corrections~\cite{nltkcorpora}.
@@
 While we have added range restrictions to symbols such as \texttt{Hour} (1--12 for 12-hour format and 1--24 for 24-hour format), \texttt{Days} (1--7), \texttt{Month} (1--12), there are special cases that need to be accounted for.
-For example, there is no  `November 31, 2015', `February 29, 2013', or `Sunday, May  5, 1965'.
+For example, there is no  \texttt{`November 31, 2015'}, \texttt{`February 29, 2013'}, or \texttt{`Sunday, May  5, 1965'}.
 November only as 30 days in any year; 2013 was not a leap year; and May 5, 1965 was a Wednesday.
@@
 
-Except February, Equation $1$ addresses January and March through July. After July, the pattern should skip one, and the rest of the months should follow the alternating pattern inversely.
+Except for \texttt{February}, Equation $1$ addresses \texttt{January} and \texttt{March} through \texttt{July}.
+After \texttt{July}, the pattern should skip one, and the rest of the months should follow the alternating pattern inversely.
 
@@
 
-In Equation $2$ the number of days for August through December is correct ($x \in [8..12]$), but not for the remaining months.
+In Equation $2$ the number of days for \texttt{August} through \texttt{December} is correct ($x \in [8..12]$), but not for the remaining months.
 We hence introduce a bit-masking function so that the equation is equal to 1 over the desired domain and 0 otherwise.
@@
 
-\item LeapYear Restriction for February:
-While the above restriction applies to all months barring February, we also apply a constraint to the number of days for February, based on whether the year is leap year or not.
+\item LeapYear Restriction for \texttt{February}:
+While the above restriction applies to all months barring \texttt{February}, we also apply a constraint to the number of days for February, based on whether the year is leap year or not.
 For this, we define a new symbol in the grammar called \texttt{LeapYear}.
-If an expression containing the month `February' or any such variant (\textit{e.g.} `Feb', `2') with the day `29' and a year, would need to resolve the \texttt{Year} symbol to be a \texttt{LeapYear}, defined as:
+If an expression containing the month \texttt{`February'} or any such variant (\textit{e.g.} \texttt{`Feb'}, \texttt{`2'}) with the day \texttt{`29'} and a year, would need to resolve the \texttt{Year} symbol to be a \texttt{LeapYear}, defined as:
 
@@
 date, time or date-time), there could be several directions that the parser resolves to.
-For example, in a date string \texttt{5/6/2015}, the pattern could either be \texttt{M/d/yyyy} or \texttt{d/M/yyyy}.
+For example, in a date string \texttt{`5/6/2015'}, the pattern could either be \texttt{M/d/yyyy} or \texttt{d/M/yyyy}.
 
@@
 The grammar parser's success is often limited by data ambiguity, and incomplete expressions.
-For example an input string `3/7/2005' could either be interpreted as as the 7th day of March or the 3rd day of July depending on the date format used.
+For example an input string \texttt{`3/7/2005'} could either be interpreted as as the 7th day of March or the 3rd day of July depending on the date format used.
 While a single entry cannot resolve the parser's ambiguity, a column of data provides more context for determining the dominant pattern prevalent in the dataset.
@@
 As a second pass of the parser, starting from the most probable parse tree, each parse tree is then applied to the entire file's columnar data to determine the dominant pattern.
-For example, another entry in the same dataset could be `25/3/2007' thus increasing the probability that the format is \texttt{dd/mm/yyyy} as opposed to \texttt{mm/dd/yyyy}.
+For example, another entry in the same dataset could be \texttt{`25/3/2007'} thus increasing the probability that the format is \texttt{dd/mm/yyyy} as opposed to \texttt{mm/dd/yyyy}.
 

### 2015-07-16 testing.tex
Some typo fixes

--- 2015-07-16 testing.tex
@@
 The minimum description length implementation also recognizes the Excel date formats, but this was not supported by the natural language implementation.
-There was one file that contained formats such as ``Fall 2000'' and ``Spring 2000'', and the two algorithms picked different seasons in their format.
-One other file contained a set of integers that are not actually dates, and one algorithm picked MMyyyy format, while the other picked HHmmss format.
+There was one file that contained formats such as \texttt{`Fall 2000'} and \texttt{`Spring 2000'}, and the two algorithms picked different seasons in their format.
+One other file contained a set of integers that are not actually dates, and one algorithm picked \texttt{MMyyyy} format, while the other picked \texttt{HHmmss} format.

### 2015-07-16 intro.tex

--- 2015-07-16 intro.tex
@@
 \item String operations are notoriously slow compared to scalar operations (typically 10-100x slower in modern RDBMSes)
-\item Default parsing of date formats is locale-dependent, and may not work when the analysis is shared across an international organization (\eg between the US and European offices)
+\item The parsing code was hard to understand and maintain because it used a verbose, general-purpose string-handling syntax instead of a specialized domain language.
+\item Default parsing of date formats is locale-dependent, and may not work when the analysis is shared across an international organization (\textit{e.g} between the US and European offices)
 \item The parsing code is hard to understand and maintain because it uses a verbose, general-purpose string-handling syntax instead of a specialized domain language.

### 2015-07-16 nlp.tex

@@
 The date-time grammar also includes a large number of syntactically correct but semantically invalid
+
 date-time expressions.

### 2015-07-16 nlp.tex
Edits to text based on Dan's comments.

@@
 The parsed output is then converted to the ICU Date Format Language described in Section 3.2 in order to effectively cross-validate results from the MDL algorithm.
+While this grammar accounts for the use of `a.m.' and `p.m.' for the 12-hour format , ICU does not support these tokens, and they are simply ignored.
 
@@
 \end{itemize}
+
+However, Equation $4$ is just an approximation.
+The Gregorian calendar also requires that a year evenly divisible by $100$ (\textit{e.g.}, 1800) is a leap year only if it is also evenly divisible by $400$.
 

### 2015-07-16 testing.tex
Trying to explain differences in histogram of formats

@@
 $1634$ unique formats were identified.
-Figure \ref{fig:NLP1} shows a histogram of the most common formats identified.
+Figure \ref{fig:NLP1} shows a histogram of the most common formats identified containing \texttt{Year}.
+There are some variations compared to Figure \ref{fig:M3} including the fact that we are not filtering the results by any error rate in the NLP parsed results.
 

### 2015-07-16 futurework.tex
Edits from Rick Cole's paper pass.

@@
-While we are satisfied with the results so far, we found a number of column that we were not able to process with these techniques.
-Many columns contained multiple date formats, and we would like to recognize this situation and generate predicate-based calculations to increase the accuracy of the results and thereby make the experience even more seamless for the user.
+While we are satisfied with the results so far, we found a number of columns that we were not able to process with these techniques.
+Many columns contained multiple date formats.
+We would like to recognize this situation and generate predicate-based calculations to increase the accuracy of the results and thereby make the experience even more seamless for the user.
 Other columns contain date ranges, which we would like to handle by generating multiple calculations, possibly by combining regular expressions with date parsing.

### 2015-07-16 intro.tex
Edits from Rick Cole's paper pass.

@@
 While the existing standards for query languages are helpful for defining a core set of row-level functions, there are many useful functions beyond the standards that are only explicitly supported by a subset of RDBMSes, often with different names.
-And even when a database does not provide an implementation of a particular function, it is usually possible to generate it as an inline combination of existing functionality.
+Even when a database does not provide an implementation of a particular function, it is usually possible to generate it as an inline combination of existing functionality.
 The calculation languages of visualization systems can thus be designed to provide a \textit{lingua franca} that tames this Babel of dialects by providing a common syntax for as many of these functions as possible.
@@
 \item String operations are notoriously slow compared to scalar operations (typically 10-100x slower in modern RDBMSes)
-\item The parsing code was hard to understand and maintain because it used a verbose, general-purpose string-handling syntax instead of a specialized domain language.
 \item Default parsing of date formats is locale-dependent, and may not work when the analysis is shared across an international organization (\textit{e.g} between the US and European offices)

### 2015-07-16 nlp.tex
Removed the spatial reference.
Edits from Rick Cole's paper pass.

@@
 
-We define a \textit{EBNF} grammar for identifying date-time strings below:
+We define a \textit{EBNF} grammar for identifying date-time strings as follows:
 
 \begin{grammar}
-<TimeGrammar> ::= <Hours> ':' <Minutes> ':' <Seconds> <TimeZone> <AMPM> (for 12-hour formats)
+<TimeGrammar> ::= <Hours> ':' <Minutes> ':' <Seconds> <TimeZone> <AMPM> (for 12-hour formats);
 
@@
 
-The date-time grammar also includes a large number of syntactically correct but semantically invalid
-
-date-time expressions.
+The date-time grammar also includes a large number of syntactically correct but semantically invalid date-time expressions.
 While we have added range restrictions to symbols such as \texttt{Hour} (1--12 for 12-hour format and 1--24 for 24-hour format), \texttt{Days} (1--7), \texttt{Month} (1--12), there are special cases that need to be accounted for.
@@
 \item LeapYear Restriction for \texttt{February}:
-While the above restriction applies to all months barring \texttt{February}, we also apply a constraint to the number of days for February, based on whether the year is leap year or not.
+While the above restriction applies to all months barring \texttt{February}, we also apply a constraint to the number of days for \texttt{February}, based on whether the year is leap year or not.
 For this, we define a new symbol in the grammar called \texttt{LeapYear}.

### 2015-07-16 testing.tex
Edits from Rick Cole's paper pass.

@@
 The PCFG grammar uses a set of $22$ phrase-level non-terminals and $30$ terminals to classify the various constituents in the corpus.
-Since the CYK algorithm finds the maximum likelihood solution, there are no search errors (rather probability $p = 0$, and a modification in the grammar is required to improve accuracy.
+Because the CYK algorithm finds the maximum likelihood solution, there are no search errors (rather probability $p = 0$, and a modification in the grammar is required to improve accuracy.
 

### 2015-07-16 conclusion.tex
weird spacing when using definition.

--- 2015-07-06 conclusion.tex
@@
 While validating the algorithms on a large corpus, we also found that the number of distinct formats in the wild is surprisingly high, and demonstrates the wisdom of including general-purpose date parsing functions in data visualization tools, data cleaning tools and RDBMSes.
-In particular, it is interesting to note that the most prominent open source RDBMSes (\eg MySQL and Postgres) both have a built-in version of \dateparse, possibly reflecting that this is a common need that gets implemented when users are empowered to add to a code base.
+In particular, it is interesting to note that the most prominent open source RDBMSes (\eg MySQL and Postgres) both have a built-in version of \texttt{DATEPARSE}, possibly reflecting that this is a common need that gets implemented when users are empowered to extend the function library of an RDBMS.

### 2015-07-16 nlp.tex
text tweaks

@@
 The parsed output is then converted to the ICU Date Format Language described in Section 3.2 in order to effectively cross-validate results from the MDL algorithm.
-While this grammar accounts for the use of `a.m.' and `p.m.' for the 12-hour format , ICU does not support these tokens, and they are simply ignored.
+While this grammar accounts for the use of meridian markers `a.m.' and `p.m.' for the 12-hour format, ICU does not support these tokens, and they are simply ignored.
 

### 2015-07-16 nlp.tex
More typo fixes

@@
 
-While there are several functionally equivalent notations for representing CFG, we use the \textit{Backus-Naur Form} (BNF) for defining the grammar rules for \dateparse formats.
+While there are several functionally equivalent notations for representing CFG, we adopt the \textit{Backus-Naur Form} (BNF) for defining the grammar rules for \texttt{DATEPARSE} formats.
 In particular, we use the \textit{Extended Backus-Naur Form} (EBNF) as the notation is more compact and readable for frequently used constructions~\cite{Grune:1990}.
-We use a Cocke-Younger-Kasami (CYK) parser for syntactically parsing a column of date-time strings based on the grammar~\cite{Cocke:1969,Younger67,Kasami:1965}.
+We employ a Cocke-Younger-Kasami (CYK) parser for syntactically parsing a column of date-time strings based on the grammar~\cite{Cocke:1969,Younger67,Kasami:1965}.
 The parser is a dynamic programming algorithm that identifies the highest probable syntactic parse trees, given the grammar and an input date-time string.
@@
 In the context of CFGs, probabilities have been used to define a probability distribution over a set of parse trees defined by the CFG, and are a useful method for modeling such ambiguity~\cite{Collins:2003,Manning:1999}.
-The resulting formalism called Probabilistic Context-Free Grammar (PCFG), extends the CFG by assigning probabilities to the production rules of the grammar.
+The resulting formalism called \textit{Probabilistic Context-Free Grammar} (PCFG), extends the CFG by assigning probabilities to the production rules of the grammar.
 During the process of parsing the date-time pattern, the probabilities are used as a filtering mechanism to rank the pattern(s) that a given string resolves to in the grammar.
@@
 Having defined initial probabilistic weights to the \texttt{PCFG}, we employ supervised learning with a known training set of date time formats to estimate the rule probabilities.
-The training corpus is a set of files obtained from an online data analysis corpus, described in Section 5.
+The training corpus is a set of files obtained from an online data analysis corpus, described in Section 6.
 

### 2015-07-16 testing.tex
More typo fixes

@@
 Many of the files in the collection are not actually dates, such as fields named ``updated by'' (which contain ``date'').
-A random sample of 850 columns named exactly ``date'', ``time'' or ``month'' (case-insensitive) were manually judged.
+A random sample of 850 columns named exactly ``date'', ``time'' or ``month'' (case-insensitive) were manually judged.\\
 
@@
 Some of the dates are clearly just numbers, but our approach is to assume that when the user tells us that the column contains dates we should find the best fit.
-The samples include dates from a wide range of historical sources (\eg Roman pottery dates) so we have elected to defer the date identification task to the user.
+The samples include dates from a wide range of historical sources (\eg Roman pottery dates) so we have elected to defer the date identification task to the user.\\\\s
 

### 2015-07-16 intro.tex
WITH TIME ZONE

@@
 
-The SQL-99 standard defines three temporal scalar types: \texttt{DATE}, \texttt{TIMESTAMP} and \texttt{TIME}.
-They are typically implemented as fixed-point types containing an offset from some epoch (\eg Julian Days.) This makes them compact to store using column store compression techniques such as those in C-Store~\cite{Stonebraker:2005} and MonetDB/X100~\cite{Zukowski:2006}, and further allows some temporal operations to be implemented very efficiently using simple arithmetic.
+The SQL-99 standard defines three temporal scalar types: \texttt{DATE}, \texttt{TIMESTAMP} and \texttt{TIME}, which can be further qualified as either \texttt{WITH} or \texttt{WITHOUT} \texttt{TIME ZONE} (the \texttt{WITHOUT} form being more common.) They are typically implemented as fixed-point types containing an offset from some epoch (\eg Julian Days.) This makes them compact to store using column store compression techniques such as those in C-Store~\cite{Stonebraker:2005} and MonetDB/X100~\cite{Zukowski:2006}, and further allows some temporal operations to be implemented very efficiently using simple arithmetic.
 Thus from the RDBMS perspective, representing dates in scalar form provides  benefits for users, both in terms of available operations and query performance.
@@
 \item String operations are notoriously slow compared to scalar operations (typically 10-100x slower in modern RDBMSes)
-\item Default parsing of date formats is locale-dependent, and may not work when the analysis is shared across an international organization (\textit{e.g} between the US and European offices)
+\item Default parsing of date formats is locale-dependent, and may not work when the analysis is shared across an international organization (\eg between the US and European offices)
 \item The parsing code is hard to understand and maintain because it uses a verbose, general-purpose string-handling syntax instead of a specialized domain language.
@@
 
-\subsubsection{	Usability}
+\subsubsection{Usability}
 After providing this new function to our user base, we waited a few months and then investigated how it was being applied in a free, cloud-based version of the system.

### 2015-09-23 icde2016/futurework.tex
moved paper draft into separate subfolders

--- 2015-07-16 futurework.tex
@@
 
-In the course of our research, we have also identified a number of date part variants (\textit{e.g.} ordinal dates, four-letter month abbreviations, alternate meridian markers and postfix quarter syntax) that we would like to commend to the ICU project and maybe provide implementations of.
+In the course of our research, we have also identified a number of date part variants (\textit{e.g.} ordinal dates, four-letter month abbreviations, alternate meridian markers and postfix quarter syntax) that we would like to commend to the ICU project, along with possible implementations.

### 2015-09-23 icde2016/testing.tex
moved paper draft into separate subfolders

--- 2015-07-16 testing.tex
@@
 
-The validation speed is the average time needed to parse a value, and provides an estimate of how fast an ICU-based implementation can process string values into scalars and works out to 620K values per core per second.
+The validation speed is the average time needed to parse a value, and provides an estimate of how fast an ICU-based implementation can process string values into scalars and works out to 620K values per core per second.\\
 

### 2015-09-23 sigmod2016/parameters.tex
moved paper draft into separate subfolders

--- 2015-07-07 parameters.tex
@@
 \subsection{Input Data}
-The training data and test data are taken from data sets published to our free, online visualization service.
+The training data and test data are taken from data sets published to a free, online data analysis website.
 The published data includes both the raw data and how it was used for analysis.
-The service limits data sets to a maximum of 100K rows stored in a database which supports an implementation of \dateparse.
+The service limits data sets to a maximum of 100K rows stored in a database which supports an implementation of \texttt{DATEPARSE}.
 We selected a large number of string columns whose names suggested that they might contain temporal data.

### 2015-09-23 sigmod2016/relatedwork.tex
moved paper draft into separate subfolders

--- 2015-07-15 relatedwork.tex
@@
-\subsection{Previous Work}
+%\subsection{Previous Work}
 
@@
 There are several bodies of research on developing semantic parsers and grammars for interpreting time and date expressions.
-Lee \textit{et al} use a combinatory categorical grammar by combining the TIMEX3 standard~\cite{timex3} with contextual cues such as document creation time to determine the reference for parsing time expressions such as `2nd Friday of July'~\cite{LeeADZ14}.
+Lee \textit{et al.} use a combinatory categorical grammar by combining the TIMEX3 standard~\cite{timex3} with contextual cues such as document creation time to determine the reference for parsing time expressions such as `2nd Friday of July'~\cite{LeeADZ14}.
 Related work by Gabor \textit{et al.} employs a probabilistic approach for learning to interpret temporal phrases~\cite{Angeli:2012}.
@@
 \item We show how to extend prior work on Minimum Descriptive Length structure extraction to generate a freely available date format domain language with over 90\% accuracy;
-\item We describe a second Natural Language Processing technique for generating the same date format domain language with similar accuracy;
+\item We describe a second Natural Language Processing technique for generating the same date format domain language with similar accuracy.
+We describe how the basic algorithm is extended to support grammar variants and constraints unique to date formats.
+The parsing algorithm is extended to compute an overall dominant pattern over a data column;
 \item We describe some limitations of this domain language that would improve its utility.

### 2015-09-28 icde2016/intro.tex
Abstract & Intro
Also created new background section to absorb parameters and remaining
intro stuff.

--- 2015-07-16 intro.tex
@@
 
-\subsubsection{Clean as You Go}
 During the analytic process, a user may discover that parts of the data are not yet suitable for analysis.
@@
 
-\subsubsection{Function Libraries}
-While the existing standards for query languages are helpful for defining a core set of row-level functions, there are many useful functions beyond the standards that are only explicitly supported by a subset of RDBMSes, often with different names.
-Even when a database does not provide an implementation of a particular function, it is usually possible to generate it as an inline combination of existing functionality.
-The calculation languages of visualization systems can thus be designed to provide a \textit{lingua franca} that tames this Babel of dialects by providing a common syntax for as many of these functions as possible.
-
-\subsection{\dateparse}
-The focus of this work is the usability of one such function from our visualization system that we will call \dateparse.
-
-\subsubsection{	Scalar Dates}
-
-The SQL-99 standard defines three temporal scalar types: \texttt{DATE}, \texttt{TIMESTAMP} and \texttt{TIME}, which can be further qualified as either \texttt{WITH} or \texttt{WITHOUT} \texttt{TIME ZONE} (the \texttt{WITHOUT} form being more common.) They are typically implemented as fixed-point types containing an offset from some epoch (\eg Julian Days.) This makes them compact to store using column store compression techniques such as those in C-Store~\cite{Stonebraker:2005} and MonetDB/X100~\cite{Zukowski:2006}, and further allows some temporal operations to be implemented very efficiently using simple arithmetic.
-Thus from the RDBMS perspective, representing dates in scalar form provides  benefits for users, both in terms of available operations and query performance.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=\columnwidth]{figures/FigureI1}
-\caption{Categorical Date Scalars.}
-\label{fig:I1}
-\end{figure}
-
-Our visualization system models the first two of these SQL-99 types natively; the third (pure time) is folded into \texttt{TIMESTAMP} by appending it to a fixed date of \texttt{1899-12-30} (a convention derived from Microsoft Excel).
-From the analytic perspective, date types are dimensional (\ie independent variables) and can be used as either \textit{categorical} (simply ordered) or \textit{quantitative} (ordered with a distance metric) fields.
-
-Categorical dates have a natural hierarchy associated with them generated by calendar \textit{binning}.
-The visualization in Figure \ref{fig:I1} shows an example of a bar chart employing binned categorical dates in a year/quarter hierarchy.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=\columnwidth]{figures/FigureI2}
-\caption{Quantitative Date Scalars.}
-\label{fig:I2}
-\end{figure}
-
-Quantitative dates are typically used for time series on an axis that maps the underlying distance measure to display pixel distance.
-The quantitative analog to categorical binning is \textit{truncation} whereby all the dates in a bin are represented by the first date in the bin.
-Truncation accurately preserves the distance semantics while enabling roll-up to coarser levels of detail.
-The visualization in Figure \ref{fig:I2} shows the same sales data in a time series rolled up to the quarter level.
-
-These types of visualizations are difficult to specify with simple categorical string representations of dates.
-Thus the conversion of unparsed strings to date scalars is also an important component of data modeling for analytics.
-
-\subsubsection{Parsing}
+\subsection{Parsing Dates}
 This parsing of date representations is one of the oldest data preparation problems we have observed.
@@
 Some are common, but others can be quite idiosyncratic.
-Table 1 shows a selection of unusual date formats found in our test data.
+Table 1 shows a selection of unusual date formats found in our online data.
 The first example shows a time zone in the middle of the date and a year after the time; the second shows a leading unmatched bracket and a colon between the date and time components; the third shows confusion between the seconds' decimal point and the time part delimiter; the fourth shows a two digit year apostrophe on a four digit year and the fifth shows a dash separating the date and time components.
@@
 %\begin{tabular}{ |l|l|l| }
-\begin{tabular}{|p{0.498\linewidth}| p{0.485\linewidth}|}
+\begin{tabular}{|p{0.4\linewidth}| p{0.4\linewidth}|}
 \hline
@@
 
-Our first attempt to help users with this problem was to add a new function to the the calculation language called \dateparse.
-This function would take a string column and convert it to a datetime using a special purpose domain language for describing dates.
-Such functions exist in a number of databases (\eg MySQL, Oracle and Postgres), so it was a natural addition to the function library.
+Several RDBMSes (\eg MySQL, Oracle and Postgres) provide row-level functions for formatted date parsing, but our analysis of user data showed a 15\% syntax error rate with these functions.
+And even with perfect formatting, the user still has to interrupt their flow to learn a formatting syntax -- which they will likely forget once their problem has been solved.
 
-We chose the date parsing syntax defined by the International Components for Unicode (or ICU) project~\cite{ICU} for the common domain language because it mirrored what was already available in our code base, and translating this syntax into the formats used by various database vendors was relatively painless.
-There are a few patches of non-overlapping functionality, but they tend to be obscure and we provide warnings when functionality is missing.
+One solution to this problem might have been to design a graphical environment that enabled users to construct valid patterns using visually compelling representations to reduce the cognitive load.
+This approach would have involved a substantial development effort with no guarantee that the result would be correct if the user misunderstood the environment.
+Moreover, this approach only solves the syntax problem -- the user is still required to switch contexts to solve their problem.
 
-\subsubsection{Usability}
-After providing this new function to our user base, we waited a few months and then investigated how it was being applied in a free, cloud-based version of the system.
-We found that although there were a number of new calculations using \dateparse, the error rate for the domain language syntax was about 15\%.
-\dateparse was capable of solving the problem, and some users were able to discover it, but the syntax did not appear to be easy to use reliably.
+Our solution wa to develop two machine learning algorithms for automatically deriving the format string from the user's data with over 90\% parsing accuracy.
+This approach allows the user to simply specify "this column is a date" and the visualisation system can respond quickly and accurately enough to avoid interrupting the user's analytic flow.
 
-\subsection{Automated Format Extraction}
-One solution to this problem might have been to design a graphical environment that enabled users to construct valid patterns.
-This would have involved a substantial development effort with no guarantee that the result would be correct if the user misunderstood the environment.
-Instead, we have developed two algorithms for automating the derivation of the format string, each of which uses a different machine learning technique.
-Both algorithms result in over 90\% parsing accuracy -- and 100\% syntactic correctness because they are machine generated.
+\subsection{Contributions}
+Specifically, our contributions include:
+
+\begin{itemize}
+\item Quantification of the surprising variety of date formats in use;
+\item Descriptions of two algorithms for automating date format extraction, at least one of which is performant enough for interactive use;
+\item Demonstration of an effective approach for integrating data cleaning into analytic processes.
+\end{itemize}
+
+\subsection{Organization}
+The rest of this paper is organized as follows: The next section introduces the parameters of the problem space.
+The following two sections describe the two different algorithms, one using Minimum Descriptive Length and the other using Natural Language Processing.
+In section 6, we evaluate the algorithms on a corpus of 30K columns, both by sampling the outputs manually and then by using the algorithms to validate each other.
+We then discuss future work in section 7 and conclude in section 8.

### 2015-09-28 icde2016/paper.tex
Changed paper to use ICDE latex template and bib
Abstract & Intro
Also created new background section to absorb parameters and remaining
intro stuff.

--- 2015-07-08 paper.tex
@@
 
-\documentclass{acm_proc_article-sp}
+%\documentclass{acm_proc_article-sp}
+\documentclass[conference]{IEEEtran}
 %
@@
 \usepackage{syntax}
+\usepackage{amsmath}% http://ctan.org/pkg/amsmath
+\usepackage{mathtools}
+\usepackage{color}
 
@@
 \def\ie{\textit{i.e.} }
-\def\dateparse{\texttt{DATEPARSE} }
+\def\dateparse{\texttt{DATEPARSE}}
+\hyphenation{time-stamp}
+
+% to add comments
+\newcommand{\vidya}[1]{\textcolor{red}{(vidya: #1)}}
+
+% formatting math symbols
+\newcommand{\Mod}[1]{\ (\text{mod}\ #1)}
+\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
+\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
+
+% tightening the whitespace for itemize
+\newenvironment{tight_itemize}{\begin{itemize} \itemsep
+-2pt}{\end{itemize}}
+
 %
@@
 
-\conferenceinfo{SIGMOD}{'16 SanFrancisco, CA USA}
-
 \title{Does Anybody Really Know What Time It Is?\\
 Automating the Extraction of Date Scalars}
-\input{anonymous}
+%\input{anonymous}
+% author names and affiliations
+% use a multiple column layout for up to three different
+% affiliations
+%\author{\IEEEauthorblockN{Michael Shell}
+%\IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
+%Georgia Institute of Technology\\
+%Atlanta, Georgia 30332--0250\\
+%Email: http://www.michaelshell.org/contact.html}
+%\and
+%\IEEEauthorblockN{Homer Simpson}
+%\IEEEauthorblockA{Twentieth Century Fox\\
+%Springfield, USA\\
+%Email: homer@thesimpsons.com}
+%\and
+%\IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
+%\IEEEauthorblockA{Starfleet Academy\\
+%San Francisco, California 96678-2391\\
+%}}
 
@@
 \begin{abstract}
-With the advent of modern data visualization tools, data preparation has become a bottleneck for analytic workflows.
-Users who are already engaged in data analysis prefer to stay in the visualization environment for cleaning and preparation tasks whenever possible, both to preserve analytic flow and to take advantage of the visualization environment to spot inconsistent data.
-This has led many visualization environments to include simple data preparation functions such as scalar parsing, pattern matching and categorical binning in their analytic toolkits.
+Interactive visual analytics is an important set of tools for the timely analysis of big data.
+Users who are already engaged in interactive data analysis prefer to stay in the visualization environment for unavoidable cleaning and preparation tasks in order to preserve their analytic flow.
+This has led visualization environments to include simple data preparation functions such as scalar parsing, pattern matching and categorical binning in their analytic toolkits.
+One of the most common scalar parsing tasks is extracting date and time data from string representations.
+Several RDBMSes include date parsing ``mini-languages'' to cover the wide range of possible formats, but analysis of user data from one visualization system shows that the parsing language syntax is difficult for users to master without disturbing their analytic flow.
 
-One of the most common scalar parsing tasks is extracting date and time data from string representations.
-Several RDBMSes include date parsing ``mini-languages'' to cover the wide range of possible formats.
-Analysis of the usage of such functions in one visualization system shows, however, that the date parsing language syntax is difficult for users to master.
-
-In this paper, we present two algorithms for automatically deriving date format strings from a column of data, one based on minimum entropy and one based on natural language modeling.
-Both have similar accuracies of over 90\% on a large corpus of date columns extracted from an online data repository and are in substantial agreement with each other.
-The minimal entropy approach is also fast enough to produce results within the user's perceptual threshold, making it suitable for interactive work.
+In this paper, we present two algorithms for automatically deriving these date formats from a column of data with minimal user disruption.
+Both have accuracies of over 90\% on a large corpus of date columns extracted from an online data repository and are in substantial agreement with each other.
+One approach is also fast enough to produce results within the user's perceptual threshold, making it suitable for interactive work.
 \end{abstract}
 
+\vidya{Need to figure out the IEEE equivalent for categories and terms}
 %D.3.3 [Data Processing]: Data Cleaning, Natural Language Processing.
 % A category with the (minimum) three required fields
-\category{D.3.3}{Data Processing}{Data Cleaning, Natural Language Processing}
+%\category{D.3.3}{Data Processing}{Data Cleaning, Natural Language Processing}
 %A category including the fourth, optional field follows...
@@
 
-\terms{Algorithms, Performance}
+%\terms{Algorithms, Performance}
 
@@
 
-\section{Parameters}
-\input{parameters}
+\section{Background}
+\input{parameters}\\\\
 
@@
 \input{testing}
+
+\section{Related Work}
+\input{relatedwork}
 
@@
 % produce the bibliography for the citations in your paper.
-\bibliographystyle{abbrv}
+\bibliographystyle{IEEEtran}
 \bibliography{dateparse}  % sigproc.bib is the name of the Bibliography in this case
@@
 
-\balancecolumns
+%\balancecolumns
 % That's all folks!

### 2015-09-28 icde2016/relatedwork.tex
Abstract & Intro
Also created new background section to absorb parameters and remaining
intro stuff.

--- 2015-09-23 sigmod2016/relatedwork.tex
@@
 \end{itemize}
-
-\subsection{Overview}
-The rest of this paper is organized as follows: The next section introduces the parameters of the problem space.
-The following two sections describe the two different algorithms, one using Minimum Descriptive Length and the other using Natural Language Processing.
-In section 6, we evaluate the algorithms on a corpus of 30K columns, both by sampling the outputs manually and then by using the algorithms to validate each other.
-We then discuss future work in section 7 and conclude in section 8.

### 2015-09-29 icde2016/background.tex
Abstract & Intro
Also created new background section to absorb parameters and remaining
intro stuff.
Move background contents
Put it in more useful locations.

--- 2015-07-16 intro.tex
@@
-In recent years, there has been a growth of interest in data visualization technologies for human-assisted data analysis using systems such as Polaris~\cite{Stolte:2008} and Spotfire~\cite{Ahlberg:1996}.
-While computers can provide high-speed and high-volume data processing, humans have domain knowledge and the ability to process data in parallel, notably by using our visual systems.
-Most importantly, humans provide the definition of what is valuable in an analysis.
-Accordingly, human/computer analytic systems are essential to extracting knowledge of value to humans and their societies from the large amounts of data being generated today.
+Before delving into detailed descriptions of the two algorithms, we describe the common elements of the problem they are intended to solve.
+These are the input data, the output language and the interpretation of the output language for partial dates.
 
-\subsection{Interactivity}
-Visualization systems are most effective when they are interactive, thereby allowing a user to explore data and connect it to their domain knowledge and sense of what is important without breaking cognitive flow.
-In recent years, a number of such systems have been developed, both by the academic community and by the commercial sector.
-Exploration of data consists not only in creating visual displays, but also in creating and modifying domain-specific computations.
-Consequently, most data visualization systems include facilities for defining such calculations as part of the data model being analyzed.
-The most effective systems allow users to define these calculations as part of the analytic interaction, which permits the user to stay in the flow of analysis~\cite{Morton:2012}.
+\subsection{Input Data}
+The primary assumption of the algorithms is that we have a column of string values that the user has indicated are really date scalars.
+The strings may also be in a non-English language, so we also extract the string collation from the column metadata and provide it as an optional secondary input.
 
-\subsubsection{Clean as You Go}
-During the analytic process, a user may discover that parts of the data are not yet suitable for analysis.
-Solutions to this problem are often provided by data preparation tools external to the visual analysis environment, which requires the user to break their analytic flow, launch another tool and reprocess their data before returning to their analysis.
-If the user does not own this process (\eg it is the responsibility of another department), then there can be significant delays (including ``never.'') More subtly, the result of updated external processing may not be compatible with the user's existing work, which can lead to more time lost reconciling the new data model with the existing analysis.
+\subsection{The ICU Date Format Language}
+The date format syntax we selected is the one provided by the ICU open-source project.
+We chose it because we were already using ICU in the code base, we had access to the source code, and it provides localized date part data for a large number of languages.
 
-From the user's perspective, the boundary between preparation and analysis is not nearly so clean cut.
-Bad data is often discovered using visual analysis techniques (e.g.
-histograms or scatter plots) and it is most natural for the user to ``clean what she sees'' instead of switching to a second tool.
-This leads to an ``adaptive'' process whereby users will prefer leveraging existing tools in the analytics environment (no matter how well suited to the task) over switching to another application.
-Thus a well-designed interactive visual analysis environment will provide tools that enable users to perform such cleaning tasks as interactively as possible.
-
-\subsubsection{Function Libraries}
-While the existing standards for query languages are helpful for defining a core set of row-level functions, there are many useful functions beyond the standards that are only explicitly supported by a subset of RDBMSes, often with different names.
-Even when a database does not provide an implementation of a particular function, it is usually possible to generate it as an inline combination of existing functionality.
-The calculation languages of visualization systems can thus be designed to provide a \textit{lingua franca} that tames this Babel of dialects by providing a common syntax for as many of these functions as possible.
-
-\subsection{\dateparse}
-The focus of this work is the usability of one such function from our visualization system that we will call \dateparse.
-
-\subsubsection{	Scalar Dates}
-
-The SQL-99 standard defines three temporal scalar types: \texttt{DATE}, \texttt{TIMESTAMP} and \texttt{TIME}.
-They are typically implemented as fixed-point types containing an offset from some epoch (\eg Julian Days.) This makes them compact to store using column store compression techniques such as those in C-Store~\cite{Stonebraker:2005} and MonetDB/X100~\cite{Zukowski:2006}, and further allows some temporal operations to be implemented very efficiently using simple arithmetic.
-Thus from the RDBMS perspective, representing dates in scalar form provides  benefits for users, both in terms of available operations and query performance.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=\columnwidth]{figures/FigureI1}
-\caption{Categorical Date Scalars.}
-\label{fig:I1}
-\end{figure}
-
-Our visualization system models the first two of these SQL-99 types natively; the third (pure time) is folded into \texttt{TIMESTAMP} by appending it to a fixed date of \texttt{1899-12-30} (a convention derived from Microsoft Excel).
-From the analytic perspective, date types are dimensional (\ie independent variables) and can be used as either \textit{categorical} (simply ordered) or \textit{quantitative} (ordered with a distance metric) fields.
-
-Categorical dates have a natural hierarchy associated with them generated by calendar \textit{binning}.
-The visualization in Figure \ref{fig:I1} shows an example of a bar chart employing binned categorical dates in a year/quarter hierarchy.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=\columnwidth]{figures/FigureI2}
-\caption{Quantitative Date Scalars.}
-\label{fig:I2}
-\end{figure}
-
-Quantitative dates are typically used for time series on an axis that maps the underlying distance measure to display pixel distance.
-The quantitative analog to categorical binning is \textit{truncation} whereby all the dates in a bin are represented by the first date in the bin.
-Truncation accurately preserves the distance semantics while enabling roll-up to coarser levels of detail.
-The visualization in Figure \ref{fig:I2} shows the same sales data in a time series rolled up to the quarter level.
-
-These types of visualizations are difficult to specify with simple categorical string representations of dates.
-Thus the conversion of unparsed strings to date scalars is also an important component of data modeling for analytics.
-
-\subsubsection{Parsing}
-This parsing of date representations is one of the oldest data preparation problems we have observed.
-A common version of this problem is how to convert columns of integers of the form \texttt{yyyyMMdd} to date scalars.
-Na\"{i}ve users often solve this problem by converting the integer to a string and performing some locale-dependent string operations before casting the string back to a date.
-Unfortunately, this approach has a number of problems:
+The syntax is documented at the ICU web site~\cite{ICU}. While fairly complete, it has a few limitations that we ran into when evaluating the algorithms:
 \begin{itemize}
-\item String operations are notoriously slow compared to scalar operations (typically 10-100x slower in modern RDBMSes)
-\item Default parsing of date formats is locale-dependent, and may not work when the analysis is shared across an international organization (\textit{e.g} between the US and European offices)
-\item The parsing code is hard to understand and maintain because it uses a verbose, general-purpose string-handling syntax instead of a specialized domain language.
+\item No support for 4 letter year abbreviations (\eg \texttt{Sept.})
+\item No support for ordinal days (\eg \texttt{July 4th})
+\item No support for quarter postfix notation (\eg \texttt{2Q})
+\item No support for variant meridian markers (\eg \texttt{a.m.})
 \end{itemize}
 
-And this is but a single date format.
-Our studies of online data collections suggest that there are hundreds of distinct temporal date formats in user data sets.
-Some are common, but others can be quite idiosyncratic.
-Table 1 shows a selection of unusual date formats found in our test data.
-The first example shows a time zone in the middle of the date and a year after the time; the second shows a leading unmatched bracket and a colon between the date and time components; the third shows confusion between the seconds' decimal point and the time part delimiter; the fourth shows a two digit year apostrophe on a four digit year and the fifth shows a dash separating the date and time components.
+These limitations did not affect the results significantly, and in the future we hope to submit ICU extensions to handle some of these issues.
 
-\begin{table}[ht]
-\centering
-%\begin{tabular}{ |l|l|l| }
-\begin{tabular}{|p{0.498\linewidth}| p{0.485\linewidth}|}
-\hline
-\centering
-\textbf{ICU Format} & \textbf{Example}\\ \hline
-\scriptsize{EEE MMM dd HH:mm:ss zzz yyyy} & \scriptsize{Fri Apr 01 02:09:27 EDT 2011}\\ \hline
-\scriptsize{[dd/MMM/yyyy:HH:mm:ss} & \scriptsize{[10/Aug/2014:09:30:40}\\ \hline
-\scriptsize{dd-MMM-yy hh.mm.ss.SSSSSS a} & \scriptsize{01-OCT-13 01.09.00.000000 PM}\\ \hline
-\scriptsize{MM 'yyyy} & \scriptsize{01 '2013}\\ \hline
-\scriptsize{MM/dd/yyyy - HH:mm} & \scriptsize{04/09/2014 - 23:47}\\ \hline
-\end{tabular}
-\label{tab:dateformats}
-\caption{Unusual Date Formats.}
-\end{table}
+One other quirk of the ICU syntax may be a contributing factor to the user confusion around writing correct ICU date formats.
+The use of lexicographical case in the meta-symbols of the format language can be confusing (\eg \texttt{y} is used for years, but \texttt{M} is used for months while \texttt{m} is used for minutes.) Another advantage of an automated algorithm is that it hides such problems from most users, significantly improving the usability of the function.
 
-Our first attempt to help users with this problem was to add a new function to the the calculation language called \dateparse.
-This function would take a string column and convert it to a datetime using a special purpose domain language for describing dates.
-Such functions exist in a number of databases (\eg MySQL, Oracle and Postgres), so it was a natural addition to the function library.
+\subsection{Partial Dates}
+Many of the date formats that we encountered were incomplete dates, which necessitated creating rules for what date scalar they represented.
 
-We chose the date parsing syntax defined by the International Components for Unicode (or ICU) project~\cite{ICU} for the common domain language because it mirrored what was already available in our code base, and translating this syntax into the formats used by various database vendors was relatively painless.
-There are a few patches of non-overlapping functionality, but they tend to be obscure and we provide warnings when functionality is missing.
+ICU's date parsing APIs allow the specification of default values for parts when a format does not contain them.
+In our implementation, all time fields are set to \texttt{0} (midnight) and the date fields are set based on whether the format contains any date part specifications.
+When date parts are present, we use \texttt{2000-01-01} as the set of default date parts as it is the start of a leap year.
+When dealing with pure time formats, we model the output as a date/time and use \texttt{1899-12-30} for the date parts.
 
-\subsubsection{	Usability}
-After providing this new function to our user base, we waited a few months and then investigated how it was being applied in a free, cloud-based version of the system.
-We found that although there were a number of new calculations using \dateparse, the error rate for the domain language syntax was about 15\%.
-\dateparse was capable of solving the problem, and some users were able to discover it, but the syntax did not appear to be easy to use reliably.
-
-\subsection{Automated Format Extraction}
-One solution to this problem might have been to design a graphical environment that enabled users to construct valid patterns.
-This would have involved a substantial development effort with no guarantee that the result would be correct if the user misunderstood the environment.
-Instead, we have developed two algorithms for automating the derivation of the format string, each of which uses a different machine learning technique.
-Both algorithms result in over 90\% parsing accuracy -- and 100\% syntactic correctness because they are machine generated.
+ICU will also parse Time Zones (which the algorithms recognize but the visualization system does not model) and Quarters (which we interpret as the first month of the period.) RDBMSes such as Oracle and Postgres that support time zones will be able to take advantage of  identified time zone fields.

### 2015-09-29 icde2016/intro.tex
Move background contents
Put it in more useful locations.

@@
 While computers can provide high-speed and high-volume data processing, humans have domain knowledge and the ability to process data in parallel, notably by using our visual systems.
-Most importantly, humans provide the definition of what is valuable in an analysis.
+Humans also provide the definition of what is valuable in an analysis.
 Accordingly, human/computer analytic systems are essential to extracting knowledge of value to humans and their societies from the large amounts of data being generated today.
@@
 During the analytic process, a user may discover that parts of the data are not yet suitable for analysis.
-Solutions to this problem are often provided by data preparation tools external to the visual analysis environment, which requires the user to break their analytic flow, launch another tool and reprocess their data before returning to their analysis.
+Solutions to this problem are often provided by data preparation tools external to the visual analysis environment, which requires the user to break their cognitive flow, launch another tool and reprocess their data before returning to their analysis.
 If the user does not own this process (\eg it is the responsibility of another department), then there can be significant delays (including ``never.'') More subtly, the result of updated external processing may not be compatible with the user's existing work, which can lead to more time lost reconciling the new data model with the existing analysis.
@@
 histograms or scatter plots) and it is most natural for the user to ``clean what she sees'' instead of switching to a second tool.
-This leads to an ``adaptive'' process whereby users will prefer leveraging existing tools in the analytics environment (no matter how well suited to the task) over switching to another application.
+This leads to an ``adaptive'' process whereby users will prefer leveraging existing tools in the analytics environment (no matter how poorly they may be suited to the task) over switching to another application.
 Thus a well-designed interactive visual analysis environment will provide tools that enable users to perform such cleaning tasks as interactively as possible.
+
+\subsection{Scalar Dates}
+
+The SQL-99 standard defines three temporal scalar types: \texttt{DATE}, \texttt{TIMESTAMP} and \texttt{TIME}, which can be further qualified as either \texttt{WITH} or \texttt{WITHOUT} \texttt{TIME ZONE} (the \texttt{WITHOUT} form being more common.) They are typically implemented as fixed-point types containing an offset from some epoch (\eg Julian Days.) This makes them compact to store using column store compression techniques such as those in C-Store~\cite{Stonebraker:2005} and MonetDB/X100~\cite{Zukowski:2006}, and further allows some temporal operations to be implemented very efficiently using simple arithmetic.
+Thus from the RDBMS perspective, representing dates in scalar form provides  benefits for users, both in terms of available operations and query performance.
+
+Our visualization system models the first two of these SQL-99 types natively; the third (pure time) is folded into \texttt{TIMESTAMP} by appending it to a fixed date of \texttt{1899-12-30} (a convention derived from Microsoft Excel).
+From the analytic perspective, date types are dimensional (\ie independent variables) and can be used as either \textit{categorical} (simply ordered) or \textit{quantitative} (ordered with a distance metric) fields.
+
+Categorical dates have a natural hierarchy associated with them generated by calendar \textit{binning}.
+The visualization in Figure \ref{fig:I1} shows an example of a bar chart employing binned categorical dates in a year/quarter hierarchy.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=\columnwidth]{figures/FigureI1}
+\caption{Categorical Date Scalars.}
+\label{fig:I1}
+\end{figure}
+
+Quantitative dates are typically used for time series on an axis that maps the underlying distance measure to display pixel distance.
+The quantitative analog to categorical binning is \textit{truncation} whereby all the dates in a bin are represented by the first date in the bin.
+Truncation accurately preserves the distance semantics while enabling roll-up to coarser levels of detail.
+The visualization in Figure \ref{fig:I2} shows the same sales data in a time series rolled up to the quarter level.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=\columnwidth]{figures/FigureI2}
+\caption{Quantitative Date Scalars.}
+\label{fig:I2}
+\end{figure}
+
+These types of visualizations are difficult to specify with simple categorical string representations of dates.
+Thus the parsing of strings to date scalars is an important form of data modeling for analytics.
 
@@
 Some are common, but others can be quite idiosyncratic.
-Table 1 shows a selection of unusual date formats found in our online data.
+Table 1 shows a selection of unusual date formats found in our corpus.
 The first example shows a time zone in the middle of the date and a year after the time; the second shows a leading unmatched bracket and a colon between the date and time components; the third shows confusion between the seconds' decimal point and the time part delimiter; the fourth shows a two digit year apostrophe on a four digit year and the fifth shows a dash separating the date and time components.
@@
 
-Several RDBMSes (\eg MySQL, Oracle and Postgres) provide row-level functions for formatted date parsing, but our analysis of user data showed a 15\% syntax error rate with these functions.
-And even with perfect formatting, the user still has to interrupt their flow to learn a formatting syntax -- which they will likely forget once their problem has been solved.
+Several RDBMSes (\eg MySQL, Oracle and Postgres) provide row-level functions for parsing formatted dates, but our analysis of user data showed a 15\% syntax error rate using these functions.
+And even with perfect syntax, the user still has to interrupt her flow to learn a formatting syntax -- which they will likely forget once this specific problem has been solved.
 
@@
 
-Our solution wa to develop two machine learning algorithms for automatically deriving the format string from the user's data with over 90\% parsing accuracy.
-This approach allows the user to simply specify "this column is a date" and the visualisation system can respond quickly and accurately enough to avoid interrupting the user's analytic flow.
+\subsection{Contributions}
+Our solution was to develop two machine learning algorithms for automatically deriving the format string from the user's data with over 95\% parsing accuracy.
+This approach allows the user to simply specify "this column is a date" and the visualisation system can respond quickly and accurately enough to avoid interrupting the user's cognitive flow.
 
-\subsection{Contributions}
 Specifically, our contributions include:
@@
 \item Descriptions of two algorithms for automating date format extraction, at least one of which is performant enough for interactive use;
-\item Demonstration of an effective approach for integrating data cleaning into analytic processes.
+\item Demonstration of an effective approach for integrating data cleaning into analytic workflows.
 \end{itemize}
@@
 \subsection{Organization}
-The rest of this paper is organized as follows: The next section introduces the parameters of the problem space.
+The rest of this paper is organized as follows: The next section provides background information for the problem space.
 The following two sections describe the two different algorithms, one using Minimum Descriptive Length and the other using Natural Language Processing.

### 2015-09-29 icde2016/testing.tex
Move background contents
Put it in more useful locations.

--- 2015-07-16 testing.tex
@@
-\subsection{Data}
+\subsection{Data Preparation}
 
-We scanned all files published to a free, online data analysis website.
+The training data and test data are taken from data sets published to a free, online data analysis website.
+The published data includes both the raw data and how it was used for analysis.
+The service limits data sets to a maximum of 100K rows stored in a database which supports an implementation of \texttt{DATEPARSE}.
+
 We collected the contents of columns with names containing any of the following strings: Date, month, created, dt (abbreviation for date), mes (month in Spanish), datum (date in German), fecha (date in Spanish), data (date in Portuguese), and the day character used in Chinese and Japanese.
@@
 Fields of any data type other than date or datetime were analyzed, including strings, integers, and floats.
-One file was created for each scanned column, containing the unique non-null values in the column.
-
-The resulting set of files were de-duplicated. This handles any duplicate files from repeated publishing events. Files over 1MB were manually reviewed and those that did not contain dates removed.
+One file containing the unique non-null values was created for each scanned column, and the column collation was stored in a second table.
 
@@
 
-\subsubsection{Training Data}
+\subsubsection{NULL Filtering}
+Each column was then reduced to a maximum sample set of 32. We excluded domain values that appeared to be representations of \texttt{NULL}:
+\begin{itemize}
+\item Values containing the substring \texttt{NULL}
+\item Values containing no digits and at most one non-whitespace character (\eg \texttt{" / / "})
+\item Values on a list of empirically determined common \texttt{NULL} values (\eg \texttt{0000-00-00}, \texttt{NaN})
+\end{itemize}
+Columns that had no remaining valid samples were discarded.
 
-Training used a subset of the online data uploaded through February 2014. There were 30,968 files in the resulting set.
+\subsubsection{Sampling}
+The remaining samples were hashed and sorted on the hash value, with the top 32 values being retained as the column's sample.
+We can increase the sample size if needed, but for dates, it appears to be an adequate number.
+In any case, the median number of non-null rows per domain in our test data is 50, so increasing the sample size would have little benefit.
 
-\subsubsection{Verification Data}
-Verification used another set of data from the same website, collected from May 2014 to April 2015.
-This set was collected similarly, but also added columns named time.
-There may be some duplication between the training and verification data.
-There were 31,546 files in the resulting verification set.
+\subsubsection{Numeric Timestamps}
+After NULL filtering, a common class of date representation (numeric timestamps) remained that was unsuited for parsing via our date format syntax.
+This included Unix epoch timestamps (expressed as second, millisecond or microsecond counts from \texttt{1970-01-01}) and Microsoft Excel timestamps (expressed fractional days since \texttt{1900-01-01}).
+A simple test to check that the column was numeric and inside a specific range of values representing recent dates allowed us to tag these columns to avoid analyzing them further (and incidentally to identify them for generating a simple date extraction calculation for the user.)
+
+\subsubsection{Training and Verification}
+
+The data was divided into two sets, one for training and one for validation. There were 30,968 files in the training set and 31,546 files in the resulting verification set.
 
@@
 
-The validation speed is the average time needed to parse a value, and provides an estimate of how fast an ICU-based implementation can process string values into scalars and works out to 620K values per core per second.
+The validation speed is the average time needed to parse a value, and provides an estimate of how fast the ICU implementation can process string values into scalars and works out to 620K values per core per second.\\
 

### 2015-09-30 icde2016/background.tex
Describe ICU

@@
-Before delving into detailed descriptions of the two algorithms, we describe the common elements of the problem they are intended to solve.
-These are the input data, the output language and the interpretation of the output language for partial dates.
+\subsection{The ICU Date Format Language}
+The choice of a date format syntax for generation is somewhat arbitrary, but for the purposes of exposition we will be using the ICU Date Format Language.
+Other syntaxes can be generated directly from the algorithms, or by translating the ICU syntax into
 
-\subsection{Input Data}
-The primary assumption of the algorithms is that we have a column of string values that the user has indicated are really date scalars.
-The strings may also be in a non-English language, so we also extract the string collation from the column metadata and provide it as an optional secondary input.
-
-\subsection{The ICU Date Format Language}
-The date format syntax we selected is the one provided by the ICU open-source project.
+The date format syntax we selected for generation is the one provided by the ICU open-source project.
 We chose it because we were already using ICU in the code base, we had access to the source code, and it provides localized date part data for a large number of languages.
 
-The syntax is documented at the ICU web site~\cite{ICU}. While fairly complete, it has a few limitations that we ran into when evaluating the algorithms:
+ICU's format syntax is typical of most syntax languages and provides a large number of codes, and a summary of it is reproduced in Table 2.
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{|p{0.2\linewidth}| p{0.6\linewidth}|}
+\hline
+\centering
+\textbf{ICU Code} & \textbf{Interpretation}\\ \hline
+
+\scriptsize{yy} & \scriptsize{year (96)}\\ \hline
+\scriptsize{yyyy} & \scriptsize{year (1996)}\\ \hline
+
+\scriptsize{QQ} & \scriptsize{quarter (02)}\\ \hline
+\scriptsize{QQQ} & \scriptsize{quarter (Q2)}\\ \hline
+\scriptsize{QQQQ} & \scriptsize{quarter (2nd quarter)}\\ \hline
+
+\scriptsize{MM} & \scriptsize{month in year (09)}\\ \hline
+\scriptsize{MMM} & \scriptsize{month in year (Sept)}\\ \hline
+\scriptsize{MMMM} & \scriptsize{month in year (September)}\\ \hline
+
+\scriptsize{d} & \scriptsize{day in month (2)}\\ \hline
+\scriptsize{dd} & \scriptsize{day in month (02)}\\ \hline
+
+\scriptsize{EEE} & \scriptsize{day of week (Tues)}\\ \hline
+\scriptsize{EEEE} & \scriptsize{day of week (Tuesday)}\\ \hline
+
+\scriptsize{a} & \scriptsize{am/pm marker (pm)}\\ \hline
+
+\scriptsize{h} & \scriptsize{hour in am/pm 1:12 (7)}\\ \hline
+\scriptsize{hh} & \scriptsize{hour in am/pm 1:12 (07)}\\ \hline
+
+\scriptsize{H} & \scriptsize{hour in day 0:23 (0)}\\ \hline
+\scriptsize{HH} & \scriptsize{hour in day 0:23 (00)}\\ \hline
+
+\scriptsize{m} & \scriptsize{minute in hour (4)}\\ \hline
+\scriptsize{mm} & \scriptsize{minute in hour (04)}\\ \hline
+
+\scriptsize{s} & \scriptsize{second in minute (5)}\\ \hline
+\scriptsize{ss} & \scriptsize{second in minute (05)}\\ \hline
+
+\scriptsize{S} & \scriptsize{millisecond (2)}\\ \hline
+\scriptsize{SS} & \scriptsize{millisecond (23)}\\ \hline
+\scriptsize{SSS} & \scriptsize{millisecond (235)}\\ \hline
+
+\scriptsize{Z} & \scriptsize{Time Zone: RFC 822 (-0800)}\\ \hline
+\scriptsize{ZZ} & \scriptsize{Time Zone: RFC 822 (-0800)}\\ \hline
+\scriptsize{ZZZ} & \scriptsize{Time Zone: RFC 822 (-0800)}\\ \hline
+\scriptsize{ZZZZ} & \scriptsize{Time Zone: localized GMT (GMT-08:00)}\\ \hline
+\scriptsize{ZZZZZ} & \scriptsize{Time Zone: ISO8601 (-08:00)}\\ \hline
+
+\scriptsize{'} & \scriptsize{escape for text (nothing)}\\ \hline
+\scriptsize{''} & \scriptsize{two single quotes produce one (')}\\ \hline
+
+\end{tabular}
+\label{tab:icuformats} \\
+\caption{ICU Format Codes.}
+\end{table}
+
+The complete syntax is documented at the ICU web site~\cite{ICU}.
+
+While fairly extensive, the ICU syntax has a few limitations that we ran into when working with real-world data:
 \begin{itemize}
-\item No support for 4 letter year abbreviations (\eg \texttt{Sept.})
+\item No support for 4 letter month abbreviations (\eg \texttt{Sept.})
 \item No support for ordinal days (\eg \texttt{July 4th})
@@
 These limitations did not affect the results significantly, and in the future we hope to submit ICU extensions to handle some of these issues.
+Moreover, all the languages we examined had their own limitations, so ICU is not unusual in this respect.
 
-One other quirk of the ICU syntax may be a contributing factor to the user confusion around writing correct ICU date formats.
-The use of lexicographical case in the meta-symbols of the format language can be confusing (\eg \texttt{y} is used for years, but \texttt{M} is used for months while \texttt{m} is used for minutes.) Another advantage of an automated algorithm is that it hides such problems from most users, significantly improving the usability of the function.
-
-\subsection{Partial Dates}
-Many of the date formats that we encountered were incomplete dates, which necessitated creating rules for what date scalar they represented.
-
-ICU's date parsing APIs allow the specification of default values for parts when a format does not contain them.
-In our implementation, all time fields are set to \texttt{0} (midnight) and the date fields are set based on whether the format contains any date part specifications.
-When date parts are present, we use \texttt{2000-01-01} as the set of default date parts as it is the start of a leap year.
-When dealing with pure time formats, we model the output as a date/time and use \texttt{1899-12-30} for the date parts.
-
-ICU will also parse Time Zones (which the algorithms recognize but the visualization system does not model) and Quarters (which we interpret as the first month of the period.) RDBMSes such as Oracle and Postgres that support time zones will be able to take advantage of  identified time zone fields.
+One other quirk of the ICU syntax may be a contributing factor to the user confusion around writing correct ICU date formats in the existing visualisation system.
+The use of lexicographical case in the meta-symbols of the format language can be confusing (\eg \texttt{y} is used for years, but \texttt{M} is used for months while \texttt{m} is used for minutes.
+A better design would have made case more consistent within dates and times.
+Another advantage of automated syntax generation is that it hides such problems from most users, significantly improving the usability of the function.

### 2015-09-30 icde2016/paper.tex
Move background contents
Put it in more useful locations.
Describe ICU

@@
 In this paper, we present two algorithms for automatically deriving these date formats from a column of data with minimal user disruption.
-Both have accuracies of over 90\% on a large corpus of date columns extracted from an online data repository and are in substantial agreement with each other.
+Both have accuracies of over 95\% on a large corpus of date columns extracted from an online data repository and are in substantial agreement with each other.
 One approach is also fast enough to produce results within the user's perceptual threshold, making it suitable for interactive work.
@@
 \section{Background}
-\input{parameters}\\\\
+\input{background}
 
@@
 
-\section{Testing}
+\section{Experiments}
 \input{testing}

### 2015-09-30 icde2016/testing.tex
Describe ICU

@@
 The published data includes both the raw data and how it was used for analysis.
-The service limits data sets to a maximum of 100K rows stored in a database which supports an implementation of \texttt{DATEPARSE}.
+The service limits data sets to a maximum of 100K rows stored in a columnar database with collated strings.
 
@@
 Any column already converted by Excel or Jet was not included in this study.
+
+The data was divided into two sets, one for training and one for validation. There were 30,968 files in the training set and 31,546 files in the resulting verification set.
 
@@
 
-\subsubsection{Training and Verification}
+\subsubsection{Partial Dates}
+Many of the date formats that we encountered were incomplete dates, which necessitated creating rules for what date scalar they represented.
 
-The data was divided into two sets, one for training and one for validation. There were 30,968 files in the training set and 31,546 files in the resulting verification set.
+ICU's date parsing APIs allow the specification of default values for parts when a format does not contain them.
+In our implementation, all time fields are set to \texttt{0} (midnight) and the date fields are set based on whether the format contains any date part specifications.
+When date parts are present, we use \texttt{2000-01-01} as the set of default date parts as it is the start of a leap year.
+When dealing with pure time formats, we model the output as a date/time and use \texttt{1899-12-30} for the date parts.
+
+ICU will also parse Time Zones (which the algorithms recognize but the visualization system does not model) and Quarters (which we interpret as the first month of the period.) RDBMSes such as Oracle and Postgres that support time zones will be able to take advantage of  identified time zone fields.
 

### 2015-10-01 icde2016/background.tex
More intro cleanup

@@
 \subsection{The ICU Date Format Language}
-The choice of a date format syntax for generation is somewhat arbitrary, but for the purposes of exposition we will be using the ICU Date Format Language.
-Other syntaxes can be generated directly from the algorithms, or by translating the ICU syntax into
 
-The date format syntax we selected for generation is the one provided by the ICU open-source project.
-We chose it because we were already using ICU in the code base, we had access to the source code, and it provides localized date part data for a large number of languages.
-
-ICU's format syntax is typical of most syntax languages and provides a large number of codes, and a summary of it is reproduced in Table 2.
+The choice of a date format syntax for generation is somewhat arbitrary, but for the purposes of exposition we will be using the formatting language defined by the ICU open-source project.
+We chose it because we were already using ICU in our system, we had access to the source code, and it provides localized date part data for a large number of languages (\eg month names).
+ICU's format syntax is typical of most syntax languages and provides a large number of date part codes, and a summary of it is reproduced in Table 2.
+The complete syntax is documented at the ICU web site~\cite{ICU}.
+Other syntaxes can either be generated directly from the algorithms, or by translating the ICU syntax into into the desired syntax.
+(In fact, this latter approach is how our visualization system internally generates formats for RDBMSes that have non-ICU syntax.)
 
@@
 \centering
-\begin{tabular}{|p{0.2\linewidth}| p{0.6\linewidth}|}
+\begin{tabular}{|p{0.15\linewidth}| p{0.5\linewidth}|}
 \hline
-\centering
 \textbf{ICU Code} & \textbf{Interpretation}\\ \hline
@@
 
-\scriptsize{d} & \scriptsize{day in month (2)}\\ \hline
 \scriptsize{dd} & \scriptsize{day in month (02)}\\ \hline
@@
 
-\scriptsize{h} & \scriptsize{hour in am/pm 1:12 (7)}\\ \hline
 \scriptsize{hh} & \scriptsize{hour in am/pm 1:12 (07)}\\ \hline
-
-\scriptsize{H} & \scriptsize{hour in day 0:23 (0)}\\ \hline
 \scriptsize{HH} & \scriptsize{hour in day 0:23 (00)}\\ \hline
-
-\scriptsize{m} & \scriptsize{minute in hour (4)}\\ \hline
 \scriptsize{mm} & \scriptsize{minute in hour (04)}\\ \hline
-
-\scriptsize{s} & \scriptsize{second in minute (5)}\\ \hline
 \scriptsize{ss} & \scriptsize{second in minute (05)}\\ \hline
@@
 
+\scriptsize{zzz} & \scriptsize{Time Zone: specific non-location (PDT)}\\ \hline
+
 \scriptsize{Z} & \scriptsize{Time Zone: RFC 822 (-0800)}\\ \hline
-\scriptsize{ZZ} & \scriptsize{Time Zone: RFC 822 (-0800)}\\ \hline
-\scriptsize{ZZZ} & \scriptsize{Time Zone: RFC 822 (-0800)}\\ \hline
 \scriptsize{ZZZZ} & \scriptsize{Time Zone: localized GMT (GMT-08:00)}\\ \hline
@@
 
-The complete syntax is documented at the ICU web site~\cite{ICU}.
+While fairly extensive, the ICU syntax has a few limitations that we ran into when working with real-world data:
 
-While fairly extensive, the ICU syntax has a few limitations that we ran into when working with real-world data:
 \begin{itemize}
@@
 
-These limitations did not affect the results significantly, and in the future we hope to submit ICU extensions to handle some of these issues.
+These limitations did not affect the results significantly, and in the future we hope to submit ICU extensions to handle some of these formats.
 Moreover, all the languages we examined had their own limitations, so ICU is not unusual in this respect.
 
-One other quirk of the ICU syntax may be a contributing factor to the user confusion around writing correct ICU date formats in the existing visualisation system.
+One other quirk of the ICU syntax may be a contributing factor to the observed user difficulties writing correct ICU date formats in our visualization system.
 The use of lexicographical case in the meta-symbols of the format language can be confusing (\eg \texttt{y} is used for years, but \texttt{M} is used for months while \texttt{m} is used for minutes.

### 2015-10-01 icde2016/intro.tex
More intro cleanup

@@
 In recent years, there has been a growth of interest in data visualization technologies for human-assisted data analysis using systems such as Polaris~\cite{Stolte:2008} and Spotfire~\cite{Ahlberg:1996}.
-While computers can provide high-speed and high-volume data processing, humans have domain knowledge and the ability to process data in parallel, notably by using our visual systems.
+While computers can provide high-speed and high-volume data processing, humans have both domain knowledge and the ability to process data in parallel by using our visual systems.
 Humans also provide the definition of what is valuable in an analysis.
-Accordingly, human/computer analytic systems are essential to extracting knowledge of value to humans and their societies from the large amounts of data being generated today.
+Accordingly, hybrid human/computer analytic systems are essential to extracting knowledge of value to humans and their societies from the large amounts of data being generated today.
 
@@
 In recent years, a number of such systems have been developed, both by the academic community and by the commercial sector.
-Exploration of data consists not only in creating visual displays, but also in creating and modifying domain-specific computations.
-Consequently, most data visualization systems include facilities for defining such calculations as part of the data model being analyzed.
+Exploration of data consists not only in creating visual displays, but also in creating and modifying domain-specific computations in the data model, and most data visualization systems include facilities for defining such calculations.
 The most effective systems allow users to define these calculations as part of the analytic interaction, which permits the user to stay in the flow of analysis~\cite{Morton:2012}.
@@
 histograms or scatter plots) and it is most natural for the user to ``clean what she sees'' instead of switching to a second tool.
-This leads to an ``adaptive'' process whereby users will prefer leveraging existing tools in the analytics environment (no matter how poorly they may be suited to the task) over switching to another application.
-Thus a well-designed interactive visual analysis environment will provide tools that enable users to perform such cleaning tasks as interactively as possible.
+This leads to an ``adaptive'' process whereby users will prefer leveraging existing tools in the analytics environment over switching to another application -- no matter how poorly these tools may be suited to the task.
+Thus a well-designed interactive visual analysis environment will provide effective tools that enable users to perform such cleaning and preparation tasks as interactively as possible.
 
@@
 
+One of the most common data preparation tasks we see users performing is the parsing of date strings into scalar date representations, and streamlining this process is the focus of this paper.
+We now provide some background on the use of dates in data analytics to provide context for the problem.
+
 The SQL-99 standard defines three temporal scalar types: \texttt{DATE}, \texttt{TIMESTAMP} and \texttt{TIME}, which can be further qualified as either \texttt{WITH} or \texttt{WITHOUT} \texttt{TIME ZONE} (the \texttt{WITHOUT} form being more common.) They are typically implemented as fixed-point types containing an offset from some epoch (\eg Julian Days.) This makes them compact to store using column store compression techniques such as those in C-Store~\cite{Stonebraker:2005} and MonetDB/X100~\cite{Zukowski:2006}, and further allows some temporal operations to be implemented very efficiently using simple arithmetic.
-Thus from the RDBMS perspective, representing dates in scalar form provides  benefits for users, both in terms of available operations and query performance.
+Thus from the RDBMS perspective, representing dates in scalar form provides  benefits for users, both in terms of analytic operations and query performance.
 
@@
 Truncation accurately preserves the distance semantics while enabling roll-up to coarser levels of detail.
-The visualization in Figure \ref{fig:I2} shows the same sales data in a time series rolled up to the quarter level.
+The visualization in Figure \ref{fig:I2} shows the same sales data in a quantitative time series rolled up to the quarter level.
 
@@
 These types of visualizations are difficult to specify with simple categorical string representations of dates.
-Thus the parsing of strings to date scalars is an important form of data modeling for analytics.
+Thus the parsing of strings to date scalars is an important form of preparation for data analytics.
 
@@
 Some are common, but others can be quite idiosyncratic.
-Table 1 shows a selection of unusual date formats found in our corpus.
+Table 1 shows a selection of unusual date formats found in our corpus (the meanings of the ICU formatting codes can be found in Table 2).
 The first example shows a time zone in the middle of the date and a year after the time; the second shows a leading unmatched bracket and a colon between the date and time components; the third shows confusion between the seconds' decimal point and the time part delimiter; the fourth shows a two digit year apostrophe on a four digit year and the fifth shows a dash separating the date and time components.
@@
 Our solution was to develop two machine learning algorithms for automatically deriving the format string from the user's data with over 95\% parsing accuracy.
-This approach allows the user to simply specify "this column is a date" and the visualisation system can respond quickly and accurately enough to avoid interrupting the user's cognitive flow.
+This approach allows the user to simply specify "this column is a date" and the visualization system can respond quickly and accurately enough to avoid interrupting the user's cognitive flow.
 
@@
 \begin{itemize}
-\item Quantification of the surprising variety of date formats in use;
-\item Descriptions of two algorithms for automating date format extraction, at least one of which is performant enough for interactive use;
+\item A qualitative description of the surprising variety of date formats in use;
+\item Two algorithms for automating date format extraction, at least one of which is performant enough for interactive use;
+\item Several performance enhancements for Minimum Descriptive Length structure pruning;
 \item Demonstration of an effective approach for integrating data cleaning into analytic workflows.

### 2015-10-08 icde2016/futurework.tex
Contributions & Related Work

@@
-While we are satisfied with the results so far, we found a number of columns that we were not able to process with these techniques.
+During testing, we found a number of columns that we were not able to process with these techniques.
 Many columns contained multiple date formats.
-We would like to recognize this situation and generate predicate-based calculations to increase the accuracy of the results and thereby make the experience even more seamless for the user.
+We would like to recognize this situation and generate predicate-based calculations (\eg if the string matches format 1 then apply format 1 else ...) to increase the accuracy of the results and thereby make the experience even more seamless for the user.
 Other columns contain date ranges, which we would like to handle by generating multiple calculations, possibly by combining regular expressions with date parsing.
 
-In this paper, we have only considered the parsing of strings, but dates are often formatted as integers (\textit{e.g.} \texttt{201507016}).
-It is significantly faster to decompose integers into date parts using arithmetic operations (\textit{e.g.} \texttt{mod} and \texttt{div}) than by using slower, locale-sensitive string parsing functions.
+In this paper, we have only considered the parsing of strings, but dates are often formatted as integers (\eg \texttt{201507016}).
+It is significantly faster to decompose integers into date parts using arithmetic operations (\eg \texttt{mod} and \texttt{div}) than by using slower, locale-sensitive string parsing functions.
 Moreover, the number of possible formats is low enough that we may be able to enumerate them.
@@
 
-In the course of our research, we have also identified a number of date part variants (\textit{e.g.} ordinal dates, four-letter month abbreviations, alternate meridian markers and postfix quarter syntax) that we would like to commend to the ICU project, along with possible implementations.
+In the course of our research, we have also identified a number of date part variants (\eg ordinal dates, four-letter month abbreviations, alternate meridian markers and postfix quarter syntax) that we would like to commend to the ICU project, along with possible implementations.

### 2015-10-08 icde2016/intro.tex
Contributions & Related Work

@@
 \scriptsize{dd-MMM-yy hh.mm.ss.SSSSSS a} & \scriptsize{01-OCT-13 01.09.00.000000 PM}\\ \hline
-\scriptsize{MM 'yyyy} & \scriptsize{01 '2013}\\ \hline
+\scriptsize{MM ''yyyy} & \scriptsize{01 '2013}\\ \hline
 \scriptsize{MM/dd/yyyy - HH:mm} & \scriptsize{04/09/2014 - 23:47}\\ \hline
@@
 
-\subsection{Contributions}
 Our solution was to develop two machine learning algorithms for automatically deriving the format string from the user's data with over 95\% parsing accuracy.
 This approach allows the user to simply specify "this column is a date" and the visualization system can respond quickly and accurately enough to avoid interrupting the user's cognitive flow.
+
+\subsection{Contributions}
 
@@
 \begin{itemize}
-\item A qualitative description of the surprising variety of date formats in use;
-\item Two algorithms for automating date format extraction, at least one of which is performant enough for interactive use;
-\item Several performance enhancements for Minimum Descriptive Length structure pruning;
-\item Demonstration of an effective approach for integrating data cleaning into analytic workflows.
+\item By analyzing an on line corpus, we provide evidence that practical date parsing requires the ability to recognize hundreds of formats;
+\item We show how to extend prior work on Minimum Descriptive Length structure extraction to generate a freely available date format domain language with over 95\% accuracy;
+\item We describe a second Natural Language Processing technique for generating the same date format domain language with similar accuracy.
+We describe how the basic algorithm is extended to support grammar variants and constraints unique to date formats.
+The parsing algorithm is extended to compute an overall dominant pattern over a data column;
+\item We describe some limitations of this domain language that would improve its utility.
 \end{itemize}
@@
 The following two sections describe the two different algorithms, one using Minimum Descriptive Length and the other using Natural Language Processing.
-In section 6, we evaluate the algorithms on a corpus of 30K columns, both by sampling the outputs manually and then by using the algorithms to validate each other.
-We then discuss future work in section 7 and conclude in section 8.
+In section V, we evaluate the algorithms on a corpus of 30K columns, both by sampling the outputs manually and then by using the algorithms to validate each other.
+We then discuss related work in section VI, future work in section VII and conclude in section VIII.

### 2015-10-14 icde2016/nlp.tex
Fixed a typo in nlp section, added opening paragraph to testing section.
Removed NLP 101 text.

--- 2015-07-16 nlp.tex
@@
-The main characteristic of formal grammar is the presence of production rules that are used to rewrite an expression, and a starting symbol which is the origin of all forms of a given expression.
-In the production rules, we find non-terminals and terminals, with parsed expressions containing only terminals.
-A formal grammar is considered ``context free'' when its production rules can be applied regardless of the context of a nonterminal.
+\subsection{Context-Free Grammar}
+ICU date formats are well defined both structurally and semantically, and can be defined by a context-free grammar (CFG).
 This has several advantages.
 First, it allows for modular syntax definitions, which simplifies grammar development and enables reuse.
-Second, it grants total freedom
-in structuring a grammar to best fit its intended use.
+Second, it grants total freedom in structuring a grammar to best fit its intended use.
 The power of such a grammar is its expressiveness, and is not constrained by pattern rigidities that are prevalent with regular expressions and text processing~\cite{Grune:1990}.
 
-\subsection{Context-Free Grammar}
-ICU date formats are well defined both structurally and semantically, and can be defined by a context-free grammar (CFG).
 A CFG consists of a set of non-terminal symbols $X$, a set of terminal symbols $\beta$, a start non-terminal symbol $S \in X$ from which the grammar generates the strings, and a set of production rules $\tau$, with each rule of the form $X \rightarrow \beta$~\cite{Hopcroft:1990}.
 
-While there are several functionally equivalent notations for representing CFG, we adopt the \textit{Backus-Naur Form} (BNF) for defining the grammar rules for \texttt{DATEPARSE} formats.
+We adopt the \textit{Backus-Naur Form} (BNF) for defining the grammar rules for \texttt{DATEPARSE} formats.
 In particular, we use the \textit{Extended Backus-Naur Form} (EBNF) as the notation is more compact and readable for frequently used constructions~\cite{Grune:1990}.
@@
 For example, there is no  \texttt{`November 31, 2015'}, \texttt{`February 29, 2013'}, or \texttt{`Sunday, May  5, 1965'}.
-November only as 30 days in any year; 2013 was not a leap year; and May 5, 1965 was a Wednesday.
+November only has 30 days in any year; 2013 was not a leap year; and May 5, 1965 was a Wednesday.
 

### 2015-10-14 icde2016/nlp.tex
Added text to justify why initial probability weights are assigned.

@@
 
-In order to bootstrap the probability weights in the \texttt{PCFG}, we manually set initial probability weights to various rules in the grammar as follows:
+Assigning initial probability weights to the grammar for computing the parse with the highest weight, tends to be a more scalable and flexible approach to disambiguation rather than building a deep knowledge ontology.
+We bootstrap the probability weights in the \texttt{PCFG}, by setting initial probability weights to various rules in the grammar as follows:
+
 \begin{itemize}

### 2015-10-15 icde2016/paper.tex
Use IEEE author syntax
ICDE fit and finish
- No keyword section
- Single-blind submissions
- Bibliography break
Comments in abstract

@@
 
-%
-\def\sharedaffiliation{%
-\end{tabular}
-\begin{tabular}{c}}
-%
 \def\eg{\textit{e.g.} }
@@
 \newcommand{\vidya}[1]{\textcolor{red}{(vidya: #1)}}
+\newcommand{\richard}[1]{\textcolor{green}{(richard: #1)}}
+\newcommand{\dan}[1]{\textcolor{blue}{(dan: #1)}}
 
@@
 Automating the Extraction of Date Scalars}
-%\input{anonymous}
+\input{authors}
 % author names and affiliations
@@
 \begin{abstract}
-Interactive visual analytics is an important set of tools for the timely analysis of big data.
+Interactive visual analytics is an \vidya{visual analytics are not tools by themselves.
+Reword to something like - ...is an efficient way to...}important set of tools for the timely analysis of big data.
 Users who are already engaged in interactive data analysis prefer to stay in the visualization environment for unavoidable cleaning and preparation tasks in order to preserve their analytic flow.
@@
 
-In this paper, we present two algorithms for automatically deriving these date formats from a column of data with minimal user disruption.
+In this paper, we present two algorithms for automatically deriving these date formats from a column of data with minimal user disruption \vidya{briefly mention what the two algorithms are}.
 Both have accuracies of over 95\% on a large corpus of date columns extracted from an online data repository and are in substantial agreement with each other.
@@
 
-\vidya{Need to figure out the IEEE equivalent for categories and terms}
+%Looks like they don't use them.
+%\vidya{Need to figure out the IEEE equivalent for categories and terms}
 %D.3.3 [Data Processing]: Data Cleaning, Natural Language Processing.
@@
 \section{Acknowledgments}
-\input{credits}
+\input{acknowledgements}
 
@@
 % produce the bibliography for the citations in your paper.
+\IEEEtriggeratref{13}
 \bibliographystyle{IEEEtran}

### 2015-10-15 icde2016/intro.tex
Comments to intro

@@
 In recent years, there has been a growth of interest in data visualization technologies for human-assisted data analysis using systems such as Polaris~\cite{Stolte:2008} and Spotfire~\cite{Ahlberg:1996}.
-While computers can provide high-speed and high-volume data processing, humans have both domain knowledge and the ability to process data in parallel by using our visual systems.
+While computers can provide high-speed and high-volume data processing, humans have both domain knowledge and the ability to process data in parallel by using our visual systems \vidya{need some references - perhaps Bill Cleveland's work or Jock's book}.
 Humans also provide the definition of what is valuable in an analysis.
-Accordingly, hybrid human/computer analytic systems are essential to extracting knowledge of value to humans and their societies from the large amounts of data being generated today.
+\vidya{Don`t really understand the point of this sentence:}Accordingly, hybrid human/computer analytic systems are essential to extracting knowledge of value to humans and their societies from the large amounts of data being generated today.
 
@@
 
-One of the most common data preparation tasks we see users performing is the parsing of date strings into scalar date representations, and streamlining this process is the focus of this paper.
+One of the most common data preparation tasks we see users performing is the parsing of date strings into scalar date representations, and streamlining this process is the focus of this paper \vidya{do we have any data to backup this?}.
 We now provide some background on the use of dates in data analytics to provide context for the problem.
@@
 
-Our solution was to develop two machine learning algorithms for automatically deriving the format string from the user's data with over 95\% parsing accuracy.
-This approach allows the user to simply specify "this column is a date" and the visualization system can respond quickly and accurately enough to avoid interrupting the user's cognitive flow.
+Our solution was to develop two machine learning algorithms for automatically deriving the format string from the user's data with over 95\% parsing accuracy \vidya{we need to mention why we chose to develop two separate algorithms.
+One reason being for cross-validation over a large set of output results from each}.
+This approach allows the user to simply specify ``this column is a date" and the visualization system can respond quickly and accurately enough to avoid interrupting the user's cognitive flow.
 
@@
 Specifically, our contributions include:
-
+\vidya{also good to mention that we are able to parse many different formats and provide a unique approach to cross-validate the the parsing accuracy by developing two independent and different algorithms}
 \begin{itemize}
-\item By analyzing an on line corpus, we provide evidence that practical date parsing requires the ability to recognize hundreds of formats;
+\item By analyzing an online corpus, we provide evidence that practical date parsing requires the ability to recognize hundreds of formats;
 \item We show how to extend prior work on Minimum Descriptive Length structure extraction to generate a freely available date format domain language with over 95\% accuracy;

### 2015-10-15 icde2016/background.tex
Improved the table format
1. Increased vertical padding in each table cell
2. Added space between the table and its caption.

@@
 The choice of a date format syntax for generation is somewhat arbitrary, but for the purposes of exposition we will be using the formatting language defined by the ICU open-source project.
-We chose it because we were already using ICU in our system, we had access to the source code, and it provides localized date part data for a large number of languages (\eg month names).
+We chose this format because we were already using ICU in our system, we had access to the source code, and it provides localized date part data for a large number of languages (\eg month names).
 ICU's format syntax is typical of most syntax languages and provides a large number of date part codes, and a summary of it is reproduced in Table 2.
@@
 \centering
+\bgroup
+\def\arraystretch{1.5}
 \begin{tabular}{|p{0.15\linewidth}| p{0.5\linewidth}|}
@@
 \end{tabular}
+\egroup
 \label{tab:icuformats} \\

### 2015-10-15 icde2016/mdl.tex
moved paper draft into separate subfolders
Small edits

--- 2015-07-16 mdl.tex
@@
 Since the ICU parser is flexible about parsing single or double digit formats, we use double-digit formats, but accept one or two digits.
-One important exception to this rule is for years, which are fixed width fields (2 or 4).
+One important exception to this rule is for years, which are fixed width fields ($2$ or $4$).
 
@@
 \begin{itemize}
-\item Since we have a set of samples, we can apply the candidate format to the strings to see how well it performs. Formats with fewer parse errors are preferred.
+\item Because we have a set of samples, we can apply the candidate format to the strings to see how well it performs. Formats with fewer parse errors are preferred.
 \item Date parts can be considered a place-value system, so we prefer ``more significant'' components (\eg month-day-year over hour-minute-second).

### 2015-10-15 icde2016/paper.tex
Improved the table format
1. Increased vertical padding in each table cell
2. Added space between the table and its caption.
Incorporate Vidya's comments

@@
 \usepackage{color}
+\usepackage{caption} 
+\captionsetup[table]{skip=10pt}
 
@@
 \begin{abstract}
-Interactive visual analytics is an \vidya{visual analytics are not tools by themselves.
-Reword to something like - ...is an efficient way to...}important set of tools for the timely analysis of big data.
+Interactive visual analytics is an effective approach for the timely analysis of big data.
 Users who are already engaged in interactive data analysis prefer to stay in the visualization environment for unavoidable cleaning and preparation tasks in order to preserve their analytic flow.
@@
 
-In this paper, we present two algorithms for automatically deriving these date formats from a column of data with minimal user disruption \vidya{briefly mention what the two algorithms are}.
+In this paper, we present two algorithms for automatically deriving these date formats from a column of data with minimal user disruption, one based on minimal entropy representations and another based on natural language processing techniques.
 Both have accuracies of over 95\% on a large corpus of date columns extracted from an online data repository and are in substantial agreement with each other.

### 2015-10-15 icde2016/intro.tex
Improved the table format
1. Increased vertical padding in each table cell
2. Added space between the table and its caption.
Cross-validation

--- 2015-10-08 icde2016/intro.tex
@@
 In recent years, there has been a growth of interest in data visualization technologies for human-assisted data analysis using systems such as Polaris~\cite{Stolte:2008} and Spotfire~\cite{Ahlberg:1996}.
-While computers can provide high-speed and high-volume data processing, humans have both domain knowledge and the ability to process data in parallel by using our visual systems.
+While computers can provide high-speed and high-volume data processing, humans have both the domain knowledge and the ability to process data in parallel by using their visual systems \cite{Cleveland.McGill1984,Mackinlay:1986}.
 Humans also provide the definition of what is valuable in an analysis.
-Accordingly, hybrid human/computer analytic systems are essential to extracting knowledge of value to humans and their societies from the large amounts of data being generated today.
+Algorithms that rely on \emph{both} the processing capability of computer systems and human feedback, are more effective in extracting useful knowledge from the large amounts of data being generated.
+\vidya{still struggling to understand how interactivity is relevant to what we're doing here...}
 
@@
 
-One of the most common data preparation tasks we see users performing is the parsing of date strings into scalar date representations, and streamlining this process is the focus of this paper.
+One of the most common data preparation tasks we see users performing is the parsing of date strings into scalar date representations, and streamlining this process is the focus of this paper \vidya{do we have any data to backup this?}\richard{Yes, I think Dan has some}.
 We now provide some background on the use of dates in data analytics to provide context for the problem.
@@
 \centering
-%\begin{tabular}{ |l|l|l| }
+\bgroup
+\def\arraystretch{1.5}
 \begin{tabular}{|p{0.4\linewidth}| p{0.4\linewidth}|}
@@
 \end{tabular}
+\egroup
 \label{tab:dateformats}
@@
 Our solution was to develop two machine learning algorithms for automatically deriving the format string from the user's data with over 95\% parsing accuracy.
-This approach allows the user to simply specify "this column is a date" and the visualization system can respond quickly and accurately enough to avoid interrupting the user's cognitive flow.
+We developed two algorithms because we were not aware of any previous work on this problem and we needed to be able to cross-validate our results on a large testing corpus.
+Either approach allows the user to simply specify ``this column is a date" and the visualization system can respond quickly and accurately enough to avoid interrupting the user's cognitive flow.
 
@@
 Specifically, our contributions include:
-
 \begin{itemize}
-\item By analyzing an on line corpus, we provide evidence that practical date parsing requires the ability to recognize hundreds of formats;
+\item By analyzing an online corpus, we provide evidence that practical date parsing requires the ability to recognize hundreds of formats;
 \item We show how to extend prior work on Minimum Descriptive Length structure extraction to generate a freely available date format domain language with over 95\% accuracy;
@@
 The parsing algorithm is extended to compute an overall dominant pattern over a data column;
+\item We provide evidence that the development of multiple, independent parsing algorithms provides an effective means of cross-validation on large corpi;
 \item We describe some limitations of this domain language that would improve its utility.

### 2015-10-15 icde2016/intro.tex
reworded the last sentence

--- 2015-10-15 icde2016/intro.tex
@@
 In recent years, there has been a growth of interest in data visualization technologies for human-assisted data analysis using systems such as Polaris~\cite{Stolte:2008} and Spotfire~\cite{Ahlberg:1996}.
-While computers can provide high-speed and high-volume data processing, humans have both domain knowledge and the ability to process data in parallel by using our visual systems \vidya{need some references - perhaps Bill Cleveland's work or Jock's book}.
+While computers can provide high-speed and high-volume data processing, humans have both the domain knowledge and the ability to process data in parallel by using their visual systems \vidya{need some references - perhaps Bill Cleveland's work or Jock's book}.
 Humans also provide the definition of what is valuable in an analysis.
-\vidya{Don`t really understand the point of this sentence:}Accordingly, hybrid human/computer analytic systems are essential to extracting knowledge of value to humans and their societies from the large amounts of data being generated today.
+Algorithms that rely on \emph{both} the processing capability of computer systems and human feedback, are more effective in extracting useful knowledge from the large amounts of data being generated.
+\vidya{still struggling to understand how interactivity is relevant to what we're doing here...}
 

### 2015-10-15 icde2016/intro.tex
Added references

@@
 In recent years, there has been a growth of interest in data visualization technologies for human-assisted data analysis using systems such as Polaris~\cite{Stolte:2008} and Spotfire~\cite{Ahlberg:1996}.
-While computers can provide high-speed and high-volume data processing, humans have both the domain knowledge and the ability to process data in parallel by using their visual systems \vidya{need some references - perhaps Bill Cleveland's work or Jock's book}.
+While computers can provide high-speed and high-volume data processing, humans have both the domain knowledge and the ability to process data in parallel by using their visual systems \cite{Cleveland.McGill1984,Mackinlay:1986}.
 Humans also provide the definition of what is valuable in an analysis.

### 2015-10-15 icde2016/background.tex
Explain DATEPARSE (briefly)

@@
 Another advantage of automated syntax generation is that it hides such problems from most users, significantly improving the usability of the function.
+
+\subsection{The \dateparse\ Function}
+
+The date format generated by the algorithms is presumed to be a string argument to a scalar function called \dateparse, which converts a string value to a date value using the format string.
+Several examples of this kind of function can be found in RDBMSes such as MySQL's \functionname{STR_TO_DATE}, Oracle's \functionname{TO_TIMESTAMP} and Postgres' \functionname{TO_TIMESTAMP}.
+There are also programming language library implementations such as Python's \functionname{strptime} and ICU's own \functionname{DateFormat::parse}.

### 2015-10-15 icde2016/intro.tex
Pronoun gender consistency

--- 2015-10-15 icde2016/intro.tex
@@
 Several RDBMSes (\eg MySQL, Oracle and Postgres) provide row-level functions for parsing formatted dates, but our analysis of user data showed a 15\% syntax error rate using these functions.
-And even with perfect syntax, the user still has to interrupt her flow to learn a formatting syntax -- which they will likely forget once this specific problem has been solved.
+And even with perfect syntax, the user still has to interrupt her flow to learn a formatting syntax -- which she will likely forget once this specific problem has been solved.
 

### 2015-10-15 icde2016/mdl.tex
Misplaced paragraph in list.
More comments to mdl section.

@@
 \subsection{Domains}
+\vidya{Explain what you mean by domain.}
 Potter's Wheel presents an algorithm for deriving a common structure for a set of strings by breaking each string down into a sequence of domains.
 These domains are described by an interface that includes:
+\vidya{can you provide brief examples for each of these functions so I can understand what they are?}
 \begin{itemize}
@@
 
-In our approach, we implement all of these functions, but with significant changes to the last one, which we will describe below.
+In our approach, we implement all of these functions, but with significant changes to the last one, which we will describe below.\vidya{can you briefly explain why we significantly change the last one here?}
 
@@
 \item Once these semantic preferences have been considered, we then prefer the more compact (MDL) representation.
+\end{itemize}
+
 The output of the algorithm is now an ordered list of formats and associated locales.
 These can then be used to drive a user interface that allows the user to choose between the possibilities or the top-ranking format can simply be used automatically.
-\end{itemize}

### 2015-10-15 icde2016/intro.tex
corpi -> corpora

@@
 The parsing algorithm is extended to compute an overall dominant pattern over a data column;
-\item We provide evidence that the development of multiple, independent parsing algorithms provides an effective means of cross-validation on large corpi;
+\item We provide evidence that the development of multiple, independent parsing algorithms provides an effective means of cross-validation on large corpora;
 \item We describe some limitations of this domain language that would improve its utility.

### 2015-10-15 icde2016/mdl.tex
Explain domains

--- 2015-10-15 icde2016/mdl.tex
@@
 \subsection{Domains}
-Potter's Wheel presents an algorithm for deriving a common structure for a set of strings by breaking each string down into a sequence of domains.
-These domains are described by an interface that includes:
+Potter's Wheel presents an algorithm for deriving a common structure for a set of strings by breaking each string down into a sequence of \textit{domains}.
+A domain is a set of strings, with a few optional statistical properties.
+They are defined by an interface that includes:
 \begin{itemize}
-\item A required inclusion function to test for membership in the domain (\texttt{match})
-\item An optional function to compute the number of values in the domain with a given length (\texttt{cardinality})
-\item An optional function to update statistics for the domain based on a given value (\texttt{updateStatistics})
-\item An optional function to prevent consideration of a domain that is redundant (\texttt{isRedundantAfter}).
+\item A required inclusion function to test for membership in the domain (\functionname{match}). 
+For example, the domain \functionname{<Digits>} would return true for the string \functionname{"123"} 
+and false for the string \functionname{"abc"}.
+\item An optional function to compute the number of values in the domain with a given length (\functionname{cardinality}).
+For example, the domain for three letter month names (\functionname{MMM}) would return $12$ for the length $3$ and $0$ otherwise.
+\item An optional function to update statistics for the domain based on a given value (\functionname{updateStatistics}).
+For example, the domain \functionname{MMM} might keep a histogram of how frequently each month was encountered.
+\item An optional function to prevent consideration of a domain that is redundant (\functionname{isRedundantAfter}).
+For example, the domain \functionname{<Digits>} would be redundant after itself 
+(\ie there is no difference between a single run of digits and two adjacent runs.)
 \end{itemize}
 
-In our approach, we implement all of these functions, but with significant changes to the last one, which we will describe below.
+In our approach, we implement all of these functions, but with significant changes to the last one to increase level of pruning applied during
+domain enumeration. These changes will be described below.
 

### 2015-10-15 icde2016/nlp.tex
Some edits including section numbering fixes.

@@
 
-The parsed output is then converted to the ICU Date Format Language described in Section 3.2 in order to effectively cross-validate results from the MDL algorithm.
+The parsed output is then converted to the ICU Date Format Language described in Section IIA in order to effectively cross-validate results from the MDL algorithm.
 While this grammar accounts for the use of meridian markers `a.m.' and `p.m.' for the 12-hour format, ICU does not support these tokens, and they are simply ignored.
@@
 \item Restriction on the distribution of $30$ and $31$:
-Months usually alternate between lengths of 30 and 31 days. We can use x mod 2 to get an alternating pattern of 1 and 0, then just add our constant base number of days:
+Months usually alternate between lengths of $30$ and $31$ days. We can use $x \Mod 2$ to get an alternating pattern of 1 and 0, then just add our constant base number of days:
 
@@
 
-Now if we substitute this expression in the x + 1 dividend of Equation $2$, we can invert the pattern using our mask:
+Now if we substitute this expression in the $x + 1$ dividend of Equation $2$, we can invert the pattern using our mask:
 
@@
 \begin{itemize}
-\item Similar to the MDL ranking properties described in Section 3.7, rules involving the non-terminal \texttt{DateGrammar} on the left-side of the rule, are assigned probability weights higher than those rules with the non-terminal being \texttt{TimeGrammar}.
+\item Similar to the MDL ranking properties described in Section IIIG, rules involving the non-terminal \texttt{DateGrammar} on the left-side of the rule, are assigned probability weights higher than those rules with the non-terminal being \texttt{TimeGrammar}.
 In practice, assigning $p(\texttt{DateGrammar} \rightarrow \beta_{1}) = 0.9$ and $p(\texttt{TimeGrammar} \rightarrow \beta_{2}) = 0.7$ leads to optimal ranking of the parse trees.
 
-\item For the non-terminal symbol \texttt{Day}, the constraint specified is a range between 01 and 28-31, depending on month/year.
+\item For the non-terminal symbol \texttt{Day}, the constraint specified is a range between $01$ and $28-31$, depending on month/year.
 However, for \texttt{dd} $> 12$, $p(\texttt{Day} \rightarrow d d) = 1.0$, and $0.5$ otherwise.

### 2015-10-15 icde2016/background.tex
Fei's latest comments

@@
 
-The choice of a date format syntax for generation is somewhat arbitrary, but for the purposes of exposition we will be using the formatting language defined by the ICU open-source project.
+The choice of a date format syntax for generation is somewhat arbitrary, but for the purposes of exposition we will be using the formatting language defined by the ICU open-source project~\cite{ICU}.
 We chose this format because we were already using ICU in our system, we had access to the source code, and it provides localized date part data for a large number of languages (\eg month names).
 ICU's format syntax is typical of most syntax languages and provides a large number of date part codes, and a summary of it is reproduced in Table 2.
-The complete syntax is documented at the ICU web site~\cite{ICU}.
+The complete syntax is documented at the ICU web site.
 Other syntaxes can either be generated directly from the algorithms, or by translating the ICU syntax into into the desired syntax.

### 2015-10-15 icde2016/mdl.tex
Fei's latest comments

@@
 Date domains typically have small widths, so we found it advantageous to provide the shortest and longest match sizes for use in structure enumeration and matching.
+For example, months have between $1$ and $2$ digits, so there is no need to test matches whose length is outside this range.
 
 Date domains are also often uniform in that adding more characters to a mismatch will not help.
-For example, a 2-digit day domain that does not match a 1-letter substring will not be able to generate a match by adding more characters.
+For example, a $2$-digit day domain that does not match a $1$-letter substring will not be able to generate a match by adding more characters.
 
@@
 In particular, they cannot appear adjacent to a two-digit field if the structure contains punctuation.
-(This can come up in some small early 21st century year domains where a two-digit year can masquerade as almost any numeric field.)
+(This can come up in some small early 21st century year domains where a two-digit year can masquerade as almost any numeric field \eg $08$ for $2008$.)
 \end{itemize}
@@
 \item Date parts can be considered a place-value system, so we prefer ``more significant'' components (\eg month-day-year over hour-minute-second).
-\item If two formats from different locales give the same results, prefer the original column locale. The sample set may have missed an example where this could be important.
+\item If two formats from different locales give the same results, prefer the original column locale.
+The sample set may have missed an example where this could be important.
+For example, the month name for "September" is the same in both German and English, but "October" and "Oktober" are different.
 \item If the format has an ambiguous date order (\eg all days are less than 12), then prefer the default date order of the locale.

### 2015-10-15 icde2016/intro.tex
Fei's latest comments
Dan's statistics, Vidya's interactivity, column balancing

@@
 Humans also provide the definition of what is valuable in an analysis.
-Algorithms that rely on \emph{both} the processing capability of computer systems and human feedback, are more effective in extracting useful knowledge from the large amounts of data being generated.
-\vidya{still struggling to understand how interactivity is relevant to what we're doing here...}
+Systems that rely on \emph{both} the processing capability of computer systems and human feedback, are more effective in extracting useful knowledge from the large amounts of data being generated than either by itself.
 
@@
 
-One of the most common data preparation tasks we see users performing is the parsing of date strings into scalar date representations, and streamlining this process is the focus of this paper \vidya{do we have any data to backup this?}\richard{Yes, I think Dan has some}.
+One of the most common data preparation tasks we see users performing is the parsing of date strings into scalar date representations, 
+and streamlining this process is the focus of this paper. An examination of the author calculations in an online repository of our
+visualization system shows that there are about as many workbooks containing date parsing calculations ($3.3\%$) as there are
+ integer type conversions of any kind ($3.4\%$).
 We now provide some background on the use of dates in data analytics to provide context for the problem.
@@
 The first example shows a time zone in the middle of the date and a year after the time; the second shows a leading unmatched bracket and a colon between the date and time components; the third shows confusion between the seconds' decimal point and the time part delimiter; the fourth shows a two digit year apostrophe on a four digit year and the fifth shows a dash separating the date and time components.
+As can be seen in Figure \ref{fig:M2}, the long tail of formats means that these idiosyncrasies are the rule, not the exception, 
+so a static enumeration of a small set of formats will not produce high reliability.
 

### 2015-10-16 icde2016/testing.tex
Fixed a typo in nlp section, added opening paragraph to testing section.
Fixed a section number typo and prettified table.
Addressed #4 and #11 of Fei's feedback.

@@
+%1 what are the questions to answer
+%2 setup (machine/code implementation/data) could be one setup per experiment or as a whole if they are identical
+%3. how to perform the experiments
+%4. what are the observed results
+%5. what are the conclusions(discussions) from the results.
+
+For each of the two algorithms described, we want to measure the fraction of candidate date columns that the algorithm is able to recognize.
+We describe a large training and experimental corpus that we collected, followed by the results of applying each algorithm to that data.
+
 \subsection{Data Preparation}
@@
 \centering
+\bgroup
+\def\arraystretch{1.5}
 \begin{tabular}{|p{0.48\linewidth}| p{0.24\linewidth}|}
@@
 \end{tabular}
+\egroup
 \label{tab:mdlstats}
@@
 One other file contained a set of integers that are not actually dates, and one algorithm picked \texttt{MMyyyy} format, while the other picked \texttt{HHmmss} format.
+
+\subsubsection{Production}
+
+Although both algorithms achieved comparable high accuracies in parsing date time strings, we utilized the minimum description length method in the production code of our system.
+This was primarily due to the fact that the MDL algorithm was written in C++, allowing for higher performance.
+We believe that NLP version if written in C++ as opposed to Python, would lead to better performance.

### 2015-10-16 icde2016/testing.tex
Added text motivating using public

@@
 The training data and test data are taken from data sets published to a free, online data analysis website.
+This website contains data sets that users analyzed and were interested in sharing with other people through the web.
+It may not be representative of data in other settings such as corporate data warehouses, but each data set is one that a user was willing to invest some time in analyzing.
 The published data includes both the raw data and how it was used for analysis.
@@
 For instance, typing the string \texttt{"12/31/1999"} into Microsoft Excel, is automatically interpreted as the date \texttt{1999-12-31}.
-The Microsoft Jet library used to read these text files detects a few date formats as well.
+The Microsoft Jet library \vidya{need a reference here} used to read these text files detects a few date formats as well.
 Any column already converted by Excel or Jet was not included in this study.
@@
 \subsubsection{Sampling}
-The remaining samples were hashed and sorted on the hash value, with the top 32 values being retained as the column's sample.
+The remaining samples were hashed and sorted on the hash value, with the top $32$ values being retained as the column's sample.
 We can increase the sample size if needed, but for dates, it appears to be an adequate number.
-In any case, the median number of non-null rows per domain in our test data is 50, so increasing the sample size would have little benefit.
+In any case, the median number of non-null rows per domain in our test data is $50$, so increasing the sample size would have little benefit.
 
@@
 \subsubsection{Natural Language Processing}
+
 We implemented the grammar, training and parsing in Python 3.4.1 using the Natural Language Toolkit libraries~\cite{nltk} on a quad core Dell T7600 running Windows 7.
@@
 
+Out of the $31,546$ files used for testing, the NLP parser identified a dominant format for $26, 534 (84.11\%)$ where $p >= 0.5$.
+$1634$ unique formats were identified.
+Figure \ref{fig:NLP1} shows a histogram of the most common formats identified containing \texttt{Year}.
+There are some variations compared to Figure \ref{fig:M3} including the fact that we are not filtering the results by any error rate in the NLP parsed results.
+
 \begin{figure}[ht]
@@
 \end{figure}
-
-Out of the $31,546$ files used for testing, the NLP parser identified a dominant format for $26, 534 (84.11\%)$ where $p >= 0.5$.
-$1634$ unique formats were identified.
-Figure \ref{fig:NLP1} shows a histogram of the most common formats identified containing \texttt{Year}.
-There are some variations compared to Figure \ref{fig:M3} including the fact that we are not filtering the results by any error rate in the NLP parsed results.
 

### 2015-10-16 icde2016/mdl.tex
grammar fix
typo fix

@@
 In our approach, we implement all of these functions, but with significant changes to the last one to increase level of pruning applied during
-domain enumeration. These changes will be described below.
+domain enumeration. These changes are described below.
 
@@
 These are mostly straightforward enumerations and numeric ranges, each tagged with the ICU format code.
-Since the ICU parser is flexible about parsing single or double digit formats, we use double-digit formats, but accept one or two digits.
+Since the ICU parser is flexible about parsing single or double-digit formats, we use double-digit formats, but accept one or two digits.
 One important exception to this rule is for years, which are fixed width fields ($2$ or $4$).

### 2016-04-04 cikm2016/intro.tex
CIKM formatting

--- 2015-10-15 icde2016/intro.tex
@@
 
-Our solution was to develop two machine learning algorithms for automatically deriving the format string from the user's data with over 95\% parsing accuracy.
+Our solution was to develop two algorithms for automatically deriving the format string from the user's data with over 95\% parsing accuracy.
+Both algorithms are built based on classical machine learning algorithms that learn from pattern recognition to make data-driven predictions.
 We developed two algorithms because we were not aware of any previous work on this problem and we needed to be able to cross-validate our results on a large testing corpus.

### 2016-04-04 cikm2016/relatedwork.tex
CIKM formatting

--- 2015-09-28 icde2016/relatedwork.tex
@@
-%\subsection{Previous Work}
-
 Data preparation has been considered an analytic bottleneck since at least the description of the Potter's Wheel system~\cite{Raman:2001}.
@@
 
-\subsection{Contributions}
-
 Our work differs from these previous methods for parsing date-time expressions as we are using an entire column of data as context without the necessary presence of rich lexical identifiers to determine temporal context.
-This sort of data tends to be prevalent in datasets that are used in visualizations.
-
-In particular, our contributions are:
-\begin{itemize}
-\item By analysing an on line corpus, we provide evidence that practical date parsing requires the ability to recognise hundreds of formats;
-\item We show how to extend prior work on Minimum Descriptive Length structure extraction to generate a freely available date format domain language with over 90\% accuracy;
-\item We describe a second Natural Language Processing technique for generating the same date format domain language with similar accuracy.
-We describe how the basic algorithm is extended to support grammar variants and constraints unique to date formats.
-The parsing algorithm is extended to compute an overall dominant pattern over a data column;
-\item We describe some limitations of this domain language that would improve its utility.
-\end{itemize}
+This kind of data tends to be prevalent in datasets that are used in visualizations.

### 2016-04-28 cikm2016/cikm_dateparse.tex
CIKM formatting
Some rewording in the abstract and removal of orphan words.

--- 2015-10-15 icde2016/paper.tex
@@
 %\documentclass{acm_proc_article-sp}
-\documentclass[conference]{IEEEtran}
+%\documentclass[conference]{IEEEtran}
+\documentclass{sig-alternate-05-2015}
 %
@@
 \usepackage{caption} 
+\usepackage{tikz}
 \captionsetup[table]{skip=10pt}
 
+%
+\def\sharedaffiliation{%
+\end{tabular}
+\begin{tabular}{c}}
+%
 \def\eg{\textit{e.g.} }
 \def\ie{\textit{i.e.} }
-\def\dateparse{\texttt{DATEPARSE}}
+\def\dateparse{\mbox{\texttt{DATEPARSE}}}
 \hyphenation{time-stamp}
@@
 \newcommand{\dan}[1]{\textcolor{blue}{(dan: #1)}}
+\newcommand\functionname[1]{\mbox{\texttt{#1}}}
 
@@
 Automating the Extraction of Date Scalars}
-\input{authors}
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
-%\author{\IEEEauthorblockN{Michael Shell}
-%\IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
-%Georgia Institute of Technology\\
-%Atlanta, Georgia 30332--0250\\
-%Email: http://www.michaelshell.org/contact.html}
-%\and
-%\IEEEauthorblockN{Homer Simpson}
-%\IEEEauthorblockA{Twentieth Century Fox\\
-%Springfield, USA\\
-%Email: homer@thesimpsons.com}
-%\and
-%\IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
-%\IEEEauthorblockA{Starfleet Academy\\
-%San Francisco, California 96678-2391\\
-%}}
+\input{anonymous}
 
@@
 Interactive visual analytics is an effective approach for the timely analysis of big data.
-Users who are already engaged in interactive data analysis prefer to stay in the visualization environment for unavoidable cleaning and preparation tasks in order to preserve their analytic flow.
-This has led visualization environments to include simple data preparation functions such as scalar parsing, pattern matching and categorical binning in their analytic toolkits.
+Users who are already engaged in interactive data analysis prefer to stay in the visualization environment for unavoidable data cleaning and preparation tasks in order to preserve their analytic flow.
+This has led visualization environments to include simple data preparation functions such as scalar parsing, pattern matching and categorical binning.
 One of the most common scalar parsing tasks is extracting date and time data from string representations.
-Several RDBMSes include date parsing ``mini-languages'' to cover the wide range of possible formats, but analysis of user data from one visualization system shows that the parsing language syntax is difficult for users to master without disturbing their analytic flow.
+Several relational database systems include date parsing ``mini-languages'' to cover the wide range of possible formats, but analysis of user data from one visualization system shows that the parsing language syntax is difficult for users to master without disturbing their analytic flow.
 
-In this paper, we present two algorithms for automatically deriving these date formats from a column of data with minimal user disruption, one based on minimal entropy representations and another based on natural language processing techniques.
+In this paper, we present two algorithms for automatically deriving date formats from a column of data with minimal user disruption, one based on minimal entropy representations and another based on natural language processing techniques.
 Both have accuracies of over 95\% on a large corpus of date columns extracted from an online data repository and are in substantial agreement with each other.
@@
 \section{Acknowledgments}
-\input{acknowledgements}
+\input{credits}
 
@@
 % produce the bibliography for the citations in your paper.
-\IEEEtriggeratref{13}
-\bibliographystyle{IEEEtran}
+\bibliographystyle{abbrv}
 \bibliography{dateparse}  % sigproc.bib is the name of the Bibliography in this case

### 2016-04-28 cikm2016/intro.tex
edits and comments in intro.

@@
 While computers can provide high-speed and high-volume data processing, humans have both the domain knowledge and the ability to process data in parallel by using their visual systems \cite{Cleveland.McGill1984,Mackinlay:1986}.
-Humans also provide the definition of what is valuable in an analysis.
+%Humans also provide the definition of what is valuable in an analysis.
 Systems that rely on \emph{both} the processing capability of computer systems and human feedback, are more effective in extracting useful knowledge from the large amounts of data being generated than either by itself.
@@
 Visualization systems are most effective when they are interactive, thereby allowing a user to explore data and connect it to their domain knowledge and sense of what is important without breaking cognitive flow.
+
+\vidya{do you have examples for this below to cite?}
 In recent years, a number of such systems have been developed, both by the academic community and by the commercial sector.
-Exploration of data consists not only in creating visual displays, but also in creating and modifying domain-specific computations in the data model, and most data visualization systems include facilities for defining such calculations.
+
+Exploration of data consists not only in creating visualizations, but also in creating and modifying domain-specific computations in the data model.
 The most effective systems allow users to define these calculations as part of the analytic interaction, which permits the user to stay in the flow of analysis~\cite{Morton:2012}.
@@
 Solutions to this problem are often provided by data preparation tools external to the visual analysis environment, which requires the user to break their cognitive flow, launch another tool and reprocess their data before returning to their analysis.
+
+\vidya{this para below either needs to be rewritten to sound less like a white paper or removed/abridged since the previous sentences already allude to this point somewhat.}
 If the user does not own this process (\eg it is the responsibility of another department), then there can be significant delays (including ``never.'') More subtly, the result of updated external processing may not be compatible with the user's existing work, which can lead to more time lost reconciling the new data model with the existing analysis.
-
 From the user's perspective, the boundary between preparation and analysis is not nearly so clean cut.
-Bad data is often discovered using visual analysis techniques (e.g.
-histograms or scatter plots) and it is most natural for the user to ``clean what she sees'' instead of switching to a second tool.
+Bad data is often discovered using visual analysis techniques (\textit{e.g.}, histograms or scatter plots) and it is most natural for the user to ``clean what she sees'' instead of switching to a second tool.
 This leads to an ``adaptive'' process whereby users will prefer leveraging existing tools in the analytics environment over switching to another application -- no matter how poorly these tools may be suited to the task.

### 2016-04-28 cikm2016/intro.tex
Tightened the spacing in bullet points.
More tightening and commented out organization and acknowledgement sections.

@@
 
-One of the most common data preparation tasks we see users performing is the parsing of date strings into scalar date representations, 
-and streamlining this process is the focus of this paper. An examination of the author calculations in an online repository of our
+One of the most common data preparation tasks we see users performing is the parsing of date strings into scalar date representations,  and streamlining this process is the focus of this paper.
+An examination of the author calculations in an online repository of our
 visualization system shows that there are about as many workbooks containing date parsing calculations ($3.3\%$) as there are
@@
 \begin{itemize}
+\setlength\itemsep{0em}
 \item String operations are notoriously slow compared to scalar operations (typically 10-100x slower in modern RDBMSes)
@@
 \begin{itemize}
+\setlength\itemsep{0em}
 \item By analyzing an online corpus, we provide evidence that practical date parsing requires the ability to recognize hundreds of formats;
@@
 \end{itemize}
-
-\subsection{Organization}
-The rest of this paper is organized as follows: The next section provides background information for the problem space.
+%% vidya: don't think we need this.
+%\subsection{Organization}
+%The rest of this paper is organized as follows: The next section provides background information for the problem space.
 The following two sections describe the two different algorithms, one using Minimum Descriptive Length and the other using Natural Language Processing.

### 2016-04-28 cikm2016/nlp.tex
Tightened spacing some more.

--- 2015-10-15 icde2016/nlp.tex
@@
+%But I do think that certain things can be described using more algorithmic way instead of text.
+
 \subsection{Context-Free Grammar}
@@
 \begin{itemize}
-\item Restriction on the distribution of $30$ and $31$:
-Months usually alternate between lengths of $30$ and $31$ days. We can use $x \Mod 2$ to get an alternating pattern of 1 and 0, then just add our constant base number of days:
-
-\begin{equation}
-\texttt{Day} = 30 + x \Mod 2
-\end{equation}
-
-where $x \in [1..12]$ for each of the $12$ months.
+\setlength\itemsep{0em}
+\item \textbf{Restriction on the distribution of $30$ and $31$}:
+Months usually alternate between lengths of $30$ and $31$ days.
+We use $x \Mod 2$ to get an alternating pattern of 1 and 0, and then add the constant base number of days, $\texttt{Day} = 30 + x \Mod 2$, where $x \in [1..12]$ months.
 
@@
 After \texttt{July}, the pattern should skip one, and the rest of the months should follow the alternating pattern inversely.
-
 To obtain an inverse pattern of alternating 0 and 1, we add 1 to the dividend:
-
-\begin{equation}
-\texttt{Day} = 30 + (x + 1) \Mod 2
-\end{equation}
+$\texttt{Day} = 30 + (x + 1) \Mod 2$.
 
@@
 Floor division by 8 works well, since $x < 16$.
+Substituting this expression in the $x + 1$ dividend of Equation $2$, we can invert the pattern using our mask: $\texttt{Day} = 30 + (x + \floor{\frac{x}{8}}) \Mod 2$.
 
-Now if we substitute this expression in the $x + 1$ dividend of Equation $2$, we can invert the pattern using our mask:
-
-\begin{equation}
-\texttt{Day} = 30 + (x + \floor{\frac{x}{8}}) \Mod 2
-\end{equation}
-
-\item LeapYear Restriction for \texttt{February}:
+\item \textbf{LeapYear Restriction for \texttt{February}}:
 While the above restriction applies to all months barring \texttt{February}, we also apply a constraint to the number of days for \texttt{February}, based on whether the year is leap year or not.
 For this, we define a new symbol in the grammar called \texttt{LeapYear}.
-If an expression containing the month \texttt{`February'} or any such variant (\textit{e.g.} \texttt{`Feb'}, \texttt{`2'}) with the day \texttt{`29'} and a year, would need to resolve the \texttt{Year} symbol to be a \texttt{LeapYear}, defined as:
-
-\begin{equation}
-\texttt{Year} \Mod 4 == 0
-\end{equation}
-
+If an expression containing the month \texttt{`February'} or any such variant (\textit{e.g.} \texttt{`Feb'}, \texttt{`2'}) with the day \texttt{`29'} and a year, would need to resolve the \texttt{Year} symbol to be a \texttt{LeapYear}, defined as $\texttt{Year} \Mod 4 == 0$.
 \end{itemize}
@@
 The Gregorian calendar also requires that a year evenly divisible by $100$ (\textit{e.g.}, 1800) is a leap year only if it is also evenly divisible by $400$.
-
 Equations $3$ and $4$ are now added as constraints to the \texttt{Days} symbol in the grammar.
@@
 \subsection{Probabilistic Context-Free Grammar}
+
+%\begin{tikzpicture}
+%	\node (is-root) {S}
+%		[sibling distance=3cm]
+%		child { node {DateGrammar 0.9} }
+%		child {
+%			node {TimeGrammar}
+%				[sibling distance=1.5cm]
+%				child { node {an} }
+%				child { node {example} }
+%				child { node {.} }
+%				child[missing]
+%		};
+%	\path (is-root) +(0,-2.5\tikzleveldistance)
+%		node {\textit{6/15/1978}};
+%\end{tikzpicture}
 
@@
 The locale of the data column is used to retrieve the corresponding corpora for helping with the parsing and for correcting syntactical errors.
-However, similar to the locale sensitivity issue described in Section 4.6, the locale and the data may not always match.
+However, similar to the locale sensitivity issue described in Section IIIF, the locale and the data may not always match.
 The dominant pattern is simply computed over the data column based on the ranked set of parse tree results.

### 2016-04-28 cikm2016/background.tex
More tightening and commented out organization and acknowledgement sections.
Tightened spacing some more.
First pass at tightening
Move some stuff around, pruned excessive verbiage, and tried to make it
read less like a white paper or documentation.

--- 2015-10-15 icde2016/background.tex
@@
 \egroup
-\label{tab:icuformats} \\
 \caption{ICU Format Codes.}
+\label{tab:icuformats}
 \end{table}
 
-While fairly extensive, the ICU syntax has a few limitations that we ran into when working with real-world data:
+While fairly extensive, the ICU syntax has a few limitations when working with real-world data:
 
 \begin{itemize}
+\setlength\itemsep{0em}
 \item No support for 4 letter month abbreviations (\eg \texttt{Sept.})
@@
 The use of lexicographical case in the meta-symbols of the format language can be confusing (\eg \texttt{y} is used for years, but \texttt{M} is used for months while \texttt{m} is used for minutes.
-A better design would have made case more consistent within dates and times.
-Another advantage of automated syntax generation is that it hides such problems from most users, significantly improving the usability of the function.
+A better design would have made case more consistent within dates and time formats.
+Another advantage of automatic syntax generation is that it hides such problems from users.
 

### 2016-04-28 cikm2016/intro.tex
First pass at tightening
Move some stuff around, pruned excessive verbiage, and tried to make it
read less like a white paper or documentation.

@@
 Visualization systems are most effective when they are interactive, thereby allowing a user to explore data and connect it to their domain knowledge and sense of what is important without breaking cognitive flow.
+Exploration of data consists not only in creating visualizations, but also in creating and modifying domain-specific computations in the data model. 
+During the analytic process, a user may discover that parts of the data are not yet suitable for analysis.
 
-\vidya{do you have examples for this below to cite?}
-In recent years, a number of such systems have been developed, both by the academic community and by the commercial sector.
+Preparing this data for analysis often uses data preparation tools external to the visual analysis environment, which requires the user to break their cognitive flow, launch another tool and reprocess her data before returning to her analysis.
+Recent work~\cite{Morton:2012} has found, however, that the most effective systems allow users to define these calculations as part of the analytic interaction, which enables the user to stay in the flow of analysis.
 
-Exploration of data consists not only in creating visualizations, but also in creating and modifying domain-specific computations in the data model.
-The most effective systems allow users to define these calculations as part of the analytic interaction, which permits the user to stay in the flow of analysis~\cite{Morton:2012}.
-
-During the analytic process, a user may discover that parts of the data are not yet suitable for analysis.
-Solutions to this problem are often provided by data preparation tools external to the visual analysis environment, which requires the user to break their cognitive flow, launch another tool and reprocess their data before returning to their analysis.
-
-\vidya{this para below either needs to be rewritten to sound less like a white paper or removed/abridged since the previous sentences already allude to this point somewhat.}
-If the user does not own this process (\eg it is the responsibility of another department), then there can be significant delays (including ``never.'') More subtly, the result of updated external processing may not be compatible with the user's existing work, which can lead to more time lost reconciling the new data model with the existing analysis.
-From the user's perspective, the boundary between preparation and analysis is not nearly so clean cut.
-Bad data is often discovered using visual analysis techniques (\textit{e.g.}, histograms or scatter plots) and it is most natural for the user to ``clean what she sees'' instead of switching to a second tool.
-This leads to an ``adaptive'' process whereby users will prefer leveraging existing tools in the analytics environment over switching to another application -- no matter how poorly these tools may be suited to the task.
-Thus a well-designed interactive visual analysis environment will provide effective tools that enable users to perform such cleaning and preparation tasks as interactively as possible.
+We have observed that one of the most common data preparation tasks users perform is parsing date strings into scalar date representations.
+For example, an examination of author calculations in a public repository using our
+visualization system shows that there are about as many workbooks containing date parsing calculations ($3.3\%$) as there are
+integer type conversions of any kind ($3.4\%$).
+Streamlining this task is the focus of this paper.
 
@@
 
-One of the most common data preparation tasks we see users performing is the parsing of date strings into scalar date representations,  and streamlining this process is the focus of this paper.
-An examination of the author calculations in an online repository of our
-visualization system shows that there are about as many workbooks containing date parsing calculations ($3.3\%$) as there are
- integer type conversions of any kind ($3.4\%$).
-We now provide some background on the use of dates in data analytics to provide context for the problem.
+The SQL-99 standard defines three temporal scalar types: \texttt{DATE}, \texttt{TIMESTAMP} and \texttt{TIME}, which can be further qualified as either \texttt{WITH} or \texttt{WITHOUT} \texttt{TIME ZONE} (the \texttt{WITHOUT} form being more common.) They are typically implemented as fixed-point types containing an offset from some epoch (\eg Julian Days.) This makes them compact to store using column store compression techniques such as those in C-Store~\cite{Stonebraker:2005} and MonetDB/X100~\cite{Zukowski:2006}, and further allows some temporal operations to be implemented very efficiently using simple arithmetic.
+Thus from the RDBMS perspective, representing dates in scalar form provides  benefits for users, both in terms of analytic richness and query performance.
 
-The SQL-99 standard defines three temporal scalar types: \texttt{DATE}, \texttt{TIMESTAMP} and \texttt{TIME}, which can be further qualified as either \texttt{WITH} or \texttt{WITHOUT} \texttt{TIME ZONE} (the \texttt{WITHOUT} form being more common.) They are typically implemented as fixed-point types containing an offset from some epoch (\eg Julian Days.) This makes them compact to store using column store compression techniques such as those in C-Store~\cite{Stonebraker:2005} and MonetDB/X100~\cite{Zukowski:2006}, and further allows some temporal operations to be implemented very efficiently using simple arithmetic.
-Thus from the RDBMS perspective, representing dates in scalar form provides  benefits for users, both in terms of analytic operations and query performance.
-
-Our visualization system models the first two of these SQL-99 types natively; the third (pure time) is folded into \texttt{TIMESTAMP} by appending it to a fixed date of \texttt{1899-12-30} (a convention derived from Microsoft Excel).
-From the analytic perspective, date types are dimensional (\ie independent variables) and can be used as either \textit{categorical} (simply ordered) or \textit{quantitative} (ordered with a distance metric) fields.
-
+From an analytic perspective, date types are dimensional (\ie independent variables) and can be used as either \textit{categorical} (simply ordered) or \textit{quantitative} (ordered with a distance metric) fields.
 Categorical dates have a natural hierarchy associated with them generated by calendar \textit{binning}.
@@
 These types of visualizations are difficult to specify with simple categorical string representations of dates.
-Thus the parsing of strings to date scalars is an important form of preparation for data analytics.
 
 \subsection{Parsing Dates}
-This parsing of date representations is one of the oldest data preparation problems we have observed.
-A common version of this problem is how to convert columns of integers of the form \texttt{yyyyMMdd} to date scalars.
-Na\"{i}ve users often solve this problem by converting the integer to a string and performing some locale-dependent string operations before casting the string back to a date.
+A common version of the date parsing problem is converting columns of integers in the form \texttt{yyyyMMdd} to date scalars.
+\Naive\ users often solve this problem by converting the integer to a string and performing some locale-dependent string operations before casting the string back to a date.
 Unfortunately, this approach has a number of problems:
@@
 \item String operations are notoriously slow compared to scalar operations (typically 10-100x slower in modern RDBMSes)
-\item Default parsing of date formats is locale-dependent, and may not work when the analysis is shared across an international organization (\eg between the US and European offices)
+\item Default parsing of date formats is locale-dependent, and may not work when the parsing expression is shared across an international organization (\eg between the US and European offices)
 \item The parsing code is hard to understand and maintain because it uses a verbose, general-purpose string-handling syntax instead of a specialized domain language.
@@
 
-And this is but a single date format.
-Our studies of online data collections suggest that there are hundreds of distinct temporal date formats in user data sets.
+And this is but a single date format. In Section~\ref{sec:experiments} we show that there are \textit{hundreds} of distinct temporal date formats in user data sets (\cf Figure~\ref{fig:M2}).
 Some are common, but others can be quite idiosyncratic. 
-Table 1 shows a selection of unusual date formats found in our corpus (the meanings of the ICU formatting codes can be found in Table 2).
-The first example shows a time zone in the middle of the date and a year after the time; the second shows a leading unmatched bracket and a colon between the date and time components; the third shows confusion between the seconds' decimal point and the time part delimiter; the fourth shows a two digit year apostrophe on a four digit year and the fifth shows a dash separating the date and time components.
-As can be seen in Figure \ref{fig:M2}, the long tail of formats means that these idiosyncrasies are the rule, not the exception, 
+Table~\ref{tab:dateformats} shows a selection of unusual date formats found in our corpus (the meanings of the ICU formatting codes can be found in Table~\ref{tab:icuformats}). 
+The first example shows a time zone in the middle of the date and a year after the time; 
+the second shows a leading unmatched bracket and a colon between the date and time components; 
+the third shows confusion between the seconds' decimal point and the time part delimiter; 
+the fourth shows a two digit year apostrophe on a four digit year and the fifth shows a dash separating the date and time components. 
+This ``long tail'' of formats means that these idiosyncrasies are the rule, not the exception, 
 so a static enumeration of a small set of formats will not produce high reliability.
@@
 \egroup
+\caption{Unusual Date Formats.}
 \label{tab:dateformats}
-\caption{Unusual Date Formats.}
 \end{table}
@@
 Several RDBMSes (\eg MySQL, Oracle and Postgres) provide row-level functions for parsing formatted dates, but our analysis of user data showed a 15\% syntax error rate using these functions.
-And even with perfect syntax, the user still has to interrupt her flow to learn a formatting syntax -- which she will likely forget once this specific problem has been solved.
+And even with perfect syntax, the user still has to interrupt her flow to learn a formatting syntax -- one that she will likely forget after this specific problem has been solved.
 
-One solution to this problem might have been to design a graphical environment that enabled users to construct valid patterns using visually compelling representations to reduce the cognitive load.
-This approach would have involved a substantial development effort with no guarantee that the result would be correct if the user misunderstood the environment.
+One solution might have been to design a graphical environment for constructing valid patterns and using visually compelling representations to reduce the cognitive load. 
+Yet this approach provide no guarantee that the result would be correct if the user misunderstood the environment. 
 Moreover, this approach only solves the syntax problem -- the user is still required to switch contexts to solve their problem.
 
-Our solution was to develop two algorithms for automatically deriving the format string from the user's data with over 95\% parsing accuracy.
+Our solution was to develop two algorithms for automatically deriving the format string from the user's data with over $95\%$ parsing accuracy.
 Both algorithms are built based on classical machine learning algorithms that learn from pattern recognition to make data-driven predictions.
-We developed two algorithms because we were not aware of any previous work on this problem and we needed to be able to cross-validate our results on a large testing corpus.
-Either approach allows the user to simply specify ``this column is a date" and the visualization system can respond quickly and accurately enough to avoid interrupting the user's cognitive flow.
+We developed two algorithms because we were unaware of any previous work on this problem and we needed to be able to cross-validate our results on a large testing corpus.
+With both approaches the user can simply specify ``this column is a date" and the visualization system can respond quickly and accurately enough to avoid interrupting the user's cognitive flow.
+
+\subsection{Related Work}
+\input{relatedwork}
 
@@
 %% vidya: don't think we need this.
+%% richard: It's kind of traditional in the DB community, but we can just leave it commented out - and put it back if they complain!
 %\subsection{Organization}

### 2016-04-28 cikm2016/testing.tex
Tightened spacing some more.
First pass at tightening
Move some stuff around, pruned excessive verbiage, and tried to make it
read less like a white paper or documentation.

--- 2015-10-16 icde2016/testing.tex
@@
 For instance, typing the string \texttt{"12/31/1999"} into Microsoft Excel, is automatically interpreted as the date \texttt{1999-12-31}.
-The Microsoft Jet library \vidya{need a reference here} used to read these text files detects a few date formats as well.
+The Microsoft Jet library \cite{Jet} used to read these text files detects a few date formats as well.
 Any column already converted by Excel or Jet was not included in this study.
@@
 \begin{itemize}
+\setlength\itemsep{0em}
 \item Values containing the substring \texttt{NULL}
@@
 \end{itemize}
-Columns that had no remaining valid samples were discarded.
+%Columns that had no remaining valid samples were discarded.
 
@@
 
-\subsection{Evaluation}
+\subsubsection{Training Data}
+
 Once the training data was analyzed, it was grouped by date format.
@@
 Many of the files in the collection are not actually dates, such as fields named ``updated by'' (which contain ``date'').
-A random sample of 850 columns named exactly ``date'', ``time'' or ``month'' (case-insensitive) were manually judged.\\
+A random sample of 850 columns named exactly ``date'', ``time'' or ``month'' (case-insensitive) were manually judged.
 
-\subsubsection{Minimum Descriptive Length}
+\subsection{Minimum Descriptive Length}
 Testing of the MDL algorithm was performed on a 24-core Dell T7610 running Windows 7 with the data stored on a 250GB SSD.
@@
 \egroup
+\caption{MDL Parsing Statistics.}
 \label{tab:mdlstats}
-\caption{MDL Parsing Statistics.}
 \end{table}
@@
 In cases where we generated multiple formats and the main format produced errors, we applied the second format to the unparsed strings.
-The summary statistics from this processing are presented in Table 2.
+The summary statistics from this processing are presented in Table~\ref{tab:mdlstats}.
 
@@
 At 2.5ms, this is well below most human perceptual thresholds for a set of 32 samples, so any latency in command execution would be restricted to the ability of the underlying database to provide the samples for analysis in a timely manner.
-(As a column store, the database can often supply such domains without a full table scan, further improving responsiveness.)
+A columnar database (such as the one underlying our visualization system) can often supply such domains without a full table scan, further improving responsiveness.
 
-The validation speed is the average time needed to parse a value, and provides an estimate of how fast the ICU implementation can process string values into scalars and works out to 620K values per core per second.\\
+The validation speed is the average time needed to parse a value, and provides an estimate of how fast the ICU implementation can process string values into scalars and works out to 620K values per core per second.
 
@@
 
-\subsubsection{Natural Language Processing}
+\subsection{Natural Language Processing}
 
@@
 
-\subsubsection{Cross-Checking}
+\subsection{Cross-Checking}
 We compared the results of the minimum description length and natural language processing algorithms.
@@
 One other file contained a set of integers that are not actually dates, and one algorithm picked \texttt{MMyyyy} format, while the other picked \texttt{HHmmss} format.
-
-\subsubsection{Production}
-
-Although both algorithms achieved comparable high accuracies in parsing date time strings, we utilized the minimum description length method in the production code of our system.
-This was primarily due to the fact that the MDL algorithm was written in C++, allowing for higher performance.
-We believe that NLP version if written in C++ as opposed to Python, would lead to better performance.

### 2016-04-29 cikm2016/background.tex
More text pruning.

@@
 
-While fairly extensive, the ICU syntax has a few limitations when working with real-world data:
-
-\begin{itemize}
-\setlength\itemsep{0em}
-\item No support for 4 letter month abbreviations (\eg \texttt{Sept.})
-\item No support for ordinal days (\eg \texttt{July 4th})
-\item No support for quarter postfix notation (\eg \texttt{2Q})
-\item No support for variant meridian markers (\eg \texttt{a.m.})
-\end{itemize}
+While fairly extensive, the ICU syntax has a few limitations when working with real-world data.
+It has no support for 4-letter month abbreviations (\eg \texttt{Sept.}), ordinal days (\eg \texttt{July 4th}), quarter postfix notation (\eg \texttt{2Q}) and variant meridian markers (\eg \texttt{a.m.}).
 

### 2016-04-29 cikm2016/intro.tex
More text pruning.

@@
 \Naive\ users often solve this problem by converting the integer to a string and performing some locale-dependent string operations before casting the string back to a date.
-Unfortunately, this approach has a number of problems:
-\begin{itemize}
-\setlength\itemsep{0em}
-\item String operations are notoriously slow compared to scalar operations (typically 10-100x slower in modern RDBMSes)
-\item Default parsing of date formats is locale-dependent, and may not work when the parsing expression is shared across an international organization (\eg between the US and European offices)
-\item The parsing code is hard to understand and maintain because it uses a verbose, general-purpose string-handling syntax instead of a specialized domain language.
-\end{itemize}
+Unfortunately, this approach has a number of problems because string operations are notoriously slow compared to scalar operations (typically 10-100x slower in modern RDBMSes).
+Default parsing of date formats is locale-dependent, and may not work when the parsing expression is shared across an international organization (\eg between the US and European offices).
+The parsing code is often hard to understand and maintain because it uses a verbose, general-purpose string-handling syntax instead of a specialized domain language.
+Further, there is no single data format.
 
-And this is but a single date format. In Section~\ref{sec:experiments} we show that there are \textit{hundreds} of distinct temporal date formats in user data sets (\cf Figure~\ref{fig:M2}).
+In Section~\ref{sec:experiments} we show that there are \textit{hundreds} of distinct temporal date formats in user data sets (\cf Figure~\ref{fig:M2}).
 Some are common, but others can be quite idiosyncratic. 
@@
 We developed two algorithms because we were unaware of any previous work on this problem and we needed to be able to cross-validate our results on a large testing corpus.
-With both approaches the user can simply specify ``this column is a date" and the visualization system can respond quickly and accurately enough to avoid interrupting the user's cognitive flow.
+With both approaches, the user can simply specify that the column is a date, and the visualization system can respond quickly and accurately enough to avoid interrupting the user's flow.
 

### 2016-04-29 cikm2016/mdl.tex
CIKM formatting
Tightened spacing some more.
More text pruning.

--- 2015-10-16 icde2016/mdl.tex
@@
 \begin{itemize}
+\setlength\itemsep{0em}
 \item A required inclusion function to test for membership in the domain (\functionname{match}). 
-For example, the domain \functionname{<Digits>} would return true for the string \functionname{"123"} 
-and false for the string \functionname{"abc"}.
+For \textit{e.g.}, the domain \functionname{<Digits>} returns true for \functionname{"123"} 
+and false for \functionname{"abc"}.
 \item An optional function to compute the number of values in the domain with a given length (\functionname{cardinality}).
-For example, the domain for three letter month names (\functionname{MMM}) would return $12$ for the length $3$ and $0$ otherwise.
+For \textit{e.g.}, the domain for three letter month names (\functionname{MMM}) would return $12$ for the length $3$ and $0$ otherwise.
 \item An optional function to update statistics for the domain based on a given value (\functionname{updateStatistics}).
-For example, the domain \functionname{MMM} might keep a histogram of how frequently each month was encountered.
+For \textit{e.g.}, the domain \functionname{MMM} might keep a histogram of how frequently each month was encountered.
 \item An optional function to prevent consideration of a domain that is redundant (\functionname{isRedundantAfter}).
-For example, the domain \functionname{<Digits>} would be redundant after itself 
+For \textit{e.g.}, the domain \functionname{<Digits>} would be redundant after itself 
 (\ie there is no difference between a single run of digits and two adjacent runs.)
@@
 This is especially true in the date format problem because there are identical domains (\eg months and meridian hours), nearly identical domains (\eg days and hours) and there are often no field delimiters (\eg \texttt{2012Mar06134427}).
-To handle this, we have extended the existing pruning API with two other sets of domain identifiers:
+To handle this, we extend the existing pruning API with two other sets of domain identifiers:
 \begin{itemize}
+\setlength\itemsep{0em}
 \item A set of \textit{prunable} identifiers, which are not allowed to precede the domain.
@@
 \begin{itemize}
+\setlength\itemsep{0em}
 \item The set of date parts cannot contain place value gaps (\eg structures that have year and day without month are removed.)
@@
 \subsection{Ranking}
-MDL structure analysis naturally produces a ranked list of format candidates, but we have found that a number of other properties of the formats should be preferred over simple compactness:
+MDL structure analysis produces a ranked list of format candidates, but we have found that a number of other properties of the formats should be preferred over simple compactness:
 \begin{itemize}
+\setlength\itemsep{0em}
 \item Because we have a set of samples, we can apply the candidate format to the strings to see how well it performs. Formats with fewer parse errors are preferred.
@@
 The sample set may have missed an example where this could be important.
-For example, the month name for "September" is the same in both German and English, but "October" and "Oktober" are different.
+For example, the month name for "September" is the same in both German and English, but ``October" and ``Oktober" are different.
 \item If the format has an ambiguous date order (\eg all days are less than 12), then prefer the default date order of the locale.

### 2016-04-29 cikm2016/nlp.tex
More text pruning.

@@
-%But I do think that certain things can be described using more algorithmic way instead of text.
-
 \subsection{Context-Free Grammar}
@@
 
-We define a \textit{EBNF} grammar for identifying date-time strings as follows:
+A portion of the \textit{EBNF} grammar is specified below:
 
@@
 
-<Day>     ::= dd (where $dd \in [01-31]$, depending on month/year);
+%<Day>     ::= dd (where $dd \in [01-31]$, depending on month/year);
 
-<Month> := <MonthWord> | <MonthNumber>;
+%<Month> := <MonthWord> | <MonthNumber>;
 
-<MonthWord> := `January' | `February' | `March' | `April' | `May' | `June' | `July' | `August' | `September' | `October' | `November' | `December';
-
-<MonthNumber> := dd (where $dd \in [01-12]$);
-
-<DayOfWeek> ::= `Monday' | `Tuesday' | `Wednesday' | `Thursday' | `Friday' | `Saturday' | `Sunday';
-
-<Hour> ::= <TwelveHour> | <TwentyFourHour>;
-
-<Quarter> ::= `Quarter' dd (where $dd \in [01 - 04]$);
-
-<TwoYear> ::= dd;
-
-<FourYear> := dddd;
-
-<TwelveHour> ::= dd (where $dd \in [00-12]$);
-
-<TwentyFourHour> ::= dd (where $dd \in [00 -23]$);
-
-<AMPM> := `a.m.' | `p.m.';
-
-<d> ::= `0' | `1' | `2' | `3' | `4' | `5' | `6' | `7' | `8' | `9';
+%<MonthWord> := `January' | `February' | `March' | `April' | `May' | `June' | `July' | `August' | `September' | `October' | `November' | `December';
+%
+%<MonthNumber> := dd (where $dd \in [01-12]$);
+%
+%<DayOfWeek> ::= `Monday' | `Tuesday' | `Wednesday' | `Thursday' | `Friday' | `Saturday' | `Sunday';
+%
+%<Hour> ::= <TwelveHour> | <TwentyFourHour>;
+%
+%<Quarter> ::= `Quarter' dd (where $dd \in [01 - 04]$);
+%
+%<TwoYear> ::= dd;
+%
+%<FourYear> := dddd;
+%
+%<TwelveHour> ::= dd (where $dd \in [00-12]$);
+%
+%<TwentyFourHour> ::= dd (where $dd \in [00 -23]$);
+%
+%<AMPM> := `a.m.' | `p.m.';
+%
+%<d> ::= `0' | `1' | `2' | `3' | `4' | `5' | `6' | `7' | `8' | `9';
 
@@
 
-%\begin{tikzpicture}
-%	\node (is-root) {S}
-%		[sibling distance=3cm]
-%		child { node {DateGrammar 0.9} }
-%		child {
-%			node {TimeGrammar}
-%				[sibling distance=1.5cm]
-%				child { node {an} }
-%				child { node {example} }
-%				child { node {.} }
-%				child[missing]
-%		};
-%	\path (is-root) +(0,-2.5\tikzleveldistance)
-%		node {\textit{6/15/1978}};
-%\end{tikzpicture}
-
 Pattern-recognition problems such as parsing date and time formats initiate from observations generated by some structured stochastic process.
-In other words, even if the initial higher-level production rule of the grammar is known (i.e.
-date, time or date-time), there could be several directions that the parser resolves to.
-For example, in a date string \texttt{`5/6/2015'}, the pattern could either be \texttt{M/d/yyyy} or \texttt{d/M/yyyy}.
+In other words, even if the initial higher-level production rule of the grammar is known (\textit{i.e.} date, time or date-time), there could be several directions that the parser resolves to.
+For example, for \texttt{`5/6/2015'}, the pattern could either be \texttt{M/d/yyyy} or \texttt{d/M/yyyy}.
 
@@
 The resulting formalism called \textit{Probabilistic Context-Free Grammar} (PCFG), extends the CFG by assigning probabilities to the production rules of the grammar.
-During the process of parsing the date-time pattern, the probabilities are used as a filtering mechanism to rank the pattern(s) that a given string resolves to in the grammar.
+During the process of parsing the date-time pattern, the probabilities are used to rank the pattern(s) that a given string resolves to in the grammar.
 
@@
 Assigning initial probability weights to the grammar for computing the parse with the highest weight, tends to be a more scalable and flexible approach to disambiguation rather than building a deep knowledge ontology.
-We bootstrap the probability weights in the \texttt{PCFG}, by setting initial probability weights to various rules in the grammar as follows:
+We bootstrap the probability weights in the \texttt{PCFG}, by setting initial probability weights to the grammar rules as follows:
 

### 2016-04-29 cikm2016/testing.tex
More text pruning.

@@
 \subsubsection{Sampling}
-The remaining samples were hashed and sorted on the hash value, with the top $32$ values being retained as the column's sample.
-We can increase the sample size if needed, but for dates, it appears to be an adequate number.
+The remaining samples were hashed and sorted on the hash value, with the top $32$ values retained as the column's sample.
+We can increase the sample size, but for dates, it appears to be an adequate number.
 In any case, the median number of non-null rows per domain in our test data is $50$, so increasing the sample size would have little benefit.
@@
 
-ICU's date parsing APIs allow the specification of default values for parts when a format does not contain them.
+ICU's date parsing APIs allow the specification of default values for parts not found in the format.
 In our implementation, all time fields are set to \texttt{0} (midnight) and the date fields are set based on whether the format contains any date part specifications.
@@
 
-ICU will also parse Time Zones (which the algorithms recognize but the visualization system does not model) and Quarters (which we interpret as the first month of the period.) RDBMSes such as Oracle and Postgres that support time zones will be able to take advantage of  identified time zone fields.
+ICU will also parse Time Zones and Quarters (which we interpret as the first month of the period.) RDBMSes such as Oracle and Postgres that support time zones will be able to take advantage of  identified time zone fields.
 

### 2016-04-30 cikm2016/nlp.tex
More text trimming

@@
+\vidya{need to add a short intro here}
+
 \subsection{Context-Free Grammar}
@@
 
-The parsed output is then converted to the ICU Date Format Language described in Section IIA in order to effectively cross-validate results from the MDL algorithm.
+The parsed output is then converted to the ICU Date Format Language in order to effectively cross-validate results from the MDL algorithm.
 While this grammar accounts for the use of meridian markers `a.m.' and `p.m.' for the 12-hour format, ICU does not support these tokens, and they are simply ignored.
@@
 If an expression containing the month \texttt{`February'} or any such variant (\textit{e.g.} \texttt{`Feb'}, \texttt{`2'}) with the day \texttt{`29'} and a year, would need to resolve the \texttt{Year} symbol to be a \texttt{LeapYear}, defined as $\texttt{Year} \Mod 4 == 0$.
+However, this is just an approximation.
+The Gregorian calendar also requires that a year evenly divisible by $100$ (\textit{e.g.}, 1800) is a leap year only if it is also evenly divisible by $400$.
+
 \end{itemize}
-
-However, Equation $4$ is just an approximation.
-The Gregorian calendar also requires that a year evenly divisible by $100$ (\textit{e.g.}, 1800) is a leap year only if it is also evenly divisible by $400$.
-Equations $3$ and $4$ are now added as constraints to the \texttt{Days} symbol in the grammar.
 
@@
 
-The production rules are of the form $ X \rightarrow \beta$ where $X$ is the left-hand side of the rule, while $\beta$ is the right-hand side.
-
-Given a CFG grammar $G$, we define  $\tau_G(s)$ as the set of all possible parse trees for input date-time string $s$.
+Given a CFG grammar $G$, the production rules are of the form $ X \rightarrow \beta$ where $X$ is the left-hand side of the rule, while $\beta$ is the right-hand side.
+We define  $\tau_G(s)$ as the set of all possible parse trees for input date-time string $s$.
 For any $X \rightarrow \beta \in \tau(G)$, the probability $p(X \rightarrow \beta) \ge 1$.
@@
 Assigning initial probability weights to the grammar for computing the parse with the highest weight, tends to be a more scalable and flexible approach to disambiguation rather than building a deep knowledge ontology.
-We bootstrap the probability weights in the \texttt{PCFG}, by setting initial probability weights to the grammar rules as follows:
+We bootstrap the probability weights in the \texttt{PCFG} as follows:
 
 \begin{itemize}
-\item Similar to the MDL ranking properties described in Section IIIG, rules involving the non-terminal \texttt{DateGrammar} on the left-side of the rule, are assigned probability weights higher than those rules with the non-terminal being \texttt{TimeGrammar}.
+\setlength\itemsep{0em}
+\item Similar to the MDL ranking properties, rules involving the non-terminal \texttt{DateGrammar} on the left-side of the rule, are assigned probability weights higher than those rules with the non-terminal being \texttt{TimeGrammar}.
 In practice, assigning $p(\texttt{DateGrammar} \rightarrow \beta_{1}) = 0.9$ and $p(\texttt{TimeGrammar} \rightarrow \beta_{2}) = 0.7$ leads to optimal ranking of the parse trees.
@@
 Having defined initial probabilistic weights to the \texttt{PCFG}, we employ supervised learning with a known training set of date time formats to estimate the rule probabilities.
-The training corpus is a set of files obtained from an online data analysis corpus, described in Section 6.
+The training corpus is a set of files obtained from an online data analysis corpus, described in Section 5.
 
@@
 The locale of the data column is used to retrieve the corresponding corpora for helping with the parsing and for correcting syntactical errors.
-However, similar to the locale sensitivity issue described in Section IIIF, the locale and the data may not always match.
+However, similar to the locale sensitivity issue described in Section 3.6, the locale and the data may not always match.
 The dominant pattern is simply computed over the data column based on the ranked set of parse tree results.

### 2016-04-30 cikm2016/testing.tex
More tightening
More text trimming

@@
 \subsubsection{NULL Filtering}
-Each column was then reduced to a maximum sample set of 32. We excluded domain values that appeared to be representations of \texttt{NULL}:
-\begin{itemize}
-\setlength\itemsep{0em}
-\item Values containing the substring \texttt{NULL}
-\item Values containing no digits and at most one non-whitespace character (\eg \texttt{" / / "})
-\item Values on a list of empirically determined common \texttt{NULL} values (\eg \texttt{0000-00-00}, \texttt{NaN})
-\end{itemize}
+Each column was then reduced to a maximum sample set of 32.
+We excluded domain values that either contain the substring \texttt{NULL}, contain no digits and at most one non-whitespace character (\eg \texttt{" / / "}), or any empirically determined common \texttt{NULL} value (\eg \texttt{0000-00-00}, \texttt{NaN}).
 %Columns that had no remaining valid samples were discarded.
@@
 The samples were judged as to whether the produced format was reasonable and were tagged with correct formats if the produced format was unreasonable.
-
+\vidya{I don't understand this para below. Can be abridged and tightened with the previous para}
 Further manual tagging focused on the files most likely to represent dates.

### 2016-04-30 cikm2016/futurework.tex
Some more text trimming.

--- 2015-10-08 icde2016/futurework.tex
@@
 
-In this paper, we have only considered the parsing of strings, but dates are often formatted as integers (\eg \texttt{201507016}).
-It is significantly faster to decompose integers into date parts using arithmetic operations (\eg \texttt{mod} and \texttt{div}) than by using slower, locale-sensitive string parsing functions.
-Moreover, the number of possible formats is low enough that we may be able to enumerate them.
+In this paper, we have considered the parsing of strings, but dates are often formatted as integers (\eg \texttt{201507016}).
+It is significantly faster to decompose integers into date parts using arithmetic operations (\eg \texttt{mod} and \texttt{div}) than by using locale-sensitive string parsing functions.
 Timestamp preparation from numeric representations is another form of preparation which we would like to automate.

### 2016-04-30 cikm2016/testing.tex
Some more text trimming.

@@
 
-Similar to the MDL algorithm described in the previous subsection, the CYK parser is run over the same set of samples from each validation file to generate a ranked list of parse trees representing the date-time formats.
+Similar to the MDL algorithm, the CYK parser is run over the same set of samples from each validation file to generate a ranked list of parse trees representing the date-time formats.
 As a second pass of the parser, starting from the most probable parse tree, each parse tree is then applied to the entire file's columnar data to determine the dominant pattern(s).

### 2016-05-02 cikm2016/testing.tex
Training Data section
Tightened confusing language

@@
 The samples were judged as to whether the produced format was reasonable and were tagged with correct formats if the produced format was unreasonable.
-\vidya{I don't understand this para below. Can be abridged and tightened with the previous para}
-Further manual tagging focused on the files most likely to represent dates.
-Many of the files in the collection are not actually dates, such as fields named ``updated by'' (which contain ``date'').
 A random sample of 850 columns named exactly ``date'', ``time'' or ``month'' (case-insensitive) were manually judged.
@@
 The error rate reflects the fact that only about 40\% of the files have an associated format that parses the non-null values without error.
-To examine the error rate in greater detail, we turn now to Figure \ref{fig:M2}.
+To examine the error rate in greater detail, we turn now to Figure~\ref{fig:M2}.
 
-On the left hand side, we can see that the algorithm found 744 distinct formats that parsed 13,424 files with no error.
-This is a remarkable number of valid formats and underscores the need for this kind of algorithm.
-Raising the error rate threshold to 5\% results in about 2500 formats found in 15,000 files, or nearly half of the files.
+The two horizontal reference lines in Figure~\ref{fig:M2} show that we found 744 distinct formats that parsed the $13,424$ associated files with no error. 
+This is a remarkable number of distinct formats and underscores the need for this kind of algorithm. 
+Raising the error rate threshold to $5\%$ results in about $2500$ formats found in $15,000$ files, or nearly half of the files in the corpus.
 
 What do these formats look like?
-Figure \ref{fig:M3} shows a histogram of the 25 most common formats containing a year format code at the 5\% threshold, color-coded by error rate.
-(A sample value is provided to the right of each bar for illustrative purposes.) The formats have also been filtered to files with at least 5 samples.
+Figure~\ref{fig:M3} shows a histogram of the $25$ most common formats containing a year format code at the $5\%$ error threshold, color-coded by error rate.
+(A sample value is provided to the right of each bar for illustrative purposes.) The formats have also been filtered to files with at least $5$ samples.
 Most of the samples are clearly dates with a wide range of formats (the format where the time zone is between the time and the year is surprisingly common.)

### 2016-05-02 cikm2016/background.tex
More text trimming
added a comment for the icu table.
Cut down ICU table

@@
 (In fact, this latter approach is how our visualization system internally generates formats for RDBMSes that have non-ICU syntax.)
-
+%\vidya{do we need the whole table here?}
+%\richard{Chopped it down some more and removed multi-line descriptions. All we really need to get across is that there are lots of different time zones, etc., so  fewer examples are OK I think.}
 \begin{table}[ht]
@@
 \def\arraystretch{1.5}
-\begin{tabular}{|p{0.15\linewidth}| p{0.5\linewidth}|}
+\begin{tabular}{|p{0.2\linewidth}| p{0.5\linewidth}|}
 \hline
@@
 
-\scriptsize{zzz} & \scriptsize{Time Zone: specific non-location (PDT)}\\ \hline
+\scriptsize{zzz} & \scriptsize{Time Zone: (PDT)}\\ \hline
 
 \scriptsize{Z} & \scriptsize{Time Zone: RFC 822 (-0800)}\\ \hline
-\scriptsize{ZZZZ} & \scriptsize{Time Zone: localized GMT (GMT-08:00)}\\ \hline
+\scriptsize{ZZZZ} & \scriptsize{Time Zone: (GMT-08:00)}\\ \hline
 \scriptsize{ZZZZZ} & \scriptsize{Time Zone: ISO8601 (-08:00)}\\ \hline
 
-\scriptsize{'} & \scriptsize{escape for text (nothing)}\\ \hline
-\scriptsize{''} & \scriptsize{two single quotes produce one (')}\\ \hline
+\scriptsize{'} & \scriptsize{escape for text}\\ \hline
 
@@
 It has no support for 4-letter month abbreviations (\eg \texttt{Sept.}), ordinal days (\eg \texttt{July 4th}), quarter postfix notation (\eg \texttt{2Q}) and variant meridian markers (\eg \texttt{a.m.}).
-
 These limitations did not affect the results significantly, and in the future we hope to submit ICU extensions to handle some of these formats.
-Moreover, all the languages we examined had their own limitations, so ICU is not unusual in this respect.
-
-One other quirk of the ICU syntax may be a contributing factor to the observed user difficulties writing correct ICU date formats in our visualization system.
-The use of lexicographical case in the meta-symbols of the format language can be confusing (\eg \texttt{y} is used for years, but \texttt{M} is used for months while \texttt{m} is used for minutes.
-A better design would have made case more consistent within dates and time formats.
-Another advantage of automatic syntax generation is that it hides such problems from users.
 

### 2016-05-02 cikm2016/intro.tex
More text trimming
Make more academic
I hope!

@@
 
-The SQL-99 standard defines three temporal scalar types: \texttt{DATE}, \texttt{TIMESTAMP} and \texttt{TIME}, which can be further qualified as either \texttt{WITH} or \texttt{WITHOUT} \texttt{TIME ZONE} (the \texttt{WITHOUT} form being more common.) They are typically implemented as fixed-point types containing an offset from some epoch (\eg Julian Days.) This makes them compact to store using column store compression techniques such as those in C-Store~\cite{Stonebraker:2005} and MonetDB/X100~\cite{Zukowski:2006}, and further allows some temporal operations to be implemented very efficiently using simple arithmetic.
+The SQL-99 standard defines three temporal scalar types: \texttt{DATE}, \texttt{TIMESTAMP} and \texttt{TIME}, which can be further qualified as either \texttt{WITH} or \texttt{WITHOUT} \texttt{TIME ZONE} (the \texttt{WITHOUT} form being more common.) They are typically implemented as fixed-point types containing an offset from some epoch (\eg Julian Days.) This makes them compact to store using columnar compression techniques such as those in C-Store~\cite{Stonebraker:2005} and MonetDB/X100~\cite{Zukowski:2006}, and further allows some temporal operations to be implemented efficiently using simple arithmetic.
 Thus from the RDBMS perspective, representing dates in scalar form provides benefits for users, both in terms of analytic richness and query performance.
@@
 The parsing code is often hard to understand and maintain because it uses a verbose, general-purpose string-handling syntax instead of a specialized domain language.
-Further, there is no single data format.
 
@@
 
-Several RDBMSes (\eg MySQL, Oracle and Postgres) provide row-level functions for parsing formatted dates, but our analysis of user data showed a 15\% syntax error rate using these functions.
+Several RDBMSes (\eg MySQL, Oracle and Postgres) provide row-level functions for parsing formatted dates, but a simple review of user attempts to use this functionality in published analyses showed a $15\%$ syntax error rate when using these functions unassisted.
 And even with perfect syntax, the user still has to interrupt her flow to learn a formatting syntax -- one that she will likely forget after this specific problem has been solved.
@@
 Both algorithms are built based on classical machine learning algorithms that learn from pattern recognition to make data-driven predictions.
-We developed two algorithms because we were unaware of any previous work on this problem and we needed to be able to cross-validate our results on a large testing corpus.
+We developed two algorithms because we were unaware of any previous work on this problem and we needed to be able to cross-validate our results on a large test corpus.
 With both approaches, the user can simply specify that the column is a date, and the visualization system can respond quickly and accurately enough to avoid interrupting the user's flow.
@@
 
-Specifically, our contributions include:
 \begin{itemize}

### 2016-05-02 cikm2016/mdl.tex
More text trimming
Remove excess verbiage
Down to 10 pages!

@@
 \item A required inclusion function to test for membership in the domain (\functionname{match}). 
-For \textit{e.g.}, the domain \functionname{<Digits>} returns true for \functionname{"123"} 
-and false for \functionname{"abc"}.
+(\eg the domain \functionname{<Digits>} returns true for \functionname{"123"} 
+and false for \functionname{"abc"}.)
 \item An optional function to compute the number of values in the domain with a given length (\functionname{cardinality}).
-For \textit{e.g.}, the domain for three letter month names (\functionname{MMM}) would return $12$ for the length $3$ and $0$ otherwise.
+(\eg the domain for three letter month names (\functionname{MMM}) would return $12$ for the length $3$ and $0$ otherwise.)
 \item An optional function to update statistics for the domain based on a given value (\functionname{updateStatistics}).
-For \textit{e.g.}, the domain \functionname{MMM} might keep a histogram of how frequently each month was encountered.
+(\eg the domain \functionname{MMM} might keep a histogram of how frequently each month was encountered.)
 \item An optional function to prevent consideration of a domain that is redundant (\functionname{isRedundantAfter}).
-For \textit{e.g.}, the domain \functionname{<Digits>} would be redundant after itself 
-(\ie there is no difference between a single run of digits and two adjacent runs.)
+(\eg  the domain \functionname{<Digits>} would be redundant after itself because there is no difference between a single run of digits and two adjacent runs.)
 \end{itemize}
 
-In our approach, we implement all of these functions, but with significant changes to the last one to increase level of pruning applied during
-domain enumeration. These changes are described below.
+In our approach, we implement all of these functions, but with significant changes to the last one (described below) to increase the level of pruning applied during
+domain enumeration.
 
@@
 
-We also included some enumerated domains for handling constant strings and some simple regular expression domains for delimiters such as spaces, punctuation or alphabetic characters.
 We found that the inclusion of arbitrary numeric domains caused the run time to grow exponentially as the number of possible matches could not be pruned intelligently. 
-This restriction extends to domains that can contain arbitrary digit sequences (such as any).
-Because of this restriction, the algorithm cannot extract non-date numeric fields.
+This restriction extends to domains that can contain arbitrary digit sequences (such as \functionname{<Any>}). Because of this restriction, the algorithm cannot extract non-date numeric fields.
 
-\subsection{Redundancy Extensions}
+\subsection{Redundancy Extensions} \label{subsec:mdl:redundancy}
 A difficulty in using this kind of structure extraction is that the algorithm for enumerating structures is exponential in the number of domains.
 This is especially true in the date format problem because there are identical domains (\eg months and meridian hours), nearly identical domains (\eg days and hours) and there are often no field delimiters (\eg \texttt{2012Mar06134427}).
-To handle this, we extend the existing pruning API with two other sets of domain identifiers:
+To handle this, we extend the original redundancy rules with two other sets of domain identifiers:
 \begin{itemize}
@@
 
-Date domains are also often uniform in that adding more characters to a mismatch will not help.
-For example, a $2$-digit day domain that does not match a $1$-letter substring will not be able to generate a match by adding more characters.
+Date domains are also often \textit{uniform} in that adding more characters to a mismatch will not help.
+\eg a $2$-digit day domain that does not match a $1$-letter substring will not be able to generate a match by adding more characters.
 
@@
 We have also identified two opportunities for parallel computation during structure enumeration.
-
-The first computationally expensive operation is the enumeration of structures for a given sample.
-To parallelize this step, each thread is given a subset of the samples and independently produces a set of candidate domains.
-When all threads have completed, the duplicates are removed to produce a single list of candidate structures.
-
-The second computationally expensive operation is the evaluation of each generated structure over the entire sample set.
-This includes computation of the MDL, recording of domain statistics and parameterizing the structure.
-These tasks are simple to parallelize, because there are no overlaps between the data for each structure.
+Enumerating the matching structures for a single sample is computationally expensive, but the samples themselves are independent. 
+Partitioning the samples allows each thread to produce a set of candidate structures, which can eventually be merged to produce a single candidate list.
+Then once generated, the evaluation of each independent structure (computation of the MDL, recording of domain statistics and parameterizing the structure) can be conducted in parallel.
 
@@
 For example, if a set of dates contains a constant month string (\eg all values are in September) it is important to keep track of the month name domain.
-Consequently, when we parameterize a constant generic \texttt{<Word>} domain, we tag it with the date part domain that it matches (if any).
+Consequently, when we parameterize a constant generic \functionname{<Word>} domain, we tag it with any date part domain that it matched. 
 We then need to apply an additional pruning step to remove any structures that also found an equivalent domain (\eg two-digit month). 
-These rules are equivalent to the context-based redundancy rules above, but have to be applied again after parameterization of generics.
+These rules are equivalent to the context-based redundancy rules in Section~\ref{subsec:mdl:redundancy}, but have to be applied again after the parameterization of generics.
 
@@
 \item Two digit years require special handling.
-In particular, they cannot appear adjacent to a two-digit field if the structure contains punctuation.
+In particular, they cannot appear adjacent to a two-digit field if the structure contains any punctuation.
 (This can come up in some small early 21st century year domains where a two-digit year can masquerade as almost any numeric field \eg $08$ for $2008$.)
 \end{itemize}
-
-These global pruning rules are simple and intuitive, but are essential to further reducing the search space.
 
@@
 
-We have built this system for the Gregorian calendar only as our visualization system does not support non-Gregorian calendars and we have little evidence of other calendars (\eg Hebrew, Islamic Civil) being used for analytics.
+We have built this system for the Gregorian calendar as we have little evidence of other calendars (\eg Hebrew, Islamic Civil) being used for analytics. 
 ICU supports non-Gregorian calendars and we expect the algorithm could be extended to them as well if needed.

### 2016-05-02 cikm2016/cikm_dateparse.tex
Tightened the spacing in bullet points.
More tightening and commented out organization and acknowledgement sections.
First pass at tightening
Move some stuff around, pruned excessive verbiage, and tried to make it
read less like a white paper or documentation.
Add cross checking to abstract

@@
 \usepackage{tikz}
+\usepackage{enumitem}
 \captionsetup[table]{skip=10pt}
 
+\newenvironment{tight_itemize}{\begin{itemize} \itemsep
+-2pt}{\end{itemize}}
 %
@@
 %
-\def\eg{\textit{e.g.} }
-\def\ie{\textit{i.e.} }
+\def\eg{\textit{e.g.\@} }
+\def\ie{\textit{i.e.\@} }
+\def\cf{\textit{cf.\@} }
+\def\etal{\textit{et al.\@} }
+\def\naive{na\"ive}
+\def\Naive{Na\"ive}
 \def\dateparse{\mbox{\texttt{DATEPARSE}}}
@@
 
-% tightening the whitespace for itemize
-\newenvironment{tight_itemize}{\begin{itemize} \itemsep
--2pt}{\end{itemize}}
+% Alter some LaTeX defaults for better treatment of figures:
+% See p.105 of "TeX Unbound" for suggested values.
+% See pp. 199-200 of Lamport's "LaTeX" book for details.
+% General parameters, for ALL pages:
+\renewcommand{\topfraction}{0.9}    % max fraction of floats at top
+\renewcommand{\bottomfraction}{0.8} % max fraction of floats at bottom
+% Parameters for TEXT pages (not float pages):
+\setcounter{topnumber}{2}
+\setcounter{bottomnumber}{2}
+\setcounter{totalnumber}{2} % 2 may work better
+\setcounter{dbltopnumber}{1} % for 2-column pages
+\renewcommand{\dbltopfraction}{0.9} % fit big float above 2-col. text
+\renewcommand{\textfraction}{0.05}  % allow minimal text w. figs
+% Parameters for FLOAT pages (not text pages):
+\renewcommand{\floatpagefraction}{0.8}  % require fuller float pages
+% N.B.: floatpagefraction MUST be less than topfraction !!
+\renewcommand{\dblfloatpagefraction}{0.8}   % require fuller float pages
 
@@
 This has led visualization environments to include simple data preparation functions such as scalar parsing, pattern matching and categorical binning.
-One of the most common scalar parsing tasks is extracting date and time data from string representations.
+One common scalar parsing tasks is extracting date and time data from string representations.
 Several relational database systems include date parsing ``mini-languages'' to cover the wide range of possible formats, but analysis of user data from one visualization system shows that the parsing language syntax is difficult for users to master without disturbing their analytic flow.
@@
 In this paper, we present two algorithms for automatically deriving date formats from a column of data with minimal user disruption, one based on minimal entropy representations and another based on natural language processing techniques.
-Both have accuracies of over 95\% on a large corpus of date columns extracted from an online data repository and are in substantial agreement with each other.
-One approach is also fast enough to produce results within the user's perceptual threshold, making it suitable for interactive work.
+Both have accuracies of over 95\% on a large corpus of date columns extracted from an online data repository, and one is also fast enough to produce results within the user's perceptual threshold.
+Moreover, we were able to avoid prohibitively expensive manual verification by using the algorithms to cross-check each other at scale.
 \end{abstract}
@@
 
-\section{Introduction}
+\section{Introduction} \label{sec:intro}
 \input{intro}
 
-\section{Background}
+\section{Background} \label{sec:background}
 \input{background}
 
-\section{Minimal Descriptive Length}
+\section{Minimal Descriptive Length} \label{sec:mdl}
 \input{mdl}
 
-\section{Natural Language Processing}
+\section{Natural Language Processing} \label{sec:nlp}
 \input{nlp}
 
-\section{Experiments}
+\section{Evaluation} \label{sec:experiments}
 \input{testing}
 
-\section{Related Work}
-\input{relatedwork}
-
-\section{Future Work}
+\section{Future Work} \label{sec:future}
 \input{futurework}
 
-\section{Conclusion}
+\section{Conclusion} \label{sec:conclusion}
 \input{conclusion}
@@
 %ACKNOWLEDGMENTS are optional
-\section{Acknowledgments}
-\input{credits}
+%\section{Acknowledgments}
+%\input{credits}
 

### 2016-05-03 cikm2016/cikm_dateparse.tex
Minor edits in abstract - typo and breaking up a very long sentence.

@@
 This has led visualization environments to include simple data preparation functions such as scalar parsing, pattern matching and categorical binning.
-One common scalar parsing tasks is extracting date and time data from string representations.
+One common scalar parsing task is extracting date and time data from string representations.
 Several relational database systems include date parsing ``mini-languages'' to cover the wide range of possible formats, but analysis of user data from one visualization system shows that the parsing language syntax is difficult for users to master without disturbing their analytic flow.
@@
 In this paper, we present two algorithms for automatically deriving date formats from a column of data with minimal user disruption, one based on minimal entropy representations and another based on natural language processing techniques.
-Both have accuracies of over 95\% on a large corpus of date columns extracted from an online data repository, and one is also fast enough to produce results within the user's perceptual threshold.
+Both have accuracies of over 95\% on a large corpus of date columns extracted from an online data repository.
+One of the methods is also fast enough to produce results within the user's perceptual threshold.
 Moreover, we were able to avoid prohibitively expensive manual verification by using the algorithms to cross-check each other at scale.

### 2016-05-03 cikm2016/cikm_dateparse.tex
some more edits and comments.

@@
 One common scalar parsing task is extracting date and time data from string representations.
-Several relational database systems include date parsing ``mini-languages'' to cover the wide range of possible formats, but analysis of user data from one visualization system shows that the parsing language syntax is difficult for users to master without disturbing their analytic flow.
+Several relational database management systems (RDBMS) include date parsing ``mini-languages'' to cover the wide range of possible formats, but analysis of user data from one visualization system shows that the parsing language syntax is difficult for users to master without disturbing their analytic flow.
 

### 2016-05-03 cikm2016/intro.tex
some more edits and comments.

@@
 \Naive\ users often solve this problem by converting the integer to a string and performing some locale-dependent string operations before casting the string back to a date.
-Unfortunately, this approach has a number of problems because string operations are notoriously slow compared to scalar operations (typically 10-100x slower in modern RDBMSes).
+Unfortunately, this approach has a number of problems because string operations are notoriously slow compared to scalar operations (typically 10-100x slower in modern RDBMSs).
 Default parsing of date formats is locale-dependent, and may not work when the parsing expression is shared across an international organization (\eg between the US and European offices).
@@
 
-In Section~\ref{sec:experiments} we show that there are \textit{hundreds} of distinct temporal date formats in user data sets (\cf Figure~\ref{fig:M2}).
+In Section~\ref{sec:experiments} we show that there are \textit{hundreds} of distinct temporal date formats in user data sets (Figure~\ref{fig:M2}).
 Some are common, but others can be quite idiosyncratic. 
@@
 
-Several RDBMSes (\eg MySQL, Oracle and Postgres) provide row-level functions for parsing formatted dates, but a simple review of user attempts to use this functionality in published analyses showed a $15\%$ syntax error rate when using these functions unassisted.
+Several RDBMSs (\eg MySQL, Oracle and Postgres) provide row-level functions for parsing formatted dates, but a simple review of user attempts to use this functionality in published analyses showed a $15\%$ syntax error rate when using these functions unassisted.
 And even with perfect syntax, the user still has to interrupt her flow to learn a formatting syntax -- one that she will likely forget after this specific problem has been solved.
@@
 
+\vidya{should this be its own section?}
 \subsection{Related Work}
@@
 
+\vidya{should this be its own section?}
 \subsection{Contributions}

### 2016-05-03 cikm2016/background.tex
small RDBMS fix.

@@
 Other syntaxes can either be generated directly from the algorithms, or by translating the ICU syntax into into the desired syntax.
-(In fact, this latter approach is how our visualization system internally generates formats for RDBMSes that have non-ICU syntax.)
+(In fact, this latter approach is how our visualization system internally generates formats for RDBMSs that have non-ICU syntax.)
 %\vidya{do we need the whole table here?}
@@
 The date format generated by the algorithms is presumed to be a string argument to a scalar function called \dateparse, which converts a string value to a date value using the format string.
-Several examples of this kind of function can be found in RDBMSes such as MySQL's \functionname{STR_TO_DATE}, Oracle's \functionname{TO_TIMESTAMP} and Postgres' \functionname{TO_TIMESTAMP}.
+Several examples of this kind of function can be found in RDBMSs such as MySQL's \functionname{STR_TO_DATE}, Oracle's \functionname{TO_TIMESTAMP} and Postgres' \functionname{TO_TIMESTAMP}.
 There are also programming language library implementations such as Python's \functionname{strptime} and ICU's own \functionname{DateFormat::parse}.

### 2016-05-03 cikm2016/cikm_dateparse.tex
CCS Codes 1

--- 2016-05-03 cikm2016/cikm_dateparse.tex
@@
 One common scalar parsing task is extracting date and time data from string representations.
-Several relational database systems include date parsing ``mini-languages'' to cover the wide range of possible formats, but analysis of user data from one visualization system shows that the parsing language syntax is difficult for users to master without disturbing their analytic flow.
+Several relational database management systems (RDBMSs) include date parsing ``mini-languages'' to cover the wide range of possible formats, but analysis of user data from one visualization system shows that the parsing language syntax is difficult for users to master without disturbing their analytic flow.
 
@@
 
-%Looks like they don't use them.
-%\vidya{Need to figure out the IEEE equivalent for categories and terms}
-%D.3.3 [Data Processing]: Data Cleaning, Natural Language Processing.
-% A category with the (minimum) three required fields
-%\category{D.3.3}{Data Processing}{Data Cleaning, Natural Language Processing}
-%A category including the fourth, optional field follows...
-%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
+%
+% The code below should be generated by the tool at
+% http://dl.acm.org/ccs.cfm
+% Please copy and paste the code instead of the example below. 
+%
+\begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10002951.10002952.10003219.10003215</concept_id>
+<concept_desc>Information systems~Extraction, transformation and loading</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10002951.10003227.10003351.10003218</concept_id>
+<concept_desc>Information systems~Data cleaning</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10002951.10003317.10003347.10003352</concept_id>
+<concept_desc>Information systems~Information extraction</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10002951.10002952.10003219.10003223</concept_id>
+<concept_desc>Information systems~Entity resolution</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
 
-%\terms{Algorithms, Performance}
+\ccsdesc[500]{Information systems~Extraction, transformation and loading}
+\ccsdesc[500]{Information systems~Data cleaning}
+\ccsdesc[500]{Information systems~Information extraction}
+\ccsdesc[300]{Information systems~Entity resolution}
+%
+% End generated code
+%
+
+%
+%  Use this command to print the description
+%
+\printccsdesc
+
+%\keywords{Algorithms, Performance}
 

### 2016-05-03 cikm2016/cikm_dateparse.tex
More verbiage trimming

@@
 One common scalar parsing task is extracting date and time data from string representations.
-Several relational database management systems (RDBMSs) include date parsing ``mini-languages'' to cover the wide range of possible formats, but analysis of user data from one visualization system shows that the parsing language syntax is difficult for users to master without disturbing their analytic flow.
+Several relational database management systems (RDBMSs) include date parsing ``mini-languages'' to cover the wide range of possible formats, but analysis of user data from one visualization system shows that the language syntax is difficult for users to master without disturbing analytic flow.
 

### 2016-05-03 cikm2016/intro.tex
CCS Codes 1
More verbiage trimming

@@
-In recent years, there has been a growth of interest in data visualization technologies for human-assisted data analysis using systems such as Polaris~\cite{Stolte:2008} and Spotfire~\cite{Ahlberg:1996}.
+In recent years, there has been growing interest in data visualization technologies for human-assisted data analysis using systems such as Polaris~\cite{Stolte:2008} and Spotfire~\cite{Ahlberg:1996}.
 While computers can provide high-speed and high-volume data processing, humans have both the domain knowledge and the ability to process data in parallel by using their visual systems \cite{Cleveland.McGill1984,Mackinlay:1986}.
@@
 
-Preparing this data for analysis often uses data preparation tools external to the visual analysis environment, which requires the user to break their cognitive flow, launch another tool and reprocess her data before returning to her analysis.
-Recent work~\cite{Morton:2012} has found, however, that the most effective systems allow users to define these calculations as part of the analytic interaction, which enables the user to stay in the flow of analysis.
+Preparing this data for analysis often requires data preparation tools external to the visual analysis environment, which forces the user to break their cognitive flow, launch another tool and reprocess her data before returning to her analysis.
+Recent work~\cite{Morton:2012} has found that the most effective systems allow users to define these calculations as part of the analytic interaction, enabling the user to stay in the flow of analysis.
 
-We have observed that one of the most common data preparation tasks users perform is parsing date strings into scalar date representations.
-For example, an examination of author calculations in a public repository using our
-visualization system shows that there are about as many workbooks containing date parsing calculations ($3.3\%$) as there are
+We have observed that one of the most common such data preparation tasks users perform is parsing date strings into scalar date representations.
+Examination of author calculations in a public repository using our
+visualization system showed that there are about as many workbooks containing date parsing calculations ($3.3\%$) as there are
 integer type conversions of any kind ($3.4\%$).
@@
 
-These types of visualizations are difficult to specify with simple categorical string representations of dates.
+These types of visualizations are easier to specify and manipulate using scalar representations of dates.
 
 \subsection{Parsing Dates}
-A common version of the date parsing problem is converting columns of integers in the form \texttt{yyyyMMdd} to date scalars.
-\Naive\ users often solve this problem by converting the integer to a string and performing some locale-dependent string operations before casting the string back to a date.
-Unfortunately, this approach has a number of problems because string operations are notoriously slow compared to scalar operations (typically 10-100x slower in modern RDBMSs).
+A common form of the date parsing problem is converting columns of integers in the form \texttt{yyyyMMdd} to date scalars.
+\Naive\ users often solve this problem by converting the integer to a string and performing some locale-dependent string operations before casting the string back to a date, but this approach has a number of drawbacks.
+String operations are notoriously slow compared to scalar operations (typically 10-100x slower in modern RDBMSs).
 Default parsing of date formats is locale-dependent, and may not work when the parsing expression is shared across an international organization (\eg between the US and European offices).
-The parsing code is often hard to understand and maintain because it uses a verbose, general-purpose string-handling syntax instead of a specialized domain language.
+Such \textit{ad hoc} parsing code is often hard to understand and maintain because it uses a verbose, general-purpose string-handling syntax instead of a specialized domain language.
 
@@
 This ``long tail'' of formats means that these idiosyncrasies are the rule, not the exception, 
-so a static enumeration of a small set of formats will not produce high reliability.
+so a small, fixed set of formats will not produce high reliability.
 
@@
 
-One solution might have been to design a graphical environment for constructing valid patterns and using visually compelling representations to reduce the cognitive load. 
-Yet this approach provide no guarantee that the result would be correct if the user misunderstood the environment. 
-Moreover, this approach only solves the syntax problem -- the user is still required to switch contexts to solve their problem.
+One approach might have been to design a graphical environment for constructing valid patterns and using visually compelling representations to reduce the cognitive load. 
+But this approach provides no guarantee that the result would be correct if the user misunderstood the environment. 
+Moreover, it only solves the syntax problem -- the user is still required to switch contexts.
 
@@
 
-\vidya{should this be its own section?}
+%\vidya{should this be its own section?}
+%\richard{Ater discussion, we decide it was OK as is.}
 \subsection{Related Work}
@@
 
-\vidya{should this be its own section?}
+%\vidya{should this be its own section?}
 \subsection{Contributions}

### 2016-05-03 cikm2016/cikm_dateparse.tex
Back to 10 pages.

--- 2016-05-03 cikm2016/cikm_dateparse.tex
@@
 \maketitle
+
 \begin{abstract}
-Interactive visual analytics is an effective approach for the timely analysis of big data.
-Users who are already engaged in interactive data analysis prefer to stay in the visualization environment for unavoidable data cleaning and preparation tasks in order to preserve their analytic flow.
+Interactive visual analytics is an effective approach for the timely analysis of data.
+Users who are engaged in interactive data analysis prefer staying in the visualization environment for unavoidable data cleaning and preparation tasks to preserve their analytic flow.
 This has led visualization environments to include simple data preparation functions such as scalar parsing, pattern matching and categorical binning.
-One common scalar parsing task is extracting date and time data from string representations.
+A common scalar parsing task is extracting date and time data from string representations.
 Several relational database management systems (RDBMSs) include date parsing ``mini-languages'' to cover the wide range of possible formats, but analysis of user data from one visualization system shows that the parsing language syntax is difficult for users to master without disturbing their analytic flow.

### 2016-05-03 cikm2016/background.tex
Background tightening
Also renamed the section “Preliminaries”

@@
 We chose this format because we were already using ICU in our system, we had access to the source code, and it provides localized date part data for a large number of languages (\eg month names).
-ICU's format syntax is typical of most syntax languages and provides a large number of date part codes, and a summary of it is reproduced in Table 2.
-The complete syntax is documented at the ICU web site.
+ICU's format syntax is typical of most syntax languages and provides a large number of date part codes, and a representative subset of it is reproduced in Table~\ref{tab:icuformats}.
+The complete syntax is documented at the ICU web site \cite{ICU}.
 Other syntaxes can either be generated directly from the algorithms, or by translating the ICU syntax into into the desired syntax.
-(In fact, this latter approach is how our visualization system internally generates formats for RDBMSs that have non-ICU syntax.)
+(The latter is how our visualization system internally generates formats for RDBMSs that have non-ICU syntax.)
 %\vidya{do we need the whole table here?}
@@
 It has no support for 4-letter month abbreviations (\eg \texttt{Sept.}), ordinal days (\eg \texttt{July 4th}), quarter postfix notation (\eg \texttt{2Q}) and variant meridian markers (\eg \texttt{a.m.}).
-These limitations did not affect the results significantly, and in the future we hope to submit ICU extensions to handle some of these formats.
+These limitations did not affect the results significantly, and in the future we hope to submit handlers for these formats to the ICU project.
 

### 2016-05-03 cikm2016/mdl.tex
API readability.

@@
 \setlength\itemsep{0em}
-\item A required inclusion function to test for membership in the domain (\functionname{match}). 
+\item A required inclusion function \functionname{match} to test for membership in the domain
 (\eg the domain \functionname{<Digits>} returns true for \functionname{"123"} 
 and false for \functionname{"abc"}.)
-\item An optional function to compute the number of values in the domain with a given length (\functionname{cardinality}).
+\item An optional function \functionname{cardinality} to compute the number of values in the domain with a given length
 (\eg the domain for three letter month names (\functionname{MMM}) would return $12$ for the length $3$ and $0$ otherwise.)
-\item An optional function to update statistics for the domain based on a given value (\functionname{updateStatistics}).
+\item An optional function \functionname{updateStatistics} to update statistics for the domain based on a given value
 (\eg the domain \functionname{MMM} might keep a histogram of how frequently each month was encountered.)
-\item An optional function to prevent consideration of a domain that is redundant (\functionname{isRedundantAfter}).
+\item An optional function \functionname{isRedundantAfter} to prevent consideration of a domain that is redundant.
 (\eg  the domain \functionname{<Digits>} would be redundant after itself because there is no difference between a single run of digits and two adjacent runs.)

### 2016-05-03 cikm2016/mdl.tex
More MDL verbiage

@@
 A domain is a set of strings, with a few optional statistical properties.
-They are defined by an interface that includes:
+Each one is defined by an interface that includes:
 \begin{itemize}
@@
 
-In our approach, we implement all of these functions, but with significant changes to the last one (described below) to increase the level of pruning applied during
+In our approach, we implement all of these functions, but with significant changes to \functionname{isRedundantAfter} (described below) to increase the level of pruning applied during
 domain enumeration.
@@
 \subsection{Performance}
-Structure enumeration is computationally expensive, so we added a number of enhancements to the original domain extraction algorithm to keep the run time low enough for interactivity.
+Structure enumeration is computationally expensive, so we have added a number of enhancements to the original domain extraction algorithm to keep the run time low enough for interactivity.
 
@@
 
-Date domains are also often \textit{uniform} in that adding more characters to a mismatch will not help.
+Date domains are also often \emph{uniform} in that adding more characters to a mismatch will not help.
 \eg a $2$-digit day domain that does not match a $1$-letter substring will not be able to generate a match by adding more characters.
@@
 Partitioning the samples allows each thread to produce a set of candidate structures, which can eventually be merged to produce a single candidate list.
-Then once generated, the evaluation of each independent structure (computation of the MDL, recording of domain statistics and parameterizing the structure) can be conducted in parallel.
+Once the list has been generated, the evaluation of each independent structure (computation of the MDL, recording of domain statistics and parameterizing the structure) can be conducted in parallel.
 
@@
 Domain parameterization is an important part of generating compact representations via MDL, but it creates problems for date recognition. 
-For example, if a set of dates contains a constant month string (\eg all values are in September) it is important to keep track of the month name domain.
-Consequently, when we parameterize a constant generic \functionname{<Word>} domain, we tag it with any date part domain that it matched. 
+If (say) a set of dates contains a constant month string (\eg all values are in September) it is important to keep track of the month name domain
+so that the \dateparse\ function will parse the month.
+When we parameterize a constant generic \functionname{<Word>} domain, we therefore tag it with any date part domain that it matched. 
 We then need to apply an additional pruning step to remove any structures that also found an equivalent domain (\eg two-digit month). 
@@
 The pruning rules used for the structure extraction reduce the search space dramatically, but they are also contextual and can only look backwards. 
-The domains also contain a fair amount of ambiguity that requires the application of domain knowledge.
-We therefore found it necessary to add some post-generation global pruning rules:
+The domains also contain a fair amount of ambiguity that requires the application of problem space knowledge. 
+This made it necessary to add some post-generation global pruning rules:
 \begin{itemize}
@@
 \item The set of date parts cannot contain place value gaps (\eg structures that have year and day without month are removed.)
-\item Similarly, the set of time parts cannot contain place value gaps and must also be in place value order (times are never written in orders such as \texttt{mhs}.)
+\item Similarly, the set of time parts cannot contain place value gaps and must also be in place value order (times are never written in orders such as \functionname{mhs}.)
 \item The existence of time parts cannot make dates incomplete (\eg patterns like year-day-hour are removed.)
@@
 
-\subsection{Locale Sensitivity}
-Providing an acceptable international user experience requires correctly handling the locale of the column text.
-Accordingly, at the start of the structure extraction we use the column locale to create a set of domains containing locale-sensitive strings such as month names.
+\subsection{Locale}
+Providing an acceptable international user experience requires correct handling of the column locale. 
+Accordingly, at the start of the structure extraction we use the locale to create a set of domains containing locale-sensitive strings such as month names.
 We also use the locale to map these strings to upper- and lower-case in addition to the ICU mixed case strings.
-This enables us to accurately compute MDL statistics without having to map the candidate strings at runtime (which would be slow).
+This enables us to accurately compute MDL statistics without having to map the input strings at runtime (which would be slow).
 
 Knowing the locale of a string is not always helpful.
-In our data set, we found numerous cases where the locale was specified (\eg Sweden) but the data was actually in English.
+In our test data, we found numerous cases where the locale was specified (\eg Sweden) but the data was actually in English.
 Accordingly, we test both locales and rank the combined results.
@@
 \setlength\itemsep{0em}
-\item Because we have a set of samples, we can apply the candidate format to the strings to see how well it performs. Formats with fewer parse errors are preferred.
+\item Because we have a set of samples, we can apply each format to the samples to see how well it performs. Formats with fewer parse errors are preferred.
 \item Date parts can be considered a place-value system, so we prefer ``more significant'' components (\eg month-day-year over hour-minute-second).
@@
 The sample set may have missed an example where this could be important.
-For example, the month name for "September" is the same in both German and English, but ``October" and ``Oktober" are different.
+For example, the month name for "September" is the same in both English and German, but ``October" and ``Oktober" are different.
 \item If the format has an ambiguous date order (\eg all days are less than 12), then prefer the default date order of the locale.

### 2016-05-03 cikm2016/testing.tex
Fix rmgw reviewer 2 Minor comments.
Evaluation tidying

@@
 The analysis speed is the average time needed per sample for structure extraction. 
-At 2.5ms, this is well below most human perceptual thresholds for a set of 32 samples, so any latency in command execution would be restricted to the ability of the underlying database to provide the samples for analysis in a timely manner.
+At $2.2ms$, this is well below most human perceptual thresholds for a set of $32$ samples, so any latency in command execution would be restricted to the ability of the underlying database to provide the samples for analysis in a timely manner.
 A columnar database (such as the one underlying our visualization system) can often supply such domains without a full table scan, further improving responsiveness.
@@
 \caption{MDL Error Rate}
-\label{fig:M2}
+\label{fig:mdlerror}
 \end{figure}
 
-The error rate reflects the fact that only about 40\% of the files have an associated format that parses the non-null values without error.
-To examine the error rate in greater detail, we turn now to Figure~\ref{fig:M2}.
+The error rate reflects the fact that only about $40\%$ of the files have an associated format that parses the non-null values without error.
+To examine the error rate in greater detail, we turn now to Figure~\ref{fig:mdlerror}.
 
-The two horizontal reference lines in Figure~\ref{fig:M2} show that we found 744 distinct formats that parsed the $13,424$ associated files with no error. 
+The two horizontal reference lines in Figure~\ref{fig:mdlerror} show that we found 744 distinct formats that parsed the $13,424$ associated files with no error. 
 This is a remarkable number of distinct formats and underscores the need for this kind of algorithm. 
@@
 What do these formats look like?
-Figure~\ref{fig:M3} shows a histogram of the $25$ most common formats containing a year format code at the $5\%$ error threshold, color-coded by error rate.
+Figure~\ref{fig:mdloutput} shows a histogram of the $25$ most common formats containing a year format code at the $5\%$ error threshold, color-coded by error rate.
 (A sample value is provided to the right of each bar for illustrative purposes.) The formats have also been filtered to files with at least $5$ samples.
@@
 \caption{MDL Output}
-\label{fig:M3}
+\label{fig:mdloutput}
 \end{figure}
@@
 Some of the dates are clearly just numbers, but our approach is to assume that when the user tells us that the column contains dates we should find the best fit. 
-The samples include dates from a wide range of historical sources (\eg Roman pottery dates) so we have elected to defer the date identification task to the user.
+The samples include dates from a wide range of historical sources (\eg Roman pottery dates) so we have elected to defer the semantic identification task to the user.
 
@@
 Figure \ref{fig:NLP1} shows a histogram of the most common formats identified containing \texttt{Year}.
-There are some variations compared to Figure \ref{fig:M3} including the fact that we are not filtering the results by any error rate in the NLP parsed results.
+There are some variations compared to Figure \ref{fig:mdloutput} including the fact that we are not filtering the results by any error rate in the NLP parsed results.
 

### 2016-05-03 cikm2016/futurework.tex
Future Work tidying

@@
 
-In this paper, we have considered the parsing of strings, but dates are often formatted as integers (\eg \texttt{201507016}).
+In this paper, we have considered the parsing of strings, but dates are often formatted as integers (\eg \functionname{201507016}). 
 It is significantly faster to decompose integers into date parts using arithmetic operations (\eg \texttt{mod} and \texttt{div}) than by using locale-sensitive string parsing functions. 
-Timestamp preparation from numeric representations is another form of preparation which we would like to automate.
+Timestamp preparation from numeric representations is a related task that we would also like to automate.
 

### 2016-05-03 cikm2016/intro.tex
minor typo fix in intro.

@@
 Preparing this data for analysis often requires data preparation tools external to the visual analysis environment, which forces the user to break their cognitive flow, launch another tool and reprocess her data before returning to her analysis.
-Recent work~\cite{Morton:2012} has found that the most effective systems allow users to define these calculations as part of the analytic interaction, enabling the user to stay in the flow of analysis.
+Recent work~\cite{Morton:2012} found that the most effective systems allow users to define these calculations as part of the analytic interaction, enabling the user to stay in the flow of analysis.
 

### 2016-05-03 cikm2016/intro.tex
Trim scalar dates
And make it less random and confusing!

@@
 
-The SQL-99 standard defines three temporal scalar types: \texttt{DATE}, \texttt{TIMESTAMP} and \texttt{TIME}, which can be further qualified as either \texttt{WITH} or \texttt{WITHOUT} \texttt{TIME ZONE} (the \texttt{WITHOUT} form being more common.) They are typically implemented as fixed-point types containing an offset from some epoch (\eg Julian Days.) This makes them compact to store using columnar compression techniques such as those in C-Store~\cite{Stonebraker:2005} and MonetDB/X100~\cite{Zukowski:2006}, and further allows some temporal operations to be implemented efficiently using simple arithmetic.
+The SQL-99 standard defines three temporal scalar types: \functionname{DATE}, \functionname{TIMESTAMP} and \functionname{TIME}, 
+which can be further qualified as either \functionname{WITH} or \functionname{WITHOUT} \functionname{TIME ZONE} (the \functionname{WITHOUT} form being more common.) 
+These types are typically implemented as fixed-point offsets from some epoch (\eg Julian Days.) 
+This makes them compact to store using columnar compression techniques such as those in C-Store~\cite{Stonebraker:2005} and MonetDB/X100~\cite{Zukowski:2006}, 
+and further allows some temporal operations to be implemented efficiently using simple arithmetic. 
 Thus from the RDBMS perspective, representing dates in scalar form provides benefits for users, both in terms of analytic richness and query performance.
 
-From an analytic perspective, date types are dimensional (\ie independent variables) and can be used as either \textit{categorical} (simply ordered) or \textit{quantitative} (ordered with a distance metric) fields.
-Categorical dates have a natural hierarchy associated with them generated by calendar \textit{binning}.
-The visualization in Figure \ref{fig:I1} shows an example of a bar chart employing binned categorical dates in a year/quarter hierarchy.
+From an analytic perspective, date types are dimensional (\ie independent variables) and can be used as either \emph{categorical} (simply ordered) or \emph{quantitative} (ordered with a distance metric) fields.
+Categorical dates have a natural hierarchy associated with them generated by calendar \emph{binning}, which is much easier to specify and compute with a scalar representation.
+Figure \ref{fig:I1} shows an example of binned categorical dates in a year/quarter hierarchy using a bar chart.
 
@@
 Quantitative dates are typically used for time series on an axis that maps the underlying distance measure to display pixel distance. 
-The quantitative analog to categorical binning is \textit{truncation} whereby all the dates in a bin are represented by the first date in the bin.
-Truncation accurately preserves the distance semantics while enabling roll-up to coarser levels of detail.
-The visualization in Figure \ref{fig:I2} shows the same sales data in a quantitative time series rolled up to the quarter level.
+Figure \ref{fig:I2} shows the same sales data in a quantitative time series rolled up to the quarter level.
 
@@
 
-These types of visualizations are easier to specify and manipulate using scalar representations of dates.
+Both of these examples are much easier for users to specify and manipulate when the data is modeled by scalar dates.
 
@@
 
-In Section~\ref{sec:experiments} we show that there are \textit{hundreds} of distinct temporal date formats in user data sets (Figure~\ref{fig:M2}).
+In Section~\ref{sec:experiments} we show that there are \emph{hundreds} of distinct temporal date formats in user data sets (Figure~\ref{fig:M2}).
 Some are common, but others can be quite idiosyncratic. 

### 2016-05-03 cikm2016/nlp.tex
Abridged month portion in NLP section.

@@
 Months usually alternate between lengths of $30$ and $31$ days.
-We use $x \Mod 2$ to get an alternating pattern of 1 and 0, and then add the constant base number of days, $\texttt{Day} = 30 + x \Mod 2$, where $x \in [1..12]$ months.
+We use $x \Mod 2$ to get an alternating pattern of 1 and 0, and then add the constant base number of days, $\texttt{Day} = 30 + (x + 1) \Mod 2$, where $x \in [1..12]$ months.
 
-Except for \texttt{February}, Equation $1$ addresses \texttt{January} and \texttt{March} through \texttt{July}.
-After \texttt{July}, the pattern should skip one, and the rest of the months should follow the alternating pattern inversely.
-To obtain an inverse pattern of alternating 0 and 1, we add 1 to the dividend:
-$\texttt{Day} = 30 + (x + 1) \Mod 2$.
-
-In Equation $2$ the number of days for \texttt{August} through \texttt{December} is correct ($x \in [8..12]$), but not for the remaining months.
+In the above equation, the number of days for \texttt{August} through \texttt{December} is correct ($x \in [8..12]$), but not for the remaining months.
 We hence introduce a bit-masking function so that the equation is equal to 1 over the desired domain and 0 otherwise.
-Multiplying a term by this expression will result in the term being cancelled out outside its domain.
-To mask the latter piece of our function, we need an expression equal to $1$ where $8 \le x\le 12$.
-Floor division by 8 works well, since $x < 16$.
-Substituting this expression in the $x + 1$ dividend of Equation $2$, we can invert the pattern using our mask: $\texttt{Day} = 30 + (x + \floor{\frac{x}{8}}) \Mod 2$.
+Substituting this expression in the $x + 1$ dividend of the equation, we can invert the pattern using our mask: $\texttt{Day} = 30 + (x + \floor{\frac{x}{8}}) \Mod 2$.
 

### 2016-05-03 cikm2016/nlp.tex
more text tightening for months part.

@@
 We use $x \Mod 2$ to get an alternating pattern of 1 and 0, and then add the constant base number of days, $\texttt{Day} = 30 + (x + 1) \Mod 2$, where $x \in [1..12]$ months.
-
-In the above equation, the number of days for \texttt{August} through \texttt{December} is correct ($x \in [8..12]$), but not for the remaining months.
-We hence introduce a bit-masking function so that the equation is equal to 1 over the desired domain and 0 otherwise.
-Substituting this expression in the $x + 1$ dividend of the equation, we can invert the pattern using our mask: $\texttt{Day} = 30 + (x + \floor{\frac{x}{8}}) \Mod 2$.
+We then add a bit-masking function to the equation to correctly account for the number of days for \texttt{August} through \texttt{December} ($x \in [8..12]$): $\texttt{Day} = 30 + (x + \floor{\frac{x}{8}}) \Mod 2$.
 

### 2016-05-09 cikm2016/nlp.tex
Added intro para to nlp section.

@@
-\vidya{need to add a short intro here}
+The second algorithm is a natural language approach to parsing date and time formats.
+While regular expressions can be used to match strings to known formats, we chose a grammar approach as a generalization of regular expressions.
+Such an approach provides for greater expressibility and less rigidity, without the expectation of the input string to exactly match against one or more of the known formats.
+We describe a particular form of grammar called a context-free grammar approach as well as extensions to the rules of the grammar that are inherent to date and time properties such as grammar variants, constraints, and assigning probabilistic weighting through a supervised learning approach.
 
@@
 
-%<Day>     ::= dd (where $dd \in [01-31]$, depending on month/year);
+<Day>     ::= dd (where $dd \in [01-31]$, depending on month/year);
 
-%<Month> := <MonthWord> | <MonthNumber>;
+<Month> := <MonthWord> | <MonthNumber>;
 
-%<MonthWord> := `January' | `February' | `March' | `April' | `May' | `June' | `July' | `August' | `September' | `October' | `November' | `December';
-%
-%<MonthNumber> := dd (where $dd \in [01-12]$);
-%
+<MonthWord> := `January' | `February' | `March' | `April' | `May' | `June' | `July' | `August' | `September' | `October' | `November' | `December';
+
+<MonthNumber> := dd (where $dd \in [01-12]$);
+
 %<DayOfWeek> ::= `Monday' | `Tuesday' | `Wednesday' | `Thursday' | `Friday' | `Saturday' | `Sunday';
-%
+
 %<Hour> ::= <TwelveHour> | <TwentyFourHour>;
-%
+
 %<Quarter> ::= `Quarter' dd (where $dd \in [01 - 04]$);
-%
+
 %<TwoYear> ::= dd;
-%
+
 %<FourYear> := dddd;
-%
+
 %<TwelveHour> ::= dd (where $dd \in [00-12]$);
-%
+
 %<TwentyFourHour> ::= dd (where $dd \in [00 -23]$);
-%
+
 %<AMPM> := `a.m.' | `p.m.';
-%
+
 %<d> ::= `0' | `1' | `2' | `3' | `4' | `5' | `6' | `7' | `8' | `9';

### 2016-05-09 cikm2016/cikm_dateparse.tex
Background tightening
Also renamed the section “Preliminaries”
Minor abstract edit.

@@
 Interactive visual analytics is an effective approach for the timely analysis of data.
-Users who are engaged in interactive data analysis prefer staying in the visualization environment for unavoidable data cleaning and preparation tasks to preserve their analytic flow.
+Users who are already engaged in interactive data analysis prefer staying in the visualization environment for unavoidable data cleaning and preparation tasks to preserve their analytic flow.
 This has led visualization environments to include simple data preparation functions such as scalar parsing, pattern matching and categorical binning.
-A common scalar parsing task is extracting date and time data from string representations.
+One common scalar parsing task is extracting date and time data from string representations.
 Several relational database management systems (RDBMSs) include date parsing ``mini-languages'' to cover the wide range of possible formats, but analysis of user data from one visualization system shows that the parsing language syntax is difficult for users to master without disturbing their analytic flow.
@@
 
-\section{Background} \label{sec:background}
+\section{Preliminaries} \label{sec:background}
 \input{background}

### 2016-05-13 cikm2016/intro.tex
Fix mdlerror reference
minor edits to intro

@@
 which can be further qualified as either \functionname{WITH} or \functionname{WITHOUT} \functionname{TIME ZONE} (the \functionname{WITHOUT} form being more common.) 
-These types are typically implemented as fixed-point offsets from some epoch (\eg Julian Days.) 
+These types are typically implemented as fixed-point offsets from some epoch (\textit{e.g.}, Julian Days.) 
 This makes them compact to store using columnar compression techniques such as those in C-Store~\cite{Stonebraker:2005} and MonetDB/X100~\cite{Zukowski:2006}, 
@@
 String operations are notoriously slow compared to scalar operations (typically 10-100x slower in modern RDBMSs).
-Default parsing of date formats is locale-dependent, and may not work when the parsing expression is shared across an international organization (\eg between the US and European offices).
+Default parsing of date formats is locale-dependent, and may not work when the parsing expression is shared across an international organization (\textit{e.g.}, between the US and European offices).
 Such \textit{ad hoc} parsing code is often hard to understand and maintain because it uses a verbose, general-purpose string-handling syntax instead of a specialized domain language.
 
-In Section~\ref{sec:experiments} we show that there are \emph{hundreds} of distinct temporal date formats in user data sets (Figure~\ref{fig:M2}).
+In Section~\ref{sec:experiments} we show that there are \emph{hundreds} of distinct temporal date formats in user data sets (Figure~\ref{fig:mdlerror}).
 Some are common, but others can be quite idiosyncratic. 
@@
 
-Several RDBMSs (\eg MySQL, Oracle and Postgres) provide row-level functions for parsing formatted dates, but a simple review of user attempts to use this functionality in published analyses showed a $15\%$ syntax error rate when using these functions unassisted.
-And even with perfect syntax, the user still has to interrupt her flow to learn a formatting syntax -- one that she will likely forget after this specific problem has been solved.
+Several RDBMSs (\textit{e.g.}, MySQL, Oracle and Postgres) provide row-level functions for parsing formatted dates, but a simple review of user attempts to use this functionality in published analyses showed a $15\%$ syntax error rate when using these functions unassisted.
+Even with perfect syntax, the user still has to interrupt her flow to learn a formatting syntax -- one that she will likely forget after this specific problem has been solved.
 
 One approach might have been to design a graphical environment for constructing valid patterns and using visually compelling representations to reduce the cognitive load. 
-But this approach provides no guarantee that the result would be correct if the user misunderstood the environment. 
+However, this approach provides no guarantee that the result would be correct if the user misunderstood the environment. 
 Moreover, it only solves the syntax problem -- the user is still required to switch contexts.

### 2016-05-13 cikm2016/intro.tex
minor edits to related work and contributions.

@@
 \setlength\itemsep{0em}
-\item By analyzing an online corpus, we provide evidence that practical date parsing requires the ability to recognize hundreds of formats;
-\item We show how to extend prior work on Minimum Descriptive Length structure extraction to generate a freely available date format domain language with over 95\% accuracy;
+\item By analyzing an online corpus, we provide evidence that practical date parsing requires the ability to recognize hundreds of formats.
+\item We show how to extend prior work on Minimum Descriptive Length structure extraction to generate a freely available date format domain language with over 95\% accuracy.
 \item We describe a second Natural Language Processing technique for generating the same date format domain language with similar accuracy.
 We describe how the basic algorithm is extended to support grammar variants and constraints unique to date formats.
-The parsing algorithm is extended to compute an overall dominant pattern over a data column;
-\item We provide evidence that the development of multiple, independent parsing algorithms provides an effective means of cross-validation on large corpora;
+The parsing algorithm is extended to compute an overall dominant pattern over a data column.
+\item We provide evidence that the development of multiple, independent parsing algorithms provides an effective means of cross-validation on large corpora.
 \item We describe some limitations of this domain language that would improve its utility.

### 2016-05-13 cikm2016/relatedwork.tex
minor edits to related work and contributions.

@@
 
-Various approaches have been described for deriving regular expressions, and a good overview is provided by Li et al.
-in their paper on the ReLIE system for deriving regular expressions given a starting expression provided by a domain expert~\cite{Li:2008}.
+Various approaches have been described for deriving regular expressions, and a good overview is provided by Li \textit{et al.} in their paper on the ReLIE system for deriving regular expressions given a starting expression provided by a domain expert~\cite{Li:2008}.
 The Minimum Descriptive Length technique first described in Rissanen~\cite{Rissanen:1978} was used in~\cite{Raman:2001} to generate regular expressions.

### 2016-05-13 cikm2016/background.tex
minor edits in preliminaries.

@@
 The choice of a date format syntax for generation is somewhat arbitrary, but for the purposes of exposition we will be using the formatting language defined by the ICU open-source project~\cite{ICU}.
-We chose this format because we were already using ICU in our system, we had access to the source code, and it provides localized date part data for a large number of languages (\eg month names).
+We chose this format because we were already using ICU in our system, we had access to the source code, and it provides localized date part data for a large number of languages (\textit{e.g.}, month names).
 ICU's format syntax is typical of most syntax languages and provides a large number of date part codes, and a representative subset of it is reproduced in Table~\ref{tab:icuformats}.
@@
 While fairly extensive, the ICU syntax has a few limitations when working with real-world data.
-It has no support for 4-letter month abbreviations (\eg \texttt{Sept.}), ordinal days (\eg \texttt{July 4th}), quarter postfix notation (\eg \texttt{2Q}) and variant meridian markers (\eg \texttt{a.m.}).
+It has no support for 4-letter month abbreviations (\textit{e.g.}, \texttt{Sept.}), ordinal days (\textit{e.g.}, \texttt{July 4th}), quarter postfix notation (\textit{e.g.}, \texttt{2Q}) and variant meridian markers (\textit{e.g.}, \texttt{a.m.}).
 These limitations did not affect the results significantly, and in the future we hope to submit handlers for these formats to the ICU project.

### 2016-05-13 cikm2016/mdl.tex
More minor edits in the MDL section.

@@
 \item A required inclusion function \functionname{match} to test for membership in the domain
-(\eg the domain \functionname{<Digits>} returns true for \functionname{"123"} 
+(\textit{e.g.}, the domain \functionname{<Digits>} returns true for \functionname{"123"} 
 and false for \functionname{"abc"}.)
 \item An optional function \functionname{cardinality} to compute the number of values in the domain with a given length
-(\eg the domain for three letter month names (\functionname{MMM}) would return $12$ for the length $3$ and $0$ otherwise.)
+(\textit{e.g.}, the domain for three letter month names (\functionname{MMM}) would return $12$ for the length $3$ and $0$ otherwise.)
 \item An optional function \functionname{updateStatistics} to update statistics for the domain based on a given value
-(\eg the domain \functionname{MMM} might keep a histogram of how frequently each month was encountered.)
+(\textit{e.g.}, the domain \functionname{MMM} might keep a histogram of how frequently each month was encountered.)
 \item An optional function \functionname{isRedundantAfter} to prevent consideration of a domain that is redundant.
-(\eg  the domain \functionname{<Digits>} would be redundant after itself because there is no difference between a single run of digits and two adjacent runs.)
+(\textit{e.g.},  the domain \functionname{<Digits>} would be redundant after itself because there is no difference between a single run of digits and two adjacent runs.)
 \end{itemize}
 
-In our approach, we implement all of these functions, but with significant changes to \functionname{isRedundantAfter} (described below) to increase the level of pruning applied during
+In our approach, we implement all of these functions, but with significant changes to \functionname{isRedundantAfter} to increase the level of pruning applied during
 domain enumeration.
@@
 This restriction extends to domains that can contain arbitrary digit sequences (such as \functionname{<Any>}). Because of this restriction, the algorithm cannot extract non-date numeric fields.
-
+%\vspace{2.1mm}
 \subsection{Redundancy Extensions} \label{subsec:mdl:redundancy}
 A difficulty in using this kind of structure extraction is that the algorithm for enumerating structures is exponential in the number of domains.
-This is especially true in the date format problem because there are identical domains (\eg months and meridian hours), nearly identical domains (\eg days and hours) and there are often no field delimiters (\eg \texttt{2012Mar06134427}).
+This is especially true in the date format problem because there are identical domains (\textit{e.g.}, months and meridian hours), nearly identical domains (\textit{e.g.}, days and hours) and there are often no field delimiters (\textit{e.g.}, \texttt{2012Mar06134427}).
 To handle this, we extend the original redundancy rules with two other sets of domain identifiers:
@@
 Date domains are also often \emph{uniform} in that adding more characters to a mismatch will not help.
-\eg a $2$-digit day domain that does not match a $1$-letter substring will not be able to generate a match by adding more characters.
+For example, a $2$-digit day domain that does not match a $1$-letter substring will not be able to generate a match by adding more characters.
 
@@
 Domain parameterization is an important part of generating compact representations via MDL, but it creates problems for date recognition. 
-If (say) a set of dates contains a constant month string (\eg all values are in September) it is important to keep track of the month name domain
+If (say) a set of dates contains a constant month string (\textit{e.g.}, all values are in September) it is important to keep track of the month name domain
 so that the \dateparse\ function will parse the month.
 When we parameterize a constant generic \functionname{<Word>} domain, we therefore tag it with any date part domain that it matched. 
-We then need to apply an additional pruning step to remove any structures that also found an equivalent domain (\eg two-digit month). 
+We then need to apply an additional pruning step to remove any structures that also found an equivalent domain (\textit{e.g.}, two-digit month). 
 These rules are equivalent to the context-based redundancy rules in Section~\ref{subsec:mdl:redundancy}, but have to be applied again after the parameterization of generics.
@@
 \setlength\itemsep{0em}
-\item The set of date parts cannot contain place value gaps (\eg structures that have year and day without month are removed.)
+\item The set of date parts cannot contain place value gaps (\textit{e.g.}, structures that have year and day without month are removed.)
 \item Similarly, the set of time parts cannot contain place value gaps and must also be in place value order (times are never written in orders such as \functionname{mhs}.)
-\item The existence of time parts cannot make dates incomplete (\eg patterns like year-day-hour are removed.)
+\item The existence of time parts cannot make dates incomplete (\textit{e.g.}, patterns like year-day-hour are removed.)
 \item Two digit years require special handling.
 In particular, they cannot appear adjacent to a two-digit field if the structure contains any punctuation.
-(This can come up in some small early 21st century year domains where a two-digit year can masquerade as almost any numeric field \eg $08$ for $2008$.)
+(This can come up in some small early 21st century year domains where a two-digit year can masquerade as almost any numeric field \textit{e.g.}, $08$ for $2008$.)
 \end{itemize}
@@
 Knowing the locale of a string is not always helpful.
-In our test data, we found numerous cases where the locale was specified (\eg Sweden) but the data was actually in English.
+In our test data, we found numerous cases where the locale was specified (\textit{e.g.}, Sweden) but the data was actually in English.
 Accordingly, we test both locales and rank the combined results.
 
-We have built this system for the Gregorian calendar as we have little evidence of other calendars (\eg Hebrew, Islamic Civil) being used for analytics. 
+We have built this system for the Gregorian calendar as we have little evidence of other calendars (\textit{e.g.}, Hebrew, Islamic Civil) being used for analytics. 
 ICU supports non-Gregorian calendars and we expect the algorithm could be extended to them as well if needed.
@@
 \setlength\itemsep{0em}
-\item Because we have a set of samples, we can apply each format to the samples to see how well it performs. Formats with fewer parse errors are preferred.
-\item Date parts can be considered a place-value system, so we prefer ``more significant'' components (\eg month-day-year over hour-minute-second).
+\item We can apply each format to the set of samples we have, to see how well the algorithm performs. Formats with fewer parse errors are preferred.
+\item Date parts can be considered a place-value system, so we prefer ``more significant'' components (\textit{e.g.}, month-day-year over hour-minute-second).
 \item If two formats from different locales give the same results, prefer the original column locale.
 The sample set may have missed an example where this could be important.
-For example, the month name for "September" is the same in both English and German, but ``October" and ``Oktober" are different.
-\item If the format has an ambiguous date order (\eg all days are less than 12), then prefer the default date order of the locale.
+For example, the month name for ``September" is the same in both English and German, but ``October" and ``Oktober" are different.
+\item If the format has an ambiguous date order (\textit{e.g.}, all days are less than 12), then prefer the default date order of the locale.
 Again, the sample set may have missed a counterexample, so this is the best option.

### 2016-05-13 cikm2016/nlp.tex
minor edits to nlp section.

@@
 While regular expressions can be used to match strings to known formats, we chose a grammar approach as a generalization of regular expressions.
-Such an approach provides for greater expressibility and less rigidity, without the expectation of the input string to exactly match against one or more of the known formats.
+Such an approach provides for greater expressibility and less rigidity, without the expectation of the input string to \textit{exactly} match against one or more of the known formats.
 We describe a particular form of grammar called a context-free grammar approach as well as extensions to the rules of the grammar that are inherent to date and time properties such as grammar variants, constraints, and assigning probabilistic weighting through a supervised learning approach.
@@
 Second, it grants total freedom in structuring a grammar to best fit its intended use.
-The power of such a grammar is its expressiveness, and is not constrained by pattern rigidities that are prevalent with regular expressions and text processing~\cite{Grune:1990}.
+The power of such a grammar is its expressiveness, and is not constrained by pattern rigidities that are prevalent in regular expressions and text processing~\cite{Grune:1990}.
 
@@
 We extend the date-time grammar by adding morpho-syntactic variants.
-These variants allows the parser to correct for syntactic errors (\textit{e.g.} usage of whitespace characters and punctuation marks), capitalization errors and abbreviations (\textit{e.g.} \texttt{`Mon'} for \texttt{Monday}).
+These variants allows the parser to correct for syntactic errors (\textit{e.g.}, usage of whitespace characters and punctuation marks), capitalization errors and abbreviations (\textit{e.g.}, \texttt{`Mon'} for \texttt{Monday}).
 We use several various external corpora for such corrections~\cite{nltkcorpora}.
@@
 For this, we define a new symbol in the grammar called \texttt{LeapYear}.
-If an expression containing the month \texttt{`February'} or any such variant (\textit{e.g.} \texttt{`Feb'}, \texttt{`2'}) with the day \texttt{`29'} and a year, would need to resolve the \texttt{Year} symbol to be a \texttt{LeapYear}, defined as $\texttt{Year} \Mod 4 == 0$.
+If an expression containing the month \texttt{February} or any such variant (\textit{e.g.}, \texttt{`Feb'}, \texttt{`2'}) with the day \texttt{`29'} and a year, would need to resolve the \texttt{Year} symbol to be a \texttt{LeapYear}, defined as $\texttt{Year} \Mod 4 == 0$.
 However, this is just an approximation.
@@
 
-The grammar parser's success is often limited by data ambiguity, and incomplete expressions.
+The parser's success is often limited by data ambiguity, and incomplete expressions.
 For example an input string \texttt{`3/7/2005'} could either be interpreted as as the 7th day of March or the 3rd day of July depending on the date format used.

### 2016-05-13 cikm2016/testing.tex
edits to evaluation section.

@@
 Each column was then reduced to a maximum sample set of 32.
-We excluded domain values that either contain the substring \texttt{NULL}, contain no digits and at most one non-whitespace character (\eg \texttt{" / / "}), or any empirically determined common \texttt{NULL} value (\eg \texttt{0000-00-00}, \texttt{NaN}).
+We excluded domain values that either contain the substring \texttt{NULL}, contain no digits and at most one non-whitespace character (\textit{e.g.}, \texttt{" / / "}), or any empirically determined common \texttt{NULL} value (\textit{e.g.}, \texttt{0000-00-00}, \texttt{NaN}).
 %Columns that had no remaining valid samples were discarded.
@@
 After NULL filtering, a common class of date representation (numeric timestamps) remained that was unsuited for parsing via our date format syntax.
-This included Unix epoch timestamps (expressed as second, millisecond or microsecond counts from \texttt{1970-01-01}) and Microsoft Excel timestamps (expressed fractional days since \texttt{1900-01-01}).
+This included Unix epoch timestamps (expressed as second, millisecond or microsecond counts from \texttt{1970-01-01}) and Microsoft Excel time-stamps (expressed as fractional days since \texttt{1900-01-01}).
 A simple test to check that the column was numeric and inside a specific range of values representing recent dates allowed us to tag these columns to avoid analyzing them further (and incidentally to identify them for generating a simple date extraction calculation for the user.)
@@
 The error rate reflects the fact that only about $40\%$ of the files have an associated format that parses the non-null values without error.
-To examine the error rate in greater detail, we turn now to Figure~\ref{fig:mdlerror}.
+To examine the error rate in greater detail, we refer to Figure~\ref{fig:mdlerror}.
 
@@
 Some of the dates are clearly just numbers, but our approach is to assume that when the user tells us that the column contains dates we should find the best fit. 
-The samples include dates from a wide range of historical sources (\eg Roman pottery dates) so we have elected to defer the semantic identification task to the user.
+The samples include dates from a wide range of historical sources (\textit{e.g.}, Roman pottery dates) so we have elected to defer the semantic identification task to the user.
 
@@
 
-The initial average parsing speed to compute the ranked set of probable parse trees is $0.93s$, while the average time taken to compute the overall dominant pattern(s) for the entire column of data is $1.4s$.
+The initial average parsing speed to compute the ranked set of probable parse trees is $0.93s$, and the average time taken to compute the overall dominant pattern(s) for the entire column of data is $1.4s$.
 While the natural language parsing implementation is in Python as the NLTK package is easily configurable, we could expect greater speed-up in parsing performance by employing C/C++ CYK parsing libraries.
 
-Out of the $31,546$ files used for testing, the NLP parser identified a dominant format for $26, 534 (84.11\%)$ where $p >= 0.5$.
+Out of the $31,546$ files used for testing, the NLP parser identified a dominant format for $26, 534$ ($84.11\%$) where $p >= 0.5$.
 $1634$ unique formats were identified.

### 2016-05-13 cikm2016/futurework.tex
minor edits to future work section.

@@
 Many columns contained multiple date formats.
-We would like to recognize this situation and generate predicate-based calculations (\eg if the string matches format 1 then apply format 1 else ...) to increase the accuracy of the results and thereby make the experience even more seamless for the user.
+We would like to recognize this situation and generate predicate-based calculations (\textit{e.g.}, if the string matches format 1 then apply format 1 else ...) to increase the accuracy of the results and thereby make the experience even more seamless for the user.
 Other columns contain date ranges, which we would like to handle by generating multiple calculations, possibly by combining regular expressions with date parsing.
 
-In this paper, we have considered the parsing of strings, but dates are often formatted as integers (\eg \functionname{201507016}). 
-It is significantly faster to decompose integers into date parts using arithmetic operations (\eg \texttt{mod} and \texttt{div}) than by using locale-sensitive string parsing functions. 
+In this paper, we have considered the parsing of strings, but dates are often formatted as integers (\textit{e.g.}, \functionname{201507016}). 
+It is significantly faster to decompose integers into date parts using arithmetic operations (\textit{e.g.}, \texttt{mod} and \texttt{div}) than by using locale-sensitive string parsing functions. 
 Timestamp preparation from numeric representations is a related task that we would also like to automate.
 
-In the course of our research, we have also identified a number of date part variants (\eg ordinal dates, four-letter month abbreviations, alternate meridian markers and postfix quarter syntax) that we would like to commend to the ICU project, along with possible implementations.
+In the course of our research, we have also identified a number of date part variants (\textit{e.g.}, ordinal dates, four-letter month abbreviations, alternate meridian markers and postfix quarter syntax) that we would like to commend to the ICU project, along with possible implementations.

### 2016-05-13 cikm2016/conclusion.tex
CIKM formatting
minor edits to conclusion.

--- 2015-07-06 conclusion.tex
@@
 Both algorithms are accurate enough to be used automatically without user involvement.
-The MDL algorithm is also fast enough to deploy in an interactive environment.
+The MDL algorithm is also fast enough to deploy in an interactive environment, freeing users from the need to interrupt their cognitive flow during analysis in order to learn a formatting language.
 
-While validating the algorithms on a large corpus, we also found that the number of distinct formats in the wild is surprisingly high, and demonstrates the wisdom of including general-purpose date parsing functions in data visualization tools, data cleaning tools and RDBMSes.
-In particular, it is interesting to note that the most prominent open source RDBMSes (\eg MySQL and Postgres) both have a built-in version of \dateparse, possibly reflecting that this is a common need that gets implemented when users are empowered to add to a code base.
+While validating the algorithms on a large corpus, we also found that the number of distinct formats in the wild is surprisingly high, and demonstrates the wisdom of including general-purpose date parsing functions in data visualization tools, data cleaning tools and RDBMSs.
+In particular, it is interesting to note that the most prominent open source RDBMSs (\eg MySQL and Postgres) both have a built-in version of \texttt{DATEPARSE}, possibly reflecting that this is a common need that gets implemented when users are empowered to extend the function library of an RDBMS.

### 2016-05-13 cikm2016/cikm_dateparse.tex
Update CSC tags

@@
 One common scalar parsing task is extracting date and time data from string representations.
-Several relational database management systems (RDBMSs) include date parsing ``mini-languages'' to cover the wide range of possible formats, but analysis of user data from one visualization system shows that the parsing language syntax is difficult for users to master without disturbing their analytic flow.
+Several relational database management systems (RDBMSs) include date parsing ``mini-languages'' to cover the wide range of possible formats, but analysis of user data from one visualization system shows that the parsing language syntax can be difficult for users to master.
 
@@
 <concept>
+<concept_id>10002951.10002952.10003219.10003223</concept_id>
+<concept_desc>Information systems~Entity resolution</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
 <concept_id>10002951.10003227.10003351.10003218</concept_id>
@@
 <concept_desc>Information systems~Information extraction</concept_desc>
-<concept_significance>500</concept_significance>
+<concept_significance>300</concept_significance>
 </concept>
 <concept>
-<concept_id>10002951.10002952.10003219.10003223</concept_id>
-<concept_desc>Information systems~Entity resolution</concept_desc>
+<concept_id>10010147.10010178.10010179.10010182</concept_id>
+<concept_desc>Computing methodologies~Natural language generation</concept_desc>
 <concept_significance>300</concept_significance>
@@
 \ccsdesc[500]{Information systems~Extraction, transformation and loading}
+\ccsdesc[500]{Information systems~Entity resolution}
 \ccsdesc[500]{Information systems~Data cleaning}
-\ccsdesc[500]{Information systems~Information extraction}
-\ccsdesc[300]{Information systems~Entity resolution}
-%
+\ccsdesc[300]{Information systems~Information extraction}
+\ccsdesc[300]{Computing methodologies~Natural language generation}%
 % End generated code

### 2016-05-16 cikm2016/cikm_dateparse.tex
CCS regeneration
Including adhering to the “one major topic” guideline.

@@
 <concept>
+<concept_id>10002951.10003227.10003351.10003218</concept_id>
+<concept_desc>Information systems~Data cleaning</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
 <concept_id>10002951.10002952.10003219.10003215</concept_id>
 <concept_desc>Information systems~Extraction, transformation and loading</concept_desc>
-<concept_significance>500</concept_significance>
+<concept_significance>300</concept_significance>
 </concept>
@@
 <concept_desc>Information systems~Entity resolution</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10002951.10003227.10003351.10003218</concept_id>
-<concept_desc>Information systems~Data cleaning</concept_desc>
-<concept_significance>500</concept_significance>
+<concept_significance>300</concept_significance>
 </concept>
@@
 
-\ccsdesc[500]{Information systems~Extraction, transformation and loading}
-\ccsdesc[500]{Information systems~Entity resolution}
 \ccsdesc[500]{Information systems~Data cleaning}
+\ccsdesc[300]{Information systems~Extraction, transformation and loading}
+\ccsdesc[300]{Information systems~Entity resolution}
 \ccsdesc[300]{Information systems~Information extraction}
 \ccsdesc[300]{Computing methodologies~Natural language generation}%
-% End generated code
-%
 

### 2016-05-16 cikm2016/intro.tex
Fixed orphan lines

@@
 and further allows some temporal operations to be implemented efficiently using simple arithmetic. 
-Thus from the RDBMS perspective, representing dates in scalar form provides benefits for users, both in terms of analytic richness and query performance.
+Thus from the RDBMS perspective, representing dates in scalar form provides benefits for users, both in terms of analytic richness and query performance.\\
 

### 2018-12-26 cikm2016/cikm_dateparse.tex
Expose true identities.
Convert to Tableau copyright and style.
Convert DOI to  Tableau Technical Note
Updated into technical report

@@
 
-%\documentclass{acm_proc_article-sp}
-%\documentclass[conference]{IEEEtran}
-\documentclass{sig-alternate-05-2015}
+\documentclass[10pt,twoside,a4paper,twocolumn]{article}
+\usepackage[left=1.75cm, right=1.75cm, top=2cm]{geometry}
+\setlength\columnsep{20pt}
+
 %
+
 \usepackage{url}
@@
 \usepackage{enumitem}
+\usepackage{hyperref}
 \captionsetup[table]{skip=10pt}
@@
 
+\setlength{\tabcolsep}{4pt}   % set length to -2pt
 %
@@
 
-\title{Does Anybody Really Know What Time It Is?\\
-Automating the Extraction of Date Scalars}
-\input{anonymous}
+%\input{authors}
 
-\maketitle
+%\maketitle
+
+\begin{titlepage}
+   \begin{center}
+       \vspace*{1cm}
+  \includegraphics[width=0.3\textwidth]{figures/tableau}
+     \vspace*{1cm}
+
+        \Large       
+ \textbf{Does Anybody Really Know What Time It Is?\\
+Automating the Extraction of Date Scalars}\\
+ TR \#2018-01
+       \vspace{1cm}
+
+\large
+\textbf{Richard Wesley\\
+Vidya Setlur\\
+Dan Cory} 
+     %  \vfill
+
+       \vspace{2cm}
+ Tableau Software\\
+     1621 North 34th Street\\
+      Seattle, WA 98103\\  
+ \vspace{2cm}
+ \small
+Copyright \copyright 2018\\
+Tableau Software All rights reserved.\\
+The information contained in this document is the exclusive property of Tableau.
+This work is protected under United States copyright law and other international copyright treaties and conventions.No part of this work may be reproduced, used, or transmitted in any form or by any means except as expressly permitted in writing by Tableau.\\
+All requests should be sent to Attention: Tableau Legal \href{mailto:legal@tableau.com}{legal@tableau.com} 
+ \end{center}
+\end{titlepage}
 
@@
 %
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10002951.10003227.10003351.10003218</concept_id>
-<concept_desc>Information systems~Data cleaning</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10002951.10002952.10003219.10003215</concept_id>
-<concept_desc>Information systems~Extraction, transformation and loading</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-<concept>
-<concept_id>10002951.10002952.10003219.10003223</concept_id>
-<concept_desc>Information systems~Entity resolution</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-<concept>
-<concept_id>10002951.10003317.10003347.10003352</concept_id>
-<concept_desc>Information systems~Information extraction</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-<concept>
-<concept_id>10010147.10010178.10010179.10010182</concept_id>
-<concept_desc>Computing methodologies~Natural language generation</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Information systems~Data cleaning}
-\ccsdesc[300]{Information systems~Extraction, transformation and loading}
-\ccsdesc[300]{Information systems~Entity resolution}
-\ccsdesc[300]{Information systems~Information extraction}
-\ccsdesc[300]{Computing methodologies~Natural language generation}%
+%\begin{CCSXML}
+%<ccs2012>
+%<concept>
+%<concept_id>10002951.10003227.10003351.10003218</concept_id>
+%<concept_desc>Information systems~Data cleaning</concept_desc>
+%<concept_significance>500</concept_significance>
+%</concept>
+%<concept>
+%<concept_id>10002951.10002952.10003219.10003215</concept_id>
+%<concept_desc>Information systems~Extraction, transformation and loading</concept_desc>
+%<concept_significance>300</concept_significance>
+%</concept>
+%<concept>
+%<concept_id>10002951.10002952.10003219.10003223</concept_id>
+%<concept_desc>Information systems~Entity resolution</concept_desc>
+%<concept_significance>300</concept_significance>
+%</concept>
+%<concept>
+%<concept_id>10002951.10003317.10003347.10003352</concept_id>
+%<concept_desc>Information systems~Information extraction</concept_desc>
+%<concept_significance>300</concept_significance>
+%</concept>
+%<concept>
+%<concept_id>10010147.10010178.10010179.10010182</concept_id>
+%<concept_desc>Computing methodologies~Natural language generation</concept_desc>
+%<concept_significance>300</concept_significance>
+%</concept>
+%</ccs2012>
+%\end{CCSXML}
+%
+%\ccsdesc[500]{Information systems~Data cleaning}
+%\ccsdesc[300]{Information systems~Extraction, transformation and loading}
+%\ccsdesc[300]{Information systems~Entity resolution}
+%\ccsdesc[300]{Information systems~Information extraction}
+%\ccsdesc[300]{Computing methodologies~Natural language generation}%
 
@@
 %
-\printccsdesc
+%\printccsdesc
 
@@
 % produce the bibliography for the citations in your paper.
-\bibliographystyle{abbrv}
+%\bibliographystyle{abbrv}
 \bibliography{dateparse}  % sigproc.bib is the name of the Bibliography in this case

### 2018-12-26 cikm2016/cikm_dateparse.tex
Added footnote

@@
 \usepackage{hyperref}
+\usepackage[hang,flushmargin]{footmisc}
 \captionsetup[table]{skip=10pt}
@@
 \setlength{\tabcolsep}{4pt}   % set length to -2pt
-%
+
+\newcommand\blfootnote[1]{%
+  \begingroup
+  \renewcommand\thefootnote{}\footnote{#1}%
+  \addtocounter{footnote}{-1}%
+  \endgroup
+}
 \begin{document}
@@
 %\keywords{Algorithms, Performance}
+\blfootnote{\tiny{Copyright \copyright 2018\\
+Tableau Software All rights reserved.\\
+The information contained in this document is the exclusive property of Tableau.
+This work is protected under United States copyright law and other international copyright treaties and conventions.No part of this work may be reproduced, used, or transmitted in any form or by any means except as expressly permitted in writing by Tableau.\\
+All requests should be sent to Attention: Tableau Legal \href{mailto:legal@tableau.com}{legal@tableau.com}}}
 
